### 2024-06-04 manuscript/appendix1.Rmd
added old MS files to change

---
title: |
     Experience shapes individual foraging specialization and success in a virtual predator-prey system: \
     Appendix 1

output: 
  word_document:
    reference_docx: appendix-styles.docx
    fig_caption: yes
csl: ecology.csl
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(data.table)
library(brms)
library(flextable)
library(officer)
library(dplyr)

```

\newpage

```{r tableS1, ft.align = "left"}

# ===========================================================================
# 1. Import the model fit
# ===========================================================================

# path
path <- file.path(here::here("outputs"))

# Import the model
fit <- readRDS(
  file.path(path, "01_outputs_models", "B1_DHMLM-no-outlier.rds")
)

# ===========================================================================
# ===========================================================================

# ===========================================================================
# 2. Fixed effects table
# ===========================================================================

# Prepare the table ---------------------------------------------------------

# Extract the fixed effects posterior samples
b_draws <- data.table(
  as_draws_df(
    fit, variable = "^b_", regex = TRUE
  )
)

b_draws[, c(34:36) := NULL]

# Back-transform the parameters ----------------------------------------------

# Take the exponent of sigma parameters
b_draws[
  , c("b_sigma_speednovice_Intercept",
      "b_sigma_speedinterm_Intercept",
      "b_sigma_speedadvanced_Intercept",
      "b_sigma_preyspeednovice_Intercept",
      "b_sigma_preyspeedinterm_Intercept",
      "b_sigma_preyspeedadvanced_Intercept",
      "b_sigma_speednovice_Zprey_avg_rank",
      "b_sigma_speedinterm_Zprey_avg_rank",
      "b_sigma_speedadvanced_Zprey_avg_rank",
      "b_sigma_preyspeednovice_Zprey_avg_rank",
      "b_sigma_preyspeedinterm_Zprey_avg_rank",
      "b_sigma_preyspeedadvanced_Zprey_avg_rank"
  )
  := lapply(.SD, function(x) {exp(x)}),
  .SDcols = c(
    "b_sigma_speednovice_Intercept",
    "b_sigma_speedinterm_Intercept",
    "b_sigma_speedadvanced_Intercept",
    "b_sigma_preyspeednovice_Intercept",
    "b_sigma_preyspeedinterm_Intercept",
    "b_sigma_preyspeedadvanced_Intercept",
    "b_sigma_speednovice_Zprey_avg_rank",
    "b_sigma_speedinterm_Zprey_avg_rank",
    "b_sigma_speedadvanced_Zprey_avg_rank",
    "b_sigma_preyspeednovice_Zprey_avg_rank",
    "b_sigma_preyspeedinterm_Zprey_avg_rank",
    "b_sigma_preyspeedadvanced_Zprey_avg_rank"
  )
]

# plogis for the success parameters
b_draws[
  , c("b_successnovice_Intercept",
      "b_successinterm_Intercept",
      "b_successadvanced_Intercept",
      "b_successnovice_Zgame_duration",
      "b_successinterm_Zgame_duration",
      "b_successadvanced_Zgame_duration",
      "b_successnovice_Zprey_avg_rank",
      "b_successinterm_Zprey_avg_rank",
      "b_successadvanced_Zprey_avg_rank"
  )
  := lapply(.SD, function(x) {plogis(x)}),
  .SDcols = c(
    "b_successnovice_Intercept",
    "b_successinterm_Intercept",
    "b_successadvanced_Intercept",
    "b_successnovice_Zgame_duration",
    "b_successinterm_Zgame_duration",
    "b_successadvanced_Zgame_duration",
    "b_successnovice_Zprey_avg_rank",
    "b_successinterm_Zprey_avg_rank",
    "b_successadvanced_Zprey_avg_rank"
  )
]

# Reshape the table -----------------------------------------------------------

fixef_tab <- melt(
  b_draws,
  variable.name = "parameter"
)

# Summarize the values (medians + 95% CI) -------------------------------------

# Intervals
lower_interval <- function(x) {
  coda::HPDinterval(as.mcmc(x), 0.95)[1]
}
upper_interval <- function(x) {
  coda::HPDinterval(as.mcmc(x), 0.95)[2]
}

# Summarize values
fixef_tab[, ":=" (median = median(value),
                  lower_ci = lower_interval(value),
                  upper_ci = upper_interval(value)),
          by = .(parameter)]

fixef_tab <- unique(fixef_tab[, .(parameter, median, lower_ci, upper_ci)])
fixef_tab[, c(2:4) := round(fixef_tab[, c(2:4)], digits = 2)]

# Rearrange the table --------------------------------------------------------

# Add a column that specifies the experience level
fixef_tab[parameter %like% "novice", xp_level := "novice"]
fixef_tab[parameter %like% "interm", xp_level := "intermediate"]
fixef_tab[parameter %like% "advanced", xp_level := "advanced"]

# Paste upper and lower ci with estimate
fixef_tab[, estimate := paste(format(median, digits = 2), "(")]
fixef_tab[, estimate := paste(estimate, format(lower_ci, digits = 2), sep = "")]
fixef_tab[, estimate := paste(estimate, ",", sep = "")]
fixef_tab[, estimate := paste(estimate, format(upper_ci, digits = 2), sep = " ")]
fixef_tab[, estimate := paste(estimate, ")", sep = "")]
fixef_tab[, c("median", "lower_ci", "upper_ci") := NULL]

# reorder columns, and then rows based on xp level
setorder(
  fixef_tab,
  cols = -"xp_level"
)

# Create trait column
fixef_tab[, trait := ""][
  parameter %like% "success", trait := "hunting success"][
    parameter %like% "speed", trait := "predator speed"][
      parameter %like% "preyspeed", trait := "prey speed"
      ]

# Rename parameter column
fixef_tab[
  , parameter := rep(
    c(
      rep(c("intercept (mean)", "intercept (sigma)"), 2),
      "intercept (mean)",
      rep(c("prey rank (mean)", "prey rank (sigma)"), 2),
      "match duration (mean)", "prey rank (mean)"
    ),
    3
  )
]

# Reshape
fixef_tab <- dcast(
  fixef_tab,
  trait + parameter ~ xp_level,
  value.var = "estimate"
)

# Reorder columns and rows
fixef_tab <- fixef_tab[, c(1, 2, 5, 4, 3)]
table <- fixef_tab[c(4, 6, 5, 7, 8, 10, 9, 11, 1, 2, 3)]

# Because I can't align the rows properly
# in the flextable object
table[c(2:4, 6:8, 10, 11), 1] <- " "

# ===========================================================================
# ===========================================================================

# ===========================================================================
# 3. Create the table using flextable
# ===========================================================================

# Prepare the table parameters ----------------------------------------------

# Custom header
my_header <- data.frame(
  col_keys = c("trait", "parameter",
               "novice", "intermediate", "advanced"),
  line1 = c("Trait",
            "Parameter",
            "Novice",
            "Intermediate",
            "Advanced"),
  stringsAsFactors = FALSE
)

# Custom theme
my_theme <- function(x, ...) {
  x <- colformat_double(x, big.mark = " ",
                        decimal.mark = ".",
                        digits = 2)
  x <- border_remove(x)
  std_border <- fp_border(width = 1, color = "black")
  x <- hline_top(x, part = "all", border = std_border)
  x <- hline_bottom(x, part = "all", border = std_border)
  x <- width(x, j = 1, width = 2.84, unit = "cm")
  x <- width(x, j = 2, width = 3.78, unit = "cm")
  x <- width(x, j = c(3:5), width = 3.3, unit = "cm")
}

# Create the table ----------------------------------------------------------

tab1 <-

  # table structure
  table %>%
  select(trait, parameter,
         novice, intermediate, advanced) %>%
  flextable(col_keys = my_header$col_keys) %>%
  set_header_df(mapping = my_header, key = "col_keys") %>%
  my_theme() %>%

  # Align repeated rows on column 1
  #merge_v(j = 1) %>%
  valign(j = 1, valign = "top", part = "body") %>%

  # Align text
  align(align = "left", part = "all", j = 1) %>%
  align(align = "left", part = "all", j = 2) %>%
  align(align = "center", part = "all", j = 3) %>%
  align(align = "center", part = "all", j = 4) %>%
  align(align = "center", part = "all", j = 5) %>%

  # Footnote
  footnote(
    i = 1, j = 1,
    part = "header",
    value = as_paragraph(
      as_sup("a "),
      "We exponentiated the dispersion parameters (i.e. sigma) which are estimated on a log scale. We back-transformed the hunting success values, estimated on a logit scale, back to a probability scale.",
      as_sup("\nb "),
      "The intercept values on the mean part of the equation for all traits indicate mean behaviour and success at the population level. The intercept values on the dispersion (i.e. sigma) part of the equation for predator speed indicate behavioural specialization at the population level."
    ),
    ref_symbols = " "
  ) %>%

  # Table caption
  set_caption(
    caption = "Table S1. Posterior medians and 95% HPD intervals of the fixed effects estimated by the MDHGLM of predator speed, prey speed, and predator hunting success.",
    style = "Table Caption"
  ) %>%

  # Font options
  font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(size = 12, part = "footer")

tab1

# ===========================================================================
# ===========================================================================

```

\newpage

```{r tableS2, ft.align = "left"}

# ===========================================================================
# 1. Import the model fit
# ===========================================================================

# path
path <- file.path(here::here("outputs"))

# Import the model
fit <- readRDS(
  file.path(path, "01_outputs_models", "B1_DHMLM-no-outlier.rds")
)

# ===========================================================================
# ===========================================================================

# ===========================================================================
# 2. Random effects table
# ===========================================================================

# Prepare data --------------------------------------------------------------

# Extract the random effect variances
sd_draws <- data.table(
   as_draws_df(
      fit, variable = "^sd_", regex = TRUE
   )
)

sd_draws[, c(28:30) := NULL]

# Calculate parameters -------------------------------------------------------

# Take the exponent of sigma parameters
sd_draws[
   , c("sd_predator_id__sigma_speednovice_Intercept",
       "sd_predator_id__sigma_speedinterm_Intercept",
       "sd_predator_id__sigma_speedadvanced_Intercept",
       "sd_predator_id__sigma_preyspeednovice_Intercept",
       "sd_predator_id__sigma_preyspeedinterm_Intercept",
       "sd_predator_id__sigma_preyspeedadvanced_Intercept"
   )
   := lapply(.SD, function(x) {exp(x)}),
   .SDcols = c(
      "sd_predator_id__sigma_speednovice_Intercept",
      "sd_predator_id__sigma_speedinterm_Intercept",
      "sd_predator_id__sigma_speedadvanced_Intercept",
      "sd_predator_id__sigma_preyspeednovice_Intercept",
      "sd_predator_id__sigma_preyspeedinterm_Intercept",
      "sd_predator_id__sigma_preyspeedadvanced_Intercept"
   )
]

# Reshape the table -----------------------------------------------------------

ranef_tab <- melt(
   sd_draws,
   variable.name = "parameter"
)

# Summarize the values (means + 95% CI) --------------------------------------

# Intervals
lower_interval <- function(x) {
  coda::HPDinterval(as.mcmc(x), 0.95)[1]
}
upper_interval <- function(x) {
  coda::HPDinterval(as.mcmc(x), 0.95)[2]
}

# Summarize values
ranef_tab[, ":=" (median = median(value),
                  lower_ci = lower_interval(value),
                  upper_ci = upper_interval(value)),
          by = .(parameter)]

ranef_tab <- unique(ranef_tab[, .(parameter, median, lower_ci, upper_ci)])
ranef_tab[, c(2:4) := round(ranef_tab[, c(2:4)], digits = 2)]

# Rearrange the table --------------------------------------------------------

# Add a column that specifies the experience level
ranef_tab[parameter %like% "novice", xp_level := "novice"]
ranef_tab[parameter %like% "interm", xp_level := "intermediate"]
ranef_tab[parameter %like% "advanced", xp_level := "advanced"]

# Paste upper and lower ci with estimate
ranef_tab[, estimate := paste(format(median, digits = 2), "(")]
ranef_tab[, estimate := paste(estimate, format(lower_ci, digits = 2), sep = "")]
ranef_tab[, estimate := paste(estimate, ",", sep = "")]
ranef_tab[, estimate := paste(estimate, format(upper_ci, digits = 2), sep = " ")]
ranef_tab[, estimate := paste(estimate, ")", sep = "")]
ranef_tab[, c("median", "lower_ci", "upper_ci") := NULL]

# Parameter column
ranef_tab[
   , parameter1 := ifelse(
       parameter %like% "predator_id__sigma",
       "predator ID (sigma)",
       "predator ID (mean)"
   )
]
ranef_tab[parameter %like% "avatar", parameter1 := "avatar (mean)"]
ranef_tab[parameter %like% "env", parameter1 := "environment (mean)"]

# Trait column
ranef_tab[
   , trait := ifelse(
       parameter %like% "preyspeed",
       "prey speed",
       "predator speed"
   )
]
ranef_tab[parameter %like% "success", trait := "hunting success"]

# Remove old parameter column and replace by new
ranef_tab[, parameter := NULL]
setnames(ranef_tab, "parameter1", "parameter")

# Reshape
ranef_tab <- dcast(
   ranef_tab,
   trait + parameter ~ xp_level,
   value.var = "estimate"
)

# Reorder columns
ranef_tab <- ranef_tab[, c(1, 2, 5, 4, 3)]

# Reorder rows
table <- ranef_tab[c(2:9, 1)]

# ===========================================================================
# ===========================================================================

# ===========================================================================
# 3. Create the table using flextable
# ===========================================================================

# Prepare the table parameters ----------------------------------------------

# Custom header
my_header <- data.frame(
  col_keys = c("trait", "parameter",
               "novice", "intermediate", "advanced"),
  line1 = c("Trait",
            "Parameter",
            "Novice",
            "Intermediate",
            "Advanced"),
  stringsAsFactors = FALSE
)

# Custom theme
my_theme <- function(x, ...) {
  x <- colformat_double(x, big.mark = " ",
                        decimal.mark = ".",
                        digits = 2)
  x <- set_table_properties(x, layout = "autofit")
  x <- border_remove(x)
  std_border <- fp_border(width = 1, color = "black")
  x <- hline_top(x, part = "all", border = std_border)
  x <- hline_bottom(x, part = "all", border = std_border)
}

# Create the table ----------------------------------------------------------

tab2 <-

  # table structure
  table %>%
  select(trait, parameter,
         novice, intermediate, advanced) %>%
  flextable(col_keys = my_header$col_keys) %>%
  set_header_df(mapping = my_header, key = "col_keys") %>%
  my_theme() %>%

  # Align repeated rows on column 1
  merge_v(j = 1) %>%
  valign(j = 1, valign = "top", part = "body") %>%
  # font
  fontsize(size = 10, part = "all") %>%
  font(fontname = "Times New Roman", part = "all") %>%

  # Align text
  align(align = "left", part = "all", j = 1) %>%
  align(align = "left", part = "all", j = 2) %>%
  align(align = "center", part = "all", j = 3) %>%
  align(align = "center", part = "all", j = 4) %>%
  align(align = "center", part = "all", j = 5) %>%

  # Add footnote
  footnote(
    i = 1, j = 1,
    part = "header",
    value = as_paragraph(
      as_sup("a "),
      "We exponentiated the dispersion parameters (i.e. sigma) which are estimated on a log scale.",
      as_sup("\nb "),
      "The standard deviation values on the mean part of the equation indicate, for all traits, among individual differences in mean behaviour, prey encountered, and success.",
      as_sup("\nc "),
      "The standard deviation values on the dispersion part of the equation (i.e. sigma) for predator speed indicate among individual differences in behavioural specialization. For prey speed, they indicate among individual differences in the variability of prey encounters."
    ),
    ref_symbols = " "
  ) %>%

  # Add table caption
  set_caption(
    caption = "Table S2. Posterior medians and 95% HPD intervals of the random effect standard deviations estimated by the MDHGLM of predator speed, prey speed, and predator hunting success.",
    style = "Table Caption"
  ) %>%

  # Font options
  font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(size = 12, part = "footer")

tab2

# ===========================================================================
# ===========================================================================

```

\newpage

```{r figS1, out.width = "100%", fig.cap = "**Figure S1.** Among individual differences in the development of hunting expertise in the model where we do not account for prey speed and average rank. The predators' hunting success (i.e. the probability of capturing the four prey) is on the y axis, and the predators' cumulative experience (i.e. the number of matches played prior to each observation) is on the x axis. Each fitted curve represents an individual predator. The individual curves are separated by differences between their first and last predicted value, and displayed as such in three distinct panels. (A) Individuals with a >0.5 unit increase in hunting success with experience (B) Individuals with a <-0.5 unit decrease in hunting success with experience. (C) Individuals that maintained a stable hunting success (between -0.5 and 0.5 unit change in hunting success)"}

knitr::include_graphics(
  here::here(
    "outputs", "04_outputs_figures",
    "appendix1_figureS1.png"
  )
)

```

```{r figS2, out.width = "100%", fig.cap = "**Figure S2.** Correlations between the predators' mean hunting success (y axis) and intra individual variance in speed (x axis) to test for differences in success between individuals along the flexible-specialist hunter continuum. Each point represents the posterior median predicted value of an individual predator along with its 95% HPD interval. Individuals with lower IIV are specialist hunters, while individuals with higher IIV are flexible hunters. (A) Correlation when predators were novice (B) Correlation when predators were advanced"}

knitr::include_graphics(
  here::here(
    "outputs", "04_outputs_figures",
    "appendix1_figureS2.png"
  )
)

```

### 2024-06-04 manuscript/manuscript.Rmd
added old MS files to change

---
title: "Experience shapes individual foraging specialization and success in a virtual predator-prey system"
output: 
  word_document:
    reference_docx: manuscript-styles.docx
    fig_caption: yes
bibliography: references.bib
csl: chicago-author-date.csl
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(data.table)
library(brms)
library(flextable)
library(officer)
library(dplyr)

```

```{r}

path <- file.path(here::here(), "data")

dat <- fread(
  file.path(path, "FraserFrancoetal2023-data.csv"),
  select = c("timestamp", "match_id", "realm_id",
             "environment_id", "avatar_id", "predator_id",
             "game_duration", "total_xp_pred", "cumul_xp_pred",
             "hunting_success", "pred_speed", "prey_avg_speed"))

dat[, c(2:6) := lapply(.SD, function(x) as.factor(x)), .SDcols = c(2:6)]

# To have a date column
dat[, date := as.POSIXct(sub(" .*", "", timestamp))]

```

# ABSTRACT {.unnumbered}

The capacity of predators to match their tactic to their prey is expected to drive the outcome of predator-prey interactions. To do so, predators need to practice their tactics extensively to develop the skills to be successful hunters, which may result in individual foraging specialization. Yet, there are limited empirical assessments showing links between experience, foraging specialization, and hunting success at the individual level, due to the challenges of monitoring direct interactions in the wild. Here, we used a virtual predator-prey system (the game *Dead by Daylight*) to investigate how experience shapes individual foraging specialization and success in predators across repeated interactions with their prey. We show that predators optimized prey consumption as they gained experience. Prey speed was an important mediator of this relationship, resulting in differences among predators in the development of expertise. Throughout experience, 68% of predators became either increasingly specialized by always moving at a fast pace, or flexible by switching between slow and fast speeds. The predators’ strategies were partially matched to the preys’ speed, suggesting that changes in hunting behaviour were driven by repeated encounters with prey. Flexible and specialist foragers achieved similar success overall. Our findings suggest that experience may be mechanism promoting behavioural diversification in predator-prey systems.

Keywords: foraging behaviour, reciprocal behavioural plasticity, learning, antipredator behaviour, virtual ecology, Dead by Daylight

# INTRODUCTION {.unnumbered}

Predators express a diverse array of foraging tactics to capture their prey. In natural populations, differences among individuals in tactics such as ambushing and fast movement speeds are associated with differences in prey consumption [@toscano2014;@patrick.weimerskirch2014a;@toscano.etal2016]. Prey usually respond to predation cues by reducing their activity, or use rapid or unpredictable movements to successfully escape when they are chased [@walker.etal2005;@kelley.magurran2011;@herbert-read.etal2017]. This results in reciprocal interactions where predators fine-tune their tactics to their prey, suggesting that they learn and become efficient hunters through repeated encounters [@estes.etal2003;@woo.etal2008;@phillips.etal2017]. These behavioural adjustments with experience can shape the outcome of predator-prey interactions, and may have larger implications for community and ecosystem dynamics [@wooster.etal2023].

Empirical studies on human and non-human hunters show that individuals optimize foraging efficiency (e.g. search and handling times, return rates) by learning associative images or prey and environmental cues [@edwards.jackson1994;@morse2000;@macdonald2007;@reid.etal2010;@wilson-rankin2015]. As predators practice and refine their tactics, they acquire expertise (e.g. skills, knowledge) that allows them to outperform novice hunters [@dukas2019;@wooster.etal2023]. Through this process, it is hypothesized that individuals will gradually specialize in a foraging tactic (e.g. always moving fast) if it allows them to maintain successful prey captures at each attempt [@tinker.etal2009;@dukas2019]. This should occur when predators learn to hunt in environments where resources are predictable (or stable) because they require less energy and time to search for, capture, and handle prey [@weimerskirch2007;@woo.etal2008;@potier.etal2015]. Such conditions favor specialist hunters because there would be higher costs to switch tactics for the same type of prey. An alternative hypothesis is that learning is a mechanism driving behavioural flexibility. For instance, predators can gain expertise and information on their prey through repeated hunting atempts, and may hunt optimally by learning to adjust their tactic to the type of prey that they encounter [@stephens1993;@ishii.shimada2010;@kelley.magurran2011]. While there are costs to attempt novel tactics by trial and error in stable environments, predators that hunt in environments under fluctuating resources can reduce the consequences of uncertainty by learning to switch hunting tactics [@dall2010;@mathot.etal2012]. Nonetheless, complex feedback between ecological interactions and learning make it challenging to predict under which circumstances predators should specialize with experience [@stephens1993;@ishii.shimada2010;@vilapouca2021;@beyts2023].

There is increasing evidence suggesting that the emergence of behavioural specialization and its fitness consequences are driven by resource fluctuations [@patrick.weimerskirch2014b;@vandenbosch.etal2019;@santoro.etal2019;@manlick.etal2021]. For example, @szopa-comley.ioannou2022 showed that predatory blue acaras (*Andinoacara pulcher*) reduce their maximum approach speed when repeatedly exposed to prey that escape unpredictably, enabling them to be as successful as those exposed to predictable prey. Prey that use fast escapes are also proven to be more difficult to hunt [@walker.etal2005;@kelley.magurran2011;@martin.etal2022], and predators that exclusively encounter such prey should benefit from learning to specialize in fast pursuits. If predators that hunt in different environments adjust their tactics to local prey fluctuations, then specialist and flexible hunters can both attain similar hunting success and coexist [@phillips.etal2017]. Therefore, empiricists should investigate the link between foraging specialization and success across time and contexts.

A recurring challenge impeding research on predator-prey behavioural interactions is the need to collect data simultaneously on both the predator and the prey. Similar to agent-based simulations, online videogames can be useful systems to mitigate these challenges and test specific ecological hypotheses. They provide controlled virtual environments with the advantage of having real players that interact in the virtual space. For example, in the predator-prey videogame *Dead by Daylight*, four prey players need to forage for resources while avoiding predation by a fifth player. In this virtual system, the predator population is composed of individuals that either ambush or hunt at high speeds (i.e. mean movement speed along a slow-fast continuum), and their success is driven by the movement of the prey [@fraserfranco.etal2022]. We found that prey can increase their chances of survival by cooperating and moving fast when they forage [@cere.etal2021;@fraserfranco.etal2022, Santostefano et al. in press]. Of course, the life of neither the predator nor the prey players are at stake, such that emerging patterns could be driven by their motivation to win and not "true" survival. Yet, the game ellicits natural reactions in players such as fear from predation (personal observations), which corroborates with another virtual ecological study showing that predation drives individual variation in risk perception [@beauchamp2020]. In a different system, @lymbery.etal2023 showed that complex environments favor strong soldiers over large armies in a virtual game as well as in the laboratory with two ant species. These observations outline how ecological patterns can emerge from interactions in virtual systems with fixed rules. Videogames also generate large volumes of data on interacting players throughout their lifetime in the game under realistic, controlled, and repeatable ecological scenarios. Hence, virtual systems such as *Dead by Daylight* offer the opportunity to tackle fundamental questions about the role of experience on predator-prey interactions.

In this study, we assess how repeated experience shapes predator foraging specialization and success using data from players in *Dead by Daylight*. Following @fraserfranco.etal2022 we use the predators' mean movement speed along a slow-fast continuum as a proxy for the foraging tactic (ambush-cursorial), and define the level of individual foraging specialization as the intra-individual variance (i.e. IIV) in movement speed across matches, where low IIV translates to individual foraging specialization (e.g. always using a cursorial tactic). First, we investigate how individual hunting success changes with experience (i.e. how individuals develop their hunting expertise). We hypothesize that the predators' success will increase with experience up to a certain level where it will stagnate. We expect this pattern to vary among individuals due to differences in prey encounters. Second, we assess how foraging specialization emerges with experience. If all predators encounter similar groups of prey, we predict that individuals (and thus the population) should all specialize in similar speeds. In contrast, if all predators encounter varying groups of prey, then they should all converge towards flexible speeds. In both scenarios, differences among individuals in IIV across experience should be low (i.e. similar individual foraging specialization), whereas the population variance would either decrease (specialization) or increase (flexibility). Alternatively, differences among individuals in foraging specialization may emerge if they experience different interactions with their prey. In this case, we expect predators that encountered similar groups of prey across experience to specialize in similar speeds, while predators that encountered heterogeneous groups across matches should adopt flexible movement speeds, resulting in an increase in among individual differences in IIV with experience. If we detect such prey-dependent fine-tuning with experience, then all hunters along the flexible-specialist continuum should attain equal success.

# MATERIALS AND METHODS {.unnumbered}

## Study system

*DBD* is a survival asymmetric multiplayer online game (i.e. a game where the gameplay mechanics differ between two groups) developed by Behaviour Interactive Inc, in which players can play either as a predator or a prey. Every match includes only one predator and four prey. The objective of the predator is to hunt and capture the prey, and the objective of the prey is to search for resources while avoiding the predator. The resources are in the form of power generators that, once all activated, will enable the prey to escape through one of two exit doors. The composition of the predator and prey group in a match is determined by a skill-based matchmaking algorithm. A match ends when the predator kills all the prey available (i.e. that have not escaped), or when the last remaining prey escapes the virtual environment. Each player, predator or prey, can choose an avatar with abilities that encourage specific play styles (e.g. bold vs cautious prey, or ambush vs roaming predator). During our study period, the game offered `r toString(format(length(unique(dat$avatar_id))))` predator avatars. The virtual environments are composed of fixed and procedurally generated habitat components, such as vegetation, mazes, and buildings. Some of these environments are larger than others, with varying structural complexity. However, predators display only minimal changes in behaviour and hunting success across the environments, probably due to a game feature enabling them to have visual cues of the generators [@fraserfranco.etal2022]. There were `r toString(format(length(unique(dat$environment_id))))` virtual game environments available for play during the study period.

## Data collection

The videogame company provided data that spanned a period of 6 months of gameplay recorded for every player from `r toString(format(min(dat$date)))` to `r toString(format(max(dat$date)))`. We only analyzed matches where players did not know each other (i.e. "Online" mode). We filtered any matches where players were inactive, such as when mean distances traveled per second (i.e. speed) were equal to, or very close to, zero. Moreover, we used our knowledge of the game to remove any matches where players were potentially hacking, or not playing the game as intended. We then sampled players that played 300 matches or more, and monitored all their matches from the first to a maximum of 500 matches.

Our population consists of `r toString(format(length(unique(dat$predator_id))))` predator players with a total record of `r toString(formatC(length(unique(dat$match_id)), big.mark = " "))` matches. The predator-players' experience varied between `r toString(format(min(dat$total_xp_pred)))` and `r toString(format(max(dat$total_xp_pred)))` matches played. These matches lasted between `r toString(format(round(min(dat$game_duration / 60))), digits = 2)` and `r toString(format(round(max(dat$game_duration / 60))), digits = 2)` min (mean = `r toString(format(round(mean(dat$game_duration / 60))), digits = 2)` min). The following information is collected and reported for every match : the player's anonymous ID, its avatar (i.e. the predator character chosen with its specific skill-gameplay mechanics), the game environment, the predator-player's experience along with its speed, the mean speed of the groups of prey the player encountered, and their mean rank (a proxy for prey skill). The ranking system in *DBD* was implemented by the company to pair players in a match based on their skill (https://deadbydaylight.fandom.com/wiki/Rank), and failing to account for it would prevent us from detecting a change in the predator's foraging success with experience.

We analyzed the predator's mean speed and the mean speed of the prey group encountered by the predator. The predator's mean speed is measured as the mean distance traveled per second during a match (mean= `r toString(format(round(mean(dat$pred_speed), digits = 2), nsmall = 2))` ± `r toString(format(round(sd(dat$pred_speed), digits = 2), nsmall = 2))` m/s). We measured the preys' speed as the mean travel speed of the four individual prey in a match (mean = `r toString(format(round(mean(dat$prey_avg_speed, na.rm = T), digits = 2), nsmall = 2))` ± `r toString(format(round(sd(dat$prey_avg_speed, na.rm = T), digits = 2), nsmall = 2))` m/s). We defined hunting success as the number of prey consumed during the match (min = `r toString(format(min(dat$hunting_success)))`, max = `r toString(format(max(dat$hunting_success)))`). Lastly, we defined the predator's cumulative experience as the number of matches played as the predator prior to the match being monitored. For example, the first match of a player would have a cumulative experience value of 0, while the tenth match would have a value of 9. We did not account for matches where predators played as the prey. We also categorized predators for each match based on their cumulated experience. We labeled predators as novices for matches where they had cumulated less than 100 matches, intermediate for matches where they had cumulated between 100 and 299 matches, and advanced for matches where they had cumulated more than 299 matches (max 499). Since our goal was to monitor predator players throughout their experience and that they all played at least 300 matches, they all appeared in each of the three experience categories.

We recognize that we could have biased our analyses by retaining only individuals who played for at least 300 matches. For example, these individuals might be experienced videogame players and could thus already be playing like experts in their first matches in *DBD*. To verify that our sample was not biased, we compared a random sample of players that played either 20 to 50 matches, 51 to 100 matches, or 101 to 300 matches during the same timeframe as our sampled population. We then took the first 20 matches played by these players, including those from our sampled population, and compared their mean speed using a Bayesian hierarchical linear model. We found that all four groups had similar mean speeds as predators (Appendix 2: Table S1-S2), which indicates an absence of bias due to data sampling.

## Statistical analyses

### Software and computer specifications

All models were fitted under a Bayesian framework in R (version 4.1.2) using Markov chain Monte Carlo (MCMC) sampling with the package "brms" version 2.16.3 [@burkner2017], an R front-end for the STAN software [@StanDevelopmentTeam2023], and "cmdstanr" version 0.4.0 [@gabry.cesnovar2021] as the back-end for estimation (cmdstan installation version 2.28.2). The models were run on Cedar (Operating system: CentOS Linux 7), a computer cluster maintained by the Digital Research Alliance of Canada (https://docs.alliancecan.ca/wiki/Cedar). Each required 64GB of RAM with 48 cores to compile within 5 days.

### Development of expertise

We tested how prey influenced the development of predator expertise by computing five Bayesian generalized additive mixed models (GAMM) with thin-plate regression splines, all of which estimated the relationship between hunting success (i.e. number of prey consumed) and the predators' cumulative experience (i.e. number of matches played before the current match). We parametrized the models following the method of @pedersen.etal2019. The first model was the simplest, with a common global smoothing function and random intercepts for the predator ID. In this model, we assume that predators have the same development of expertise, with the model estimating a trend for the average individual (i.e. global smoother). For the second model, we included varying individual smoothers for the predator ID. Here, we assume that individual predators share a similar relationship between success and experience, but that this relationship can vary among them (e.g. predator 1 has a steeper curve than predator 2). This enabled us to test whether predators differed in the development of their expertise. In the third model, we kept the individual smoothers for the predators, but removed the global smoother. This model assumes that predators do not share a common relationship between success and experience. The fourth and fifth models were reproductions of the second and third models respectively, where we included the prey speed to assess its effect on the relationship between success and experience. We included the standardized match duration and prey rank as covariates in all five models.

We computed the five models using a modified version of the beta-binomial distribution implemented in "brms". Hunting success was estimated as the probability of consuming the four prey ($\mu_i$), drawn from a Beta distribution ($Beta(\mu_i, \phi)$) with mean ($\mu \in [0,1]$) and precision ($\phi > 0$) parameters. We used a logit link function to estimate $\mu_i$ where $\mu_i = \frac{exp(\eta_i)}{exp(\eta_i) + 1}$ and $\eta_i$ is the linear predictor, while the precision parameter ($\phi$) was estimated with an identity link. We used the default number of basis functions (K) in "brms" for the models to estimate the relationship between hunting success and experience. We assumed that the random intercepts for the predator ID ($id$) followed a Gaussian distribution with estimated standard deviation ($id \sim N(0,\ I\ \sigma_{id}^{2})$). We used weakly informative Gaussian priors for the intercept ($N(0,\ 0.5^{2})$) and the global trend of cumulative experience ($N(0,\ 2^{2}))$). Following @fraserfranco.etal2022, we defined a positive Gaussian prior on the precision parameter ($N(2,\ 0.5^{2})$), a positive Gaussian prior ($N(1,\ 0.5^{2}))$) on the game duration because longer trials lead to greater success, and a negative Gaussian prior on prey speed ($N(-1,\ 0.5^{2})$) because encountering faster prey is associated with lower success in this system. We employed weakly informative half-Gaussian priors on all the standard deviation parameters ($N(0,\ 0.5^{2})$). We compared the models' predictive accuracies using approximate leave-one-out cross-validation with Pareto-smoothed importance sampling [@vehtari.etal2017;@piironen.vehtari2017;@vehtari.etal2022].

### Foraging behaviour and predator-prey interactions across experience

We tested whether foraging behaviour changed with experience, and compared the success of predators along the flexible-specialist hunter continuum, using a multivariate double-hierarchical generalized linear model (MDHGLM) [@lee.nelder2006;@cleasby.etal2015;@mitchell.etal2016]. For each experience level, we fitted the model on three traits: the predator's speed ($Y_{1}$) and the prey's speed ($Y_{2}$) using a Gaussian error distribution, and the predator's hunting success ($Y_{3}$) using a beta-binomial error distribution. Fixed and random effect structure were trait-specific and the same for all experience levels. The predator and the prey speed included a mean and a dispersion part (eqn 1.1, 1.2 and 2.1, 2.2). For both traits, we controlled for the standardized prey rank on both the mean and the dispersion parts of the model. We included game environment, predator avatar, and individual predator identity random effects on the mean part of the model. The dispersion part included individual predator identity random effects exclusively. For hunting success, we modeled the mean part only (eqn 3) and controlled for the standardized prey rank and game duration, and included an individual identity random effect.

The model simultaneously estimated five among-individual variance components with their covariances for each experience level, resulting in a 15x15 (co)variance matrix. For each experience level, the model estimated among individual variance in mean predator speed, mean prey speed, and mean hunting success. This allows us to test for individual differences in tactic use (i.e. along the slow-fast continuum), in the mean speed of the prey encountered, and in mean prey consumption. The model also estimated among individual variance in intra-individual variation (IIV) for both prey and predator speed. This allows us to test if predators do not experience the same degree of prey heterogeneity and if they differ in individual specialization. Finally, the model estimated all pairwise covariances at the individual level among traits, among experience, and across traits and experience. The resulting 15x15 (co)variance matrix thus provides the structure of the predator-prey trait interactions as well as the relationship between specialization and success across experience at the individual-level.

The model can be written as:

```{r model_equations, ft.align = "left"}

# Prepare equations
#eqs <- c(
#  "Y_{1}(T_{j}) = X_{1}(T_{j})\\ b_{1}(T_{j}) + M_{1}(T_{j})\\ en_{1}(T_{j}) + N_{1}(T_{j})\\ av_{1}(T_{j}) + Z_{1}(T_{j})\\ #id_{1}(T_{j}) + e_{1}(T_{j})",
#  "log(\\sigma_{e_{Y_{1}(T_{j})}}) = X_{d1}(T_{j})\\ b_{d1}(T_{j}) + Z_{d1}(T_{j})\\ id_{d1}(T_{j})",
#  "Y_{2}(T_{j}) = X_{2}(T_{j})\\ b_{2}(T_{j}) + M_{2}(T_{j})\\ en_{2}(T_{j}) + N_{2}(T_{j})\\ av_{2}(T_{j}) + Z_{2}(T_{j})\\ #id_{2}(T_{j}) + e_{2}(T_{j})",
#  "log(\\sigma_{e_{Y_{2}(T_{j})}}) = X_{d2}(T_{j})\\ b_{d2}(T_{j}) + Z_{d2}(T_{j})\\ id_{d2}(T_{j})",
#  "Y_{3}(T_{j}) = X_{3}(T_{j})\\ b_{3}(T_{j}) + Z_{3}(T_{j})\\ id_{3}(T_{j}) + e_{3}(T_{j})"
#)

#eqs <- c(
#  "Y_{1}(T_{j}) = T^2_{j}(X_{1}\\ b_{1}) + T^2_{j}(M_{1}\\ en_{1}) + T^2_{j}(N_{1}\\ av_{1}) + T^2_{j}(Z_{1}\\ id_{1}) + e_{1}(T_{j})",
#  "log(\\sigma_{e_{Y_{1}(T_{j})}}) = T^2_{j}(X_{d1}\\ b_{d1}) + T^2_{j}(Z_{d1}\\ id_{d1})",
#  "Y_{2}(T_{j}) = T^2_{j}(X_{2}\\ b_{2}) + T^2_{j}(M_{2}\\ en_{2}) + T^2_{j}(N_{2}\\ av_{2}) + T^2_{j}(Z_{2}\\ id_{2}) + e_{2}(T_{j})",
#  "log(\\sigma_{e_{Y_{2}(T_{j})}}) = T^2_{j}(X_{d2}\\ b_{d2}) + T^2_{j}(Z_{d2}\\ id_{d2})",
#  "Y_{3}(T_{j}) = T^2_{j}(X_{3}\\ b_{3}) + T^2_{j}(Z_{3}\\ id_{3}) + e_{3}(T_{j})"
#)

eqs <- c(
  "Y_{1j} = X_{1j}\\ b_{1j} + M_{1j}\\ en_{1j} + N_{1j}\\ av_{1j} + Z_{1j}\\ id_{1j} + e_{1j}",
  "log(\\sigma_{e_{Y_{1j}}}) = X_{d1j}\\ b_{d1j} + Z_{d1j}\\ id_{d1j}",
  "Y_{2j} = X_{2j}\\ b_{2j} + M_{2j}\\ en_{2j} + N_{2j}\\ av_{2j} + Z_{2j}\\ id_{2j} + e_{2j}",
  "log(\\sigma_{e_{Y_{2j}}}) = X_{d2j}\\ b_{d2j} + Z_{d2j}\\ id_{d2j}",
  "Y_{3j} = X_{3j}\\ b_{3j} + Z_{3j}\\ id_{3j} + e_{3j}"
)

eqs_num <- c(
    "(eqn. 1.1)",
    "(eqn. 1.2)",
    "(eqn. 2.1)",
    "(eqn. 2.2)",
    "(eqn. 3)"
)

# Combine into a dataframe
df <- data.frame(
    formula = eqs,
    number = eqs_num
)

# Create a table to display on Word
ft <- flextable(df)

# Transform values as a mathematical formula
ft <- compose(
  x = ft, j = 1,
  value = as_paragraph(
    as_equation(formula, height = .5)
  )
)

# Customize table
my_theme <- function(x, ...) {
  x <- colformat_double(x, big.mark = " ",
                        decimal.mark = ".",
                        digits = 2)
  x <- border_remove(x)
  x <- delete_part(x, part = "header")
  x <- align(x, j = 1, align = "left", part = "body")
  x <- align(x, j = 2, align = "right", part = "body")
  x <- font(x, fontname = "Times New Roman", part = "body")
  x <- fontsize(x, size = 12, part = "body")
  x <- width(x, j = 1, width = 13, unit = "cm")
  x <- width(x, j = 2, width = 2.26, unit = "cm")
}

# Apply theme and print
ft <- my_theme(x = ft)
ft

```

where $j$ is the index of the three experience levels (i.e. novice, intermediate, advanced). The $d$ subscript indicates that a term is from the dispersion part of the model. The $b$ terms are vectors of fixed effects with their design matrices $X$ specific to each trait and experience level. The terms $en$, $av$, and $id$ are the vectors of random environment, random avatar, and random individual identity effects associated with their design matrices $M$, $N$, and $Z$ specific to each trait and experience level. For all experience levels of $Y_{1}$ and $Y_{2}$, we assumed that the game environment and avatar random effects followed a Gaussian distribution with estimated standard deviation ($en_{12j} \sim N(0,\ I\ \sigma_{en_{12j}}^{2})$ and $av_{12j} \sim N(0,\ I\ \sigma_{av_{12j}}^{2})$). $I$ is the identity matrix, with $I\sigma^{2}$ indicating that the random effects are independently and identically distributed. In addition, we assumed that the residuals follow a Gaussian distribution with observation-specific variance $e_{12j} \sim N(0,\ Diag\{\sigma_{e_{12j}}^{2}\})$. The 15 individual identity random effects $id_{ij}$ across all response variables and experience levels (five for each of the three experience levels) are indexed in vector $\alpha$. The $\alpha$ vector follows a multivariate Gaussian distribution $\alpha \sim MVN(0,\ \Sigma)$ where $\Sigma$ is a 15x15 (co)variance matrix of among individual variation across all response variables and experience levels, with $\sigma_{id_{ij}}^{2}$ on the diagonal and covariances off-diagonal.

We used weakly informative Gaussian priors for the prey rank ($N(0,\ 1^{2})$: eqn. 1.1 to 2.2) and game duration ($N(0.5,\ 0.5^{2})$: eqn. 3). Based on previous results showing that the mean predator speed revolves around 3 m/s [@fraserfranco.etal2022], we applied a weakly informative Gaussian prior on the intercept for predator and prey speed ($N(3,\ 0.5^{2})$: eqn. 1.1 and 2.1). Similarly, we used a weakly informative Gaussian prior on the intercept for the dispersion part ($N(0,\ 1^{2})$: eqn 1.2 and 2.2). For hunting success, we applied a positive Gaussian prior on the precision parameter ($N(2,\ 1^{2})$: eqn. 3) and a negative Gaussian prior ($N(-4,\ 1^{2})$: eqn 3) on the intercept because hunting success should be close to 0 at the lowest game duration. For every predicted variable, we applied a weakly informative half-Gaussian prior ($N(0,\ 1^{2})$) on the random effects for the mean part and the dispersion part of the model. We used a regularizing $LKJ(2)$ prior on the correlation matrix.

After inspection of the results, we found that key parameter values for our hypotheses were in the direction of our predictions, but they were largely driven by the very slow speed of one predator. Removing this individual reduced the estimates by a significant margin. To remain conservative, we thus present the results of the model without this individual in the main text, and provide the results of the model including it in the Appendix S3.

### Markov Chain Monte Carlo settings

We parametrized the GAMMs and the MDHGLM to run four MCMC chains with 1000 posterior samples for each parameter. We ran 2500 iterations with a thinning set to eight for the additive model with a global smoother only (see Table 1), and 1500 iterations with a thinning set to four for the other additive models. We set the first 500 iterations in each model as warm ups. We parametrized the MDHGLM to run 2500 iterations with a thinning set to eight, with the first 500 iterations used as warm ups. We assessed the convergence of the MCMC chains using trace plots, R-hat diagnostics with a threshold of <1.01, and effective sample sizes (ESS) with a threshold of >100 [@vehtari.etal2021]. We also performed posterior predictive checks which showed an adequate fit of the models (for details, see https://github.com/quantitative-ecologist/experience-hunting-tactics). We report all the posterior parameter estimates using the median of the posterior distribution with the highest posterior density (HPD) intervals at 50%, 80%, and 95%.

# RESULTS {.unnumbered}

## Development of expertise

```{r}

# Open percentages table
path <- file.path(here::here("outputs", "03_outputs_model-processing"))
percent_tab1 <- data.table(readRDS(file.path(path, "GAMM_percentages.rds")))
percent_tab2 <- data.table(readRDS(file.path(path, "DHMLM_percentages.rds")))

#fit1 <- readRDS(here::here("outputs", "01_outputs_models", "A2_GAMM-rank.rds"))
#fit2 <- readRDS(here::here("outputs", "01_outputs_models", "A2_GAMM-speed-rank.rds"))
#
#fit1
#fit2
```

Out of all five GAMM models, two predicted the data the best and achieved similar expected log pointwise densities (elpd). Both models accounted for the prey group's rank and speed (Table 1). Models in which these effects were not accounted for resulted in no change in hunting success with experience (i.e. expertise) for the average individual (results not shown). Accounting for the prey rank resulted in a concave-shaped relationship, with the highest success ranging between 100 and 400 matches (Figure 1A). In the model where we additionally included the prey's speed, the effect of experience on hunting success for the average individual followed a diminishing returns curve, with predators optimizing their success after playing ~300 matches (Figure 1C). The curve shows there was a 38% increase in the probability of consuming all prey for the average individual between the first and the ~330$^{th}$ match where success reached a plateau (Figure 1C).

```{r table1, ft.align = "left"}

# ===========================================================================
# 1. Import the model fit
# ===========================================================================

# Open the loo table
path <- file.path(here::here("outputs", "02_outputs_model-validation"))
loo_tab <- data.table(readRDS(file.path(path, "GAMM_loo-cv-table.rds")))

# Add model column
loo_tab[
  , model := c(
    "predator xp + ID smoothers + prey rank + prey speed",
    "ID smoothers + prey rank + prey speed",
    "ID smoothers + prey rank",
    "predator xp + ID smoothers + prey rank",
    "predator xp + prey rank"
    )
 ]

# Reorder columns
loo_tab <- loo_tab[, c(9, 1:8)]

# ===========================================================================
# ===========================================================================

# ===========================================================================
# 2. Compute the table using flextable
# ===========================================================================

# Prepare the table parameters ----------------------------------------------

# Custom header
my_header <- data.frame(
   col_keys = c("model",
                "elpd_diff",
                "se_diff",
                "elpd_loo",
                "se_elpd_loo"),
   line1 = c("model",
             "elpd \ndifference",
             "sd \ndifference",
             "elpd loo \nvalue",
             "elpd loo \nstandard error"),
   stringsAsFactors = FALSE
)

# Custom theme
my_theme <- function(x, ...) {
   x <- colformat_double(x, big.mark = " ",
                         decimal.mark = ".",
                         digits = 2)
   x <- set_table_properties(x, layout = "autofit")
   x <- border_remove(x)
   std_border <- fp_border(width = 1, color = "black")
   x <- hline_top(x, part = "all", border = std_border)
   x <- hline_bottom(x, part = "all", border = std_border)
}

# Create the table ----------------------------------------------------------

tab1 <- loo_tab %>%
  select(model, elpd_diff, se_diff, elpd_loo, se_elpd_loo) %>%
  flextable(col_keys = my_header$col_keys) %>%
  set_header_df(mapping = my_header, key = "col_keys") %>%
  my_theme() %>%
  align(align = "left", part = "all", j = 1) %>%
  align(align = "center", part = "all", j = 2) %>%
  align(align = "center", part = "all", j = 3) %>%
  align(align = "center", part = "all", j = 4) %>%
  align(align = "center", part = "all", j = 5) %>%
  footnote(
    i = 1, j = 1,
    part = "header",
    value = as_paragraph(
      as_sup("a "),
      "'elpd' refers to the expected log pointwise density and is the value chosen to select the best model.",
      as_sup("\nb "),
      "'xp' is an acronym for experience"),
    ref_symbols = " "
  ) %>%

  # Table caption
  set_caption(
    caption = "Table 1. Leave-one-out cross-validation table of the five GAMMs relating hunting success to predator experience.",
    style = "Table Caption"
  ) %>%

  # Font options
  font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(size = 12, part = "footer")

tab1
# ===========================================================================
# ===========================================================================

```

The relationship between hunting success and cumulative experience differed among predators (Figure 1B-D). Only `r toString(scales::percent(percent_tab1[2,]$percentage, accuracy = 0.1))` of the population had an increase in success from the first match to the last in the model where we did not account for the prey's speed and rank (Appendix S1: Figure S1A). In contrast, accounting for both effects resulted in `r toString(scales::percent(percent_tab1[5,]$percentage, accuracy = 0.1))` of the population increasing its success with experience from the first match to the last. The prey speed alone also mediated individual differences in the relationship between success and experience. The standard deviation of the individual slopes component of the model accounting only for prey rank was equal to 9.72 (9.30, 10.15 $95\%\ CI$), while the one for the model accounting for prey rank and speed was equal to 3.36 (3.04, 3.71 $95\%\ CI$), indicating that individual differences in the development of expertise decreased when accounting for prey speed.

```{r fig1, out.width = "100%", fig.cap = "**Figure 1**. Median posterior predictions of the development of predator hunting expertise. The predators' hunting success (i.e. the probability of consuming the four prey) is on the y axis, and the predators' cumulative experience (i.e. the number of matches played prior to each observation) is on the x axis. Panels A and C show the development of expertise for the average individual with the vertical dashed lines on the left representing the lowest predicted values. For panel A, the right-side vertical dashed line shows the highest predicted success. For panel C, the right-side dashed line represents the point on the curve where success was optimized, which we calculated using the finite differences method to obtain the first derivative of the predicted values. Panels B and D show among individual differences in the development of expertise, with each curve representing an individual predator. (A-B) GAMM where we control for the prey rank (C-D) GAMM where we control for the prey rank and the speed of the prey group."}

knitr::include_graphics(
  here::here("outputs", "04_outputs_figures", "figure1.png")
)

```

## Changes in foraging behaviour, prey behaviour, and hunting success

The mean speed of the predator population remained stable with experience (Figure 2 and Appendix S1: Table S1). Likewise, the variation in speed did not change with experience, indicating that foraging specialization remained stable at the population level (Figure 2 and Appendix S1: Table S1). At the individual level, predators differed slightly in their mean speed across experience levels, indicating marginal differences in hunting tactics (Figure 2 and Appendix S1: Table S2). However, they displayed important differences in foraging specialization (Appendix S1: Table S2) which also increased slightly with experience as there was a 0.11 unit increase in among individual differences in IIV from novice to advanced (Figure 2). `r toString(scales::percent(percent_tab2[1,]$percentage, accuracy = 1))` of the population switched from a flexible to a specialized hunting tactic, and vice-versa, as predators gained experience (i.e. >0.2 change in standard deviation; Figure 3). In contrast, `r toString(scales::percent(percent_tab2[2,]$percentage, accuracy = 1))` displayed lower changes (i.e. >0.05 and <0.2 change in standard deviation) and `r toString(scales::percent(percent_tab2[3,]$percentage, accuracy = 1))` showed almost no change with experience (i.e. <0.05 change in standard deviation).

```{r fig2, out.width = "100%", fig.cap = "**Figure 2**. Median posterior differences and HPD intervals comparing predator foraging behaviour at the population and individual level across experience stages. The behavioural parameter predicted by the MDHGLM being compared is on the y axis, and the difference in parameter values between two experience stages is on the x axis. The pairwise comparisons between experience stages are displayed across the three panels A, B, and C. We compare the mean speed of the predator population, the speed of the prey it encountered, and its hunting success across experience (population mean). We also compare the population's foraging specialization and variation in speed of the prey it encountered across experience (population variance). At the individual level, we compare among individual differences in speed, the speed of the prey encountered, and hunting success (individual variation mean). Lastly, we compare individual variation in IIV across experience, indicating whether individual differences in foraging specialization and in the variation of the groups of prey encountered changed with experience (individual variation IIV)."}

knitr::include_graphics(
  here::here("outputs", "04_outputs_figures", "figure2.png")
)

```

As the predator population gained experience, there was strong evidence that the prey increased their speed, although slightly (Figure 2 and Appendix S1: Table S1). The population also encountered groups of prey using different speeds through time (Figure 2 and Appendix S1: Table S1). At the individual level, all predators encountered prey with similar average speeds throughout experience (Figure 2 and Appendix S1: Table S2). Yet, even if they experienced similar average prey speeds across all groups encountered, there were important differences among predators in how (dis)similar were the groups they encountered (Appendix S1: Table S2). Thus, some predators encountered similar groups (i.e. all groups using similar speeds) while others encountered distinct groups (i.e. some slower and some faster). These differences among individuals in IIV for prey encounters increased only marginally with experience (Figure 2).

Similar to the GAMMs, the predator population's mean hunting success increased with experience (Figure 2 and Appendix S1: Table S1), with mean prey consumption revolving around two prey. There were also important differences among individuals in mean hunting success (Appendix S1: Table S2), but they remained stable across experience (Figure 2).

```{r fig3, out.width = "100%", fig.cap = "**Figure 3**. Individual behavioural distribution of the predators' speed drawn from the MDHGLM estimates for individuals that had (A) the greatest increase in specialization and (B) the greatest increase in flexibility. The background (gray) distributions are for when individuals were novice hunters, and the foreground (blue) distributions with solid contour lines are for when they were advanced hunters. Both figure panels are ordered by ascending degree of increase in either specialization or flexibility. We built the figure by first subtracting the estimated standard deviation of all individuals as novices with their standard deviation as advanced hunters, and then selected those with the greatest difference using the 25% and 75% quantiles. Individuals on panel A have an increase in specialization equal or greater than 0.2 standard deviations, while those on panel B have an increase in flexibility equal to or greater than 0.2 standard deviations."}

knitr::include_graphics(
  here::here("outputs", "04_outputs_figures", "figure3.png")
)

```

## Predator-prey interactions across experience

Predators that were on average faster had lower IIV in speed (Figure 4A-B and Appendix S4: Table S1). Thus, faster predators specialized on fast movement, while slower predators were more flexible in their movement. As predators gained experience, these strategies were increasingly defined as the correlation changed by 0.16 units, from -0.62 (-0.91, -0.41 $95\%\ CI$) for when predators were novices to -0.78 (-1.00, -0.53 $95\%\ CI$) as they reached the advanced stage (Figure 4C). Predators that were on average faster also tended to encounter groups of prey that were on average faster and more similar (Figure 4A-B and Appendix S4: Table S1). There was however no evidence that these correlations increased with experience (Figure 4C)

There was no correlation (-0.01 (-0.17, 0.09 $95\%\ CI$)) between mean prey speed and IIV in predator speed when predators were novice (Figure 4A), but the correlation increased to -0.16 (-0.33, -0.04 $95\%\ CI$) at the advanced stage (Figure 4B), such that advanced hunters that encountered faster prey tended to be more specialized (Figure 4C). The predators' IIV in speed and IIV in prey speed were weakly positively correlated (Figure 4A-B and Appendix S4: Table S1). Hence, predators that encountered groups of prey with similar speeds tended to be more specialized in their tactic, while those that encountered groups of prey with varying speeds tended to be more flexible. We did not find evidence that this correlation increased with experience (Figure 4C).

```{r fig4, out.width = "100%", fig.cap = "**Figure 4**. Correlation matrices of the MDHGLM relating the mean and IIV in speed of the predators, the mean and IIV in speed of the prey they encountered, and mean hunting success. Larger dark blue circles indicate stronger positive correlations (positive values on the colour legend), while larger dark red circles indicate stronger negative correlations (negative values on the colour legend). A) Correlations when predators were novice hunters. B) Correlations when predators were advanced hunters. C) Posterior median differences and HPD intervals comparing the trait correlations predicted by the MDHGLM between the advanced and novice levels. The difference is displayed on the y axis and the correlated traits being compared are displayed on the x axis. Positive values indicate an increase in correlation strength from novice to the advanced stage, while negative values indicate stronger correlations when predators were novice."}

knitr::include_graphics(
  here::here("outputs", "04_outputs_figures", "figure4.png")
)

```

## Success along the foraging specialization continuum

The predator's mean speed and IIV were only weakly negatively correlated with hunting success across experience (Figure 4 and Appendix S4: Table S1), suggesting that slower/flexible and faster/specialist predators were overall both equally successful. There was however a slight increase in the correlation between the predator's IIV in speed and success (Figure 4C), suggesting that flexible foragers at the advanced stage were marginally less successful (Appendix S1: Figure S2). We also found strong evidence that individuals who encountered groups of prey with faster speeds were less successful (Figure 4A-B and Appendix S4: Table S1). Predators also had greater success across the study when encountering more variable groups of prey (Figure 4A-B and Appendix S4: Table S1). Both relationships remained structurally stable across experience (Figure 4C).

# DISCUSSION

A general assumption of predator-prey studies is that predators maximize success by matching their tactic to their prey [@abrams2000], yet, it has remained largely unknown whether this results from predators learning how to hunt their prey in part because of the challenges of investigating direct interactions in the wild. By capitalizing on a virtual predator-prey system where interactions were directly monitored, we found that predators in *Dead by Daylight* increased their hunting success with experience, suggesting that predator expertise was honed through extensive practice. While there was no directional increase in either behavioural specialization or flexibility with experience at the population level, individual predators developed their own behavioural trajectories over successive encounters with the prey. This was characterized by a continuum of hunting styles ranging from a slower and flexible strategy to a highly specialized fast-paced strategy, with some predators switching between the two with experience. The behavioural strategies of predators and prey tended to match across experience, suggesting that both were adjusting to each other. Specialized cursorial hunters encountered similar groups of prey, while slower and flexible hunters encountered varying groups of prey. Overall, predators along the flexible-specialist continuum achieved similar success, although flexible hunters were slightly less successful as advanced hunters.

## The development of expertise

Our results suggest that predator expertise is honed through extensive practice. The predator population increased and stabilized its success with experience in a diminishing returns fashion as it is typically found in empirical studies of expertise [reviewed in @dukas2019]. The prey were important in mediating this pattern, probably because they increased their speed as predators gained experience. We previously showed in *DBD* that faster movement is an effective strategy used by the prey to avoid predation [@fraserfranco.etal2022], and other studies have found that as well [@walker.etal2005;@kelley.magurran2011;@martin.etal2022]. This resulted in discernable differences in the relationship between success and experience among predators, implying that the development of expertise depends on an individual's capacity to adjust to difficulty. Hunting faster prey requires costly and specialized cognitive abilities and coordination [@kelley.magurran2011], and that could explain why such prey were more difficult to capture. Thus, predators that couldn't properly hunt at high speeds or develop counter-strategies for faster prey were likely at a disadvantage.

However, even after we took the prey's speed into account, there were still important differences in expertise acquisition among predators, suggesting that other antipredator tactics were potentially involved. For example, if the prey used camouflage or hiding to avoid being detected, then predators may have found them only if they learned how to exploit the visual cues that facilitate cryptic prey detection, such as habitat characteristics or prey colour patterns [@ehlinger1989;@hughes.etal1992;@warburton2003;@szopa-comley.etal2020]. In addition, longer time intervals between hunting events may delay or even impede learning due to forgetting important information [@endler1991;@wright.etal2022]. In *DBD*, a predator that played 300 matches in the span of six months might forget more critical information related to prey detection or escape patterns than one that played 300 matches in the span of six days. Investigating the impact of such time lags in future analyses may reveal important insights on the outcome of predator-prey interactions.

## Predator-prey interactions across experience

The predator population maintained its speed and degree of specialization throughout the study period, and individuals differed only slightly in their average speeds. However, differences in specialization among individuals were important and increased slightly with experience. Predators hunting at higher speeds specialized in this tactic and tended to encounter groups of prey with similar speeds, while those hunting at slower speeds were flexible and tended to encounter varying groups of prey. On one hand, this pattern could arise if some predators adjusted their strategy to their prey, because fast-paced hunting is a specialized tactic suited for prey that use rapid evasive movements, while a slower and flexible tactic can be useful to minimize the consequences of uncertainty when prey escape unpredictably [@endler1991;@bro-jorgensen2013;@wilson.etal2018;@szopa-comley.ioannou2022]. However, we observed that 44% of the predator population showed close to no change in specialization with experience. Since these individuals were already specialized as novices (Fraser Franco, personal observation), there would be no point in them changing their tactic with experience if they were already successful, or there could have been costs to switch their strategy if they more often encountered faster prey. This could explain why the behavioural interactions between the predator and the prey remained stable across the predator's experience. Yet, because prey can also learn how to avoid predation [@kelley2003;@turner2006;@mccomb2011], we believe that the predator-prey phenotype matching more likely emerged from reciprocal adjustments in speed by predators and prey as they interacted [@kishida.etal2006;@kishida.etal2009;@edgell.rochette2009;@mcghee.etal2013]. Indeed, 56% of the predator population altered its behaviour with experience to different degrees, including 12% that even switched between flexibility and specialization. Hence, if the prey also learned through repeated interactions with the predators, it is possible that experience contributed in stabilizing the system as both were adjusting to each other, similar to Red Queen dynamics [@brockhurst2014].

Classical theory predicts that individual foraging specialization should maximize prey consumption by optimizing foraging efficiency [@stephens1986], but other work suggests that flexibility and specialization could both be adaptive as resources fluctuate [@woo.etal2008;@phillips.etal2017]. Our results support empirical findings indicating that flexible and specialist foragers obtained similar success [@weimerskirch2007;@woo.etal2008;@potier.etal2015;@phillips.etal2017;@courbin.etal2018], although flexible foragers were slightly less successful. Both strategies appeared to emerge in part from individuals learning how to hunt their prey, resulting in a general increase in success in the population. However, there were still considerable differences in success among individuals through time, suggesting that some predators were limited in their capacity to match their tactic to their prey or to increase their success through other means. Specialist foragers were faster, and thus, probably better equipped to hunt the more difficult faster prey in *DBD*. However, if the prey responded to fast predators by also being faster, then hunting at high speeds resulted in more difficult encounters for these predators, thereby decreasing the benefits of using this tactic (Figure 4). Thus, specializing probably compensated for the difficulty of hunting prey at high speeds by helping predators to better predict the location and movement of their prey. On the other hand, flexible foragers encountered variable prey with slower speeds. Yet, because the prey increased their speed with experience, the benefits of being able to hunt multiple prey types for flexible hunters may have come at the cost of not being adept at capturing faster prey [@pintor.etal2014]. Such trade-offs may reflect limitations in learning all the skills required to successfully hunt all types of prey [@healy1992;@belisle1997;@dukas2019], particularly if the skills required to hunt slower prey are nontransferable to faster prey. As the benefits of each hunting style changed under different scenarios, the combination of predatory styles in this virtual system reflects how resource fluctuations can maintain fitness equilibrium within populations, resulting in behavioural diversity in predator-prey systems [@tinker.etal2008;@woo.etal2008;@mougi.kishida2009].

## Conclusions

We found support of our prediction that prey were driving individual differences in expertise in a predator population. We also confirmed that experience and predator-prey encounters drove the patterns of individual foraging specialization. Our results suggest that predators learned with experience, as their success increased and their speed remained matched to the speed of their prey.  Even though individuals were not all equally successful, both specialized and flexible hunters achieved similar success. A potential caveat is that the more flexible hunters might have experimented with various tactics out of boredom, which could impede ecologically realistic interpretations of our data. However, the consistent association between this tactic and heterogeneous prey groups gives us confidence that it emerged from their interactions. One limitation of our study was that we couldn't monitor all the matches of the prey, which prevented us from assessing their responses to the predator through their experience. Hence, future studies should aim at monitoring reciprocal behavioural dynamics through time, which may reveal important insights on the mechanisms driving stable equilibria in predator-prey systems. Lastly, virtual systems are increasingly recognized as being useful to test hypotheses on ecological interactions [@beauchamp2020;@cere.etal2021;@fraserfranco.etal2022;@lymbery.etal2023]. We therefore hope that our study will inspire more collaborations between scientists and the videogame industry to tackle fundamental questions in ecology.

# REFERENCES {.unnumbered}

<div id="refs"></div>

### 2024-06-04 code/code_figures/appendix1_figureS1.R
added old code files to update

@@ -1,0 +1,290 @@
+# ==========================================================================
+
+#                Code to produce Appendix 1 : Figure S1                    #
+
+# ==========================================================================
+
+
+
+
+
+# ==========================================================================
+# 1. Prepare script
+# ==========================================================================
+
+
+# Load libraries and models ------------------------------------------------
+
+library(brms)
+library(data.table)
+library(ggplot2)
+library(ggpubr)
+
+path <- file.path(getwd(), "outputs", "01_outputs_models")
+
+fit1 <- readRDS(file.path(path, "A2_GAMM.rds"))
+#fit1 <- readRDS(file.path(path, "A2_GAMM-speed-rank.rds"))
+
+
+
+# Load the data ------------------------------------------------------------
+
+data <- fread("./data/FraserFrancoetal2023-data.csv",
+              select = c("predator_id",
+                         "game_duration",
+                         "pred_speed",
+                         "prey_avg_speed",
+                         "cumul_xp_pred",
+                         "total_xp_pred",
+                         "hunting_success"))
+
+data[, predator_id := as.factor(predator_id)]
+
+
+
+# Post processing preparations for custom family ---------------------------
+
+expose_functions(fit1, vectorize = TRUE)
+
+# Define the log likelihood function
+log_lik_beta_binomial2 <- function(i, prep) {
+  mu <- brms::get_dpar(prep, "mu", i = i)
+  phi <- brms::get_dpar(prep, "phi", i = i)
+  trials <- prep$data$vint1[i]
+  y <- prep$data$Y[i]
+  beta_binomial2_lpmf(y, mu, phi, trials)
+}
+
+# Define function for posterior_predict
+posterior_predict_beta_binomial2 <- function(i, prep, ...) {
+  mu <- brms::get_dpar(prep, "mu", i = i)
+  phi <- brms::get_dpar(prep, "phi", i = i)
+  trials <- prep$data$vint1[i]
+  beta_binomial2_rng(mu, phi, trials)
+}
+
+# Define function for posterior_epred
+posterior_epred_beta_binomial2 <- function(prep) {
+  mu <- brms::get_dpar(prep, "mu")
+  trials <- prep$data$vint1
+  trials <- matrix(trials, nrow = nrow(mu), ncol = ncol(mu), byrow = TRUE)
+  mu * trials
+}
+
+
+
+# Setup a custom theme for the plot ----------------------------------------
+
+custom_theme <- theme(
+  # axis values size
+  axis.text = element_text(face = "plain",
[…]

### 2024-06-04 code/code_figures/figure1.R

@@ -1,0 +1,413 @@
+# ==========================================================================
+
+#                           Plot GAMM models
+
+# ==========================================================================
+
+
+
+# ==========================================================================
+# 1. Prepare script
+# ==========================================================================
+
+
+
+# Load libraries and model -------------------------------------------------
+
+ options(mc.cores = parallel::detectCores()) 
+
+ library(parallel)
+ library(brms)
+ library(data.table)
+ library(ggplot2)
+ library(ggpubr)
+ library(viridis)
+
+ path <- file.path(getwd(), "outputs", "01_outputs_models")
+
+ modA2 <- readRDS(file.path(path, "A2_GAMM-rank.rds"))
+ modA2_prey <- readRDS(file.path(path, "A2_GAMM-speed-rank.rds"))
+
+
+
+# Load the data ------------------------------------------------------------
+
+ data <- fread("./data/FraserFrancoetal2023-data.csv",
+               select = c("predator_id",
+                          "game_duration",
+                          "pred_speed",
+                          "prey_avg_speed",
+                          "prey_avg_rank",
+                          "cumul_xp_pred",
+                          "total_xp_pred",
+                          "hunting_success"))
+
+ data[, predator_id := as.factor(predator_id)]
+
+
+
+# Post processing preparations for custom family ---------------------------
+
+ expose_functions(modA2, vectorize = TRUE)
+
+ # Define the log likelihood function
+ log_lik_beta_binomial2 <- function(i, prep) {
+   mu <- brms::get_dpar(prep, "mu", i = i)
+   phi <- brms::get_dpar(prep, "phi", i = i)
+   trials <- prep$data$vint1[i]
+   y <- prep$data$Y[i]
+   beta_binomial2_lpmf(y, mu, phi, trials)
+ }
+
+ # Define function for posterior_predict
+ posterior_predict_beta_binomial2 <- function(i, prep, ...) {
+   mu <- brms::get_dpar(prep, "mu", i = i)
+   phi <- brms::get_dpar(prep, "phi", i = i)
+   trials <- prep$data$vint1[i]
+   beta_binomial2_rng(mu, phi, trials)
+ }
+
+ # Define function for posterior_epred
+ posterior_epred_beta_binomial2 <- function(prep) {
+   mu <- brms::get_dpar(prep, "mu")
[…]

### 2024-06-04 code/code_model-processing/GAMM_percentages.R

@@ -1,0 +1,242 @@
+# ==========================================================================
+
+#                       Code to calculate percentages                      #
+
+# ==========================================================================
+
+
+
+
+
+# ==========================================================================
+# 1. Prepare script
+# ==========================================================================
+
+
+# Load libraries and models ------------------------------------------------
+
+library(brms)
+library(data.table)
+
+path <- file.path(getwd(), "outputs", "01_outputs_models")
+
+fit2 <- readRDS(file.path(path, "A2_GAMM.rds"))
+fit2p <- readRDS(file.path(path, "A2_GAMM-speed-rank.rds"))
+
+
+
+# Load the data ------------------------------------------------------------
+
+data <- fread("./data/FraserFrancoetal2023-data.csv",
+              select = c("predator_id",
+                         "game_duration",
+                         "pred_speed",
+                         "prey_avg_speed",
+                         "cumul_xp_pred",
+                         "total_xp_pred",
+                         "hunting_success"))
+
+data[, predator_id := as.factor(predator_id)]
+
+
+
+# Post processing preparations for custom family ---------------------------
+
+expose_functions(fit2, vectorize = TRUE)
+
+# Define the log likelihood function
+log_lik_beta_binomial2 <- function(i, prep) {
+  mu <- brms::get_dpar(prep, "mu", i = i)
+  phi <- brms::get_dpar(prep, "phi", i = i)
+  trials <- prep$data$vint1[i]
+  y <- prep$data$Y[i]
+  beta_binomial2_lpmf(y, mu, phi, trials)
+}
+
+# Define function for posterior_predict
+posterior_predict_beta_binomial2 <- function(i, prep, ...) {
+  mu <- brms::get_dpar(prep, "mu", i = i)
+  phi <- brms::get_dpar(prep, "phi", i = i)
+  trials <- prep$data$vint1[i]
+  beta_binomial2_rng(mu, phi, trials)
+}
+
+# Define function for posterior_epred
+posterior_epred_beta_binomial2 <- function(prep) {
+  mu <- brms::get_dpar(prep, "mu")
+  trials <- prep$data$vint1
+  trials <- matrix(trials, nrow = nrow(mu), ncol = ncol(mu), byrow = TRUE)
+  mu * trials
+}
+
[…]

### 2024-06-04 code/code_model-validation/GAMM-diagnostics.R

@@ -1,0 +1,221 @@
+# =======================================================================
+
+#                       GAMM model diagnostics                          #
+
+# =======================================================================
+
+
+
+
+
+# =======================================================================
+# 1. Load libraries, datasets, and models
+# =======================================================================
+
+
+
+# Librairies ------------------------------------------------------------
+
+ # Detect cores
+ options(mc.cores = parallel::detectCores())
+
+ library(data.table)
+ library(brms)
+ library(parallel)
+ library(ggpubr)
+
+
+
+# Import the model ------------------------------------------------------
+
+ path <- file.path(getwd(), "outputs", "01_outputs_models")
+ #path <- getwd()
+
+ # Load models
+ mod1 <- readRDS(file.path(path, "A1_GAMM-rank.rds"))
+ mod2 <- readRDS(file.path(path, "A2_GAMM-rank.rds"))
+ mod3 <- readRDS(file.path(path, "A3_GAMM-rank.rds"))
+ mod4 <- readRDS(file.path(path, "A2_GAMM-speed-rank.rds"))
+ mod5 <- readRDS(file.path(path, "A3_GAMM-speed-rank.rds"))
+
+ # Check object size
+ #print(object.size(mod1), units = "MB")
+ #print(object.size(mod2), units = "MB")
+ #print(object.size(mod3), units = "MB")
+ #print(object.size(mod4), units = "MB")
+ #print(object.size(mod5), units = "MB")
+
+# =======================================================================
+# =======================================================================
+
+
+
+
+# =======================================================================
+# 2. Prepare functions for post-processing (beta-binom custom family)
+# =======================================================================
+
+ # Expose functions
+ expose_functions(mod1, vectorize = TRUE)
+
+ # Define the log likelihood function
+ log_lik_beta_binomial2 <- function(i, prep) {
+   mu <- brms::get_dpar(prep, "mu", i = i)
[…]

### 2024-06-04 code/code_model-validation/GAMM-loocv.R

@@ -1,0 +1,144 @@
+# =======================================================================
+
+#                 GAMM leave-one-out cross-validation                   #
+
+# =======================================================================
+
+
+
+
+
+# =======================================================================
+# 1. Load libraries, datasets, and models
+# =======================================================================
+
+
+
+# Librairies ------------------------------------------------------------
+
+ # Detect cores
+ options(mc.cores = parallel::detectCores())
+
+ library(data.table)
+ library(brms)
+ library(parallel)
+
+
+
+# Import the model ------------------------------------------------------
+
+ path <- file.path(getwd(), "outputs", "01_outputs_models")
+
+ # Load models
+ mod1 <- readRDS(file.path(path, "A1_GAMM-rank.rds"))
+ mod2 <- readRDS(file.path(path, "A2_GAMM-rank.rds"))
+ mod3 <- readRDS(file.path(path, "A3_GAMM-rank.rds"))
+ mod4 <- readRDS(file.path(path, "A2_GAMM-speed-rank.rds"))
+ mod5 <- readRDS(file.path(path, "A3_GAMM-speed-rank.rds"))
+
+ # Check object size
+ #print(object.size(mod1), units = "MB")
+ #print(object.size(mod2), units = "MB")
+ #print(object.size(mod3), units = "MB")
+ #print(object.size(mod4), units = "MB")
+ #print(object.size(mod5), units = "MB")
+
+# =======================================================================
+# =======================================================================
+
+
+
+
+
+# =======================================================================
+# 2. Prepare functions for post-processing (beta-binom custom family)
+# =======================================================================
+
+ # Expose functions
+ expose_functions(mod1, vectorize = TRUE)
+
+ # Define the log likelihood function
+ log_lik_beta_binomial2 <- function(i, prep) {
+   mu <- brms::get_dpar(prep, "mu", i = i)
+   phi <- brms::get_dpar(prep, "phi", i = i)
+   trials <- prep$data$vint1[i]
+   y <- prep$data$Y[i]
+   beta_binomial2_lpmf(y, mu, phi, trials)
+ }
+
+ # Define function for posterior_predict
+ posterior_predict_beta_binomial2 <- function(i, prep, ...) {
+   mu <- brms::get_dpar(prep, "mu", i = i)
+   phi <- brms::get_dpar(prep, "phi", i = i)
[…]

### 2024-06-04 code/code_model-validation/GAMM_models-draws.R

@@ -1,0 +1,263 @@
+# =============================================================================
+
+#                       Inspect prior and posterior draws
+#                               from GAMM-GLMM models
+
+# =============================================================================
+
+
+
+
+
+# =============================================================================
+# Prepare session
+# =============================================================================
+
+
+# Load libraries --------------------------------------------------------------
+
+ library(brms)
+ library(ggplot2)
+
+
+
+# Prepare model draws ---------------------------------------------------------
+
+ path <- file.path(getwd(), "outputs", "01_outputs_models")
+
+ # Import model in R session
+ mod1 <- readRDS(file.path(path, "A1_GAMM-rank.rds"))
+ mod2 <- readRDS(file.path(path, "A2_GAMM-rank.rds"))
+ mod3 <- readRDS(file.path(path, "A3_GAMM-rank.rds"))
+ mod2p <- readRDS(file.path(path, "A2_GAMM-speed-rank.rds"))
+ mod3p <- readRDS(file.path(path, "A3_GAMM-speed-rank.rds"))
+
+
+ # Extract posterior draws
+ posterior_fit1 <- as_draws_df(mod1)
+ posterior_fit2 <- as_draws_df(mod2)
+ posterior_fit3 <- as_draws_df(mod3)
+ posterior_fit4 <- as_draws_df(mod2p)
+ posterior_fit5 <- as_draws_df(mod3p)
+
+# =============================================================================
+# =============================================================================
+
+
+
+
+
+# =============================================================================
+# 1. Model 1
+# =============================================================================
+
+
+
+# Plot prior and posterior draws ----------------------------------------------
+
+ # Intercept
+ ggplot(posterior_fit1) +
+   geom_density(aes(prior_Intercept),
+                fill = "steelblue",
+                color = "black",
+                alpha = 0.6) +
+   geom_density(aes(Intercept),
+                fill = "#FC4E07",
+                color = "black",
+                alpha = 0.6) + 
+   theme_classic()
+
+
+ # Game duration
+ ggplot(posterior_fit1) +
+   geom_density(aes(prior_b_Zgame_duration),
+                fill = "steelblue",
+                color = "black",
+                alpha = 0.6) +
+   geom_density(aes(b_Zgame_duration),
+                fill = "#FC4E07",
+                color = "black",
+                alpha = 0.6) +
+   theme_classic()
+
+ # slope term of cumul xp
+ ggplot(posterior_fit1) +
+   geom_density(aes(prior_bs_sZcumul_xp_1),
[…]

### 2024-06-04 code/code_models/A1_GAMM-rank.R

@@ -1,0 +1,178 @@
+# ==========================================================================
+
+#            GAMM model G - single common smoother - Hunting success
+
+# ==========================================================================
+
+
+
+
+
+# ==========================================================================
+# 1. Prepare the data
+# ==========================================================================
+
+
+
+# Load packages ------------------------------------------------------------
+
+# Detect cores
+options(mc.cores = parallel::detectCores())
+
+# Load libraries
+library(data.table)
+library(brms)
+library(parallel)
+
+
+
+# Load data ----------------------------------------------------------------
+
+# Folder path Compute Canada
+folder <- file.path("/home", "maxime11", "projects", "def-monti", 
+                    "maxime11", "phd_project", "data")
+
+# Load data on compute canada
+data <- fread(
+  file.path(folder, "FraserFrancoetalXXXX-data.csv"),
+  select = c("predator_id",
+             "hunting_success",
+             "prey_avg_speed",
+             "game_duration",
+             "cumul_xp_pred",
+             "prey_avg_rank")
+)
+
+# Predator id as factor
+data[, predator_id := as.factor(predator_id)]
+
+# Remove any NAs
+data <- data[complete.cases(data)]
+
+
+
+# Standardise the variables (Z-scores) -------------------------------------
+
+standardize <- function(x) {
+  (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)
+}
+
+data[, c(
+    "Zprey_speed",
+    "Zgame_duration",
+    "Zcumul_xp",
+    "Zprey_avg_rank"
+    ) := lapply(.SD, standardize),
+       .SDcols = c(3:6)]
+
+# ==========================================================================
+# ==========================================================================
+
+
+
+
+
+# ==========================================================================
+# 2. Parametrize the model
+# ==========================================================================
+
+
+
+# Compute the custom family ------------------------------------------------
+
+beta_binomial2 <- custom_family(
+  "beta_binomial2", dpars = c("mu", "phi"),
+  links = c("logit", "log"), lb = c(NA, 0),
+  type = "int", vars = "vint1[n]"
+)
+
+
+
+# Function -----------------------------------------------------------------
+
+stan_funs <- "
+  real beta_binomial2_lpmf(int y, real mu, real phi, int T) {
+    return beta_binomial_lpmf(y | T, mu * phi, (1 - mu) * phi);
+  }
[…]

### 2024-06-04 code/code_models/A2_GAMM-rank.R

@@ -1,0 +1,178 @@
+# ==========================================================================
+
+#           GAMM model GS - group-level smoother - hunting success
+
+# ==========================================================================
+
+
+
+
+
+# ==========================================================================
+# 1. Prepare the data
+# ==========================================================================
+
+
+
+# Load packages ------------------------------------------------------------
+
+# Detect cores
+options(mc.cores = parallel::detectCores())
+
+# Load libraries
+library(data.table)
+library(brms)
+library(parallel)
+
+
+
+# Load data ----------------------------------------------------------------
+
+# Folder path Compute Canada
+folder <- file.path("/home", "maxime11", "projects", "def-monti", 
+                    "maxime11", "phd_project", "data")
+
+# Load data
+data <- fread(
+  file.path(folder, "FraserFrancoetalXXXX-data.csv"),
+  select = c("predator_id",
+             "hunting_success",
+             "prey_avg_speed",
+             "game_duration",
+             "cumul_xp_pred",
+             "prey_avg_rank")
+)
+
+# Predator id as factor
+data[, predator_id := as.factor(predator_id)]
+
+# Remove any NAs
+data <- data[complete.cases(data)]
+
+
+
+# Standardise the variables (Z-scores) -------------------------------------
+
+standardize <- function(x) {
+  (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)
+}
+
+data[, c(
+    "Zprey_speed",
+    "Zgame_duration",
+    "Zcumul_xp",
+    "Zprey_avg_rank"
+    ) := lapply(.SD, standardize),
+       .SDcols = c(3:6)]
+
+
+# ==========================================================================
+# ==========================================================================
+
+
+
+
+
+# ==========================================================================
+# 2. Parametrize the model
+# ==========================================================================
+
+
+
+# Compute the custom family ------------------------------------------------
+
+beta_binomial2 <- custom_family(
+  "beta_binomial2", dpars = c("mu", "phi"),
+  links = c("logit", "log"), lb = c(NA, 0),
+  type = "int", vars = "vint1[n]"
+)
+
+
+
+# Function -----------------------------------------------------------------
+
+stan_funs <- "
+  real beta_binomial2_lpmf(int y, real mu, real phi, int T) {
+    return beta_binomial_lpmf(y | T, mu * phi, (1 - mu) * phi);
+  }
[…]

### 2024-06-04 code/code_models/A2_GAMM-speed-rank.R

@@ -1,0 +1,181 @@
+# ==========================================================================
+
+#           GAMM model GS - group-level smoother - hunting success
+
+# ==========================================================================
+
+
+
+
+
+# ==========================================================================
+# 1. Prepare the data
+# ==========================================================================
+
+
+
+# Load packages ------------------------------------------------------------
+
+# Detect cores
+options(mc.cores = parallel::detectCores())
+
+# Load libraries
+library(data.table)
+library(brms)
+library(parallel)
+
+
+
+# Load data ----------------------------------------------------------------
+
+# Folder path Compute Canada
+folder <- file.path("/home", "maxime11", "projects", "def-monti", 
+                    "maxime11", "phd_project", "data")
+
+# Load data
+data <- fread(
+  file.path(folder, "FraserFrancoetalXXXX-data.csv"),
+  select = c("predator_id",
+             "hunting_success",
+             "prey_avg_speed",
+             "game_duration",
+             "cumul_xp_pred",
+             "prey_avg_rank")
+)
+
+# Predator id as factor
+data[, predator_id := as.factor(predator_id)]
+
+# Remove any NAs
+data <- data[complete.cases(data)]
+
+
+
+# Standardise the variables (Z-scores) -------------------------------------
+
+standardize <- function(x) {
+  (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)
+}
+
+data[, c(
+    "Zprey_speed",
+    "Zgame_duration",
+    "Zcumul_xp",
+    "Zprey_avg_rank"
+    ) := lapply(.SD, standardize),
+       .SDcols = c(3:6)]
+
+
+# ==========================================================================
+# ==========================================================================
+
+
+
+
+
+# ==========================================================================
+# 2. Parametrize the model
+# ==========================================================================
+
+
+
+# Compute the custom family ------------------------------------------------
+
+beta_binomial2 <- custom_family(
+  "beta_binomial2", dpars = c("mu", "phi"),
+  links = c("logit", "log"), lb = c(NA, 0),
+  type = "int", vars = "vint1[n]"
+)
+
+
+
+# Function -----------------------------------------------------------------
+
+stan_funs <- "
+  real beta_binomial2_lpmf(int y, real mu, real phi, int T) {
+    return beta_binomial_lpmf(y | T, mu * phi, (1 - mu) * phi);
+  }
+  int beta_binomial2_rng(real mu, real phi, int T) {
+    return beta_binomial_rng(T, mu * phi, (1 - mu) * phi);
[…]

### 2024-06-04 code/code_models/A3_GAMM-rank.R

@@ -1,0 +1,173 @@
+# ==========================================================================
+
+#         GAMM model S - individual curves without common smoother
+#                              hunting success
+
+# ==========================================================================
+
+
+
+
+
+# ==========================================================================
+# 1. Prepare the data
+# ==========================================================================
+
+
+
+# Load packages ------------------------------------------------------------
+
+# Detect cores
+options(mc.cores = parallel::detectCores())
+
+# Load libraries
+library(data.table)
+library(brms)
+library(parallel)
+
+
+
+# Load data ----------------------------------------------------------------
+
+# Folder path Compute Canada
+folder <- file.path("/home", "maxime11", "projects", "def-monti", 
+                    "maxime11", "phd_project", "data")
+
+# Load data
+data <- fread(
+  file.path(folder, "FraserFrancoetalXXXX-data.csv"),
+  select = c("predator_id",
+             "hunting_success",
+             "prey_avg_speed",
+             "game_duration",
+             "cumul_xp_pred",
+             "prey_avg_rank")
+)
+
+# Predator id as factor
+data[, predator_id := as.factor(predator_id)]
+
+# Remove any NAs
+data <- data[complete.cases(data)]
+
+
+
+# Standardise the variables (Z-scores) -------------------------------------
+
+standardize <- function(x) {
+  (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)
+}
+
+data[, c(
+    "Zprey_speed",
+    "Zgame_duration",
+    "Zcumul_xp",
+    "Zprey_avg_rank"
+    ) := lapply(.SD, standardize),
+       .SDcols = c(3:6)]
+
+# ==========================================================================
+# ==========================================================================
+
+
+
+
+
+# ==========================================================================
+# 2. Parametrize the model
+# ==========================================================================
+
+
+
+# Compute the custom family ------------------------------------------------
+
+beta_binomial2 <- custom_family(
+  "beta_binomial2", dpars = c("mu", "phi"),
+  links = c("logit", "log"), lb = c(NA, 0),
+  type = "int", vars = "vint1[n]"
+)
+
+
+
+# Function -----------------------------------------------------------------
+
+stan_funs <- "
+  real beta_binomial2_lpmf(int y, real mu, real phi, int T) {
+    return beta_binomial_lpmf(y | T, mu * phi, (1 - mu) * phi);
+  }
[…]

### 2024-06-04 code/code_models/A3_GAMM-speed-rank.R

@@ -1,0 +1,177 @@
+# ==========================================================================
+
+#         GAMM model S - individual curves without common smoother
+#                              hunting success
+
+# ==========================================================================
+
+
+
+
+
+# ==========================================================================
+# 1. Prepare the data
+# ==========================================================================
+
+
+
+# Load packages ------------------------------------------------------------
+
+# Detect cores
+options(mc.cores = parallel::detectCores())
+
+# Load libraries
+library(data.table)
+library(brms)
+library(parallel)
+
+
+
+# Load data ----------------------------------------------------------------
+
+# Folder path Compute Canada
+folder <- file.path("/home", "maxime11", "projects", "def-monti", 
+                    "maxime11", "phd_project", "data")
+
+# Load data
+data <- fread(
+  file.path(folder, "FraserFrancoetalXXXX-data.csv"),
+  select = c("predator_id",
+             "hunting_success",
+             "prey_avg_speed",
+             "game_duration",
+             "cumul_xp_pred",
+             "prey_avg_rank")
+)
+
+# Predator id as factor
+data[, predator_id := as.factor(predator_id)]
+
+# Remove any NAs
+data <- data[complete.cases(data)]
+
+
+
+# Standardise the variables (Z-scores) -------------------------------------
+
+standardize <- function(x) {
+  (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)
+}
+
+data[, c(
+    "Zprey_speed",
+    "Zgame_duration",
+    "Zcumul_xp",
+    "Zprey_avg_rank"
+    ) := lapply(.SD, standardize),
+       .SDcols = c(3:6)]
+
+# ==========================================================================
+# ==========================================================================
+
+
+
+
+
+# ==========================================================================
+# 2. Parametrize the model
+# ==========================================================================
+
+
+
+# Compute the custom family ------------------------------------------------
+
+beta_binomial2 <- custom_family(
+  "beta_binomial2", dpars = c("mu", "phi"),
+  links = c("logit", "log"), lb = c(NA, 0),
+  type = "int", vars = "vint1[n]"
+)
+
+
+
+# Function -----------------------------------------------------------------
+
+stan_funs <- "
+  real beta_binomial2_lpmf(int y, real mu, real phi, int T) {
+    return beta_binomial_lpmf(y | T, mu * phi, (1 - mu) * phi);
+  }
+  int beta_binomial2_rng(real mu, real phi, int T) {
[…]

### 2024-06-04 manuscript/manuscript.Rmd
updated intro

@@
 ---
-title: "Experience shapes individual foraging specialization and success in a virtual predator-prey system"
+title: "Prey movement shapes the development of predator expertise in a virtual bi-trophic system"
 output: 
@@
 ```{r setup, include = FALSE}
-
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
@@
 library(dplyr)
-
 ```
@@
 ```{r}
-
 path <- file.path(here::here(), "data")
@@
 dat <- fread(
-  file.path(path, "FraserFrancoetal2023-data.csv"),
-  select = c("timestamp", "match_id", "realm_id",
+  file.path(
+    path, "FraserFrancoetal2023-data.csv"
+  ),
+  select = c(
+    "timestamp", "match_id", "realm_id",
     "environment_id", "avatar_id", "predator_id",
     "game_duration", "total_xp_pred", "cumul_xp_pred",
-             "hunting_success", "pred_speed", "prey_avg_speed"))
+    "hunting_success", "pred_speed", "prey_avg_speed"
+  )
+)
 
@@
 dat[, date := as.POSIXct(sub(" .*", "", timestamp))]
-
 ```
@@
 
-The capacity of predators to match their tactic to their prey is expected to drive the outcome of predator-prey interactions.
-To do so, predators need to practice their tactics extensively to develop the skills to be successful hunters, which may result in individual foraging specialization.
-Yet, there are limited empirical assessments showing links between experience, foraging specialization, and hunting success at the individual level, due to the challenges of monitoring direct interactions in the wild.
-Here, we used a virtual predator-prey system (the game *Dead by Daylight*) to investigate how experience shapes individual foraging specialization and success in predators across repeated interactions with their prey.
+The capacity of predators to match their hunting tactic to their prey is expected to drive the outcome of predator-prey interactions.
+To do so, predators need to practice their tactics extensively to develop the expertise to be successful hunters.
+Yet, there are limited empirical assessments showing links between experience and hunting success at the individual and population levels due to the challenges of monitoring direct interactions in the wild.
+Here, we use a virtual predator-prey system (the game *Dead by Daylight*) to investigate how experience shapes individual and population hunting success in predators across repeated interactions with their prey.
 We show that predators optimized prey consumption as they gained experience.
 Prey speed was an important mediator of this relationship, resulting in differences among predators in the development of expertise.
-Throughout experience, 68% of predators became either increasingly specialized by always moving at a fast pace, or flexible by switching between slow and fast speeds.
-The predators’ strategies were partially matched to the preys’ speed, suggesting that changes in hunting behaviour were driven by repeated encounters with prey.
-Flexible and specialist foragers achieved similar success overall.
-Our findings suggest that experience may be mechanism promoting behavioural diversification in predator-prey systems.
+**Décrire le reste des résultats...**
 
-Keywords: foraging behaviour, reciprocal behavioural plasticity, learning, antipredator behaviour, virtual ecology, Dead by Daylight
+Keywords: foraging behaviour, experience, learning, antipredator behaviour, virtual ecology, Dead by Daylight
 
@@
 
-Predators express a diverse array of foraging tactics to capture their prey.
-In natural populations, differences among individuals in tactics such as ambushing and fast movement speeds are associated with differences in prey consumption [@toscano2014;@patrick.weimerskirch2014a;@toscano.etal2016].
-Prey usually respond to predation cues by reducing their activity, or use rapid or unpredictable movements to successfully escape when they are chased [@walker.etal2005;@kelley.magurran2011;@herbert-read.etal2017].
-This results in reciprocal interactions where predators fine-tune their tactics to their prey, suggesting that they learn and become efficient hunters through repeated encounters [@estes.etal2003;@woo.etal2008;@phillips.etal2017].
-These behavioural adjustments with experience can shape the outcome of predator-prey interactions, and may have larger implications for community and ecosystem dynamics [@wooster.etal2023].
+The ecological consequences of predation are defined by the sum of consumptive and nonconsumptive effects (CEs and NCEs) during predator-prey interactions [refs].
+Recent advances in our understanding of the behavioural mechanisms (e.g.
+reciprocal plasticity, foraging specialization) driving prey consumption at the individual and population level have allowed ecologists to rely on foraging strategies to predict the outcome of (N)CEs and their effect on prey population dynamics [refs].
+For example, empirical studies show that prey tend to increase vigilance at the expense of foraging in the presence of ambush predators (NCEs) [@preisser2007].
+Because ambush predators are opportunistic hunters, they also tend to consume a greater range of prey types (CEs) compared to cursorial predators [@glaudas2019].
+
+While the foraging mode of predators is a reliable predictor of prey consumption, predators need to acquire expertise by practicing and learning the proper skills to successfuly capture their prey [@dukas2019;wooster.etal2023].
+The aquisition of expertise is predicted to be paramount in many predator systems where hunting tactics are learned (e.g.
+orca whales).
+Yet, there is only fraction of evidence showing the link between experience and success in natural predators, impeding our ability to accurately predict the consequences of predation on prey populations.
 
 Empirical studies on human and non-human hunters show that individuals optimize foraging efficiency (e.g.
-search and handling times, return rates) by learning associative images or prey and environmental cues [@edwards.jackson1994;@morse2000;@macdonald2007;@reid.etal2010;@wilson-rankin2015].
-As predators practice and refine their tactics, they acquire expertise (e.g.
-skills, knowledge) that allows them to outperform novice hunters [@dukas2019;@wooster.etal2023].
-Through this process, it is hypothesized that individuals will gradually specialize in a foraging tactic (e.g.
-always moving fast) if it allows them to maintain successful prey captures at each attempt [@tinker.etal2009;@dukas2019].
-This should occur when predators learn to hunt in environments where resources are predictable (or stable) because they require less energy and time to search for, capture, and handle prey [@weimerskirch2007;@woo.etal2008;@potier.etal2015].
-Such conditions favor specialist hunters because there would be higher costs to switch tactics for the same type of prey.
-An alternative hypothesis is that learning is a mechanism driving behavioural flexibility.
-For instance, predators can gain expertise and information on their prey through repeated hunting atempts, and may hunt optimally by learning to adjust their tactic to the type of prey that they encounter [@stephens1993;@ishii.shimada2010;@kelley.magurran2011].
-While there are costs to attempt novel tactics by trial and error in stable environments, predators that hunt in environments under fluctuating resources can reduce the consequences of uncertainty by learning to switch hunting tactics [@dall2010;@mathot.etal2012].
-Nonetheless, complex feedback between ecological interactions and learning make it challenging to predict under which circumstances predators should specialize with experience [@stephens1993;@ishii.shimada2010;@vilapouca2021;@beyts2023].
-
-There is increasing evidence suggesting that the emergence of behavioural specialization and its fitness consequences are driven by resource fluctuations [@patrick.weimerskirch2014b;@vandenbosch.etal2019;@santoro.etal2019;@manlick.etal2021].
-For example, @szopa-comley.ioannou2022 showed that predatory blue acaras (*Andinoacara pulcher*) reduce their maximum approach speed when repeatedly exposed to prey that escape unpredictably, enabling them to be as successful as those exposed to predictable prey.
-Prey that use fast escapes are also proven to be more difficult to hunt [@walker.etal2005;@kelley.magurran2011;@martin.etal2022], and predators that exclusively encounter such prey should benefit from learning to specialize in fast pursuits.
-If predators that hunt in different environments adjust their tactics to local prey fluctuations, then specialist and flexible hunters can both attain similar hunting success and coexist [@phillips.etal2017].
-Therefore, empiricists should investigate the link between foraging specialization and success across time and contexts.
+search and handling times, return rates) by learning associative images or using cues from their prey and their environment [@edwards.jackson1994;@morse2000;@macdonald2007;@reid.etal2010;@wilson-rankin2015].
+Predators can integrate this information to use the appropriate hunting tactic (e.g.
+ambushing or fast movement speeds) for the type of prey that they encounter [@toscano2014;@patrick.weimerskirch2014a;@toscano.etal2016].
+However, prey often use antipredator tactics such as rapid or unpredictable movements which are effective to successfully escape predators [@walker.etal2005;@kelley.magurran2011;@herbert-read.etal2017;@szopa-comley.ioannou2022].
+This results in reciprocal interactions where predators need to learn to predict the movement of their prey to capture them.
+Therefore, antipredator tactics may impede the acquisition of hunting expertise by interfering with learning processes and reducing the success of predators.
 
@@
 Similar to agent-based simulations, online videogames can be useful systems to mitigate these challenges and test specific ecological hypotheses.
-They provide controlled virtual environments with the advantage of having real players that interact in the virtual space.
+They provide controlled virtual environments to test ecological hypotheses [see @lymbery.etal2023] with the advantage of having real players that interact in the virtual space.
 For example, in the predator-prey videogame *Dead by Daylight*, four prey players need to forage for resources while avoiding predation by a fifth player.
@@
 mean movement speed along a slow-fast continuum), and their success is driven by the movement of the prey [@fraserfranco.etal2022].
-We found that prey can increase their chances of survival by cooperating and moving fast when they forage [@cere.etal2021;@fraserfranco.etal2022, Santostefano et al.
-in press].
-Of course, the life of neither the predator nor the prey players are at stake, such that emerging patterns could be driven by their motivation to win and not "true" survival.
-Yet, the game ellicits natural reactions in players such as fear from predation (personal observations), which corroborates with another virtual ecological study showing that predation drives individual variation in risk perception [@beauchamp2020].
-In a different system, @lymbery.etal2023 showed that complex environments favor strong soldiers over large armies in a virtual game as well as in the laboratory with two ant species.
+The prey can increase their chances of survival by cooperating and moving fast when they forage [@cere.etal2021;@fraserfranco.etal2022;@santostefano.etal2024].
+The game also ellicits natural reactions in players such as fear from predation (personal observations), which corroborates with another virtual ecological study showing that predation drives individual variation in risk perception [@beauchamp2020].
 These observations outline how ecological patterns can emerge from interactions in virtual systems with fixed rules.
@@
 
-In this study, we assess how repeated experience shapes predator foraging specialization and success using data from players in *Dead by Daylight*.
-Following @fraserfranco.etal2022 we use the predators' mean movement speed along a slow-fast continuum as a proxy for the foraging tactic (ambush-cursorial), and define the level of individual foraging specialization as the intra-individual variance (i.e.
-IIV) in movement speed across matches, where low IIV translates to individual foraging specialization (e.g.
-always using a cursorial tactic).
+In this study, we assess how repeated experience shapes predator foraging success using data from players in *Dead by Daylight*.
 First, we investigate how individual hunting success changes with experience (i.e.
@@
 We expect this pattern to vary among individuals due to differences in prey encounters.
-Second, we assess how foraging specialization emerges with experience.
-If all predators encounter similar groups of prey, we predict that individuals (and thus the population) should all specialize in similar speeds.
-In contrast, if all predators encounter varying groups of prey, then they should all converge towards flexible speeds.
+Second, we investigate how prey movement influences the development of expertise.
+If all individual predators encounter groups of prey with similar speeds, we predict that their gain in expertise will be similar.
+
+In contrast, based on previous results showing that faster prey are more difficult to hunt, then...
+
+if all predators encounter varying groups of prey, then they should all converge towards flexible speeds.
 In both scenarios, differences among individuals in IIV across experience should be low (i.e.
 similar individual foraging specialization), whereas the population variance would either decrease (specialization) or increase (flexibility).
-Alternatively, differences among individuals in foraging specialization may emerge if they experience different interactions with their prey.
-In this case, we expect predators that encountered similar groups of prey across experience to specialize in similar speeds, while predators that encountered heterogeneous groups across matches should adopt flexible movement speeds, resulting in an increase in among individual differences in IIV with experience.
-If we detect such prey-dependent fine-tuning with experience, then all hunters along the flexible-specialist continuum should attain equal success.
 
@@
 The following information is collected and reported for every match : the player's anonymous ID, its avatar (i.e.
-the predator character chosen with its specific skill-gameplay mechanics), the game environment, the predator-player's experience along with its speed, the mean speed of the groups of prey the player encountered, and their mean rank (a proxy for prey skill).
+the predator character chosen with its specific skill-gameplay mechanics), the game environment, the predator-player's experience, the mean speed of the groups of prey the player encountered, and their mean rank (a proxy for prey skill).
 The ranking system in *DBD* was implemented by the company to pair players in a match based on their skill (https://deadbydaylight.fandom.com/wiki/Rank), and failing to account for it would prevent us from detecting a change in the predator's foraging success with experience.
 
-We analyzed the predator's mean speed and the mean speed of the prey group encountered by the predator.
-The predator's mean speed is measured as the mean distance traveled per second during a match (mean= `r toString(format(round(mean(dat$pred_speed), digits = 2), nsmall = 2))` ± `r toString(format(round(sd(dat$pred_speed), digits = 2), nsmall = 2))` m/s).
+We analyzed the mean speed of the prey group encountered by the predator.
 We measured the preys' speed as the mean travel speed of the four individual prey in a match (mean = `r toString(format(round(mean(dat$prey_avg_speed, na.rm = T), digits = 2), nsmall = 2))` ± `r toString(format(round(sd(dat$prey_avg_speed, na.rm = T), digits = 2), nsmall = 2))` m/s).
@@
 
-### Development of expertise
+### Model specification
 
@@
 
-### Foraging behaviour and predator-prey interactions across experience
-
-We tested whether foraging behaviour changed with experience, and compared the success of predators along the flexible-specialist hunter continuum, using a multivariate double-hierarchical generalized linear model (MDHGLM) [@lee.nelder2006;@cleasby.etal2015;@mitchell.etal2016].
-For each experience level, we fitted the model on three traits: the predator's speed ($Y_{1}$) and the prey's speed ($Y_{2}$) using a Gaussian error distribution, and the predator's hunting success ($Y_{3}$) using a beta-binomial error distribution.
-Fixed and random effect structure were trait-specific and the same for all experience levels.
-The predator and the prey speed included a mean and a dispersion part (eqn 1.1, 1.2 and 2.1, 2.2).
-For both traits, we controlled for the standardized prey rank on both the mean and the dispersion parts of the model.
-We included game environment, predator avatar, and individual predator identity random effects on the mean part of the model.
-The dispersion part included individual predator identity random effects exclusively.
-For hunting success, we modeled the mean part only (eqn 3) and controlled for the standardized prey rank and game duration, and included an individual identity random effect.
-
-The model simultaneously estimated five among-individual variance components with their covariances for each experience level, resulting in a 15x15 (co)variance matrix.
-For each experience level, the model estimated among individual variance in mean predator speed, mean prey speed, and mean hunting success.
-This allows us to test for individual differences in tactic use (i.e.
-along the slow-fast continuum), in the mean speed of the prey encountered, and in mean prey consumption.
-The model also estimated among individual variance in intra-individual variation (IIV) for both prey and predator speed.
-This allows us to test if predators do not experience the same degree of prey heterogeneity and if they differ in individual specialization.
-Finally, the model estimated all pairwise covariances at the individual level among traits, among experience, and across traits and experience.
-The resulting 15x15 (co)variance matrix thus provides the structure of the predator-prey trait interactions as well as the relationship between specialization and success across experience at the individual-level.
-
-The model can be written as:
-
-```{r model_equations, ft.align = "left"}
-
-# Prepare equations
-#eqs <- c(
-#  "Y_{1}(T_{j}) = X_{1}(T_{j})\\ b_{1}(T_{j}) + M_{1}(T_{j})\\ en_{1}(T_{j}) + N_{1}(T_{j})\\ av_{1}(T_{j}) + Z_{1}(T_{j})\\ #id_{1}(T_{j}) + e_{1}(T_{j})",
-#  "log(\\sigma_{e_{Y_{1}(T_{j})}}) = X_{d1}(T_{j})\\ b_{d1}(T_{j}) + Z_{d1}(T_{j})\\ id_{d1}(T_{j})",
-#  "Y_{2}(T_{j}) = X_{2}(T_{j})\\ b_{2}(T_{j}) + M_{2}(T_{j})\\ en_{2}(T_{j}) + N_{2}(T_{j})\\ av_{2}(T_{j}) + Z_{2}(T_{j})\\ #id_{2}(T_{j}) + e_{2}(T_{j})",
-#  "log(\\sigma_{e_{Y_{2}(T_{j})}}) = X_{d2}(T_{j})\\ b_{d2}(T_{j}) + Z_{d2}(T_{j})\\ id_{d2}(T_{j})",
-#  "Y_{3}(T_{j}) = X_{3}(T_{j})\\ b_{3}(T_{j}) + Z_{3}(T_{j})\\ id_{3}(T_{j}) + e_{3}(T_{j})"
-#)
-
-#eqs <- c(
-#  "Y_{1}(T_{j}) = T^2_{j}(X_{1}\\ b_{1}) + T^2_{j}(M_{1}\\ en_{1}) + T^2_{j}(N_{1}\\ av_{1}) + T^2_{j}(Z_{1}\\ id_{1}) + e_{1}(T_{j})",
-#  "log(\\sigma_{e_{Y_{1}(T_{j})}}) = T^2_{j}(X_{d1}\\ b_{d1}) + T^2_{j}(Z_{d1}\\ id_{d1})",
-#  "Y_{2}(T_{j}) = T^2_{j}(X_{2}\\ b_{2}) + T^2_{j}(M_{2}\\ en_{2}) + T^2_{j}(N_{2}\\ av_{2}) + T^2_{j}(Z_{2}\\ id_{2}) + e_{2}(T_{j})",
-#  "log(\\sigma_{e_{Y_{2}(T_{j})}}) = T^2_{j}(X_{d2}\\ b_{d2}) + T^2_{j}(Z_{d2}\\ id_{d2})",
-#  "Y_{3}(T_{j}) = T^2_{j}(X_{3}\\ b_{3}) + T^2_{j}(Z_{3}\\ id_{3}) + e_{3}(T_{j})"
-#)
-
-eqs <- c(
-  "Y_{1j} = X_{1j}\\ b_{1j} + M_{1j}\\ en_{1j} + N_{1j}\\ av_{1j} + Z_{1j}\\ id_{1j} + e_{1j}",
-  "log(\\sigma_{e_{Y_{1j}}}) = X_{d1j}\\ b_{d1j} + Z_{d1j}\\ id_{d1j}",
-  "Y_{2j} = X_{2j}\\ b_{2j} + M_{2j}\\ en_{2j} + N_{2j}\\ av_{2j} + Z_{2j}\\ id_{2j} + e_{2j}",
-  "log(\\sigma_{e_{Y_{2j}}}) = X_{d2j}\\ b_{d2j} + Z_{d2j}\\ id_{d2j}",
-  "Y_{3j} = X_{3j}\\ b_{3j} + Z_{3j}\\ id_{3j} + e_{3j}"
-)
-
-eqs_num <- c(
-    "(eqn. 1.1)",
-    "(eqn. 1.2)",
-    "(eqn. 2.1)",
-    "(eqn. 2.2)",
-    "(eqn. 3)"
-)
-
-# Combine into a dataframe
-df <- data.frame(
-    formula = eqs,
-    number = eqs_num
-)
-
-# Create a table to display on Word
-ft <- flextable(df)
-
-# Transform values as a mathematical formula
-ft <- compose(
-  x = ft, j = 1,
-  value = as_paragraph(
-    as_equation(formula, height = .5)
-  )
-)
-
-# Customize table
-my_theme <- function(x, ...) {
-  x <- colformat_double(x, big.mark = " ",
-                        decimal.mark = ".",
-                        digits = 2)
-  x <- border_remove(x)
-  x <- delete_part(x, part = "header")
-  x <- align(x, j = 1, align = "left", part = "body")
-  x <- align(x, j = 2, align = "right", part = "body")
-  x <- font(x, fontname = "Times New Roman", part = "body")
-  x <- fontsize(x, size = 12, part = "body")
-  x <- width(x, j = 1, width = 13, unit = "cm")
-  x <- width(x, j = 2, width = 2.26, unit = "cm")
-}
-
-# Apply theme and print
-ft <- my_theme(x = ft)
-ft
-
-```
-
-where $j$ is the index of the three experience levels (i.e.
-novice, intermediate, advanced).
-The $d$ subscript indicates that a term is from the dispersion part of the model.
-The $b$ terms are vectors of fixed effects with their design matrices $X$ specific to each trait and experience level.
-The terms $en$, $av$, and $id$ are the vectors of random environment, random avatar, and random individual identity effects associated with their design matrices $M$, $N$, and $Z$ specific to each trait and experience level.
-For all experience levels of $Y_{1}$ and $Y_{2}$, we assumed that the game environment and avatar random effects followed a Gaussian distribution with estimated standard deviation ($en_{12j} \sim N(0,\ I\ \sigma_{en_{12j}}^{2})$ and $av_{12j} \sim N(0,\ I\ \sigma_{av_{12j}}^{2})$).
-$I$ is the identity matrix, with $I\sigma^{2}$ indicating that the random effects are independently and identically distributed.
-In addition, we assumed that the residuals follow a Gaussian distribution with observation-specific variance $e_{12j} \sim N(0,\ Diag\{\sigma_{e_{12j}}^{2}\})$.
-The 15 individual identity random effects $id_{ij}$ across all response variables and experience levels (five for each of the three experience levels) are indexed in vector $\alpha$.
-The $\alpha$ vector follows a multivariate Gaussian distribution $\alpha \sim MVN(0,\ \Sigma)$ where $\Sigma$ is a 15x15 (co)variance matrix of among individual variation across all response variables and experience levels, with $\sigma_{id_{ij}}^{2}$ on the diagonal and covariances off-diagonal.
-
-We used weakly informative Gaussian priors for the prey rank ($N(0,\ 1^{2})$: eqn.
-1.1 to 2.2) and game duration ($N(0.5,\ 0.5^{2})$: eqn.
-3).
-Based on previous results showing that the mean predator speed revolves around 3 m/s [@fraserfranco.etal2022], we applied a weakly informative Gaussian prior on the intercept for predator and prey speed ($N(3,\ 0.5^{2})$: eqn.
-1.1 and 2.1).
-Similarly, we used a weakly informative Gaussian prior on the intercept for the dispersion part ($N(0,\ 1^{2})$: eqn 1.2 and 2.2).
-For hunting success, we applied a positive Gaussian prior on the precision parameter ($N(2,\ 1^{2})$: eqn.
-3) and a negative Gaussian prior ($N(-4,\ 1^{2})$: eqn 3) on the intercept because hunting success should be close to 0 at the lowest game duration.
-For every predicted variable, we applied a weakly informative half-Gaussian prior ($N(0,\ 1^{2})$) on the random effects for the mean part and the dispersion part of the model.
-We used a regularizing $LKJ(2)$ prior on the correlation matrix.
-
-After inspection of the results, we found that key parameter values for our hypotheses were in the direction of our predictions, but they were largely driven by the very slow speed of one predator.
-Removing this individual reduced the estimates by a significant margin.
-To remain conservative, we thus present the results of the model without this individual in the main text, and provide the results of the model including it in the Appendix S3.
-
 ### Markov Chain Monte Carlo settings
@@
 ```{r}
-
 # Open percentages table
-path <- file.path(here::here("outputs", "03_outputs_model-processing"))
+path <- file.path(here::here("outputs", "outputs_model-processing"))
 percent_tab1 <- data.table(readRDS(file.path(path, "GAMM_percentages.rds")))
-percent_tab2 <- data.table(readRDS(file.path(path, "DHMLM_percentages.rds")))
 
@@
 # Open the loo table
-path <- file.path(here::here("outputs", "02_outputs_model-validation"))
+path <- file.path(here::here("outputs", "outputs_model-validation"))
 loo_tab <- data.table(readRDS(file.path(path, "GAMM_loo-cv-table.rds")))
@@
 my_header <- data.frame(
-   col_keys = c("model",
+  col_keys = c(
+    "model",
     "elpd_diff",
@@
     "elpd_loo",
-                "se_elpd_loo"),
-   line1 = c("model",
+    "se_elpd_loo"
+  ),
+  line1 = c(
+    "model",
     "elpd \ndifference",
@@
     "elpd loo \nvalue",
-             "elpd loo \nstandard error"),
+    "elpd loo \nstandard error"
+  ),
   stringsAsFactors = FALSE
@@
 my_theme <- function(x, ...) {
-   x <- colformat_double(x, big.mark = " ",
+  x <- colformat_double(
+    x, big.mark = " ",
     decimal.mark = ".",
-                         digits = 2)
+    digits = 2
+  )
   x <- set_table_properties(x, layout = "autofit")
@@
 knitr::include_graphics(
-  here::here("outputs", "04_outputs_figures", "figure1.png")
+  here::here("outputs", "outputs_figures", "figure1.png")
 )
-
 ```
-
-## Changes in foraging behaviour, prey behaviour, and hunting success
-
-The mean speed of the predator population remained stable with experience (Figure 2 and Appendix S1: Table S1).
-Likewise, the variation in speed did not change with experience, indicating that foraging specialization remained stable at the population level (Figure 2 and Appendix S1: Table S1).
-At the individual level, predators differed slightly in their mean speed across experience levels, indicating marginal differences in hunting tactics (Figure 2 and Appendix S1: Table S2).
-However, they displayed important differences in foraging specialization (Appendix S1: Table S2) which also increased slightly with experience as there was a 0.11 unit increase in among individual differences in IIV from novice to advanced (Figure 2).
-`r toString(scales::percent(percent_tab2[1,]$percentage, accuracy = 1))` of the population switched from a flexible to a specialized hunting tactic, and vice-versa, as predators gained experience (i.e.
->0.2 change in standard deviation; Figure 3).
-In contrast, `r toString(scales::percent(percent_tab2[2,]$percentage, accuracy = 1))` displayed lower changes (i.e.
->0.05 and <0.2 change in standard deviation) and `r toString(scales::percent(percent_tab2[3,]$percentage, accuracy = 1))` showed almost no change with experience (i.e.
-<0.05 change in standard deviation).
-
-```{r fig2, out.width = "100%", fig.cap = "**Figure 2**.
-Median posterior differences and HPD intervals comparing predator foraging behaviour at the population and individual level across experience stages.
-The behavioural parameter predicted by the MDHGLM being compared is on the y axis, and the difference in parameter values between two experience stages is on the x axis.
-The pairwise comparisons between experience stages are displayed across the three panels A, B, and C.
-We compare the mean speed of the predator population, the speed of the prey it encountered, and its hunting success across experience (population mean).
-We also compare the population's foraging specialization and variation in speed of the prey it encountered across experience (population variance).
-At the individual level, we compare among individual differences in speed, the speed of the prey encountered, and hunting success (individual variation mean).
-Lastly, we compare individual variation in IIV across experience, indicating whether individual differences in foraging specialization and in the variation of the groups of prey encountered changed with experience (individual variation IIV)."}
-
-knitr::include_graphics(
-  here::here("outputs", "04_outputs_figures", "figure2.png")
-)
-
-```
-
-As the predator population gained experience, there was strong evidence that the prey increased their speed, although slightly (Figure 2 and Appendix S1: Table S1).
-The population also encountered groups of prey using different speeds through time (Figure 2 and Appendix S1: Table S1).
-At the individual level, all predators encountered prey with similar average speeds throughout experience (Figure 2 and Appendix S1: Table S2).
-Yet, even if they experienced similar average prey speeds across all groups encountered, there were important differences among predators in how (dis)similar were the groups they encountered (Appendix S1: Table S2).
-Thus, some predators encountered similar groups (i.e.
-all groups using similar speeds) while others encountered distinct groups (i.e.
-some slower and some faster).
-These differences among individuals in IIV for prey encounters increased only marginally with experience (Figure 2).
-
-Similar to the GAMMs, the predator population's mean hunting success increased with experience (Figure 2 and Appendix S1: Table S1), with mean prey consumption revolving around two prey.
-There were also important differences among individuals in mean hunting success (Appendix S1: Table S2), but they remained stable across experience (Figure 2).
-
-```{r fig3, out.width = "100%", fig.cap = "**Figure 3**.
-Individual behavioural distribution of the predators' speed drawn from the MDHGLM estimates for individuals that had (A) the greatest increase in specialization and (B) the greatest increase in flexibility.
-The background (gray) distributions are for when individuals were novice hunters, and the foreground (blue) distributions with solid contour lines are for when they were advanced hunters.
-Both figure panels are ordered by ascending degree of increase in either specialization or flexibility.
-We built the figure by first subtracting the estimated standard deviation of all individuals as novices with their standard deviation as advanced hunters, and then selected those with the greatest difference using the 25% and 75% quantiles.
-Individuals on panel A have an increase in specialization equal or greater than 0.2 standard deviations, while those on panel B have an increase in flexibility equal to or greater than 0.2 standard deviations."}
-
-knitr::include_graphics(
-  here::here("outputs", "04_outputs_figures", "figure3.png")
-)
-
-```
-
-## Predator-prey interactions across experience
-
-Predators that were on average faster had lower IIV in speed (Figure 4A-B and Appendix S4: Table S1).
-Thus, faster predators specialized on fast movement, while slower predators were more flexible in their movement.
-As predators gained experience, these strategies were increasingly defined as the correlation changed by 0.16 units, from -0.62 (-0.91, -0.41 $95\%\ CI$) for when predators were novices to -0.78 (-1.00, -0.53 $95\%\ CI$) as they reached the advanced stage (Figure 4C).
-Predators that were on average faster also tended to encounter groups of prey that were on average faster and more similar (Figure 4A-B and Appendix S4: Table S1).
-There was however no evidence that these correlations increased with experience (Figure 4C)
-
-There was no correlation (-0.01 (-0.17, 0.09 $95\%\ CI$)) between mean prey speed and IIV in predator speed when predators were novice (Figure 4A), but the correlation increased to -0.16 (-0.33, -0.04 $95\%\ CI$) at the advanced stage (Figure 4B), such that advanced hunters that encountered faster prey tended to be more specialized (Figure 4C).
-The predators' IIV in speed and IIV in prey speed were weakly positively correlated (Figure 4A-B and Appendix S4: Table S1).
-Hence, predators that encountered groups of prey with similar speeds tended to be more specialized in their tactic, while those that encountered groups of prey with varying speeds tended to be more flexible.
-We did not find evidence that this correlation increased with experience (Figure 4C).
-
-```{r fig4, out.width = "100%", fig.cap = "**Figure 4**.
-Correlation matrices of the MDHGLM relating the mean and IIV in speed of the predators, the mean and IIV in speed of the prey they encountered, and mean hunting success.
-Larger dark blue circles indicate stronger positive correlations (positive values on the colour legend), while larger dark red circles indicate stronger negative correlations (negative values on the colour legend).
-A) Correlations when predators were novice hunters.
-B) Correlations when predators were advanced hunters.
-C) Posterior median differences and HPD intervals comparing the trait correlations predicted by the MDHGLM between the advanced and novice levels.
-The difference is displayed on the y axis and the correlated traits being compared are displayed on the x axis.
-Positive values indicate an increase in correlation strength from novice to the advanced stage, while negative values indicate stronger correlations when predators were novice."}
-
-knitr::include_graphics(
-  here::here("outputs", "04_outputs_figures", "figure4.png")
-)
-
-```
-
-## Success along the foraging specialization continuum
-
-The predator's mean speed and IIV were only weakly negatively correlated with hunting success across experience (Figure 4 and Appendix S4: Table S1), suggesting that slower/flexible and faster/specialist predators were overall both equally successful.
-There was however a slight increase in the correlation between the predator's IIV in speed and success (Figure 4C), suggesting that flexible foragers at the advanced stage were marginally less successful (Appendix S1: Figure S2).
-We also found strong evidence that individuals who encountered groups of prey with faster speeds were less successful (Figure 4A-B and Appendix S4: Table S1).
-Predators also had greater success across the study when encountering more variable groups of prey (Figure 4A-B and Appendix S4: Table S1).
-Both relationships remained structurally stable across experience (Figure 4C).
 
@@
 By capitalizing on a virtual predator-prey system where interactions were directly monitored, we found that predators in *Dead by Daylight* increased their hunting success with experience, suggesting that predator expertise was honed through extensive practice.
-While there was no directional increase in either behavioural specialization or flexibility with experience at the population level, individual predators developed their own behavioural trajectories over successive encounters with the prey.
-This was characterized by a continuum of hunting styles ranging from a slower and flexible strategy to a highly specialized fast-paced strategy, with some predators switching between the two with experience.
-The behavioural strategies of predators and prey tended to match across experience, suggesting that both were adjusting to each other.
-Specialized cursorial hunters encountered similar groups of prey, while slower and flexible hunters encountered varying groups of prey.
-Overall, predators along the flexible-specialist continuum achieved similar success, although flexible hunters were slightly less successful as advanced hunters.
+**parler de l'effet des proies**
 
@@
 
+## The role of prey behaviour
+
 However, even after we took the prey's speed into account, there were still important differences in expertise acquisition among predators, suggesting that other antipredator tactics were potentially involved.
 For example, if the prey used camouflage or hiding to avoid being detected, then predators may have found them only if they learned how to exploit the visual cues that facilitate cryptic prey detection, such as habitat characteristics or prey colour patterns [@ehlinger1989;@hughes.etal1992;@warburton2003;@szopa-comley.etal2020].
+
 In addition, longer time intervals between hunting events may delay or even impede learning due to forgetting important information [@endler1991;@wright.etal2022].
@@
 
-## Predator-prey interactions across experience
-
-The predator population maintained its speed and degree of specialization throughout the study period, and individuals differed only slightly in their average speeds.
-However, differences in specialization among individuals were important and increased slightly with experience.
-Predators hunting at higher speeds specialized in this tactic and tended to encounter groups of prey with similar speeds, while those hunting at slower speeds were flexible and tended to encounter varying groups of prey.
-On one hand, this pattern could arise if some predators adjusted their strategy to their prey, because fast-paced hunting is a specialized tactic suited for prey that use rapid evasive movements, while a slower and flexible tactic can be useful to minimize the consequences of uncertainty when prey escape unpredictably [@endler1991;@bro-jorgensen2013;@wilson.etal2018;@szopa-comley.ioannou2022].
-However, we observed that 44% of the predator population showed close to no change in specialization with experience.
-Since these individuals were already specialized as novices (Fraser Franco, personal observation), there would be no point in them changing their tactic with experience if they were already successful, or there could have been costs to switch their strategy if they more often encountered faster prey.
-This could explain why the behavioural interactions between the predator and the prey remained stable across the predator's experience.
 Yet, because prey can also learn how to avoid predation [@kelley2003;@turner2006;@mccomb2011], we believe that the predator-prey phenotype matching more likely emerged from reciprocal adjustments in speed by predators and prey as they interacted [@kishida.etal2006;@kishida.etal2009;@edgell.rochette2009;@mcghee.etal2013].
-Indeed, 56% of the predator population altered its behaviour with experience to different degrees, including 12% that even switched between flexibility and specialization.
 Hence, if the prey also learned through repeated interactions with the predators, it is possible that experience contributed in stabilizing the system as both were adjusting to each other, similar to Red Queen dynamics [@brockhurst2014].
 
-Classical theory predicts that individual foraging specialization should maximize prey consumption by optimizing foraging efficiency [@stephens1986], but other work suggests that flexibility and specialization could both be adaptive as resources fluctuate [@woo.etal2008;@phillips.etal2017].
-Our results support empirical findings indicating that flexible and specialist foragers obtained similar success [@weimerskirch2007;@woo.etal2008;@potier.etal2015;@phillips.etal2017;@courbin.etal2018], although flexible foragers were slightly less successful.
-Both strategies appeared to emerge in part from individuals learning how to hunt their prey, resulting in a general increase in success in the population.
 However, there were still considerable differences in success among individuals through time, suggesting that some predators were limited in their capacity to match their tactic to their prey or to increase their success through other means.
@@
 Such trade-offs may reflect limitations in learning all the skills required to successfully hunt all types of prey [@healy1992;@belisle1997;@dukas2019], particularly if the skills required to hunt slower prey are nontransferable to faster prey.
-As the benefits of each hunting style changed under different scenarios, the combination of predatory styles in this virtual system reflects how resource fluctuations can maintain fitness equilibrium within populations, resulting in behavioural diversity in predator-prey systems [@tinker.etal2008;@woo.etal2008;@mougi.kishida2009].
+
+**Of course, the life of neither the predator nor the prey players are at stake, such that emerging patterns could be driven by their motivation to win and not "true" survival. Yet,**
+**A potential caveat is that the more flexible hunters might have experimented with various tactics out of boredom, which could impede ecologically realistic interpretations of our data.
+However, the consistent association between this tactic and heterogeneous prey groups gives us confidence that it emerged from their interactions.
+One limitation of our study was that we couldn't monitor all the matches of the prey, which prevented us from assessing their responses to the predator through their experience.** Hence, future studies should aim at monitoring reciprocal behavioural dynamics through time, which may reveal important insights on the mechanisms driving stable equilibria in predator-prey systems.
 
@@
 We found support of our prediction that prey were driving individual differences in expertise in a predator population.
-We also confirmed that experience and predator-prey encounters drove the patterns of individual foraging specialization.
-Our results suggest that predators learned with experience, as their success increased and their speed remained matched to the speed of their prey.
-Even though individuals were not all equally successful, both specialized and flexible hunters achieved similar success.
-A potential caveat is that the more flexible hunters might have experimented with various tactics out of boredom, which could impede ecologically realistic interpretations of our data.
-However, the consistent association between this tactic and heterogeneous prey groups gives us confidence that it emerged from their interactions.
-One limitation of our study was that we couldn't monitor all the matches of the prey, which prevented us from assessing their responses to the predator through their experience.
-Hence, future studies should aim at monitoring reciprocal behavioural dynamics through time, which may reveal important insights on the mechanisms driving stable equilibria in predator-prey systems.
+Our results suggest that predators learned with experience, as their success increased...
+Even though individuals were not all equally successful.
 Lastly, virtual systems are increasingly recognized as being useful to test hypotheses on ecological interactions [@beauchamp2020;@cere.etal2021;@fraserfranco.etal2022;@lymbery.etal2023].

### 2024-06-05 manuscript/manuscript.Rmd
Ms up to date. discussion à finir

@@
 
-The ecological consequences of predation are defined by the sum of consumptive and nonconsumptive effects (CEs and NCEs) during predator-prey interactions [refs].
+The ecological consequences of predation accrue to the sum of consumptive and nonconsumptive effects (CEs and NCEs) during predator-prey interactions [@preisser2007].
 Recent advances in our understanding of the behavioural mechanisms (e.g.
-reciprocal plasticity, foraging specialization) driving prey consumption at the individual and population level have allowed ecologists to rely on foraging strategies to predict the outcome of (N)CEs and their effect on prey population dynamics [refs].
+foraging mode, foraging specialization, reciprocal plasticity) driving (N)CEs at the individual and population level have allowed ecologists to predict their outcome as well as their effect on prey population dynamics [@wirsing.etal2021].
 For example, empirical studies show that prey tend to increase vigilance at the expense of foraging in the presence of ambush predators (NCEs) [@preisser2007].
-Because ambush predators are opportunistic hunters, they also tend to consume a greater range of prey types (CEs) compared to cursorial predators [@glaudas2019].
+Because ambush predators are opportunistic hunters, they tend to consume a greater range of prey types (CEs) compared to cursorial predators [@glaudas2019].
 
-While the foraging mode of predators is a reliable predictor of prey consumption, predators need to acquire expertise by practicing and learning the proper skills to successfuly capture their prey [@dukas2019;wooster.etal2023].
-The aquisition of expertise is predicted to be paramount in many predator systems where hunting tactics are learned (e.g.
-orca whales).
-Yet, there is only fraction of evidence showing the link between experience and success in natural predators, impeding our ability to accurately predict the consequences of predation on prey populations.
-
+There are, however, internal mechanisms shaping predator foraging decision-making that are often overlooked when studying (N)CEs in predator-prey interactions [@wooster.etal2023].
+For instance, predators need to acquire expertise by practicing and learning the proper skills to successfully locate, select, and capture their prey [@dukas2019;@wooster.etal2023].
+The aquisition of expertise is predicted to be paramount in systems where predators learn different hunting tactics (e.g.
+orca whales) [@heithaus2018].
 Empirical studies on human and non-human hunters show that individuals optimize foraging efficiency (e.g.
 search and handling times, return rates) by learning associative images or using cues from their prey and their environment [@edwards.jackson1994;@morse2000;@macdonald2007;@reid.etal2010;@wilson-rankin2015].
-Predators can integrate this information to use the appropriate hunting tactic (e.g.
-ambushing or fast movement speeds) for the type of prey that they encounter [@toscano2014;@patrick.weimerskirch2014a;@toscano.etal2016].
-However, prey often use antipredator tactics such as rapid or unpredictable movements which are effective to successfully escape predators [@walker.etal2005;@kelley.magurran2011;@herbert-read.etal2017;@szopa-comley.ioannou2022].
-This results in reciprocal interactions where predators need to learn to predict the movement of their prey to capture them.
-Therefore, antipredator tactics may impede the acquisition of hunting expertise by interfering with learning processes and reducing the success of predators.
+Thus, predators can integrate this information to increase their chances of locating prey or to use the tactic most suited for the type of prey encountered [refs].
+
+Conversely, prey can use antipredator tactics that are effective to evade predators, such as rapid or unpredictable movements [@walker.etal2005;@kelley.magurran2011;@herbert-read.etal2017;@szopa-comley.ioannou2022].
+This results in reciprocal interactions where predators need to learn to predict the movement of their prey to successfully capture them.
+Yet, it may be impossible for a predator to learn unpredictable prey escape trajectories such that these antipredator tactics may interfere with a predator's acquisition of hunting expertise [troscianko.etal2018].
+Unfortunately, there is only scarce evidence showing links between experience, prey antipredator behaviour, and prey consumption in natural predators, impeding our ability to accurately predict the consequences of predation on prey populations.
 
@@
 mean movement speed along a slow-fast continuum), and their success is driven by the movement of the prey [@fraserfranco.etal2022].
-The prey can increase their chances of survival by cooperating and moving fast when they forage [@cere.etal2021;@fraserfranco.etal2022;@santostefano.etal2024].
+The prey can increase their chances of survival by cooperating and moving fast to escape the predator [@cere.etal2021;@fraserfranco.etal2022;@santostefano.etal2024].
 The game also ellicits natural reactions in players such as fear from predation (personal observations), which corroborates with another virtual ecological study showing that predation drives individual variation in risk perception [@beauchamp2020].
@@
 
-In this study, we assess how repeated experience shapes predator foraging success using data from players in *Dead by Daylight*.
-First, we investigate how individual hunting success changes with experience (i.e.
+In this study, we assess how repeated encounters with prey shapes predator hunting success using data from players in *Dead by Daylight*.
+First, we investigate how individual hunting success changes with repeated encounters (i.e.
 how individuals develop their hunting expertise).
-We hypothesize that the predators' success will increase with experience up to a certain level where it will stagnate.
-We expect this pattern to vary among individuals due to differences in prey encounters.
-Second, we investigate how prey movement influences the development of expertise.
-If all individual predators encounter groups of prey with similar speeds, we predict that their gain in expertise will be similar.
-
-In contrast, based on previous results showing that faster prey are more difficult to hunt, then...
-
-if all predators encounter varying groups of prey, then they should all converge towards flexible speeds.
-In both scenarios, differences among individuals in IIV across experience should be low (i.e.
-similar individual foraging specialization), whereas the population variance would either decrease (specialization) or increase (flexibility).
+At the population level, we hypothesize that the predators' success will increase with experience up to a certain level where it will stabilize [@dukas2019].
+However, we expect this pattern to vary among individuals due to differences in the movement of prey encountered.
+Therefore, we then investigate how prey movement influences the development of expertise.
+If predators encounter groups of prey with similar speeds, we predict that the gain in expertise will be similar among individuals.
+Alternatively, if predators encounter groups of prey with varying speeds, then the acquisition of expertise will vary among individuals.
 
@@
 that have not escaped), or when the last remaining prey escapes the virtual environment.
-Each player, predator or prey, can choose an avatar with abilities that encourage specific play styles (e.g.
+
+Before the start of a match, players (predator or prey) can choose an avatar with unique abilities that encourage specific play styles (e.g.
 bold vs cautious prey, or ambush vs roaming predator).
@@
 Some of these environments are larger than others, with varying structural complexity.
-However, predators display only minimal changes in behaviour and hunting success across the environments, probably due to a game feature enabling them to have visual cues of the generators [@fraserfranco.etal2022].
+However, predators display only minimal changes in behaviour and hunting success across the environments, probably due to a game feature enabling them to have visual cues on the resources [@fraserfranco.etal2022].
 There were `r toString(format(length(unique(dat$environment_id))))` virtual game environments available for play during the study period.
+Details on the basic characteristics of predator avatars are available at https://deadbydaylight.fandom.com/wiki/Killers.
+Details on the size and structure of the different virtual environments are available at https://dbdmaps.com/ and https://deadbydaylight.fandom.com/wiki/Realms.
 
@@
 
-Our population consists of `r toString(format(length(unique(dat$predator_id))))` predator players with a total record of `r toString(formatC(length(unique(dat$match_id)), big.mark = " "))` matches.
+Our population consists of `r toString(format(length(unique(dat$predator_id))))` players that played strictly as the predator, with a total record of `r toString(formatC(length(unique(dat$match_id)), big.mark = " "))` matches.
 The predator-players' experience varied between `r toString(format(min(dat$total_xp_pred)))` and `r toString(format(max(dat$total_xp_pred)))` matches played.
@@
 The following information is collected and reported for every match : the player's anonymous ID, its avatar (i.e.
-the predator character chosen with its specific skill-gameplay mechanics), the game environment, the predator-player's experience, the mean speed of the groups of prey the player encountered, and their mean rank (a proxy for prey skill).
+the predator character chosen with its specific skill-gameplay mechanics), the game environment, the predator-player's experience, the mean speed of the groups of prey that the predator player encountered, and the mean rank of the prey encountered (a proxy for prey skill).
 The ranking system in *DBD* was implemented by the company to pair players in a match based on their skill (https://deadbydaylight.fandom.com/wiki/Rank), and failing to account for it would prevent us from detecting a change in the predator's foraging success with experience.
@@
 We did not account for matches where predators played as the prey.
-We also categorized predators for each match based on their cumulated experience.
-We labeled predators as novices for matches where they had cumulated less than 100 matches, intermediate for matches where they had cumulated between 100 and 299 matches, and advanced for matches where they had cumulated more than 299 matches (max 499).
-Since our goal was to monitor predator players throughout their experience and that they all played at least 300 matches, they all appeared in each of the three experience categories.
 
-We recognize that we could have biased our analyses by retaining only individuals who played for at least 300 matches.
+We recognize that we could have introduced a bias in our analyses by retaining only individuals who played for at least 300 matches.
 For example, these individuals might be experienced videogame players and could thus already be playing like experts in their first matches in *DBD*.
 To verify that our sample was not biased, we compared a random sample of players that played either 20 to 50 matches, 51 to 100 matches, or 101 to 300 matches during the same timeframe as our sampled population.
-We then took the first 20 matches played by these players, including those from our sampled population, and compared their mean speed using a Bayesian hierarchical linear model.
-We found that all four groups had similar mean speeds as predators (Appendix 2: Table S1-S2), which indicates an absence of bias due to data sampling.
+We then took the first 20 matches played by these players, including those from our sampled population, and compared their mean hunting success using a Bayesian hierarchical linear model.
+We found that all four groups had similar successes as predators (Appendix 2: Table S1-S2), which suggests an absence of bias due to data sampling.
 
@@
 
-### Software and computer specifications
-
-All models were fitted under a Bayesian framework in R (version 4.1.2) using Markov chain Monte Carlo (MCMC) sampling with the package "brms" version 2.16.3 [@burkner2017], an R front-end for the STAN software [@StanDevelopmentTeam2023], and "cmdstanr" version 0.4.0 [@gabry.cesnovar2021] as the back-end for estimation (cmdstan installation version 2.28.2).
-The models were run on Cedar (Operating system: CentOS Linux 7), a computer cluster maintained by the Digital Research Alliance of Canada (https://docs.alliancecan.ca/wiki/Cedar).
-Each required 64GB of RAM with 48 cores to compile within 5 days.
-
 ### Model specification
 
-We tested how prey influenced the development of predator expertise by computing five Bayesian generalized additive mixed models (GAMM) with thin-plate regression splines, all of which estimated the relationship between hunting success (i.e.
+We tested how predators developed their expertise by computing five Bayesian generalized additive mixed models (GAMM) with thin-plate regression splines, all of which estimated the relationship between hunting success (i.e.
 number of prey consumed) and the predators' cumulative experience (i.e.
@@
 We employed weakly informative half-Gaussian priors on all the standard deviation parameters ($N(0,\ 0.5^{2})$).
-We compared the models' predictive accuracies using approximate leave-one-out cross-validation with Pareto-smoothed importance sampling [@vehtari.etal2017;@piironen.vehtari2017;@vehtari.etal2022].
+We compared the predictive accuracy of all five models using approximate leave-one-out cross-validation with Pareto-smoothed importance sampling [@vehtari.etal2017;@piironen.vehtari2017;@vehtari.etal2022].
 
@@
 We assessed the convergence of the MCMC chains using trace plots, R-hat diagnostics with a threshold of <1.01, and effective sample sizes (ESS) with a threshold of >100 [@vehtari.etal2021].
-We also performed posterior predictive checks which showed an adequate fit of the models (for details, see https://github.com/quantitative-ecologist/experience-hunting-tactics).
+We also performed posterior predictive checks which showed an adequate fit of the models.
 We report all the posterior parameter estimates using the median of the posterior distribution with the highest posterior density (HPD) intervals at 50%, 80%, and 95%.
+
+### Software and computer specifications
+
+All models were fitted in R (version 4.1.2) using Markov chain Monte Carlo (MCMC) sampling with the package "brms" version 2.16.3 [@burkner2017], an R front-end for the STAN software [@StanDevelopmentTeam2023], and "cmdstanr" version 0.4.0 [@gabry.cesnovar2021] as the back-end for parameter estimation (cmdstan installation version 2.28.2).
+The models were run on Cedar (Operating system: CentOS Linux 7), a computer cluster maintained by the Digital Research Alliance of Canada (https://docs.alliancecan.ca/wiki/Cedar).
+Each required 64GB of RAM with 48 cores to compile within 5 days.
+For details on how to reproduce our analyses, please consult the GitHub repository of this project (https://github.com/quantitative-ecologist/predator-expertise).
 
@@
 
-Out of all five GAMM models, two predicted the data the best and achieved similar expected log pointwise densities (elpd).
-Both models accounted for the prey group's rank and speed (Table 1).
-Models in which these effects were not accounted for resulted in no change in hunting success with experience (i.e.
-expertise) for the average individual (results not shown).
+Out of all five GAMM models, the two that accounted for the prey group's rank and speed were the best at predicting the data, achieving similar expected log pointwise densities (Table 1).
+Models in which prey effects were not accounted for resulted in no change in hunting success with experience (i.e.
+expertise) for the average individual (**Appendix XX**).
 Accounting for the prey rank resulted in a concave-shaped relationship, with the highest success ranging between 100 and 400 matches (Figure 1A).
-In the model where we additionally included the prey's speed, the effect of experience on hunting success for the average individual followed a diminishing returns curve, with predators optimizing their success after playing ~300 matches (Figure 1C).
+In the model where we additionally included the prey's speed, the effect of experience on hunting success for the average individual followed a diminishing returns curve as we predicted, with predators optimizing their success after playing ~300 matches (Figure 1C).
 The curve shows there was a 38% increase in the probability of consuming all prey for the average individual between the first and the ~330$^{th}$ match where success reached a plateau (Figure 1C).
@@
 In contrast, accounting for both effects resulted in `r toString(scales::percent(percent_tab1[5,]$percentage, accuracy = 0.1))` of the population increasing its success with experience from the first match to the last.
-The prey speed alone also mediated individual differences in the relationship between success and experience.
+The prey speed mediated individual differences in the relationship between success and experience.
 The standard deviation of the individual slopes component of the model accounting only for prey rank was equal to 9.72 (9.30, 10.15 $95\%\ CI$), while the one for the model accounting for prey rank and speed was equal to 3.36 (3.04, 3.71 $95\%\ CI$), indicating that individual differences in the development of expertise decreased when accounting for prey speed.
@@
 
-A general assumption of predator-prey studies is that predators maximize success by matching their tactic to their prey [@abrams2000], yet, it has remained largely unknown whether this results from predators learning how to hunt their prey in part because of the challenges of investigating direct interactions in the wild.
-By capitalizing on a virtual predator-prey system where interactions were directly monitored, we found that predators in *Dead by Daylight* increased their hunting success with experience, suggesting that predator expertise was honed through extensive practice.
-**parler de l'effet des proies**
-
-## The development of expertise
+It is generally assumed that predators increase prey consumption through time by acquiring hunting expertise [@dukas2019;@wooster.etal2023].
+Yet, few studies have empirically tested this hypothesis in part because of the challenges of investigating direct predator-prey interactions in the wild.
+By capitalizing on a virtual predator-prey system where interactions were directly monitored, we found that a predator-player population in *Dead by Daylight* increased their hunting success with experience, suggesting that predator expertise was honed through extensive practice.
+Yet, there were important differences among individuals in the development of expertise, which was in part due to differences among individuals in the movement of the prey encountered.
 
 Our results suggest that predator expertise is honed through extensive practice.
-The predator population increased and stabilized its success with experience in a diminishing returns fashion as it is typically found in empirical studies of expertise [reviewed in @dukas2019].
-The prey were important in mediating this pattern, probably because they increased their speed as predators gained experience.
+The predator population displayed an asymptotic relationship between experience and success, wherein initial gains in success were significant but gradually stabilized as experience accumulated, consistent empirical observations of expertise in both humans and nonhuman animals [reviewed in @dukas2019] The prey's speed was important in mediating this pattern at the population level, probably because they also increased their speed as predators gained experience.
 We previously showed in *DBD* that faster movement is an effective strategy used by the prey to avoid predation [@fraserfranco.etal2022], and other studies have found that as well [@walker.etal2005;@kelley.magurran2011;@martin.etal2022].
-This resulted in discernable differences in the relationship between success and experience among predators, implying that the development of expertise depends on an individual's capacity to adjust to difficulty.
-Hunting faster prey requires costly and specialized cognitive abilities and coordination [@kelley.magurran2011], and that could explain why such prey were more difficult to capture.
-Thus, predators that couldn't properly hunt at high speeds or develop counter-strategies for faster prey were likely at a disadvantage.
+This would explain why the relationship was concave in the model where we did not account for prey speed, because adept predators were probably encountering prey with higher speeds.
+Thus, our results suggest that predators can gain expertise and maintain success when they encounter prey that move at speeds lower or close to the population-average.
 
-## The role of prey behaviour
+Prey speed also mediated a portion of the differences among predator players in the development of their expertise, implying that prey antipredator behaviour can impair the development of an individual predator's expertise.
+Individual predator players probably differed in their capacity to adjust to difficult prey.
+For example, hunting faster prey requires specialized cognitive abilities and coordination that are energetically costly [@kelley.magurran2011].
+Thus, predators that couldn't develop counter-strategies for faster prey were likely at a disadvantage.
 
-However, even after we took the prey's speed into account, there were still important differences in expertise acquisition among predators, suggesting that other antipredator tactics were potentially involved.
-For example, if the prey used camouflage or hiding to avoid being detected, then predators may have found them only if they learned how to exploit the visual cues that facilitate cryptic prey detection, such as habitat characteristics or prey colour patterns [@ehlinger1989;@hughes.etal1992;@warburton2003;@szopa-comley.etal2020].
+**Such trade-offs may reflect limitations in learning all the skills required to successfully hunt all types of prey [@healy1992;@belisle1997;@dukas2019], particularly if the skills required to hunt slower prey are nontransferable to faster prey.**
 
-In addition, longer time intervals between hunting events may delay or even impede learning due to forgetting important information [@endler1991;@wright.etal2022].
+However, even after we took the prey's speed into account, there were still important differences in expertise acquisition among predator players, suggesting that other antipredator tactics were potentially involved.
+For example, the prey can use camouflage and hide to avoid detection (not currently measurable in *DBD*), requiring predators to learn how to exploit visual cues that facilitate cryptic prey detection, such as habitat characteristics or prey colour patterns [@ehlinger1989;@hughes.etal1992;@warburton2003;@szopa-comley.etal2020].
+Furthermore, it is hypothesized that longer time intervals between hunting events could hinder or delay the acquisition of expertise as individuals may forget information with longer delays [@endler1991;@wright.etal2022].
 In *DBD*, a predator that played 300 matches in the span of six months might forget more critical information related to prey detection or escape patterns than one that played 300 matches in the span of six days.
 Investigating the impact of such time lags in future analyses may reveal important insights on the outcome of predator-prey interactions.
-
-Yet, because prey can also learn how to avoid predation [@kelley2003;@turner2006;@mccomb2011], we believe that the predator-prey phenotype matching more likely emerged from reciprocal adjustments in speed by predators and prey as they interacted [@kishida.etal2006;@kishida.etal2009;@edgell.rochette2009;@mcghee.etal2013].
-Hence, if the prey also learned through repeated interactions with the predators, it is possible that experience contributed in stabilizing the system as both were adjusting to each other, similar to Red Queen dynamics [@brockhurst2014].
-
-However, there were still considerable differences in success among individuals through time, suggesting that some predators were limited in their capacity to match their tactic to their prey or to increase their success through other means.
-Specialist foragers were faster, and thus, probably better equipped to hunt the more difficult faster prey in *DBD*.
-However, if the prey responded to fast predators by also being faster, then hunting at high speeds resulted in more difficult encounters for these predators, thereby decreasing the benefits of using this tactic (Figure 4).
-Thus, specializing probably compensated for the difficulty of hunting prey at high speeds by helping predators to better predict the location and movement of their prey.
-On the other hand, flexible foragers encountered variable prey with slower speeds.
-Yet, because the prey increased their speed with experience, the benefits of being able to hunt multiple prey types for flexible hunters may have come at the cost of not being adept at capturing faster prey [@pintor.etal2014].
-Such trade-offs may reflect limitations in learning all the skills required to successfully hunt all types of prey [@healy1992;@belisle1997;@dukas2019], particularly if the skills required to hunt slower prey are nontransferable to faster prey.
 
@@
 One limitation of our study was that we couldn't monitor all the matches of the prey, which prevented us from assessing their responses to the predator through their experience.** Hence, future studies should aim at monitoring reciprocal behavioural dynamics through time, which may reveal important insights on the mechanisms driving stable equilibria in predator-prey systems.
+
+**Comment by reviewer:**
+
+Some characteristics to be aware of: 
+- Predators do not compete.
+- Both predators and prey can fail in a round yet still carry that experience and knowledge on to the next round with no real penalty (think of the movie Groundhog Day).
+- Predator and prey are surely playing a game with each other, but prey may also be playing a game among themselves, which may affect their game with the predator. 
+- This particular video game appears to have a few unusual characteristics.
+Prey do not appear to forage for resources.
+but rather they create a public good.
+Furthermore, handling time for the predator is extremely complicated and lengthy.
+- Finally, it may be rather complicated to discern all of the rules of how the different virtual environments work and what  all the aptitudes of all the different predators and prey are.
 

### 2024-07-08 code/code_figures/figure1.R
updated figure 1

@@ -14,66 +14,68 @@
 
 # Load libraries and model -------------------------------------------------
 
- options(mc.cores = parallel::detectCores()) 
+options(mc.cores = parallel::detectCores())
 
- library(parallel)
- library(brms)
- library(data.table)
- library(ggplot2)
- library(ggpubr)
- library(viridis)
+library(parallel)
+library(brms)
+library(data.table)
+library(ggplot2)
+library(ggpubr)
+library(viridis)
 
- path <- file.path(getwd(), "outputs", "01_outputs_models")
+path <- file.path(getwd(), "outputs", "outputs_models")
 
- modA2 <- readRDS(file.path(path, "A2_GAMM-rank.rds"))
- modA2_prey <- readRDS(file.path(path, "A2_GAMM-speed-rank.rds"))
+modA2 <- readRDS(file.path(path, "A2_GAMM-rank.rds"))
+modA2_prey <- readRDS(file.path(path, "A2_GAMM-speed-rank.rds"))
 
 
 
 # Load the data ------------------------------------------------------------
 
- data <- fread("./data/FraserFrancoetal2023-data.csv",
-               select = c("predator_id",
-                          "game_duration",
-                          "pred_speed",
-                          "prey_avg_speed",
-                          "prey_avg_rank",
-                          "cumul_xp_pred",
-                          "total_xp_pred",
-                          "hunting_success"))
+data <- fread(
+  "./data/FraserFrancoetal2023-data.csv",
+  select = c("predator_id",
+             "game_duration",
+             "pred_speed",
+             "prey_avg_speed",
+             "prey_avg_rank",
+             "cumul_xp_pred",
+             "total_xp_pred",
+             "hunting_success")
+)
 
- data[, predator_id := as.factor(predator_id)]
+data[, predator_id := as.factor(predator_id)]
 
 
 
 # Post processing preparations for custom family ---------------------------
 
- expose_functions(modA2, vectorize = TRUE)
+expose_functions(modA2, vectorize = TRUE)
 
- # Define the log likelihood function
- log_lik_beta_binomial2 <- function(i, prep) {
-   mu <- brms::get_dpar(prep, "mu", i = i)
-   phi <- brms::get_dpar(prep, "phi", i = i)
-   trials <- prep$data$vint1[i]
-   y <- prep$data$Y[i]
[…]

### 2024-07-08 code/code_figures/figure2.R
added figure 2

@@ -1,0 +1,243 @@
+# ==========================================================================
+
+#                         Plot GAMM sds parameters
+
+# ==========================================================================
+
+# Figure of the posterior parameters for:
+#- Intercept standard deviation
+#- Slope standard deviation
+#- Wiggliness standard deviation
+
+
+
+# ==========================================================================
+# 1. Prepare script
+# ==========================================================================
+
+
+
+# Load libraries and model -------------------------------------------------
+
+library(brms)
+library(data.table)
+library(ggplot2)
+
+path <- file.path(getwd(), "outputs", "outputs_models")
+
+fit1 <- readRDS(file.path(path, "A2_GAMM-rank.rds"))
+fit2 <- readRDS(file.path(path, "A2_GAMM-speed-rank.rds"))
+
+# ==========================================================================
+# ==========================================================================
+
+
+
+
+
+# ==========================================================================
+# 2. Extract posterior draws into a tidy table
+# ==========================================================================
+
+# Extract posterior draws fit1
+post1 <- data.table(
+  as_draws_df(
+    fit1,
+    variable = c(
+      "sds_sZcumul_xppredator_id_1",
+      "sds_sZcumul_xppredator_id_2",
+      "sds_sZcumul_xppredator_id_3"
+    )
+  )
+)
+
+# Extract posterior draws fit2
+post2 <- data.table(
+  as_draws_df(
+    fit2,
+    variable = c(
+      "sds_sZcumul_xppredator_id_1",
+      "sds_sZcumul_xppredator_id_2",
+      "sds_sZcumul_xppredator_id_3"
+    )
+  )
+)
+
+# Clear memory
+rm(fit1)
+rm(fit2)
+
+# Combine posterior draws
+figdat <- rbind(
+  post1,
+  post2
+)
+
+# Add model variable
+figdat[, model := c(rep("fit1", 1000), rep("fit2", 1000))]
+
+# ==========================================================================
+# ==========================================================================
+
+
+
+
+
+# ==========================================================================
+# 3. Summarize results from the posterior
+# ==========================================================================
+
+# Get intervals function
+get_HPD_interval <- function(x, prob = 0.95) {
+  interval <- coda::HPDinterval(as.mcmc(x), prob)
+  list(lower = interval[1], upper = interval[2])
+}
+
[…]

### 2024-07-08 code/code_figures/appendix1_figureS1.R
clean appendix code. need to run it

@@ -142,10 +142,13 @@ tab1 <- tab1[cumul_xp <= total_xp_pred, ]
 # Extract players with greatest increase and decrease ---------------------
 
 # Calculate 1st and final value
-tab1[, ":=" (
-  first = min(Zcumul_xp),
-  last = max(Zcumul_xp)),
-  by = predator_id]
+tab1[
+  , ":=" (
+    first = min(Zcumul_xp),
+    last = max(Zcumul_xp)
+  )
+  , by = predator_id
+]
 
 tab1[, equal1 := ifelse(Zcumul_xp == first, 1, 0), by = predator_id]
 tab1[, equal2 := ifelse(Zcumul_xp == last, 1, 0), by = predator_id]
@@ -159,7 +162,7 @@ tab2 <- dcast(
   tab2,
   predator_id ~ range,
   value.var = "estimate__"
-  )
+)
 
 
 # Calculate the difference in predicted success between minimum and maximum xp
@@ -269,22 +272,24 @@ gamm_plot3 <- ggplot(table[difference %between% c(-0.05, 0.05), ],
 # Prepare figure --------------------------------------------------------
 
 # Arrange paneled figure
- figure <- ggarrange(
-    NULL, gamm_plot1, NULL, gamm_plot2, NULL, gamm_plot3,
-    ncol = 6, nrow = 1,
-    labels = c("(A)", "", "(B)", "", "(C)"),
-    widths = c(0.15, 1.5, 0.15, 1.5, 0.15, 1.5)
- )
+figure <- ggarrange(
+   NULL, gamm_plot1, NULL, gamm_plot2, NULL, gamm_plot3,
+   ncol = 6, nrow = 1,
+   labels = c("(A)", "", "(B)", "", "(C)"),
+   widths = c(0.15, 1.5, 0.15, 1.5, 0.15, 1.5)
+)
 
 # Export the figure -----------------------------------------------------
 
- path <- file.path(getwd(), "outputs", "04_outputs_figures")
+path <- file.path(getwd(), "outputs", "outputs_figures")
 
- ggexport(figure,
-          filename = file.path(path, "appendix1_figureS1.png"),
-          width = 4500,
-          height = 1200,
-          res = 300)
+ggexport(
+  figure,
+  filename = file.path(path, "appendix1_figureS1.png"),
+  width = 4500,
+  height = 1200,
+  res = 300
+)
 
 # =======================================================================
 # =======================================================================

### 2024-07-09 manuscript/manuscript.Rmd
intro up to date.

@@
 
-The capacity of predators to match their hunting tactic to their prey is expected to drive the outcome of predator-prey interactions.
-To do so, predators need to practice their tactics extensively to develop the expertise to be successful hunters.
-Yet, there are limited empirical assessments showing links between experience and hunting success at the individual and population levels due to the challenges of monitoring direct interactions in the wild.
-Here, we use a virtual predator-prey system (the game *Dead by Daylight*) to investigate how experience shapes individual and population hunting success in predators across repeated interactions with their prey.
-We show that predators optimized prey consumption as they gained experience.
-Prey speed was an important mediator of this relationship, resulting in differences among predators in the development of expertise.
-**Décrire le reste des résultats...**
+**revoir abstract pour être moins centré sur la tactique**
 
-Keywords: foraging behaviour, experience, learning, antipredator behaviour, virtual ecology, Dead by Daylight
+**revoir la phrase sur ce qu'on fait dans l'étude. Intégrer l'aspect des proies**
+
+The acquisition of expertise is crucial for predators to be successful hunters.
+To do so, predators need to hone their skills and acquire knowledge through repeated and extensive practice.
+Yet, there are limited empirical assessments showing how predators acquire expertise through repeated encounters with their prey due to the challenges of monitoring direct interactions in the wild.
+Here, we use a virtual predator-prey system (the game *Dead by Daylight*) to investigate how experience shapes individual and population hunting success in human predators across repeated interactions with their prey.
+We show that predators optimized prey consumption as they gained experience, indicating that they acquired expertise with extensive practice.
+Prey speed was an important mediator of this relationship, driving differences among predators in the development of expertise.
+At the population-level, we found that faster prey could impair the acquisition of expertise by reducing hunting success.
+At the individual-level, our results show that prey speed shapes how individual predators acquire expertise.
+Therefore, our study outlines how prey antipredator behaviour can mediate the acquisition of expertise in predator populations.
+
+Keywords: predator-prey, experience, learning, antipredator behaviour, virtual ecology, Dead by Daylight
 
@@
 The ecological consequences of predation accrue to the sum of consumptive and nonconsumptive effects (CEs and NCEs) during predator-prey interactions [@preisser2007].
-Recent advances in our understanding of the behavioural mechanisms (e.g.
-foraging mode, foraging specialization, reciprocal plasticity) driving (N)CEs at the individual and population level have allowed ecologists to predict their outcome as well as their effect on prey population dynamics [@wirsing.etal2021].
-For example, empirical studies show that prey tend to increase vigilance at the expense of foraging in the presence of ambush predators (NCEs) [@preisser2007].
-Because ambush predators are opportunistic hunters, they tend to consume a greater range of prey types (CEs) compared to cursorial predators [@glaudas2019].
+Recent advances in our understanding of the mechanisms (e.g.
+foraging mode, reciprocal plasticity, size-dependent foraging) driving (N)CEs at the individual and population level have allowed ecologists to predict their outcome as well as their effect on prey population dynamics [@Peckarsky.etal2008;@gravel.etal2013;@wirsing.etal2021].
+For example, empirical studies show that prey tend to increase vigilance at the expense of foraging in the presence of ambush predators (NCEs) [@preisser2007], with such hunters consuming a greater range of prey types (CEs) compared to cursorial hunters [@glaudas2019].
 
-There are, however, internal mechanisms shaping predator foraging decision-making that are often overlooked when studying (N)CEs in predator-prey interactions [@wooster.etal2023].
+There are, however, internal mechanisms contributing to predator foraging success that are often overlooked when studying (N)CEs in predator-prey interactions [@wooster.etal2023].
 For instance, predators need to acquire expertise by practicing and learning the proper skills to successfully locate, select, and capture their prey [@dukas2019;@wooster.etal2023].
-The aquisition of expertise is predicted to be paramount in systems where predators learn different hunting tactics (e.g.
-orca whales) [@heithaus2018].
 Empirical studies on human and non-human hunters show that individuals optimize foraging efficiency (e.g.
-search and handling times, return rates) by learning associative images or using cues from their prey and their environment [@edwards.jackson1994;@morse2000;@macdonald2007;@reid.etal2010;@wilson-rankin2015].
-Thus, predators can integrate this information to increase their chances of locating prey or to use the tactic most suited for the type of prey encountered [refs].
+search and handling times, return rates) by learning associative images, or by using cues from their prey and their environment [@edwards.jackson1994;@morse2000;@macdonald2007;@reid.etal2010;@wilson-rankin2015].
+Thus, predators can integrate this information to increase their chances of locating prey, or to use the attack tactic most suited for a given context.
 
-Conversely, prey can use antipredator tactics that are effective to evade predators, such as rapid or unpredictable movements [@walker.etal2005;@kelley.magurran2011;@herbert-read.etal2017;@szopa-comley.ioannou2022].
-This results in reciprocal interactions where predators need to learn to predict the movement of their prey to successfully capture them.
-Yet, it may be impossible for a predator to learn unpredictable prey escape trajectories such that these antipredator tactics may interfere with a predator's acquisition of hunting expertise [troscianko.etal2018].
-Unfortunately, there is only scarce evidence showing links between experience, prey antipredator behaviour, and prey consumption in natural predators, impeding our ability to accurately predict the consequences of predation on prey populations.
+On the other hand, prey use antipredator tactics such as rapid escapes to evade predators [@walker.etal2005;@kelley.magurran2011;@herbert-read.etal2017].
+This compels predators to develop swift motor skills to succesfully capture such prey.
+However, predators may be limited in their capacity to develop the necessary attributes (e.g.
+physical, physiological, neurological) for fast-paced hunting, which may impair the development of their hunting expertise.
+For example, experimental studies have shown that certain camouflage strategies, such as disruptive coloration, can impair expertise acquisition in humans and birds [@martin.etal2012;@troscianko.etal2013;@troscianko.etal2018].
+Unfortunately, while our understanding of how predation risk influences information acquisition in prey has increased over the years [reviewed in @crane.etal2024], there is only scarce evidence showing links between prey antipredator strategies and the development of expertise in human and nonhuman predators.
+Therefore, our ability to accurately predict the (N)CEs of predation on prey populations remain limited.
 
@@
 The prey can increase their chances of survival by cooperating and moving fast to escape the predator [@cere.etal2021;@fraserfranco.etal2022;@santostefano.etal2024].
-The game also ellicits natural reactions in players such as fear from predation (personal observations), which corroborates with another virtual ecological study showing that predation drives individual variation in risk perception [@beauchamp2020].
-These observations outline how ecological patterns can emerge from interactions in virtual systems with fixed rules.
+The game also ellicits natural reactions in players such as freezing when predation is imminent (personal observations), which corroborates with another virtual ecological study showing that predation drives individual variation in risk perception [@beauchamp2020].
+These observations outline how fundamental ecological patterns can emerge from human interactions in virtual systems with fixed rules [@brosnan.etal2017;@kasumovic.etal2017].
 Videogames also generate large volumes of data on interacting players throughout their lifetime in the game under realistic, controlled, and repeatable ecological scenarios.
@@
 Alternatively, if predators encounter groups of prey with varying speeds, then the acquisition of expertise will vary among individuals.
+We predict that faster prey will impair expertise acquisition, such that predators that encountered faster prey will have lower success throughout experience.
 
@@
 Some of these environments are larger than others, with varying structural complexity.
-However, predators display only minimal changes in behaviour and hunting success across the environments, probably due to a game feature enabling them to have visual cues on the resources [@fraserfranco.etal2022].
+However, predators display only minimal changes in behaviour and hunting success across these environments, probably due to a game feature enabling them to have visual cues on the resources [@fraserfranco.etal2022].
 There were `r toString(format(length(unique(dat$environment_id))))` virtual game environments available for play during the study period.
@@
 
-#fit1 <- readRDS(here::here("outputs", "01_outputs_models", "A2_GAMM-rank.rds"))
-#fit2 <- readRDS(here::here("outputs", "01_outputs_models", "A2_GAMM-speed-rank.rds"))
-#
-#fit1
-#fit2
+# Import top 2 model fits
+fit1 <- readRDS(here::here("outputs", "outputs_models", "A2_GAMM-rank.rds"))
+fit2 <- readRDS(here::here("outputs", "outputs_models", "A2_GAMM-speed-rank.rds"))
 ```
@@
 
+```{r fig2, out.width = "100%", fig.cap = "**Figure 2**.
+Posterior distributions of the individual-level standard deviation parameters.
+The first two panels represent the intercept and slope standard deviations of the linear components relating hunting success to cumulative experience, and the third panel represents the standard deviation of the wiggliness component of the curves.
+The light coloured distributions are for the model with a shared trend where we did not account for prey speed, and the darker coloured distributions are for the model with a shared trend where we accounted for prey speed (i.e.
+the model with the highest predictive power)"}
+
+knitr::include_graphics(
+  here::here("outputs", "outputs_figures", "figure2.png")
+)
+```
+
 # DISCUSSION

### 2024-07-09 code/code_figures/figure2.R
updated fig2

@@ -181,16 +181,15 @@ figdat_long[
 # Custom theme
 custom_theme <- theme(
   # axis values size
-  axis.text = element_text(size = 14, color = "black"),
+  axis.text = element_text(size = 15, color = "black"),
   # axis titles size
-  axis.title = element_text(size = 16),
-  strip.text.x = element_text(size = 14),
+  axis.title = element_text(size = 17),
+  strip.text.x = element_text(size = 15),
   panel.grid = element_blank(),
   panel.background = element_blank(),
-  legend.title = element_text(size = 14),
-  legend.text = element_text(size = 14),
-  legend.position = "top",
-  legend
+  legend.title = element_text(size = 15),
+  legend.text = element_text(size = 15),
+  legend.position = "top"
 )
 
 # Basic plot variables
@@ -235,8 +234,8 @@ ggsave(
   filename = file.path(path, "figure2.png"),
   units = "px",
   width = 3600,
-  height = 1400,
-  dpi = 300
+  height = 1500,
+  dpi = 400
 )
 
 # ==========================================================================

### 2024-07-09 code/code_figures/figure2.R
updated fig2

@@ -233,8 +233,8 @@ ggsave(
   p,
   filename = file.path(path, "figure2.png"),
   units = "px",
-  width = 3600,
-  height = 1500,
+  width = 3800,
+  height = 1700,
   dpi = 400
 )
 

### 2024-07-09 code/code_model-processing/GAMM_percentages.R
updated gamm percentages

@@ -18,10 +18,10 @@
 library(brms)
 library(data.table)
 
-path <- file.path(getwd(), "outputs", "01_outputs_models")
+path <- file.path(getwd(), "outputs", "outputs_models")
 
-fit2 <- readRDS(file.path(path, "A2_GAMM.rds"))
-fit2p <- readRDS(file.path(path, "A2_GAMM-speed-rank.rds"))
+fit1 <- readRDS(file.path(path, "A2_GAMM-rank.rds"))
+fit2 <- readRDS(file.path(path, "A2_GAMM-speed-rank.rds"))
 
 
 
@@ -80,19 +80,19 @@ posterior_epred_beta_binomial2 <- function(prep) {
 # 2. Prepare data
 # ==========================================================================
 
+dat1 <- conditional_effects(
+  fit1, method = "fitted", effects = "Zcumul_xp:predator_id",
+  robust = FALSE, re_formula = NULL
+)
+
 dat2 <- conditional_effects(
   fit2, method = "fitted", effects = "Zcumul_xp:predator_id",
   robust = FALSE, re_formula = NULL
 )
 
-dat2p <- conditional_effects(
-  fit2p, method = "fitted", effects = "Zcumul_xp:predator_id",
-  robust = FALSE, re_formula = NULL
-)
-
 # Extract values in a table
+tab1 <- data.table(dat1[[1]])
 tab2 <- data.table(dat2[[1]])
-tab2p <- data.table(dat2p[[1]])
 
 
 # Back transform x-axis values
@@ -108,22 +108,30 @@ scaled_breaks <- sequence / standev
 xp <- unique(data[, .(predator_id, total_xp_pred)])
 
 # Compute non standardized cumulative XP
+tab1[, cumul_xp := (Zcumul_xp * standev) + mean(data$cumul_xp_pred)]
 tab2[, cumul_xp := (Zcumul_xp * standev) + mean(data$cumul_xp_pred)]
-tab2p[, cumul_xp := (Zcumul_xp * standev) + mean(data$cumul_xp_pred)]
 
 # Now merge the two tables adding the total XP
+tab1 <- merge(tab1, xp, by = "predator_id")
 tab2 <- merge(tab2, xp, by = "predator_id")
-tab2p <- merge(tab2p, xp, by = "predator_id")
 
 # Cut all matches where fitted values are above total XP
[…]

### 2024-07-09 manuscript/manuscript.Rmd
minor corrections to text

@@
 
+**arranger les appendices pour inclure la table qui ne montre que le succès**
+
+**il n'y aurait qu'un seul appendice**
+
 The acquisition of expertise is crucial for predators to be successful hunters.
@@
 In this study, we assess how repeated encounters with prey shapes predator hunting success using data from players in *Dead by Daylight*.
-First, we investigate how individual hunting success changes with repeated encounters (i.e.
-how individuals develop their hunting expertise).
-At the population level, we hypothesize that the predators' success will increase with experience up to a certain level where it will stabilize [@dukas2019].
-However, we expect this pattern to vary among individuals due to differences in the movement of prey encountered.
-Therefore, we then investigate how prey movement influences the development of expertise.
-If predators encounter groups of prey with similar speeds, we predict that the gain in expertise will be similar among individuals.
-Alternatively, if predators encounter groups of prey with varying speeds, then the acquisition of expertise will vary among individuals.
-We predict that faster prey will impair expertise acquisition, such that predators that encountered faster prey will have lower success throughout experience.
+First, we investigate the development of expertise at the population level.
+We hypothesize that the predator population's success will increase with experience up to a certain level where it will stabilize [@dukas2019].
+However, we expect this pattern to change depending on the movement of the prey encountered.
+We hypothesize that prey will influence the development of expertise, and predict that faster prey will reduce the gain in expertise.
+Therefore, we then investigate how prey movement influences the development of expertise at the individual level.
+If prey speed does not influence hunting success, we predict that the gain in expertise will be similar among individuals.
+Alternatively, if prey speed influences hunting success, then the acquisition of expertise will vary among individuals.
 
@@
 
-We parametrized the GAMMs and the MDHGLM to run four MCMC chains with 1000 posterior samples for each parameter.
-We ran 2500 iterations with a thinning set to eight for the additive model with a global smoother only (see Table 1), and 1500 iterations with a thinning set to four for the other additive models.
-We set the first 500 iterations in each model as warm ups.
-We parametrized the MDHGLM to run 2500 iterations with a thinning set to eight, with the first 500 iterations used as warm ups.
+We parametrized the GAMMs to run four MCMC chains with 1000 posterior samples for each parameter.
+We ran 2500 iterations with a thinning set to eight for the additive model with a global smoother only (see Table 1), and 1500 iterations with a thinning set to four for the remaining additive models.
+We set the first 500 iterations of each model as warm ups.
 We assessed the convergence of the MCMC chains using trace plots, R-hat diagnostics with a threshold of <1.01, and effective sample sizes (ESS) with a threshold of >100 [@vehtari.etal2021].
 We also performed posterior predictive checks which showed an adequate fit of the models.
-We report all the posterior parameter estimates using the median of the posterior distribution with the highest posterior density (HPD) intervals at 50%, 80%, and 95%.
 
@@
 
-## Development of expertise
+## Development of expertise at the population level
 
 ```{r}
-# Open percentages table
-path <- file.path(here::here("outputs", "outputs_model-processing"))
-percent_tab1 <- data.table(readRDS(file.path(path, "GAMM_percentages.rds")))
-
 # Import top 2 model fits
@@
 fit2 <- readRDS(here::here("outputs", "outputs_models", "A2_GAMM-speed-rank.rds"))
+
+sum1 <- summary(fit1, robust = TRUE)$splines
+sum2 <- summary(fit2, robust = TRUE)$splines
+
+# Clear memory
+rm(fit1)
+rm(fit2)
 ```
@@
 Models in which prey effects were not accounted for resulted in no change in hunting success with experience (i.e.
-expertise) for the average individual (**Appendix XX**).
+expertise) for the average individual.
 Accounting for the prey rank resulted in a concave-shaped relationship, with the highest success ranging between 100 and 400 matches (Figure 1A).
-In the model where we additionally included the prey's speed, the effect of experience on hunting success for the average individual followed a diminishing returns curve as we predicted, with predators optimizing their success after playing ~300 matches (Figure 1C).
+In the model where we included the prey's speed, the effect of experience on hunting success for the average individual followed a diminishing returns curve as we predicted, with predators optimizing their success after playing ~300 matches (Figure 1C).
 The curve shows there was a 38% increase in the probability of consuming all prey for the average individual between the first and the ~330$^{th}$ match where success reached a plateau (Figure 1C).
+
+\newpage
 
@@
 # ===========================================================================
-
 ```
 
-The relationship between hunting success and cumulative experience differed among predators (Figure 1B-D).
-Only `r toString(scales::percent(percent_tab1[2,]$percentage, accuracy = 0.1))` of the population had an increase in success from the first match to the last in the model where we did not account for the prey's speed and rank (Appendix S1: Figure S1A).
-In contrast, accounting for both effects resulted in `r toString(scales::percent(percent_tab1[5,]$percentage, accuracy = 0.1))` of the population increasing its success with experience from the first match to the last.
-The prey speed mediated individual differences in the relationship between success and experience.
-The standard deviation of the individual slopes component of the model accounting only for prey rank was equal to 9.72 (9.30, 10.15 $95\%\ CI$), while the one for the model accounting for prey rank and speed was equal to 3.36 (3.04, 3.71 $95\%\ CI$), indicating that individual differences in the development of expertise decreased when accounting for prey speed.
+## Development of expertise at the individual level
+
+Prey speed did not influence among individual differences in average hunting success as the posterior distributions of the standard deviations of individual intercepts were almost completely overlapping (Figure 2, median = `r toString(round(sum1[2,]$Estimate, digits = 2))` vs median = `r toString(round(sum2[2,]$Estimate, digits = 2))`).
+However, individuals differed in the development of their hunting expertise (Figure 1B-D).
+We found strong evidence that the speed of the prey mediated among individual differences in the linear relationship between success and experience, as there were substantial differences in the standard deviations of the individual slopes between the two models (Figure 2, median = `r toString(round(sum1[3,]$Estimate, digits = 2))` vs median = `r toString(round(sum2[3,]$Estimate, digits = 2))`).
+Differences among individuals in the direction of the linear relationship between success and experience were `r toString(round(sum1[3,]$Estimate / sum2[3,]$Estimate, digits = 2))` times lower when we removed the effect of prey speed (i.e.
+accounting for it in the model).
 
@@
 
+We also found strong evidence that the speed of the prey mediated the form of the relationship between hunting success and experience at the individual level.
+We detected large differences in the standard deviations of the wiggliness (Figure 2, median = `r toString(round(sum1[4,]$Estimate, digits = 2))` vs median = `r toString(round(sum2[4,]$Estimate, digits = 2))`).
+The lower standard deviation for the model were we accounted for prey speed suggests that the form of the relationship between success and experience was more similar among individuals.
+
 ```{r fig2, out.width = "100%", fig.cap = "**Figure 2**.
@@
 The first two panels represent the intercept and slope standard deviations of the linear components relating hunting success to cumulative experience, and the third panel represents the standard deviation of the wiggliness component of the curves.
+The vertical dashed lines are the medians of the posterior distributions.
 The light coloured distributions are for the model with a shared trend where we did not account for prey speed, and the darker coloured distributions are for the model with a shared trend where we accounted for prey speed (i.e.
-the model with the highest predictive power)"}
+the model with the highest predictive accuracy)"}
 
@@
 
+*revoir la première phrase, devrait intégrer les proies*
+
 It is generally assumed that predators increase prey consumption through time by acquiring hunting expertise [@dukas2019;@wooster.etal2023].
-Yet, few studies have empirically tested this hypothesis in part because of the challenges of investigating direct predator-prey interactions in the wild.
-By capitalizing on a virtual predator-prey system where interactions were directly monitored, we found that a predator-player population in *Dead by Daylight* increased their hunting success with experience, suggesting that predator expertise was honed through extensive practice.
-Yet, there were important differences among individuals in the development of expertise, which was in part due to differences among individuals in the movement of the prey encountered.
+Yet, few studies have empirically tested how prey antipredator strategies mediate the acquisitio of hunting expertise because of the challenges of investigating direct predator-prey interactions in the wild.
+By capitalizing on a virtual predator-prey system where interactions were directly monitored, we found that a predator-player population in *Dead by Daylight* increased their hunting success with experience.
+Yet, there were important differences among individuals in the development of said expertise, which we found to be in part due to the movement of the prey encountered.
 
-Our results suggest that predator expertise is honed through extensive practice.
-The predator population displayed an asymptotic relationship between experience and success, wherein initial gains in success were significant but gradually stabilized as experience accumulated, consistent empirical observations of expertise in both humans and nonhuman animals [reviewed in @dukas2019] The prey's speed was important in mediating this pattern at the population level, probably because they also increased their speed as predators gained experience.
+Our results suggest that the acquisition of hunting expertise was honed through extensive practice.
+The predator population displayed an asymptotic relationship between experience and success, wherein initial gains in success were significant but gradually stabilized as experience accumulated.
+Our observations are consistent with empirical studies of expertise in both humans and nonhuman animals [reviewed in @dukas2019].
+However, the prey's speed was important in mediating this pattern at the population level, probably because prey also increased their speed as predators gained experience.
 We previously showed in *DBD* that faster movement is an effective strategy used by the prey to avoid predation [@fraserfranco.etal2022], and other studies have found that as well [@walker.etal2005;@kelley.magurran2011;@martin.etal2022].
-This would explain why the relationship was concave in the model where we did not account for prey speed, because adept predators were probably encountering prey with higher speeds.
+This would explain why the relationship was concave in the model where we did not account for prey speed, because adept predators were encountering prey with higher speeds.
 Thus, our results suggest that predators can gain expertise and maintain success when they encounter prey that move at speeds lower or close to the population-average.
 
-Prey speed also mediated a portion of the differences among predator players in the development of their expertise, implying that prey antipredator behaviour can impair the development of an individual predator's expertise.
+Prey speed also mediated differences among predator players in the development of their expertise, implying that prey antipredator behaviour can impair the development of an individual predator's expertise.
 Individual predator players probably differed in their capacity to adjust to difficult prey.
@@
 Thus, predators that couldn't develop counter-strategies for faster prey were likely at a disadvantage.
+
+*ici, répéter les études utilisant le camouflage, ensuite inclure la phrase des compromis, et conclure avec une avenue évolutive.*
 

### 2024-07-09 code/code_models/DHMLM-RandomSample.R
DHMLM script and output of random sample

@@ -1,0 +1,272 @@
+# =======================================================================
+
+#            Compare results of our data to random samples              #
+
+# =======================================================================
+
+
+
+
+
+# =======================================================================
+# 1. Load libraries, and import dataset
+# =======================================================================
+
+
+# When working on my local computer
+# renv::activate()
+
+
+
+# Detect number of cores ------------------------------------------------
+
+options(mc.cores = parallel::detectCores())
+
+
+
+# Load libraries --------------------------------------------------------
+
+library(data.table)
+library(brms)
+library(parallel)
+
+
+
+# Import the data -------------------------------------------------------
+
+# Folder path Compute Canada
+folder <- file.path("/home", "maxime11", "projects", "def-monti",
+                    "maxime11", "phd_project", "data")
+
+# Load data on compute canada
+data <- fread(
+  file.path(folder, "FraserFrancoetal2023-data-random.csv"),
+  select = c("predator_id",
+             "avatar_id",
+             "hunting_success",
+             "cumul_xp_pred",
+             "total_xp_pred",
+             "game_duration",
+             "pred_speed",
+             "prey_avg_speed",
+             "prey_avg_rank")
+)
+
+# Predator id as factor
+data[, predator_id := as.factor(predator_id)]
+data[, predator_avatar_id := as.factor(avatar_id)]
+
+# Group players by experience
+data[, player_type := "empty"]
+data[total_xp_pred < 50, player_type := "early_quitters"]
+data[total_xp_pred %between% c(50, 99), player_type := "mid_quitters"]
+data[total_xp_pred %between% c(100, 299), player_type := "slightly_engaged"]
+data[total_xp_pred >= 300, player_type := "engaged"]
+
+data[, player_type := as.factor(player_type)]
+data[, total_xp_pred := player_type]
+
+# =======================================================================
+# =======================================================================
+
+
+
+
+
+# =======================================================================
+# 2. Prepare the data for the model
+# =======================================================================
+
+
+# Standardise the variables (Z-scores) ----------------------------------
+
+# Apply the function and create new columns
+# The function standardizes the variables by group :
+# in this case, by level of experience
+
+standardize <- function(x) {
[…]

### 2024-07-09 manuscript/appendix.Rmd
added appendix

---
title: |
     Prey movement shapes the development of predator expertise in a virtual bi-trophic system: \
     Appendix

output: 
  word_document:
    reference_docx: appendix-styles.docx
    fig_caption: yes
csl: chicago-author-date.csl
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(data.table)
library(brms)
library(flextable)
library(officer)
library(dplyr)
```

\newpage

```{r tableprep, include = FALSE}
# ===========================================================================
# 1. Extract the fixed effects from the MV model
# ===========================================================================

# Prepare the table ---------------------------------------------------------

# Model fit
fit <- readRDS(
  here::here("outputs", "outputs_models", "DHMLM-RandomSample.rds")
)

# Fixed effects table
results <- data.table(
  round(summary(fit, robust = TRUE, prob = 0.89)$fixed, digits = 2)[, 1:4],
  keep.rownames = TRUE
)
setnames(results, "rn", "parameter")

# Trait variable (to remove after)
results[parameter %like% "predspeed", trait := "predator speed"][
  parameter %like% "preyavgspeed", trait := "prey speed"][
    parameter %like% "success", trait := "hunting success"
  ]

# mu parameters table
tab <- results[
    !(parameter %like% "sigma"),
    c(6, 1, 2, 4, 5)
]
tab <- tab[trait == "hunting success"]

# Rename parameter values ---------------------------------------------------

# Rename
tab[,
  parameter := 
      c("game duration",
        "cumulative experience",
        "group 1",
        "group 4",
        "group 2",
        "group 3")
]
# Reorder rows
tab <- tab[
  c(1,2,3,5,6,4)
]
tab[, trait := NULL]

# ===========================================================================
# ===========================================================================

# ===========================================================================
# 2. Compute the table using flextable
# ===========================================================================

# Prepare the table parameters ----------------------------------------------

# Custom header
my_header <- data.frame(
   col_keys = c("parameter",
                "Estimate",
                "l-89% CI",
                "u-89% CI"),
   line1 = c("Parameter",
             "Estimate",
             "lower 89% CI",
             "upper 89% CI"),
   stringsAsFactors = FALSE
)

# Custom theme
my_theme <- function(x, ...) {
   x <- colformat_double(x, big.mark = " ",
                         decimal.mark = ".",
                         digits = 2)
   x <- set_table_properties(x, layout = "autofit")
   x <- border_remove(x)
   std_border <- fp_border(width = 1, color = "black")
   x <- hline_top(x, part = "all", border = std_border)
   x <- hline_bottom(x, part = "all", border = std_border)
   x <- width(x, j = 1, width = 3.78, unit = "cm")
   x <- width(x, j = c(3:4), width = 3.3, unit = "cm")
}
```

```{r tableS1, ft.align = "left"}
# Create the table ----------------------------------------------------------

tabf <- tab %>%
  flextable(col_keys = my_header$col_keys) %>%
  set_header_df(mapping = my_header, key = "col_keys") %>%
  my_theme() %>%
  merge_v(j = 1) %>%
  valign(j = 1, valign = "top", part = "body") %>%
  align(align = "left", part = "all", j = 1) %>%
  align(align = "center", part = "all", j = c(2:4)) %>%
  footnote(
    i = 1, j = 1,
    part = "header",
    value = as_paragraph(
      as_sup("a "),
      "Group 1: <50 matches, Group 2: between 50 and 99 matches, Group 3: between 100 and 299 matches, Group 4: > 299 (i.e. group in the main text)"
    ),
    ref_symbols = " "
  ) %>%
  set_caption(
    tab,
    caption = "Table S1. Fixed effects table comparing the success of random groups of players with different amounts of matches played to the group presented in the main text.",
    style = "Table Caption"
  ) %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(size = 12, part = "footer") %>%
  font(fontname = "Times New Roman", part = "all")

tabf

# ===========================================================================
# ===========================================================================
```

\newpage

```{r figS1, out.width = "100%", fig.cap = "**Figure S1.** Among individual differences in the development of hunting expertise in the model where we do not account for prey speed and average rank. The predators' hunting success (i.e. the probability of capturing the four prey) is on the y axis, and the predators' cumulative experience (i.e. the number of matches played prior to each observation) is on the x axis. Each fitted curve represents an individual predator. The individual curves are separated by differences between their first and last predicted value, and displayed as such in three distinct panels. (A) Individuals with a >0.5 unit increase in hunting success with experience (B) Individuals with a <-0.5 unit decrease in hunting success with experience. (C) Individuals that maintained a stable hunting success (between -0.5 and 0.5 unit change in hunting success)"}

knitr::include_graphics(
  here::here(
    "outputs", "outputs_figures",
    "appendix1_figureS1.png"
  )
)
```

### 2024-07-09 code/code_figures/appendix1_figureS1.R
removed old appendix files

@@ -1,295 +1,0 @@
-# ==========================================================================
-
-#                Code to produce Appendix 1 : Figure S1                    #
-
-# ==========================================================================
-
-
-
-
-
-# ==========================================================================
-# 1. Prepare script
-# ==========================================================================
-
-
-# Load libraries and models ------------------------------------------------
-
-library(brms)
-library(data.table)
-library(ggplot2)
-library(ggpubr)
-
-path <- file.path(getwd(), "outputs", "01_outputs_models")
-
-fit1 <- readRDS(file.path(path, "A2_GAMM.rds"))
-#fit1 <- readRDS(file.path(path, "A2_GAMM-speed-rank.rds"))
-
-
-
-# Load the data ------------------------------------------------------------
-
-data <- fread("./data/FraserFrancoetal2023-data.csv",
-              select = c("predator_id",
-                         "game_duration",
-                         "pred_speed",
-                         "prey_avg_speed",
-                         "cumul_xp_pred",
-                         "total_xp_pred",
-                         "hunting_success"))
-
-data[, predator_id := as.factor(predator_id)]
-
-
-
-# Post processing preparations for custom family ---------------------------
-
-expose_functions(fit1, vectorize = TRUE)
-
-# Define the log likelihood function
-log_lik_beta_binomial2 <- function(i, prep) {
-  mu <- brms::get_dpar(prep, "mu", i = i)
-  phi <- brms::get_dpar(prep, "phi", i = i)
-  trials <- prep$data$vint1[i]
-  y <- prep$data$Y[i]
-  beta_binomial2_lpmf(y, mu, phi, trials)
-}
-
-# Define function for posterior_predict
-posterior_predict_beta_binomial2 <- function(i, prep, ...) {
-  mu <- brms::get_dpar(prep, "mu", i = i)
-  phi <- brms::get_dpar(prep, "phi", i = i)
-  trials <- prep$data$vint1[i]
-  beta_binomial2_rng(mu, phi, trials)
-}
-
-# Define function for posterior_epred
-posterior_epred_beta_binomial2 <- function(prep) {
-  mu <- brms::get_dpar(prep, "mu")
-  trials <- prep$data$vint1
-  trials <- matrix(trials, nrow = nrow(mu), ncol = ncol(mu), byrow = TRUE)
-  mu * trials
-}
-
-
-
-# Setup a custom theme for the plot ----------------------------------------
-
-custom_theme <- theme(
-  # axis values size
-  axis.text = element_text(face = "plain",
[…]

### 2024-07-09 code/code_figures/figureS1.R
arranged figure S1

@@ -1,0 +1,177 @@
+# ==========================================================================
+
+#                               Figure S1
+
+# ==========================================================================
+
+# Figure of the posterior parameters for:
+# - Groups of players that are randomly sampled
+# - The group studied in the paper
+
+
+
+# ==========================================================================
+# 1. Prepare script
+# ==========================================================================
+
+
+# Load libraries -----------------------------------------------------------
+
+library(data.table)
+library(brms)
+library(ggplot2)
+library(viridis)
+
+
+
+# Import model -------------------------------------------------------------
+
+path <- file.path(getwd(), "ouputs", "outputs_models")
+fit <- readRDS(
+  file.path(path, "DHMLM-RandomSample.rds")
+)
+# ==========================================================================
+# ==========================================================================
+
+
+
+
+
+# ==========================================================================
+# 2. Extract posterior draws
+# ==========================================================================
+
+# Extraction with specified parameters
+draws <- data.table(
+  brms::as_draws_df(
+    x = fit,
+    variable = c(
+      "b_huntingsuccess_total_xp_predearly_quitters",
+      "b_huntingsuccess_total_xp_predengaged",
+      "b_huntingsuccess_total_xp_predmid_quitters",
+      "b_huntingsuccess_total_xp_predslightly_engaged"
+    )
+  )
+)
+
+# ==========================================================================
+# ==========================================================================
+
+
+
+
+
+# ==========================================================================
+# 3. Reshape posterior draws table for plotting
+# ==========================================================================
+
+# Convert to long format
+drawsl <- melt(
+  draws,
+  measure.vars = patterns("^b_huntingsuccess"),
+  variable.name = "group",
+  value.name = "value"
+)
+drawsl <- drawsl[, c(4,5)] # delete the chain info
+
+# Relabel
+drawsl[
+  , group := factor(
+    group,
+    levels = c("b_huntingsuccess_total_xp_predearly_quitters",
+               "b_huntingsuccess_total_xp_predengaged",
+               "b_huntingsuccess_total_xp_predmid_quitters",
+               "b_huntingsuccess_total_xp_predslightly_engaged"),
+    labels = c("Group 1",
+               "Group 4",
+               "Group 2",
+               "Group 3")
+  )
+]
+
[…]

### 2024-07-09 manuscript/appendix.Rmd
appendix up to date

@@
 
-```{r figS1, out.width = "100%", fig.cap = "**Figure S1.** Among individual differences in the development of hunting expertise in the model where we do not account for prey speed and average rank.
-The predators' hunting success (i.e.
-the probability of capturing the four prey) is on the y axis, and the predators' cumulative experience (i.e.
-the number of matches played prior to each observation) is on the x axis.
-Each fitted curve represents an individual predator.
-The individual curves are separated by differences between their first and last predicted value, and displayed as such in three distinct panels.
-(A) Individuals with a >0.5 unit increase in hunting success with experience (B) Individuals with a <-0.5 unit decrease in hunting success with experience.
-(C) Individuals that maintained a stable hunting success (between -0.5 and 0.5 unit change in hunting success)"}
+```{r figS1, out.width = "100%", fig.cap = "Figure S1.
+Posterior distributions of the median hunting success per group of players.
+The values were back transformed to probability scale and can be interpreted as the probability of capturing the four prey.
+Group 1: <50 matches, Group 2: between 50 and 99 matches, Group 3: between 100 and 299 matches, Group 4: > 299 (i.e.
+group in the main text)."}
 
@@
     "outputs", "outputs_figures",
-    "appendix1_figureS1.png"
+    "figureS1.png"
   )

### 2024-07-10 code/code_figures/figureS2.R
added figure S2

@@ -1,0 +1,198 @@
+# ==========================================================================
+
+#                               Figure S2
+
+# ==========================================================================
+
+# Figure of the relationship between prey speed and predator XP
+
+
+
+# ==========================================================================
+# 1. Prepare script
+# ==========================================================================
+
+
+# Load libraries -----------------------------------------------------------
+
+library(data.table)
+library(brms)
+library(ggplot2)
+library(viridis)
+
+
+
+# Import model -------------------------------------------------------------
+
+path <- file.path(getwd(), "outputs", "outputs_models")
+fit <- readRDS(file.path(path, "A2_GAMM-speed-rank.rds"))
+
+
+
+# Load the data ------------------------------------------------------------
+
+data <- fread(
+  "./data/FraserFrancoetal2023-data.csv",
+  select = c("predator_id",
+             "game_duration",
+             "pred_speed",
+             "prey_avg_speed",
+             "prey_avg_rank",
+             "cumul_xp_pred",
+             "total_xp_pred",
+             "hunting_success")
+)
+
+data[, predator_id := as.factor(predator_id)]
+# ==========================================================================
+# ==========================================================================
+
+
+
+
+
+# ==========================================================================
+# 2. Compute the predictions
+# ==========================================================================
+
+# Post processing preparations for custom family ---------------------------
+
+expose_functions(fit, vectorize = TRUE)
+
+# Define the log likelihood function
+log_lik_beta_binomial2 <- function(i, prep) {
+  mu <- brms::get_dpar(prep, "mu", i = i)
+  phi <- brms::get_dpar(prep, "phi", i = i)
+  trials <- prep$data$vint1[i]
+  y <- prep$data$Y[i]
+  beta_binomial2_lpmf(y, mu, phi, trials)
+}
+
+# Define function for posterior_predict
+posterior_predict_beta_binomial2 <- function(i, prep, ...) {
+  mu <- brms::get_dpar(prep, "mu", i = i)
+  phi <- brms::get_dpar(prep, "phi", i = i)
+  trials <- prep$data$vint1[i]
+  beta_binomial2_rng(mu, phi, trials)
+}
+
+# Define function for posterior_epred
+posterior_epred_beta_binomial2 <- function(prep) {
+  mu <- brms::get_dpar(prep, "mu")
[…]

### 2024-07-10 code/code_model-validation/GAMM-diagnostics.R
update diagnostics

@@ -16,27 +16,26 @@
 
 # Librairies ------------------------------------------------------------
 
- # Detect cores
- options(mc.cores = parallel::detectCores())
+# Detect cores
+options(mc.cores = parallel::detectCores())
 
- library(data.table)
- library(brms)
- library(parallel)
- library(ggpubr)
+library(data.table)
+library(brms)
+library(parallel)
+library(ggpubr)
 
 
 
 # Import the model ------------------------------------------------------
 
- path <- file.path(getwd(), "outputs", "01_outputs_models")
- #path <- getwd()
+path <- file.path(getwd(), "outputs", "outputs_models")
 
- # Load models
- mod1 <- readRDS(file.path(path, "A1_GAMM-rank.rds"))
- mod2 <- readRDS(file.path(path, "A2_GAMM-rank.rds"))
- mod3 <- readRDS(file.path(path, "A3_GAMM-rank.rds"))
- mod4 <- readRDS(file.path(path, "A2_GAMM-speed-rank.rds"))
- mod5 <- readRDS(file.path(path, "A3_GAMM-speed-rank.rds"))
+# Load models
+mod1 <- readRDS(file.path(path, "A1_GAMM-rank.rds"))
+mod2 <- readRDS(file.path(path, "A2_GAMM-rank.rds"))
+mod3 <- readRDS(file.path(path, "A3_GAMM-rank.rds"))
+mod4 <- readRDS(file.path(path, "A2_GAMM-speed-rank.rds"))
+mod5 <- readRDS(file.path(path, "A3_GAMM-speed-rank.rds"))
 
  # Check object size
  #print(object.size(mod1), units = "MB")
@@ -100,122 +99,139 @@
 
 # Trace plots and parameter distributions -------------------------------
 
- #plot(mod1)
- #plot(mod2)
- #plot(mod3)
- #plot(mod4)
- #plot(mod5)
+plot(mod1)
+plot(mod2)
+plot(mod3)
+plot(mod4)
+plot(mod5)
 
 
 
 # Posterior predictive checks -------------------------------------------
 
- # Check distributions
- pp1 <- pp_check(mod1)
- pp1 <- pp1 + ggtitle("GAMM - A1") +
-  theme(legend.position = "bottom")
[…]

### 2024-07-10 code/code_figures/figureS1.R
switched appendix fig order

@@ -4,9 +4,7 @@
 
 # ==========================================================================
 
-# Figure of the posterior parameters for:
-# - Groups of players that are randomly sampled
-# - The group studied in the paper
+# Figure of the relationship between prey speed and predator XP
 
 
 
@@ -26,34 +24,26 @@ library(viridis)
 
 # Import model -------------------------------------------------------------
 
-path <- file.path(getwd(), "ouputs", "outputs_models")
-fit <- readRDS(
-  file.path(path, "DHMLM-RandomSample.rds")
-)
-# ==========================================================================
-# ==========================================================================
+path <- file.path(getwd(), "outputs", "outputs_models")
+fit <- readRDS(file.path(path, "A2_GAMM-speed-rank.rds"))
 
 
 
+# Load the data ------------------------------------------------------------
 
-
-# ==========================================================================
-# 2. Extract posterior draws
-# ==========================================================================
-
-# Extraction with specified parameters
-draws <- data.table(
-  brms::as_draws_df(
-    x = fit,
-    variable = c(
-      "b_huntingsuccess_total_xp_predearly_quitters",
-      "b_huntingsuccess_total_xp_predengaged",
-      "b_huntingsuccess_total_xp_predmid_quitters",
-      "b_huntingsuccess_total_xp_predslightly_engaged"
-    )
-  )
+data <- fread(
+  "./data/FraserFrancoetal2023-data.csv",
+  select = c("predator_id",
+             "game_duration",
+             "pred_speed",
+             "prey_avg_speed",
+             "prey_avg_rank",
+             "cumul_xp_pred",
+             "total_xp_pred",
+             "hunting_success")
 )
 
+data[, predator_id := as.factor(predator_id)]
 # ==========================================================================
 # ==========================================================================
 
@@ -62,39 +52,76 @@ draws <- data.table(
 
 
 # ==========================================================================
-# 3. Reshape posterior draws table for plotting
+# 2. Compute the predictions
 # ==========================================================================
 
-# Convert to long format
[…]

### 2024-07-10 code/code_figures/figureS2.R

@@ -4,7 +4,9 @@
 
 # ==========================================================================
 
-# Figure of the relationship between prey speed and predator XP
+# Figure of the posterior parameters for:
+# - Groups of players that are randomly sampled
+# - The group studied in the paper
 
 
 
@@ -24,26 +26,34 @@ library(viridis)
 
 # Import model -------------------------------------------------------------
 
-path <- file.path(getwd(), "outputs", "outputs_models")
-fit <- readRDS(file.path(path, "A2_GAMM-speed-rank.rds"))
+path <- file.path(getwd(), "ouputs", "outputs_models")
+fit <- readRDS(
+  file.path(path, "DHMLM-RandomSample.rds")
+)
+# ==========================================================================
+# ==========================================================================
 
 
 
-# Load the data ------------------------------------------------------------
 
-data <- fread(
-  "./data/FraserFrancoetal2023-data.csv",
-  select = c("predator_id",
-             "game_duration",
-             "pred_speed",
-             "prey_avg_speed",
-             "prey_avg_rank",
-             "cumul_xp_pred",
-             "total_xp_pred",
-             "hunting_success")
+
+# ==========================================================================
+# 2. Extract posterior draws
+# ==========================================================================
+
+# Extraction with specified parameters
+draws <- data.table(
+  brms::as_draws_df(
+    x = fit,
+    variable = c(
+      "b_huntingsuccess_total_xp_predearly_quitters",
+      "b_huntingsuccess_total_xp_predengaged",
+      "b_huntingsuccess_total_xp_predmid_quitters",
+      "b_huntingsuccess_total_xp_predslightly_engaged"
+    )
+  )
 )
 
-data[, predator_id := as.factor(predator_id)]
 # ==========================================================================
 # ==========================================================================
 
@@ -52,76 +62,39 @@ data[, predator_id := as.factor(predator_id)]
 
 
 # ==========================================================================
-# 2. Compute the predictions
+# 3. Reshape posterior draws table for plotting
 # ==========================================================================
 
[…]

### 2024-07-10 code/code_figures/figureS1.R
switched back appendix fig order

@@ -4,7 +4,9 @@
 
 # ==========================================================================
 
-# Figure of the relationship between prey speed and predator XP
+# Figure of the posterior parameters for:
+# - Groups of players that are randomly sampled
+# - The group studied in the paper
 
 
 
@@ -24,26 +26,34 @@ library(viridis)
 
 # Import model -------------------------------------------------------------
 
-path <- file.path(getwd(), "outputs", "outputs_models")
-fit <- readRDS(file.path(path, "A2_GAMM-speed-rank.rds"))
+path <- file.path(getwd(), "ouputs", "outputs_models")
+fit <- readRDS(
+  file.path(path, "DHMLM-RandomSample.rds")
+)
+# ==========================================================================
+# ==========================================================================
 
 
 
-# Load the data ------------------------------------------------------------
 
-data <- fread(
-  "./data/FraserFrancoetal2023-data.csv",
-  select = c("predator_id",
-             "game_duration",
-             "pred_speed",
-             "prey_avg_speed",
-             "prey_avg_rank",
-             "cumul_xp_pred",
-             "total_xp_pred",
-             "hunting_success")
+
+# ==========================================================================
+# 2. Extract posterior draws
+# ==========================================================================
+
+# Extraction with specified parameters
+draws <- data.table(
+  brms::as_draws_df(
+    x = fit,
+    variable = c(
+      "b_huntingsuccess_total_xp_predearly_quitters",
+      "b_huntingsuccess_total_xp_predengaged",
+      "b_huntingsuccess_total_xp_predmid_quitters",
+      "b_huntingsuccess_total_xp_predslightly_engaged"
+    )
+  )
 )
 
-data[, predator_id := as.factor(predator_id)]
 # ==========================================================================
 # ==========================================================================
 
@@ -52,76 +62,39 @@ data[, predator_id := as.factor(predator_id)]
 
 
 # ==========================================================================
-# 2. Compute the predictions
+# 3. Reshape posterior draws table for plotting
 # ==========================================================================
 
[…]

### 2024-07-10 code/code_figures/figureS2.R

@@ -4,9 +4,7 @@
 
 # ==========================================================================
 
-# Figure of the posterior parameters for:
-# - Groups of players that are randomly sampled
-# - The group studied in the paper
+# Figure of the relationship between prey speed and predator XP
 
 
 
@@ -26,34 +24,26 @@ library(viridis)
 
 # Import model -------------------------------------------------------------
 
-path <- file.path(getwd(), "ouputs", "outputs_models")
-fit <- readRDS(
-  file.path(path, "DHMLM-RandomSample.rds")
-)
-# ==========================================================================
-# ==========================================================================
+path <- file.path(getwd(), "outputs", "outputs_models")
+fit <- readRDS(file.path(path, "A2_GAMM-speed-rank.rds"))
 
 
 
+# Load the data ------------------------------------------------------------
 
-
-# ==========================================================================
-# 2. Extract posterior draws
-# ==========================================================================
-
-# Extraction with specified parameters
-draws <- data.table(
-  brms::as_draws_df(
-    x = fit,
-    variable = c(
-      "b_huntingsuccess_total_xp_predearly_quitters",
-      "b_huntingsuccess_total_xp_predengaged",
-      "b_huntingsuccess_total_xp_predmid_quitters",
-      "b_huntingsuccess_total_xp_predslightly_engaged"
-    )
-  )
+data <- fread(
+  "./data/FraserFrancoetal2023-data.csv",
+  select = c("predator_id",
+             "game_duration",
+             "pred_speed",
+             "prey_avg_speed",
+             "prey_avg_rank",
+             "cumul_xp_pred",
+             "total_xp_pred",
+             "hunting_success")
 )
 
+data[, predator_id := as.factor(predator_id)]
 # ==========================================================================
 # ==========================================================================
 
@@ -62,39 +52,76 @@ draws <- data.table(
 
 
 # ==========================================================================
-# 3. Reshape posterior draws table for plotting
+# 2. Compute the predictions
 # ==========================================================================
 
-# Convert to long format
[…]

### 2024-07-10 manuscript/appendix.Rmd
update appendix

@@
 ```
+
+```{r figS2, out.width = "100%", fig.cap = "Figure S2.
+Median posterior predictions of the relationship between hunting success (i.e.
+the probability of capturing the four prey) and the average speed of the prey group.
+The black line represents the median posterior predictions, and the gray band the the 89% compatibility interval.
+The predictions are from the model with the highest predictive accuracy (see Table 1 in main text)."}
+
+knitr::include_graphics(
+  here::here(
+    "outputs", "outputs_figures",
+    "figureS2.png"
+  )
+)
+```

### 2024-07-10 manuscript/manuscript.Rmd
results section up to date

@@
 **revoir la phrase sur ce qu'on fait dans l'étude. Intégrer l'aspect des proies**
-
-**arranger les appendices pour inclure la table qui ne montre que le succès**
-
-**il n'y aurait qu'un seul appendice**
 
@@
 To verify that our sample was not biased, we compared a random sample of players that played either 20 to 50 matches, 51 to 100 matches, or 101 to 300 matches during the same timeframe as our sampled population.
-We then took the first 20 matches played by these players, including those from our sampled population, and compared their mean hunting success using a Bayesian hierarchical linear model.
-We found that all four groups had similar successes as predators (Appendix 2: Table S1-S2), which suggests an absence of bias due to data sampling.
+We then took the first 20 matches played by these players, including those from our sampled population, and compared their median hunting success using a Bayesian hierarchical linear model.
+We found that all four groups had similar success as predators (Appendix 1: Table S1 and Figure S1), suggesting an absence of bias due to data sampling.
 
@@
 
-We parametrized the GAMMs to run four MCMC chains with 1000 posterior samples for each parameter.
+We parametrized the GAMMs to run four MCMC chains.
 We ran 2500 iterations with a thinning set to eight for the additive model with a global smoother only (see Table 1), and 1500 iterations with a thinning set to four for the remaining additive models.
 We set the first 500 iterations of each model as warm ups.
+For each model, we obtained 1000 posterior samples per parameter.
 We assessed the convergence of the MCMC chains using trace plots, R-hat diagnostics with a threshold of <1.01, and effective sample sizes (ESS) with a threshold of >100 [@vehtari.etal2021].
@@
 # RESULTS {.unnumbered}
-
-## Development of expertise at the population level
 
@@
 
+## Development of expertise at the population level
+
 Out of all five GAMM models, the two that accounted for the prey group's rank and speed were the best at predicting the data, achieving similar expected log pointwise densities (Table 1).
-Models in which prey effects were not accounted for resulted in no change in hunting success with experience (i.e.
-expertise) for the average individual.
-Accounting for the prey rank resulted in a concave-shaped relationship, with the highest success ranging between 100 and 400 matches (Figure 1A).
+Models in which prey effects were not accounted for resulted in no change in hunting success with experience for the average individual (i.e.
+no gain in expertise).
+Accounting for the prey rank resulted in a concave-shaped relationship, with the highest success ranging between ~200 and ~300 matches (Figure 1A).
+
+We found strong evidence of a negative relationship between hunting success and prey speed (Figure S2).
 In the model where we included the prey's speed, the effect of experience on hunting success for the average individual followed a diminishing returns curve as we predicted, with predators optimizing their success after playing ~300 matches (Figure 1C).
-The curve shows there was a 38% increase in the probability of consuming all prey for the average individual between the first and the ~330$^{th}$ match where success reached a plateau (Figure 1C).
+The curve shows there was a 38% increase in the probability of consuming all prey for the average individual between the first and the ~330$^{th}$ match, where success reached a plateau (Figure 1C).
 
@@
 We detected large differences in the standard deviations of the wiggliness (Figure 2, median = `r toString(round(sum1[4,]$Estimate, digits = 2))` vs median = `r toString(round(sum2[4,]$Estimate, digits = 2))`).
-The lower standard deviation for the model were we accounted for prey speed suggests that the form of the relationship between success and experience was more similar among individuals.
+The lower standard deviation for the model where we accounted for prey speed suggests that the form of the relationship between success and experience was more similar among individuals.
 
@@
 
-It is generally assumed that predators increase prey consumption through time by acquiring hunting expertise [@dukas2019;@wooster.etal2023].
-Yet, few studies have empirically tested how prey antipredator strategies mediate the acquisitio of hunting expertise because of the challenges of investigating direct predator-prey interactions in the wild.
+Predators usually increase prey consumption through time by acquiring hunting expertise [@dukas2019;@wooster.etal2023].
+Yet, few studies have empirically tested how prey antipredator strategies mediate the acquisition of hunting expertise because of the challenges of investigating direct predator-prey interactions in the wild.
 By capitalizing on a virtual predator-prey system where interactions were directly monitored, we found that a predator-player population in *Dead by Daylight* increased their hunting success with experience.

### 2024-07-10 manuscript/manuscript.Rmd
manuscript completed

@@
 In this study, we assess how repeated encounters with prey shapes predator hunting success using data from players in *Dead by Daylight*.
-First, we investigate the development of expertise at the population level.
+First, we investigate how the predator population develops its hunting expertise.
 We hypothesize that the predator population's success will increase with experience up to a certain level where it will stabilize [@dukas2019].
@@
 
-*revoir la première phrase, devrait intégrer les proies*
-
-Predators usually increase prey consumption through time by acquiring hunting expertise [@dukas2019;@wooster.etal2023].
-Yet, few studies have empirically tested how prey antipredator strategies mediate the acquisition of hunting expertise because of the challenges of investigating direct predator-prey interactions in the wild.
+Predators usually optimize prey consumption through time by acquiring hunting expertise [@dukas2019;@wooster.etal2023].
+Yet, few studies have empirically tested how prey antipredator strategies mediate the acquisition of hunting expertise due to the challenges of investigating direct predator-prey interactions in the wild.
 By capitalizing on a virtual predator-prey system where interactions were directly monitored, we found that a predator-player population in *Dead by Daylight* increased their hunting success with experience.
@@
 Our observations are consistent with empirical studies of expertise in both humans and nonhuman animals [reviewed in @dukas2019].
-However, the prey's speed was important in mediating this pattern at the population level, probably because prey also increased their speed as predators gained experience.
+However, the prey's speed was important in mediating this pattern at the population level.
+Encountering faster prey led to more difficult encounters for predators.
 We previously showed in *DBD* that faster movement is an effective strategy used by the prey to avoid predation [@fraserfranco.etal2022], and other studies have found that as well [@walker.etal2005;@kelley.magurran2011;@martin.etal2022].
-This would explain why the relationship was concave in the model where we did not account for prey speed, because adept predators were encountering prey with higher speeds.
+This could explain why the relationship was concave in the model where we did not account for prey speed, because the prey were probably increasing their speed as they gained experience.
+Because the matchmaking algorithm pairs players based on their skill, adept predators were probably encountering prey moving at higher speeds.
 Thus, our results suggest that predators can gain expertise and maintain success when they encounter prey that move at speeds lower or close to the population-average.
 
-Prey speed also mediated differences among predator players in the development of their expertise, implying that prey antipredator behaviour can impair the development of an individual predator's expertise.
+Prey speed also mediated differences among predator players in the development of their expertise.
 Individual predator players probably differed in their capacity to adjust to difficult prey.
-For example, hunting faster prey requires specialized cognitive abilities and coordination that are energetically costly [@kelley.magurran2011].
-Thus, predators that couldn't develop counter-strategies for faster prey were likely at a disadvantage.
+Animals are expected to have limited attention that constrain diet choice and image search formation [@dukas.etal2001].
+Moreover, hunting faster prey requires specialized cognitive abilities and coordination that are energetically costly [@kelley.magurran2011].
+Thus, predators that couldn't develop counter-strategies to detect or chase faster prey were likely at a disadvantage.
+Parallel observations have been outlined in studies of prey camouflage strategies.
+For example, @troscianko.etal2018 showed in a computer experiment involving humans that disruptive camouflage was efficient at preventing the acquisition of expertise during search image formation.
+Human subjects exposed to a restricted set of strategies were also less efficient compared to those exposed to a variety of strategies.
+Therefore, our observations suggest that prey antipredator behaviour can also impair predator expertise acquisition, with potential implications for predator-prey dynamics.
 
-*ici, répéter les études utilisant le camouflage, ensuite inclure la phrase des compromis, et conclure avec une avenue évolutive.*
-
-**Such trade-offs may reflect limitations in learning all the skills required to successfully hunt all types of prey [@healy1992;@belisle1997;@dukas2019], particularly if the skills required to hunt slower prey are nontransferable to faster prey.**
-
-However, even after we took the prey's speed into account, there were still important differences in expertise acquisition among predator players, suggesting that other antipredator tactics were potentially involved.
-For example, the prey can use camouflage and hide to avoid detection (not currently measurable in *DBD*), requiring predators to learn how to exploit visual cues that facilitate cryptic prey detection, such as habitat characteristics or prey colour patterns [@ehlinger1989;@hughes.etal1992;@warburton2003;@szopa-comley.etal2020].
-Furthermore, it is hypothesized that longer time intervals between hunting events could hinder or delay the acquisition of expertise as individuals may forget information with longer delays [@endler1991;@wright.etal2022].
+Despite adjusting for the prey's speed, discernible variations in expertise acquisition among predator players persisted, indicating the possible influence of other factors.
+It is hypothesized that longer time intervals between hunting events can hinder or delay the acquisition of expertise because individuals may forget information when delays are longer [@endler1991;@wright.etal2022].
 In *DBD*, a predator that played 300 matches in the span of six months might forget more critical information related to prey detection or escape patterns than one that played 300 matches in the span of six days.
 Investigating the impact of such time lags in future analyses may reveal important insights on the outcome of predator-prey interactions.
-
-**Of course, the life of neither the predator nor the prey players are at stake, such that emerging patterns could be driven by their motivation to win and not "true" survival. Yet,**
-**A potential caveat is that the more flexible hunters might have experimented with various tactics out of boredom, which could impede ecologically realistic interpretations of our data.
-However, the consistent association between this tactic and heterogeneous prey groups gives us confidence that it emerged from their interactions.
-One limitation of our study was that we couldn't monitor all the matches of the prey, which prevented us from assessing their responses to the predator through their experience.** Hence, future studies should aim at monitoring reciprocal behavioural dynamics through time, which may reveal important insights on the mechanisms driving stable equilibria in predator-prey systems.
-
-**Comment by reviewer:**
-
-Some characteristics to be aware of: 
-- Predators do not compete.
-- Both predators and prey can fail in a round yet still carry that experience and knowledge on to the next round with no real penalty (think of the movie Groundhog Day).
-- Predator and prey are surely playing a game with each other, but prey may also be playing a game among themselves, which may affect their game with the predator. 
-- This particular video game appears to have a few unusual characteristics.
-Prey do not appear to forage for resources.
-but rather they create a public good.
-Furthermore, handling time for the predator is extremely complicated and lengthy.
-- Finally, it may be rather complicated to discern all of the rules of how the different virtual environments work and what  all the aptitudes of all the different predators and prey are.
+Another limitation of our study is that the life of neither the predator nor the prey players are at stake in *DBD*, such that emerging patterns could be driven by their motivation to win and not "true" survival.
+For example, some players could experiment with the game out of boredom, which could impede ecologically realistic interpretations of our data.
+Lastly, the prey can be consumed in a match but carry that experience on to the next round without any penalty.
+Thus, interpretations of our results must be made with these caveats in mind.
 
@@
 
-We found support of our prediction that prey were driving individual differences in expertise in a predator population.
-Our results suggest that predators learned with experience, as their success increased...
-Even though individuals were not all equally successful.
-Lastly, virtual systems are increasingly recognized as being useful to test hypotheses on ecological interactions [@beauchamp2020;@cere.etal2021;@fraserfranco.etal2022;@lymbery.etal2023].
+We found support of our hypothesis that prey antipredator behaviour was driving individual differences in expertise in a predator population.
+Our results suggest that predators gained expertise, as their success increased wit extensive practice, but that prey speed impaired expertise acquisition.
+Future analyses should investigate how efficient antipredator tactics are associated with prey experience, as this may reveal important insights on the eco-evolutionary dynamics of predator prey interactions.
+Lastly, virtual systems are increasingly used as systems to test hypotheses on ecological interactions [@beauchamp2020;@cere.etal2021;@fraserfranco.etal2022;@lymbery.etal2023;@santostefano.etal2024].
 We therefore hope that our study will inspire more collaborations between scientists and the videogame industry to tackle fundamental questions in ecology.

### 2024-07-10 manuscript/manuscript.Rmd
manuscript truly finished (forgot abstract)

@@
 
-**revoir abstract pour être moins centré sur la tactique**
-
-**revoir la phrase sur ce qu'on fait dans l'étude. Intégrer l'aspect des proies**
-
 The acquisition of expertise is crucial for predators to be successful hunters.
 To do so, predators need to hone their skills and acquire knowledge through repeated and extensive practice.
+Recent studies suggest that such acquisition of expertise may not follow expected patterns when prey use antipredator tactics to evade detection and pursuit.
+Such tactics may even impair the acquisition of predator expertise.
 Yet, there are limited empirical assessments showing how predators acquire expertise through repeated encounters with their prey due to the challenges of monitoring direct interactions in the wild.
 Here, we use a virtual predator-prey system (the game *Dead by Daylight*) to investigate how experience shapes individual and population hunting success in human predators across repeated interactions with their prey.
-We show that predators optimized prey consumption as they gained experience, indicating that they acquired expertise with extensive practice.
+We show that predators optimized prey consumption as they gained experience, indicating that they acquired expertise through extensive practice.
 Prey speed was an important mediator of this relationship, driving differences among predators in the development of expertise.

### 2024-08-02 manuscript/manuscript.Rmd
separated a paragraph to ease reading
aranged intro and updated refs. discussion left to polish.

@@
 
-The ecological consequences of predation accrue to the sum of consumptive and nonconsumptive effects (CEs and NCEs) during predator-prey interactions [@preisser2007].
+Predation is a fundamental biological process driving ecosystem functioning.
 Recent advances in our understanding of the mechanisms (e.g.
-foraging mode, reciprocal plasticity, size-dependent foraging) driving (N)CEs at the individual and population level have allowed ecologists to predict their outcome as well as their effect on prey population dynamics [@Peckarsky.etal2008;@gravel.etal2013;@wirsing.etal2021].
-For example, empirical studies show that prey tend to increase vigilance at the expense of foraging in the presence of ambush predators (NCEs) [@preisser2007], with such hunters consuming a greater range of prey types (CEs) compared to cursorial hunters [@glaudas2019].
+foraging mode, reciprocal plasticity, size-dependent foraging) driving predator foraging success at the individual and population level have allowed ecologists to predict their consequences on prey behaviour and population dynamics [@Peckarsky.etal2008;@gravel.etal2013;@wirsing.etal2021].
+Prey typically respond to predation by managing tradeoffs between vigilance and foraging depending on the type of predation regime [@preisser2007].
+Such prey antipredator resonses are hypothesized to drive differences among predators in their hunting success by disrupting the acquisition of hunting expertise [@wooster.etal2023], which is deemed essential for many predators to reach adulthood and survive [@woo.etal2008;@wooster.etal2023].
 
-There are, however, internal mechanisms contributing to predator foraging success that are often overlooked when studying (N)CEs in predator-prey interactions [@wooster.etal2023].
-For instance, predators need to acquire expertise by practicing and learning the proper skills to successfully locate, select, and capture their prey [@dukas2019;@wooster.etal2023].
+Expertise can be defined as the characteristics, skills, and knowledge allowing individuals with extensive practice to outperform novices on complex tasks [@dukas2017].
+In the context of hunting, predators need to acquire expertise by practicing and learning the proper skills to successfully locate, select, and capture their prey.
 Empirical studies on human and non-human hunters show that individuals optimize foraging efficiency (e.g.
-search and handling times, return rates) by learning associative images, or by using cues from their prey and their environment [@edwards.jackson1994;@morse2000;@macdonald2007;@reid.etal2010;@wilson-rankin2015].
-Thus, predators can integrate this information to increase their chances of locating prey, or to use the attack tactic most suited for a given context.
+search and handling times, return rates) by associative learning, by developing search images, or by exploiting cues from their prey and their environment [@edwards.jackson1994;@morse2000;@macdonald2007;@reid.etal2010;@wilson-rankin2015].
+Through these processes, expert predators should have greater knowledge, better energy management, and accute motor skills to increase their chances of locating and capturing prey [@dukas2019].
+Therefore, differences among predators in their capacity to acquire expertise throughout their lifetime may underlie differences in hunting success, with potential consequences for predator-prey dynamics [@wooster.etal2023].
 
-On the other hand, prey use antipredator tactics such as rapid escapes to evade predators [@walker.etal2005;@kelley.magurran2011;@herbert-read.etal2017].
-This compels predators to develop swift motor skills to succesfully capture such prey.
+Prey are known to use evasive tactics such as rapid escapes and movements to avoid being captured [@walker.etal2005;@kelley.magurran2011;@herbert-read.etal2017], compelling predators to develop swift motor skills to succesfully capture such prey.
 However, predators may be limited in their capacity to develop the necessary attributes (e.g.
 physical, physiological, neurological) for fast-paced hunting, which may impair the development of their hunting expertise.
-For example, experimental studies have shown that certain camouflage strategies, such as disruptive coloration, can impair expertise acquisition in humans and birds [@martin.etal2012;@troscianko.etal2013;@troscianko.etal2018].
-Unfortunately, while our understanding of how predation risk influences information acquisition in prey has increased over the years [reviewed in @crane.etal2024], there is only scarce evidence showing links between prey antipredator strategies and the development of expertise in human and nonhuman predators.
-Therefore, our ability to accurately predict the (N)CEs of predation on prey populations remain limited.
+There is only scarce evidence showing links between prey antipredator strategies and the development of expertise in human and nonhuman predators.
+Experimental studies have shown that certain camouflage strategies in prey can impair expertise acquisition in humans and birds [@martin.etal2012;@troscianko.etal2013].
+For instance, @troscianko.etal2018 found that disruptive coloration interfered with the formation of search images in human subjects, hindering improvement in detection times after repeated attempts.
+Given these examples, antipredator behaviour may also hold the potential to hinder the development of predator expertise.
 
 A recurring challenge impeding research on predator-prey behavioural interactions is the need to collect data simultaneously on both the predator and the prey.
-Similar to agent-based simulations, online videogames can be useful systems to mitigate these challenges and test specific ecological hypotheses.
-They provide controlled virtual environments to test ecological hypotheses [see @lymbery.etal2023] with the advantage of having real players that interact in the virtual space.
-For example, in the predator-prey videogame *Dead by Daylight*, four prey players need to forage for resources while avoiding predation by a fifth player.
-In this virtual system, the predator population is composed of individuals that either ambush or hunt at high speeds (i.e.
+Here we mitigate these challenges by using *Dead by Daylight* as our study system, a videogame where four prey players need to forage for resources while avoiding predation by a fifth player.
+Similar to agent-based simulations, *Dead by Daylight* provides controlled virtual environments to test ecological hypotheses [see @lymbery.etal2023 for an example with *Age of Empires II*], but with the advantage of having real players that interact in the virtual space.
+In this game, the predator population is composed of individuals that either ambush or hunt at high speeds (i.e.
 mean movement speed along a slow-fast continuum), and their success is driven by the movement of the prey [@fraserfranco.etal2022].
@@
 The game also ellicits natural reactions in players such as freezing when predation is imminent (personal observations), which corroborates with another virtual ecological study showing that predation drives individual variation in risk perception [@beauchamp2020].
-These observations outline how fundamental ecological patterns can emerge from human interactions in virtual systems with fixed rules [@brosnan.etal2017;@kasumovic.etal2017].
-Videogames also generate large volumes of data on interacting players throughout their lifetime in the game under realistic, controlled, and repeatable ecological scenarios.
-Hence, virtual systems such as *Dead by Daylight* offer the opportunity to tackle fundamental questions about the role of experience on predator-prey interactions.
+These observations outline how ecological phenomena can emerge from human interactions in virtual systems with fixed rules [@brosnan.etal2017;@kasumovic.etal2017].
+Videogames also generate large volumes of data on thousands of interacting players throughout their lifetime in the game under realistic, controlled, and repeatable ecological scenarios.
+Hence, *Dead by Daylight* allows us to tackle fundamental questions about the role of antipredator behaviour and experience on predator-prey interactions.
 
@@
 We assumed that the random intercepts for the predator ID ($id$) followed a Gaussian distribution with estimated standard deviation ($id \sim N(0,\ I\ \sigma_{id}^{2})$).
+
 We used weakly informative Gaussian priors for the intercept ($N(0,\ 0.5^{2})$) and the global trend of cumulative experience ($N(0,\ 2^{2}))$).
@@
 This could explain why the relationship was concave in the model where we did not account for prey speed, because the prey were probably increasing their speed as they gained experience.
-Because the matchmaking algorithm pairs players based on their skill, adept predators were probably encountering prey moving at higher speeds.
+Because the matchmaking algorithm pairs players based on their skill, experienced predators were probably encountering prey moving at higher speeds.
 Thus, our results suggest that predators can gain expertise and maintain success when they encounter prey that move at speeds lower or close to the population-average.
@@
 Investigating the impact of such time lags in future analyses may reveal important insights on the outcome of predator-prey interactions.
-Another limitation of our study is that the life of neither the predator nor the prey players are at stake in *DBD*, such that emerging patterns could be driven by their motivation to win and not "true" survival.
+Another potential reason why differences among individuals persisted is that the life of neither the predator nor the prey players are at stake in *DBD*, such that emerging patterns could be driven by their motivation to win and not "true" survival.
 For example, some players could experiment with the game out of boredom, which could impede ecologically realistic interpretations of our data.
@@
 Thus, interpretations of our results must be made with these caveats in mind.
+**inclure l'idée du prédateur qui peut aussi driver ça**
 

### 2024-08-16 manuscript/manuscript.Rmd
updated abstract

@@
 To do so, predators need to hone their skills and acquire knowledge through repeated and extensive practice.
-Recent studies suggest that such acquisition of expertise may not follow expected patterns when prey use antipredator tactics to evade detection and pursuit.
-Such tactics may even impair the acquisition of predator expertise.
-Yet, there are limited empirical assessments showing how predators acquire expertise through repeated encounters with their prey due to the challenges of monitoring direct interactions in the wild.
+However, recent studies suggest that predators may not necessarily increase their success with expertise when prey use antipredator tactics to evade detection and pursuit.
+Therefore, antipredator tactics may impair the acquisition of predator expertise, but there are limited empirical assessments showing how predators acquire expertise through repeated encounters with their prey due to the challenges of monitoring direct interactions in the wild.
 Here, we use a virtual predator-prey system (the game *Dead by Daylight*) to investigate how experience shapes individual and population hunting success in human predators across repeated interactions with their prey.
 We show that predators optimized prey consumption as they gained experience, indicating that they acquired expertise through extensive practice.
-Prey speed was an important mediator of this relationship, driving differences among predators in the development of expertise.
-At the population-level, we found that faster prey could impair the acquisition of expertise by reducing hunting success.
-At the individual-level, our results show that prey speed shapes how individual predators acquire expertise.
-Therefore, our study outlines how prey antipredator behaviour can mediate the acquisition of expertise in predator populations.
+At the population-level, we found that faster prey impair the acquisition of expertise by reducing hunting success.
+Prey speed was also an important mediator of this relationship at the individual level, driving differences among predators in the development of expertise.
+Our study outlines how prey antipredator behaviour can mediate the acquisition of expertise in predator populations.
 
@@
 
-Predation is a fundamental biological process driving ecosystem functioning.
+Predation is a fundamental biological process involved in ecosystem functioning.
 Recent advances in our understanding of the mechanisms (e.g.

### 2024-08-16 manuscript/manuscript.Rmd
updated discussion

@@
 There is only scarce evidence showing links between prey antipredator strategies and the development of expertise in human and nonhuman predators.
-Experimental studies have shown that certain camouflage strategies in prey can impair expertise acquisition in humans and birds [@martin.etal2012;@troscianko.etal2013].
+Experimental studies have shown that certain prey camouflage strategies can impair expertise acquisition in humans and birds [@martin.etal2012;@troscianko.etal2013].
 For instance, @troscianko.etal2018 found that disruptive coloration interfered with the formation of search images in human subjects, hindering improvement in detection times after repeated attempts.
@@
 
-```{r}
+```{r summary, cache = TRUE}
 # Import top 2 model fits
@@
 By capitalizing on a virtual predator-prey system where interactions were directly monitored, we found that a predator-player population in *Dead by Daylight* increased their hunting success with experience.
-Yet, there were important differences among individuals in the development of said expertise, which we found to be in part due to the movement of the prey encountered.
+Yet, there were important differences among individuals in the development of said expertise, which we found to be in part due to the speed of the prey encountered.
+Hence, our study provides empirical evidence that prey antipredator behaviour can impair the development of hunting expertise.
 
@@
 The predator population displayed an asymptotic relationship between experience and success, wherein initial gains in success were significant but gradually stabilized as experience accumulated.
-Our observations are consistent with empirical studies of expertise in both humans and nonhuman animals [reviewed in @dukas2019].
+These observations are consistent with empirical studies of expertise in both humans and nonhuman animals [reviewed in @dukas2019].
 However, the prey's speed was important in mediating this pattern at the population level.
@@
 
-Prey speed also mediated differences among predator players in the development of their expertise.
-Individual predator players probably differed in their capacity to adjust to difficult prey.
-Animals are expected to have limited attention that constrain diet choice and image search formation [@dukas.etal2001].
-Moreover, hunting faster prey requires specialized cognitive abilities and coordination that are energetically costly [@kelley.magurran2011].
+Prey speed also mediated differences among predator players in the development of their expertise, suggesting that individual predator players differed in their capacity to adjust to difficult prey.
+Animals are expected to have limited attention that constrain diet choice and image search formation, and hunting faster prey requires specialized cognitive abilities and coordination that are energetically costly [@dukas.etal2001;@kelley.magurran2011].
 Thus, predators that couldn't develop counter-strategies to detect or chase faster prey were likely at a disadvantage.
@@
 
-Despite adjusting for the prey's speed, discernible variations in expertise acquisition among predator players persisted, indicating the possible influence of other factors.
-It is hypothesized that longer time intervals between hunting events can hinder or delay the acquisition of expertise because individuals may forget information when delays are longer [@endler1991;@wright.etal2022].
+Despite adjusting for the prey's speed, discernible variations in expertise acquisition among predator players persisted.
+A potential explanation is that the predators' hunting tactic may drive the antripredator tactic of the prey.
+We know from a previous study that predators specialize as cursorial or ambush hunters [@fraserfranco.etal2022].
+Therefore, predators using a cursorial tactic may drive prey to move faster at the cost of having a slower acquisition of expertise.
+
+Despite adjusting for the prey's speed, noticeable differences in the predators' expertise acquisition still emerged.
+One possible explanation is that the predators' hunting tactics may indirectly shape their own expertise through changes in prey behaviour.
+Previous research indicates that predators tend to specialize as either cursorial or ambush hunters in *DBD* [@fraserfranco.etal2022].
+Consequently, predators employing a cursorial tactic may push prey to move faster, which, in turn, may hinder their own expertise acquisition due to the increased difficulty of hunting such prey.
+An alternative explanation is that longer time intervals between hunting events are hypothesized to hinder or delay the acquisition of expertise because individuals may forget information when delays are longer [@endler1991;@wright.etal2022].
 In *DBD*, a predator that played 300 matches in the span of six months might forget more critical information related to prey detection or escape patterns than one that played 300 matches in the span of six days.
 Investigating the impact of such time lags in future analyses may reveal important insights on the outcome of predator-prey interactions.
-Another potential reason why differences among individuals persisted is that the life of neither the predator nor the prey players are at stake in *DBD*, such that emerging patterns could be driven by their motivation to win and not "true" survival.
-For example, some players could experiment with the game out of boredom, which could impede ecologically realistic interpretations of our data.
-Lastly, the prey can be consumed in a match but carry that experience on to the next round without any penalty.
-Thus, interpretations of our results must be made with these caveats in mind.
-**inclure l'idée du prédateur qui peut aussi driver ça**
+Another potential reason for the persistent differences among individuals is that neither the predator nor prey players' lives are at stake in *DBD*.
+As a result, emerging patterns may be driven more by the players' motivation to win rather than "true" survival.
+For example, some players could experiment with the game out of boredom, which could shape how expertise is honed and impede ecologically realistic interpretations of our data.
 
@@
 
-We found support of our hypothesis that prey antipredator behaviour was driving individual differences in expertise in a predator population.
-Our results suggest that predators gained expertise, as their success increased wit extensive practice, but that prey speed impaired expertise acquisition.
-Future analyses should investigate how efficient antipredator tactics are associated with prey experience, as this may reveal important insights on the eco-evolutionary dynamics of predator prey interactions.
-Lastly, virtual systems are increasingly used as systems to test hypotheses on ecological interactions [@beauchamp2020;@cere.etal2021;@fraserfranco.etal2022;@lymbery.etal2023;@santostefano.etal2024].
+We found support of our hypothesis that prey antipredator behaviour was driving individual differences in expertise acquisition in a human predator population in the game *Dead by Daylight*.
+As we did here with predators, future analyses should investigate how antipredator tactics are developped with experience, as it may reveal important insights on the eco-evolutionary dynamics of predator prey interactions.
+Using a video game system, our study demonstrates that prey antipredator behavior can impair the acquisition of hunting expertise.
+This adds to a growing body of research showing how virtual systems can be used to test hypotheses on ecological interactions [@beauchamp2020;@cere.etal2021;@fraserfranco.etal2022;@lymbery.etal2023;@santostefano.etal2024].
 We therefore hope that our study will inspire more collaborations between scientists and the videogame industry to tackle fundamental questions in ecology.

### 2024-08-16 manuscript/manuscript.Rmd
minor edit to a discussion sentence and knit doc

@@
 Because the matchmaking algorithm pairs players based on their skill, experienced predators were probably encountering prey moving at higher speeds.
-Thus, our results suggest that predators can gain expertise and maintain success when they encounter prey that move at speeds lower or close to the population-average.
+Thus, our results suggest that predators can gain expertise and maintain success when they encounter prey that move at speeds lower than or closer to the population-average.
 

### 2024-08-23 manuscript/manuscript.Rmd
updated abstract

@@
 The acquisition of expertise is crucial for predators to be successful hunters.
-To do so, predators need to hone their skills and acquire knowledge through repeated and extensive practice.
-However, recent studies suggest that predators may not necessarily increase their success with expertise when prey use antipredator tactics to evade detection and pursuit.
-Therefore, antipredator tactics may impair the acquisition of predator expertise, but there are limited empirical assessments showing how predators acquire expertise through repeated encounters with their prey due to the challenges of monitoring direct interactions in the wild.
+To achieve this, predators must hone their skills and gain knowledge through repeated and extensive practice.
+On the other hand, prey may hinder the development of predator expertise by employing antipredator tactics to evade detection and pursuit.
+However, empirical evidence on how predators acquire expertise through repeated encounters with their prey remains limited, largely due to the challenges of monitoring direct interactions in the wild.
 Here, we use a virtual predator-prey system (the game *Dead by Daylight*) to investigate how experience shapes individual and population hunting success in human predators across repeated interactions with their prey.
 We show that predators optimized prey consumption as they gained experience, indicating that they acquired expertise through extensive practice.
-At the population-level, we found that faster prey impair the acquisition of expertise by reducing hunting success.
+At the population-level, we found that faster prey impaired the acquisition of expertise by reducing hunting success.
 Prey speed was also an important mediator of this relationship at the individual level, driving differences among predators in the development of expertise.

### 2024-08-23 manuscript/manuscript.Rmd
update title to "acquisition"

@@
 ---
-title: "Prey movement shapes the development of predator expertise in a virtual bi-trophic system"
+title: "Prey movement shapes the acquisition of predator expertise in a virtual bi-trophic system"
 output: 

### 2024-08-23 manuscript/manuscript.Rmd
update intro. change from development to acquisition.

@@
 To achieve this, predators must hone their skills and gain knowledge through repeated and extensive practice.
-On the other hand, prey may hinder the development of predator expertise by employing antipredator tactics to evade detection and pursuit.
+On the other hand, prey may hinder the acquisition of predator expertise by employing antipredator tactics to evade detection and pursuit.
 However, empirical evidence on how predators acquire expertise through repeated encounters with their prey remains limited, largely due to the challenges of monitoring direct interactions in the wild.
@@
 At the population-level, we found that faster prey impaired the acquisition of expertise by reducing hunting success.
-Prey speed was also an important mediator of this relationship at the individual level, driving differences among predators in the development of expertise.
+Prey speed was also an important mediator of this relationship at the individual level, driving differences among predators in the acquisition of expertise.
 Our study outlines how prey antipredator behaviour can mediate the acquisition of expertise in predator populations.
@@
 
-Predation is a fundamental biological process involved in ecosystem functioning.
-Recent advances in our understanding of the mechanisms (e.g.
-foraging mode, reciprocal plasticity, size-dependent foraging) driving predator foraging success at the individual and population level have allowed ecologists to predict their consequences on prey behaviour and population dynamics [@Peckarsky.etal2008;@gravel.etal2013;@wirsing.etal2021].
-Prey typically respond to predation by managing tradeoffs between vigilance and foraging depending on the type of predation regime [@preisser2007].
-Such prey antipredator resonses are hypothesized to drive differences among predators in their hunting success by disrupting the acquisition of hunting expertise [@wooster.etal2023], which is deemed essential for many predators to reach adulthood and survive [@woo.etal2008;@wooster.etal2023].
-
-Expertise can be defined as the characteristics, skills, and knowledge allowing individuals with extensive practice to outperform novices on complex tasks [@dukas2017].
-In the context of hunting, predators need to acquire expertise by practicing and learning the proper skills to successfully locate, select, and capture their prey.
+Antipredation responses by prey are hypothesized to drive differences among predators in their hunting success by disrupting the acquisition of hunting expertise [@wooster.etal2023].
+Expertise can be defined as the characteristics, skills, and knowledge that provide individuals with the ability to outperform novices on complex tasks [@dukas2017].
+In the context of hunting, predators need to acquire expertise by practising and learning the proper skills to successfully locate, select, and capture their prey, a process deemed essential for many predators to reach adulthood and survive [@woo.etal2008;@wooster.etal2023].
 Empirical studies on human and non-human hunters show that individuals optimize foraging efficiency (e.g.
 search and handling times, return rates) by associative learning, by developing search images, or by exploiting cues from their prey and their environment [@edwards.jackson1994;@morse2000;@macdonald2007;@reid.etal2010;@wilson-rankin2015].
-Through these processes, expert predators should have greater knowledge, better energy management, and accute motor skills to increase their chances of locating and capturing prey [@dukas2019].
+Through these processes, expert predators should have greater knowledge, better energy management, and acute motor skills to increase their chances of locating and capturing prey [@dukas2019].
 Therefore, differences among predators in their capacity to acquire expertise throughout their lifetime may underlie differences in hunting success, with potential consequences for predator-prey dynamics [@wooster.etal2023].
 
-Prey are known to use evasive tactics such as rapid escapes and movements to avoid being captured [@walker.etal2005;@kelley.magurran2011;@herbert-read.etal2017], compelling predators to develop swift motor skills to succesfully capture such prey.
+Prey are known to use evasive tactics such as rapid escapes to avoid being captured [@walker.etal2005;@kelley.magurran2011;@herbert-read.etal2017], driving predators to develop the motor skills required to capture such prey succesfully.
 However, predators may be limited in their capacity to develop the necessary attributes (e.g.
-physical, physiological, neurological) for fast-paced hunting, which may impair the development of their hunting expertise.
-There is only scarce evidence showing links between prey antipredator strategies and the development of expertise in human and nonhuman predators.
-Experimental studies have shown that certain prey camouflage strategies can impair expertise acquisition in humans and birds [@martin.etal2012;@troscianko.etal2013].
-For instance, @troscianko.etal2018 found that disruptive coloration interfered with the formation of search images in human subjects, hindering improvement in detection times after repeated attempts.
-Given these examples, antipredator behaviour may also hold the potential to hinder the development of predator expertise.
+physical, physiological, neurological) for fast-paced hunting, which may impair the acquisition of hunting expertise.
+There is limited empirical evidence showing links between prey antipredator strategies and the acquisition of expertise in human and nonhuman predators.
+Experimental studies show that certain camouflage strategies of prey can impair expertise acquisition in humans and birds [@martin.etal2012;@troscianko.etal2013].
+For instance, @troscianko.etal2018 found that disruptive colouration interfered with the formation of search images in human subjects, hindering improvement in detection times after repeated attempts.
+Given these examples, antipredator behaviour may also hold the potential to hinder the acquisition of predator expertise.
+This is an important topic of study since predator cognition can maintain behavioural variation, shape prey phenotypic composition, and destabilize predator-prey systems [@kondoh2010;skelhornandcandy2016;@kikuchi.etal2023].
 
 A recurring challenge impeding research on predator-prey behavioural interactions is the need to collect data simultaneously on both the predator and the prey.
-Here we mitigate these challenges by using *Dead by Daylight* as our study system, a videogame where four prey players need to forage for resources while avoiding predation by a fifth player.
+Here we mitigate these challenges by using *Dead by Daylight* as our study system, a videogame where four prey players must forage for resources while avoiding predation by a fifth player.
 Similar to agent-based simulations, *Dead by Daylight* provides controlled virtual environments to test ecological hypotheses [see @lymbery.etal2023 for an example with *Age of Empires II*], but with the advantage of having real players that interact in the virtual space.
-In this game, the predator population is composed of individuals that either ambush or hunt at high speeds (i.e.
+In this game, the predator population comprises individuals that either ambush or hunt at high speeds (i.e.
 mean movement speed along a slow-fast continuum), and their success is driven by the movement of the prey [@fraserfranco.etal2022].
 The prey can increase their chances of survival by cooperating and moving fast to escape the predator [@cere.etal2021;@fraserfranco.etal2022;@santostefano.etal2024].
-The game also ellicits natural reactions in players such as freezing when predation is imminent (personal observations), which corroborates with another virtual ecological study showing that predation drives individual variation in risk perception [@beauchamp2020].
+The game also elicits natural reactions in players such as freezing when predation is imminent (personal observations), corroborating another virtual ecological study showing that predation drives individual variation in risk perception [@beauchamp2020].
 These observations outline how ecological phenomena can emerge from human interactions in virtual systems with fixed rules [@brosnan.etal2017;@kasumovic.etal2017].
 Videogames also generate large volumes of data on thousands of interacting players throughout their lifetime in the game under realistic, controlled, and repeatable ecological scenarios.
-Hence, *Dead by Daylight* allows us to tackle fundamental questions about the role of antipredator behaviour and experience on predator-prey interactions.
+Hence, *Dead by Daylight* allows us to tackle fundamental questions about the role of antipredator behaviour and experience in predator-prey interactions.
 
 In this study, we assess how repeated encounters with prey shapes predator hunting success using data from players in *Dead by Daylight*.
-First, we investigate how the predator population develops its hunting expertise.
-We hypothesize that the predator population's success will increase with experience up to a certain level where it will stabilize [@dukas2019].
+We quantify expertise acquisition as the relationship between hunting success (i.e.
+the probability of capturing all prey) and repeated experience (i.e.
+cumulated atempts).
+First, test the hypothesis that predator success will increase with experience up to some point at which it will stabilize [@dukas2019].
 However, we expect this pattern to change depending on the movement of the prey encountered.
-We hypothesize that prey will influence the development of expertise, and predict that faster prey will reduce the gain in expertise.
-Therefore, we then investigate how prey movement influences the development of expertise at the individual level.
+We hypothesize that prey will influence expertise acquisition, and predict that faster prey will reduce the gain in expertise.
+Therefore, we also investigate how prey movement influences expertise acquisition at the individual level.
 If prey speed does not influence hunting success, we predict that the gain in expertise will be similar among individuals.
@@
 The first model was the simplest, with a common global smoothing function and random intercepts for the predator ID.
-In this model, we assume that predators have the same development of expertise, with the model estimating a trend for the average individual (i.e.
+In this model, we assume that predators have the same acquisition of expertise, with the model estimating a trend for the average individual (i.e.
 global smoother).
@@
 
-## Development of expertise at the population level
+## Acquisition of expertise at the population level
 
@@
 
-## Development of expertise at the individual level
+## Acquisition of expertise at the individual level
 
 Prey speed did not influence among individual differences in average hunting success as the posterior distributions of the standard deviations of individual intercepts were almost completely overlapping (Figure 2, median = `r toString(round(sum1[2,]$Estimate, digits = 2))` vs median = `r toString(round(sum2[2,]$Estimate, digits = 2))`).
-However, individuals differed in the development of their hunting expertise (Figure 1B-D).
+However, individuals differed in the acquisition of their hunting expertise (Figure 1B-D).
 We found strong evidence that the speed of the prey mediated among individual differences in the linear relationship between success and experience, as there were substantial differences in the standard deviations of the individual slopes between the two models (Figure 2, median = `r toString(round(sum1[3,]$Estimate, digits = 2))` vs median = `r toString(round(sum2[3,]$Estimate, digits = 2))`).
@@
 ```{r fig1, out.width = "100%", fig.cap = "**Figure 1**.
-Median posterior predictions of the development of predator hunting expertise.
+Median posterior predictions of the acquisition of predator hunting expertise.
 The predators' hunting success (i.e.
@@
 the number of matches played prior to each observation) is on the x axis.
-Panels A and C show the development of expertise for the average individual with the vertical dashed lines on the left representing the lowest predicted values.
+Panels A and C show the acquisition of expertise for the average individual with the vertical dashed lines on the left representing the lowest predicted values.
 For panel A, the right-side vertical dashed line shows the highest predicted success.
 For panel C, the right-side dashed line represents the point on the curve where success was optimized, which we calculated using the finite differences method to obtain the first derivative of the predicted values.
-Panels B and D show among individual differences in the development of expertise, with each curve representing an individual predator.
+Panels B and D show among individual differences in the acquisition of expertise, with each curve representing an individual predator.
 (A-B) GAMM where we control for the prey rank (C-D) GAMM where we control for the prey rank and the speed of the prey group."}
@@
 By capitalizing on a virtual predator-prey system where interactions were directly monitored, we found that a predator-player population in *Dead by Daylight* increased their hunting success with experience.
-Yet, there were important differences among individuals in the development of said expertise, which we found to be in part due to the speed of the prey encountered.
-Hence, our study provides empirical evidence that prey antipredator behaviour can impair the development of hunting expertise.
+Yet, there were important differences among individuals in the acquisition of said expertise, which we found to be in part due to the speed of the prey encountered.
+Hence, our study provides empirical evidence that prey antipredator behaviour can impair the acquisition of hunting expertise.
 
@@
 
-Prey speed also mediated differences among predator players in the development of their expertise, suggesting that individual predator players differed in their capacity to adjust to difficult prey.
+Prey speed also mediated differences among predator players in the acquisition of their expertise, suggesting that individual predator players differed in their capacity to adjust to difficult prey.
 Animals are expected to have limited attention that constrain diet choice and image search formation, and hunting faster prey requires specialized cognitive abilities and coordination that are energetically costly [@dukas.etal2001;@kelley.magurran2011].
@@
 As we did here with predators, future analyses should investigate how antipredator tactics are developped with experience, as it may reveal important insights on the eco-evolutionary dynamics of predator prey interactions.
-Using a video game system, our study demonstrates that prey antipredator behavior can impair the acquisition of hunting expertise.
+Using a video game system, our study demonstrates that prey antipredator behaviour can impair the acquisition of hunting expertise.
 This adds to a growing body of research showing how virtual systems can be used to test hypotheses on ecological interactions [@beauchamp2020;@cere.etal2021;@fraserfranco.etal2022;@lymbery.etal2023;@santostefano.etal2024].

### 2024-09-02 manuscript/manuscript.Rmd
typo in intro
address Clint and Fran comments up to discussion

@@
 cumulated atempts).
-First, test the hypothesis that predator success will increase with experience up to some point at which it will stabilize [@dukas2019].
+First, we test the hypothesis that predator success will increase with experience up to some point at which it will stabilize [@dukas2019].
 However, we expect this pattern to change depending on the movement of the prey encountered.
@@
 
-*DBD* is a survival asymmetric multiplayer online game (i.e.
-a game where the gameplay mechanics differ between two groups) developed by Behaviour Interactive Inc, in which players can play either as a predator or a prey.
+*DBD* is a survival asymmetric (i.e.
+gameplay mechanics differ between two groups) multiplayer online game  developed by Behaviour Interactive Inc., in which players can play either as predators or prey.
 Every match includes only one predator and four prey.
@@
 The resources are in the form of power generators that, once all activated, will enable the prey to escape through one of two exit doors.
-The composition of the predator and prey group in a match is determined by a skill-based matchmaking algorithm.
+A skill-based matchmaking algorithm determines the composition of the predator and prey group in a match.
 A match ends when the predator kills all the prey available (i.e.
@@
 During our study period, the game offered `r toString(format(length(unique(dat$avatar_id))))` predator avatars.
-The virtual environments are composed of fixed and procedurally generated habitat components, such as vegetation, mazes, and buildings.
+The virtual environments comprise fixed and procedurally generated habitat components, such as vegetation, mazes, and buildings.
 Some of these environments are larger than others, with varying structural complexity.
-However, predators display only minimal changes in behaviour and hunting success across these environments, probably due to a game feature enabling them to have visual cues on the resources [@fraserfranco.etal2022].
+However, predators display only minimal changes in behaviour and hunting success across these environments [@fraserfranco.etal2022].
 There were `r toString(format(length(unique(dat$environment_id))))` virtual game environments available for play during the study period.
@@
 
-The videogame company provided data that spanned a period of 6 months of gameplay recorded for every player from `r toString(format(min(dat$date)))` to `r toString(format(max(dat$date)))`.
-We only analyzed matches where players did not know each other (i.e.
-"Online" mode).
-We filtered any matches where players were inactive, such as when mean distances traveled per second (i.e.
+Behaviour Interactive Inc.
+provided data that spanned six months of gameplay recorded for every player from `r toString(format(min(dat$date)))` to `r toString(format(max(dat$date)))`.
+We analyzed only matches where players did not know each other.
+We filtered any matches where players were inactive, such as when mean distances travelled per second (i.e.
 speed) were equal to, or very close to, zero.
 Moreover, we used our knowledge of the game to remove any matches where players were potentially hacking, or not playing the game as intended.
-We then sampled players that played 300 matches or more, and monitored all their matches from the first to a maximum of 500 matches.
 
-Our population consists of `r toString(format(length(unique(dat$predator_id))))` players that played strictly as the predator, with a total record of `r toString(formatC(length(unique(dat$match_id)), big.mark = " "))` matches.
+We sampled players that played 300 matches or more and monitored all their matches from the first to a maximum of 500 matches.
+We recognize that we could have introduced a bias by retaining only those individuals, as they might already be seasoned video game enthusiasts and exhibit expert-level performance in their early matches in *DBD*.
+Thus, we verified that our sample was not biased by comparing a random sample of players that played either 20 to 50 matches, 51 to 100 matches, or 101 to 300 matches during the same timeframe as our sampled population.
+We then took the first 20 matches played by these players, including those from our sampled population, and compared their median hunting success using a Bayesian hierarchical linear model.
+We found that all four groups had similar success as predators (Appendix 1: Table S1 and Figure S1), suggesting an absence of bias due to data sampling.
+
+Our population consists of `r toString(format(length(unique(dat$predator_id))))` players who played as the predator, with a total record of `r toString(formatC(length(unique(dat$match_id)), big.mark = " "))` matches.
 The predator-players' experience varied between `r toString(format(min(dat$total_xp_pred)))` and `r toString(format(max(dat$total_xp_pred)))` matches played.
@@
 We analyzed the mean speed of the prey group encountered by the predator.
-We measured the preys' speed as the mean travel speed of the four individual prey in a match (mean = `r toString(format(round(mean(dat$prey_avg_speed, na.rm = T), digits = 2), nsmall = 2))` ± `r toString(format(round(sd(dat$prey_avg_speed, na.rm = T), digits = 2), nsmall = 2))` m/s).
+We measured the prey's speed as the mean travel speed of the four individual prey in a match (mean = `r toString(format(round(mean(dat$prey_avg_speed, na.rm = T), digits = 2), nsmall = 2))` ± `r toString(format(round(sd(dat$prey_avg_speed, na.rm = T), digits = 2), nsmall = 2))` m/s).
 We defined hunting success as the number of prey consumed during the match (min = `r toString(format(min(dat$hunting_success)))`, max = `r toString(format(max(dat$hunting_success)))`).
@@
 We did not account for matches where predators played as the prey.
-
-We recognize that we could have introduced a bias in our analyses by retaining only individuals who played for at least 300 matches.
-For example, these individuals might be experienced videogame players and could thus already be playing like experts in their first matches in *DBD*.
-To verify that our sample was not biased, we compared a random sample of players that played either 20 to 50 matches, 51 to 100 matches, or 101 to 300 matches during the same timeframe as our sampled population.
-We then took the first 20 matches played by these players, including those from our sampled population, and compared their median hunting success using a Bayesian hierarchical linear model.
-We found that all four groups had similar success as predators (Appendix 1: Table S1 and Figure S1), suggesting an absence of bias due to data sampling.
 
@@
 We parametrized the models following the method of @pedersen.etal2019.
-The first model was the simplest, with a common global smoothing function and random intercepts for the predator ID.
+The first model (I) was the simplest, with a common global smoothing function and random intercepts for the predator ID.
 In this model, we assume that predators have the same acquisition of expertise, with the model estimating a trend for the average individual (i.e.
 global smoother).
-For the second model, we included varying individual smoothers for the predator ID.
+The second model (II) included varying individual smoothers for the predator ID.
 Here, we assume that individual predators share a similar relationship between success and experience, but that this relationship can vary among them (e.g.
@@
 This enabled us to test whether predators differed in the development of their expertise.
-In the third model, we kept the individual smoothers for the predators, but removed the global smoother.
+In the third model (III), we kept the individual smoothers for the predators, but removed the global smoother.
 This model assumes that predators do not share a common relationship between success and experience.
-The fourth and fifth models were reproductions of the second and third models respectively, where we included the prey speed to assess its effect on the relationship between success and experience.
+The fourth (IV) and fifth (V) models were expansions of the third and second models respectively, where we included the prey speed to assess its effect on the relationship between success and experience.
 We included the standardized match duration and prey rank as covariates in all five models.
 
-We computed the five models using a modified version of the beta-binomial distribution implemented in "brms".
+We computed the five models using a modified version of the beta-binomial distribution.
 Hunting success was estimated as the probability of consuming the four prey ($\mu_i$), drawn from a Beta distribution ($Beta(\mu_i, \phi)$) with mean ($\mu \in [0,1]$) and precision ($\phi > 0$) parameters.
 We used a logit link function to estimate $\mu_i$ where $\mu_i = \frac{exp(\eta_i)}{exp(\eta_i) + 1}$ and $\eta_i$ is the linear predictor, while the precision parameter ($\phi$) was estimated with an identity link.
-We used the default number of basis functions (K) in "brms" for the models to estimate the relationship between hunting success and experience.
+We used ten basis functions (K = 10) for the models to estimate the relationship between hunting success and experience.
 We assumed that the random intercepts for the predator ID ($id$) followed a Gaussian distribution with estimated standard deviation ($id \sim N(0,\ I\ \sigma_{id}^{2})$).
+We fitted all models in R (version 4.1.2) using Markov chain Monte Carlo (MCMC) sampling with the package "brms" version 2.16.3 [@burkner2017], an R front-end for the STAN software [@StanDevelopmentTeam2023], and "cmdstanr" version 0.4.0 [@gabry.cesnovar2021] as the back-end for parameter estimation (cmdstan installation version 2.28.2).
+For further details, please consult the GitHub repository of this project (https://github.com/quantitative-ecologist/predator-expertise).
 
@@
 We parametrized the GAMMs to run four MCMC chains.
-We ran 2500 iterations with a thinning set to eight for the additive model with a global smoother only (see Table 1), and 1500 iterations with a thinning set to four for the remaining additive models.
-We set the first 500 iterations of each model as warm ups.
+We ran 2500 iterations with a thinning set to eight for model I (see Table 1), and 1500 iterations with a thinning set to four for the remaining additive models (models II to V).
+We set the first 500 iterations of each model as warm-ups.
 For each model, we obtained 1000 posterior samples per parameter.
@@
 We also performed posterior predictive checks which showed an adequate fit of the models.
-
-### Software and computer specifications
-
-All models were fitted in R (version 4.1.2) using Markov chain Monte Carlo (MCMC) sampling with the package "brms" version 2.16.3 [@burkner2017], an R front-end for the STAN software [@StanDevelopmentTeam2023], and "cmdstanr" version 0.4.0 [@gabry.cesnovar2021] as the back-end for parameter estimation (cmdstan installation version 2.28.2).
-The models were run on Cedar (Operating system: CentOS Linux 7), a computer cluster maintained by the Digital Research Alliance of Canada (https://docs.alliancecan.ca/wiki/Cedar).
-Each required 64GB of RAM with 48 cores to compile within 5 days.
-For details on how to reproduce our analyses, please consult the GitHub repository of this project (https://github.com/quantitative-ecologist/predator-expertise).
 
@@
 
-Out of all five GAMM models, the two that accounted for the prey group's rank and speed were the best at predicting the data, achieving similar expected log pointwise densities (Table 1).
+Out of all five GAMM models, the two that accounted for the prey group's rank and speed were the best at predicting the data with similar predictive accuracies (Table 1).
 Models in which prey effects were not accounted for resulted in no change in hunting success with experience for the average individual (i.e.
-no gain in expertise).
+no gain in expertise, results not shown).
 Accounting for the prey rank resulted in a concave-shaped relationship, with the highest success ranging between ~200 and ~300 matches (Figure 1A).
@@
 We found strong evidence of a negative relationship between hunting success and prey speed (Figure S2).
-In the model where we included the prey's speed, the effect of experience on hunting success for the average individual followed a diminishing returns curve as we predicted, with predators optimizing their success after playing ~300 matches (Figure 1C).
-The curve shows there was a 38% increase in the probability of consuming all prey for the average individual between the first and the ~330$^{th}$ match, where success reached a plateau (Figure 1C).
+As predicted, the effect of experience on hunting success for the average individual followed a diminishing returns curve when controlling for prey speed (model IV), with predators optimizing their success after playing ~300 matches (Figure 1C).
+The curve shows a 38% increase in the probability of consuming all prey for the average individual between the first and the ~330$^{th}$ match, where success reached a plateau (Figure 1C).
 
@@
   , model := c(
-    "predator xp + ID smoothers + prey rank + prey speed",
-    "ID smoothers + prey rank + prey speed",
-    "ID smoothers + prey rank",
-    "predator xp + ID smoothers + prey rank",
-    "predator xp + prey rank"
+    "(V) predator xp + ID smoothers + prey rank + prey speed",
+    "(IV) ID smoothers + prey rank + prey speed",
+    "(III) ID smoothers + prey rank",
+    "(II) predator xp + ID smoothers + prey rank",
+    "(I) predator xp + ID intercepts + prey rank"
   )
@@
 
-Prey speed did not influence among individual differences in average hunting success as the posterior distributions of the standard deviations of individual intercepts were almost completely overlapping (Figure 2, median = `r toString(round(sum1[2,]$Estimate, digits = 2))` vs median = `r toString(round(sum2[2,]$Estimate, digits = 2))`).
+Prey speed did not influence among-individual differences in average hunting success as the posterior distributions of the standard deviations of individual intercepts were almost completely overlapping (Figure 2, median = `r toString(round(sum1[2,]$Estimate, digits = 2))` vs median = `r toString(round(sum2[2,]$Estimate, digits = 2))`).
 However, individuals differed in the acquisition of their hunting expertise (Figure 1B-D).
-We found strong evidence that the speed of the prey mediated among individual differences in the linear relationship between success and experience, as there were substantial differences in the standard deviations of the individual slopes between the two models (Figure 2, median = `r toString(round(sum1[3,]$Estimate, digits = 2))` vs median = `r toString(round(sum2[3,]$Estimate, digits = 2))`).
+We found strong evidence that the speed of the prey mediated among-individual differences in the linear relationship between success and experience, as there were substantial differences in the standard deviations of the individual slopes between the two models (Figure 2, median = `r toString(round(sum1[3,]$Estimate, digits = 2))` vs median = `r toString(round(sum2[3,]$Estimate, digits = 2))`).
 Differences among individuals in the direction of the linear relationship between success and experience were `r toString(round(sum1[3,]$Estimate / sum2[3,]$Estimate, digits = 2))` times lower when we removed the effect of prey speed (i.e.
@@
 The predators' hunting success (i.e.
-the probability of consuming the four prey) is on the y axis, and the predators' cumulative experience (i.e.
-the number of matches played prior to each observation) is on the x axis.
+the probability of consuming the four prey) is on the y-axis, and the predators' cumulative experience (i.e.
+the number of matches played before each observation) is on the x-axis.
 Panels A and C show the acquisition of expertise for the average individual with the vertical dashed lines on the left representing the lowest predicted values.
@@
 Panels B and D show among individual differences in the acquisition of expertise, with each curve representing an individual predator.
-(A-B) GAMM where we control for the prey rank (C-D) GAMM where we control for the prey rank and the speed of the prey group."}
+(A-B) model II where we control for the prey rank (C-D) model V where we control for the prey rank and the speed of the prey group."}
 
@@
 We detected large differences in the standard deviations of the wiggliness (Figure 2, median = `r toString(round(sum1[4,]$Estimate, digits = 2))` vs median = `r toString(round(sum2[4,]$Estimate, digits = 2))`).
-The lower standard deviation for the model where we accounted for prey speed suggests that the form of the relationship between success and experience was more similar among individuals.
+The lower standard deviation for the model where we accounted for prey speed (model V) suggests that the form of the relationship between success and experience was more similar among individuals.
 
@@
 The vertical dashed lines are the medians of the posterior distributions.
-The light coloured distributions are for the model with a shared trend where we did not account for prey speed, and the darker coloured distributions are for the model with a shared trend where we accounted for prey speed (i.e.
-the model with the highest predictive accuracy)"}
+For each panel, the light-coloured distributions (i.e.
+green) are for the model with a shared trend where we did not account for prey speed (model II), and the darker-coloured distributions (i.e.
+purple) are for the model with a shared trend where we accounted for prey speed (model IV)."}
 
@@
 
-Predators usually optimize prey consumption through time by acquiring hunting expertise [@dukas2019;@wooster.etal2023].
-Yet, few studies have empirically tested how prey antipredator strategies mediate the acquisition of hunting expertise due to the challenges of investigating direct predator-prey interactions in the wild.
-By capitalizing on a virtual predator-prey system where interactions were directly monitored, we found that a predator-player population in *Dead by Daylight* increased their hunting success with experience.
-Yet, there were important differences among individuals in the acquisition of said expertise, which we found to be in part due to the speed of the prey encountered.
-Hence, our study provides empirical evidence that prey antipredator behaviour can impair the acquisition of hunting expertise.
+Using a virtual predator-prey system where we monitored hunting success across experience, we provide rare empirical support for the hypothesis that prey antipredator behaviour can impair the acquisition of hunting expertise.
+Our observations show that the predator population increased their hunting success with experience.
+We found, however, important differences among individuals in expertise acquisition that were related to the speed of the prey encountered.
 
-Our results suggest that the acquisition of hunting expertise was honed through extensive practice.
+Our results suggest that predators hone their hunting expertise through extensive practice.
 The predator population displayed an asymptotic relationship between experience and success, wherein initial gains in success were significant but gradually stabilized as experience accumulated.
 These observations are consistent with empirical studies of expertise in both humans and nonhuman animals [reviewed in @dukas2019].
-However, the prey's speed was important in mediating this pattern at the population level.
-Encountering faster prey led to more difficult encounters for predators.
-We previously showed in *DBD* that faster movement is an effective strategy used by the prey to avoid predation [@fraserfranco.etal2022], and other studies have found that as well [@walker.etal2005;@kelley.magurran2011;@martin.etal2022].
-This could explain why the relationship was concave in the model where we did not account for prey speed, because the prey were probably increasing their speed as they gained experience.
+However, prey speed played a crucial role in shaping this pattern at the population level as encounters with faster prey resulted in lower hunting success (Figure S2).
+We previously showed in *DBD* that faster movement is effective for prey to evade predation [@fraserfranco.etal2022], a finding consistent with studies using other systems [@walker.etal2005;@kelley.magurran2011;@martin.etal2022].
+We suspect that experienced prey may increasingly rely on this strategy, which could explain why the relationship between hunting success and experience was concave when we did not control for prey speed (model II, Figure 1A).
+
 Because the matchmaking algorithm pairs players based on their skill, experienced predators were probably encountering prey moving at higher speeds.
+
 Thus, our results suggest that predators can gain expertise and maintain success when they encounter prey that move at speeds lower than or closer to the population-average.
+
+Additionally, since the matchmaking algorithm pairs players based on skill, experienced predators were likely facing prey with increasingly higher movement speeds.
 

### 2024-09-03 manuscript/manuscript.Rmd
integrated all of Clint and Fran's comments

@@
 Through these processes, expert predators should have greater knowledge, better energy management, and acute motor skills to increase their chances of locating and capturing prey [@dukas2019].
-Therefore, differences among predators in their capacity to acquire expertise throughout their lifetime may underlie differences in hunting success, with potential consequences for predator-prey dynamics [@wooster.etal2023].
+Therefore, differences among predators in their capacity to acquire expertise throughout their lifetime may underlie differences in hunting success.
+This is an important topic of study since predator cognition can maintain behavioural variation, shape prey phenotypic composition, and destabilize predator-prey systems [@kondoh2010;@skelhornandcandy2016;@kikuchi.etal2023].
 
 Prey are known to use evasive tactics such as rapid escapes to avoid being captured [@walker.etal2005;@kelley.magurran2011;@herbert-read.etal2017], driving predators to develop the motor skills required to capture such prey succesfully.
-However, predators may be limited in their capacity to develop the necessary attributes (e.g.
-physical, physiological, neurological) for fast-paced hunting, which may impair the acquisition of hunting expertise.
-There is limited empirical evidence showing links between prey antipredator strategies and the acquisition of expertise in human and nonhuman predators.
-Experimental studies show that certain camouflage strategies of prey can impair expertise acquisition in humans and birds [@martin.etal2012;@troscianko.etal2013].
+Yet, predators may be limited in their capacity to develop the necessary attributes (e.g.
+physical, physiological, neurological) for fast-paced hunting, which may impair their acquisition of hunting expertise.
+To our knowledge, there is no empirical evidence showing links between prey antipredator behaviour and the acquisition of expertise in human and nonhuman predators.
+However, experimental studies show that certain camouflage strategies of prey can impair expertise acquisition in humans and birds [@martin.etal2012;@troscianko.etal2013].
 For instance, @troscianko.etal2018 found that disruptive colouration interfered with the formation of search images in human subjects, hindering improvement in detection times after repeated attempts.
 Given these examples, antipredator behaviour may also hold the potential to hinder the acquisition of predator expertise.
-This is an important topic of study since predator cognition can maintain behavioural variation, shape prey phenotypic composition, and destabilize predator-prey systems [@kondoh2010;skelhornandcandy2016;@kikuchi.etal2023].
+This represents a significant gap in our understanding of predator-prey interactions, as individual differences in predator expertise may structure food webs via changes in foraging behaviour and frequency- or trait-dependent prey selection [@allen.etal1988;@ishii.shimada2012;@wooster.etal2023].
 
@@
 
-*DBD* is a survival asymmetric (i.e.
+*Dead by Daylight* is a survival asymmetric (i.e.
 gameplay mechanics differ between two groups) multiplayer online game  developed by Behaviour Interactive Inc., in which players can play either as predators or prey.
@@
 We sampled players that played 300 matches or more and monitored all their matches from the first to a maximum of 500 matches.
-We recognize that we could have introduced a bias by retaining only those individuals, as they might already be seasoned video game enthusiasts and exhibit expert-level performance in their early matches in *DBD*.
+We recognize that we could have introduced a bias by retaining only those individuals, as they might already be seasoned video game enthusiasts and exhibit expert-level performance in their early matches in *Dead by Daylight*.
 Thus, we verified that our sample was not biased by comparing a random sample of players that played either 20 to 50 matches, 51 to 100 matches, or 101 to 300 matches during the same timeframe as our sampled population.
@@
 the predator character chosen with its specific skill-gameplay mechanics), the game environment, the predator-player's experience, the mean speed of the groups of prey that the predator player encountered, and the mean rank of the prey encountered (a proxy for prey skill).
-The ranking system in *DBD* was implemented by the company to pair players in a match based on their skill (https://deadbydaylight.fandom.com/wiki/Rank), and failing to account for it would prevent us from detecting a change in the predator's foraging success with experience.
+The ranking system in *Dead by Daylight* was implemented by the company to pair players in a match based on their skill (https://deadbydaylight.fandom.com/wiki/Rank), and failing to account for it would prevent us from detecting a change in the predator's foraging success with experience.
 
@@
 
+### Hypothesis testing
+
+We tested the hypothesis that antipredator behaviour impairs the acquisition of expertise at the population level by visually comparing the global trends of model II (not controlling for prey speed) and model V (controlling for prey speed) relating hunting success to cumulated experience (Figure 1).
+At the individual level, we tested our hypothesis that antipredator behaviour generates differences among predators in expertise acquisition by comparing the standard deviation of individual-level parameters relating experience to hunting success.
+Using model II and model V, we compared the standard deviations of 1) the random intercepts (i.e.
+mean differences in hunting success), 2) the random slopes (i.e.
+linear component relating hunting success with experience), and 3) the curve wiggliness (i.e.
+nonlinear component relating hunting success with experience).
+
 # RESULTS {.unnumbered}
@@
 Out of all five GAMM models, the two that accounted for the prey group's rank and speed were the best at predicting the data with similar predictive accuracies (Table 1).
-Models in which prey effects were not accounted for resulted in no change in hunting success with experience for the average individual (i.e.
+Models in which prey rank was not accounted for resulted in no change in hunting success with experience for the average individual (i.e.
 no gain in expertise, results not shown).
-Accounting for the prey rank resulted in a concave-shaped relationship, with the highest success ranging between ~200 and ~300 matches (Figure 1A).
+Accounting for the prey rank resulted in a concave-shaped relationship (model II), with the highest success ranging between ~200 and ~300 matches (Figure 1A).
 
 We found strong evidence of a negative relationship between hunting success and prey speed (Figure S2).
-As predicted, the effect of experience on hunting success for the average individual followed a diminishing returns curve when controlling for prey speed (model IV), with predators optimizing their success after playing ~300 matches (Figure 1C).
+As predicted, the effect of experience on hunting success for the average individual followed a diminishing returns curve when controlling for prey speed (model V), with predators optimizing their success after playing ~300 matches (Figure 1C).
 The curve shows a 38% increase in the probability of consuming all prey for the average individual between the first and the ~330$^{th}$ match, where success reached a plateau (Figure 1C).
@@
 However, prey speed played a crucial role in shaping this pattern at the population level as encounters with faster prey resulted in lower hunting success (Figure S2).
-We previously showed in *DBD* that faster movement is effective for prey to evade predation [@fraserfranco.etal2022], a finding consistent with studies using other systems [@walker.etal2005;@kelley.magurran2011;@martin.etal2022].
+We previously showed in *Dead by Daylight* that faster movement is effective for prey to evade predation [@fraserfranco.etal2022], a finding consistent with studies using other systems [@walker.etal2005;@kelley.magurran2011;@martin.etal2022].
 We suspect that experienced prey may increasingly rely on this strategy, which could explain why the relationship between hunting success and experience was concave when we did not control for prey speed (model II, Figure 1A).
-
-Because the matchmaking algorithm pairs players based on their skill, experienced predators were probably encountering prey moving at higher speeds.
-
 Thus, our results suggest that predators can gain expertise and maintain success when they encounter prey that move at speeds lower than or closer to the population-average.
 
-Additionally, since the matchmaking algorithm pairs players based on skill, experienced predators were likely facing prey with increasingly higher movement speeds.
-
-Prey speed also mediated differences among predator players in the acquisition of their expertise, suggesting that individual predator players differed in their capacity to adjust to difficult prey.
-Animals are expected to have limited attention that constrain diet choice and image search formation, and hunting faster prey requires specialized cognitive abilities and coordination that are energetically costly [@dukas.etal2001;@kelley.magurran2011].
-Thus, predators that couldn't develop counter-strategies to detect or chase faster prey were likely at a disadvantage.
+Prey speed also mediated differences among predator players in the acquisition of expertise, suggesting that individual predators varied in their capacity to adjust to challenging prey.
+Animals are expected to have limited attention spans, which restricts diet choice and the formation of search images [@dukars.etal2001].
+Hunting faster prey demands specialized cognitive abilities and coordination that are energetically costly [@kelley.magurran2011].
+Thus, predators that failed to develop counter-strategies for detecting or chasing faster prey were likely at a disadvantage.
 Parallel observations have been outlined in studies of prey camouflage strategies.
@@
 Human subjects exposed to a restricted set of strategies were also less efficient compared to those exposed to a variety of strategies.
-Therefore, our observations suggest that prey antipredator behaviour can also impair predator expertise acquisition, with potential implications for predator-prey dynamics.
+Therefore, our observations show that prey antipredator behaviour can also impair predator expertise acquisition.
 
-Despite adjusting for the prey's speed, discernible variations in expertise acquisition among predator players persisted.
-A potential explanation is that the predators' hunting tactic may drive the antripredator tactic of the prey.
-We know from a previous study that predators specialize as cursorial or ambush hunters [@fraserfranco.etal2022].
-Therefore, predators using a cursorial tactic may drive prey to move faster at the cost of having a slower acquisition of expertise.
-
-Despite adjusting for the prey's speed, noticeable differences in the predators' expertise acquisition still emerged.
+Despite adjusting for the prey's speed (i.e.
+model V), noticeable differences in expertise acquisition among predators persited.
 One possible explanation is that the predators' hunting tactics may indirectly shape their own expertise through changes in prey behaviour.
-Previous research indicates that predators tend to specialize as either cursorial or ambush hunters in *DBD* [@fraserfranco.etal2022].
-Consequently, predators employing a cursorial tactic may push prey to move faster, which, in turn, may hinder their own expertise acquisition due to the increased difficulty of hunting such prey.
+Predators tend to specialize as cursorial or ambush hunters in *Dead by Daylight* [@fraserfranco.etal2022].
+Consequently, those employing a cursorial tactic may push prey to move faster, which, in turn, may hinder their own expertise acquisition due to the increased difficulty of hunting such prey.
 An alternative explanation is that longer time intervals between hunting events are hypothesized to hinder or delay the acquisition of expertise because individuals may forget information when delays are longer [@endler1991;@wright.etal2022].
-In *DBD*, a predator that played 300 matches in the span of six months might forget more critical information related to prey detection or escape patterns than one that played 300 matches in the span of six days.
-Investigating the impact of such time lags in future analyses may reveal important insights on the outcome of predator-prey interactions.
-Another potential reason for the persistent differences among individuals is that neither the predator nor prey players' lives are at stake in *DBD*.
+For example, a predator that played 300 matches in six months might forget critical information related to prey detection or escape patterns compared to one that played 300 matches in six days.
+Investigating the impact of such time lags in future analyses may reveal important insights into the outcome of predator-prey interactions.
+Another potential reason for the persistent differences among individuals is that neither the predator nor prey players' lives are at stake in the game.
 As a result, emerging patterns may be driven more by the players' motivation to win rather than "true" survival.
@@
 
-We found support of our hypothesis that prey antipredator behaviour was driving individual differences in expertise acquisition in a human predator population in the game *Dead by Daylight*.
-As we did here with predators, future analyses should investigate how antipredator tactics are developped with experience, as it may reveal important insights on the eco-evolutionary dynamics of predator prey interactions.
-Using a video game system, our study demonstrates that prey antipredator behaviour can impair the acquisition of hunting expertise.
-This adds to a growing body of research showing how virtual systems can be used to test hypotheses on ecological interactions [@beauchamp2020;@cere.etal2021;@fraserfranco.etal2022;@lymbery.etal2023;@santostefano.etal2024].
+We found support of our hypothesis that prey antipredator behaviour drives individual differences in expertise acquisition in a human predator population in the game *Dead by Daylight*.
+Future analyses should investigate how antipredator tactics developp with experience, as it may reveal important insights on the eco-evolutionary dynamics of predator-prey interactions.
+Our study demonstrating that prey antipredator behaviour can impair the acquisition of hunting expertise adds to a growing body of research showing how virtual systems can be used to test hypotheses on ecological interactions [@beauchamp2020;@cere.etal2021;@fraserfranco.etal2022;@lymbery.etal2023;@santostefano.etal2024].
 We therefore hope that our study will inspire more collaborations between scientists and the videogame industry to tackle fundamental questions in ecology.

### 2024-09-03 manuscript/manuscript.Rmd
minor edit to hypothesis testing section

@@
 We tested the hypothesis that antipredator behaviour impairs the acquisition of expertise at the population level by visually comparing the global trends of model II (not controlling for prey speed) and model V (controlling for prey speed) relating hunting success to cumulated experience (Figure 1).
-At the individual level, we tested our hypothesis that antipredator behaviour generates differences among predators in expertise acquisition by comparing the standard deviation of individual-level parameters relating experience to hunting success.
-Using model II and model V, we compared the standard deviations of 1) the random intercepts (i.e.
+At the individual level, we tested our hypothesis that antipredator behaviour generates differences among predators in expertise acquisition by comparing individual-level parameters of model II and model V.
+Specifically, we compared the standard deviations of 1) the random intercepts (i.e.
 mean differences in hunting success), 2) the random slopes (i.e.

### 2024-09-03 manuscript/manuscript.Rmd
MS up to date (corrected ref typo)
minor edit to discussion

@@
 Prey speed also mediated differences among predator players in the acquisition of expertise, suggesting that individual predators varied in their capacity to adjust to challenging prey.
-Animals are expected to have limited attention spans, which restricts diet choice and the formation of search images [@dukars.etal2001].
+Animals are expected to have limited attention spans, which restricts diet choice and the formation of search images [@dukas.etal2001].
 Hunting faster prey demands specialized cognitive abilities and coordination that are energetically costly [@kelley.magurran2011].
@@
 As a result, emerging patterns may be driven more by the players' motivation to win rather than "true" survival.
-For example, some players could experiment with the game out of boredom, which could shape how expertise is honed and impede ecologically realistic interpretations of our data.
+For example, some players could experiment with the game out of boredom, which could also contribute in shaping how expertise is honed in this particular system.
 

### 2024-09-07 code/code_models/GAMM-I.R
GAMM naming conventions to reflect paper

@@ -1,6 +1,7 @@
 # ==========================================================================
 
-#           GAMM model GS - group-level smoother - hunting success
+#            GAMM model G - single common smoother - Hunting success
+#                             Model I in article
 
 # ==========================================================================
 
@@ -32,7 +33,7 @@ library(parallel)
 folder <- file.path("/home", "maxime11", "projects", "def-monti", 
                     "maxime11", "phd_project", "data")
 
-# Load data
+# Load data on compute canada
 data <- fread(
   file.path(folder, "FraserFrancoetalXXXX-data.csv"),
   select = c("predator_id",
@@ -64,7 +65,6 @@ data[, c(
     "Zprey_avg_rank"
     ) := lapply(.SD, standardize),
        .SDcols = c(3:6)]
-
 
 # ==========================================================================
 # ==========================================================================
@@ -113,9 +113,8 @@ stanvars <- stanvar(scode = stan_funs, block = "functions")
 model_formula <- brmsformula(
   hunting_success | vint(4) ~
     s(Zcumul_xp) +
-    s(Zcumul_xp, predator_id, bs = "fs") +
+    s(predator_id, bs = "re") +
     Zgame_duration +
-    Zprey_speed +
     Zprey_avg_rank
 )
 
@@ -128,9 +127,6 @@ priors <- c(
   set_prior("normal(1, 0.5)",
             class = "b",
             coef = "Zgame_duration"),
-  set_prior("normal(-1, 0.5)",
-            class = "b",
-            coef = "Zprey_speed"),
   set_prior("normal(0, 1)",
             class = "b",
             coef = "Zprey_avg_rank"),
@@ -146,7 +142,9 @@ priors <- c(
   # priors on phi
   set_prior("normal(2, 0.5)",
             class = "phi")
-)
+            )
+
+
 
 # ==========================================================================
 # ==========================================================================
@@ -159,23 +157,25 @@ priors <- c(
 # 3. Run the model
 # ==========================================================================
 
-model_gs <- brm(formula = model_formula,
-                family = beta_binomial2,
-                warmup = 500,
-                iter = 1500,
-                thin = 4,
-                chains = 4,
-                threads = threading(12),
-                backend = "cmdstanr",
-                inits = "0",
-                seed = 123,
-                prior = priors,
-                sample_prior = TRUE,
-                control = list(adapt_delta = 0.99),
-                data = data,
[…]

### 2024-09-07 code/code_models/GAMM-II.R

@@ -1,7 +1,7 @@
 # ==========================================================================
 
-#         GAMM model S - individual curves without common smoother
-#                              hunting success
+#           GAMM model GS - group-level smoother - hunting success
+#                        Model II in article
 
 # ==========================================================================
 
@@ -58,13 +58,16 @@ standardize <- function(x) {
   (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)
 }
 
-data[, c(
+data[
+  , c(
     "Zprey_speed",
     "Zgame_duration",
     "Zcumul_xp",
     "Zprey_avg_rank"
-    ) := lapply(.SD, standardize),
-       .SDcols = c(3:6)]
+  ) := lapply(.SD, standardize),
+  .SDcols = c(3:6)
+]
+
 
 # ==========================================================================
 # ==========================================================================
@@ -112,9 +115,9 @@ stanvars <- stanvar(scode = stan_funs, block = "functions")
 
 model_formula <- brmsformula(
   hunting_success | vint(4) ~
-    s(Zcumul_xp, predator_id, bs = "fs", m = 2) +
+    s(Zcumul_xp) +
+    s(Zcumul_xp, predator_id, bs = "fs") +
     Zgame_duration +
-    Zprey_speed +
     Zprey_avg_rank
 )
 
@@ -130,9 +133,9 @@ priors <- c(
   set_prior("normal(0, 1)",
             class = "b",
             coef = "Zprey_avg_rank"),
-  set_prior("normal(-1, 0.5)",
+  set_prior("normal(0, 2)",
             class = "b",
-            coef = "Zprey_speed"),
+            coef = "sZcumul_xp_1"),
   # prior on the intercept
   set_prior("normal(0, 0.5)",
             class = "Intercept"),
@@ -155,23 +158,25 @@ priors <- c(
 # 3. Run the model
 # ==========================================================================
 
-model_s <- brm(formula = model_formula,
-                family = beta_binomial2,
-                warmup = 500,
-                iter = 1500,
-                thin = 4,
-                chains = 4,
-                threads = threading(12),
-                backend = "cmdstanr",
-                inits = "0",
-                seed = 123,
-                prior = priors,
-                sample_prior = TRUE,
-                control = list(adapt_delta = 0.99),
-                data = data,
-                stanvars = stanvars)
+fit <- brm(
+  formula = model_formula,
+  family = beta_binomial2,
[…]

### 2024-09-07 code/code_models/GAMM-III.R

@@ -1,6 +1,8 @@
 # ==========================================================================
 
-#            GAMM model G - single common smoother - Hunting success
+#         GAMM model S - individual curves without common smoother
+#                              hunting success
+#                           Model III in article
 
 # ==========================================================================
 
@@ -32,7 +34,7 @@ library(parallel)
 folder <- file.path("/home", "maxime11", "projects", "def-monti", 
                     "maxime11", "phd_project", "data")
 
-# Load data on compute canada
+# Load data
 data <- fread(
   file.path(folder, "FraserFrancoetalXXXX-data.csv"),
   select = c("predator_id",
@@ -57,13 +59,15 @@ standardize <- function(x) {
   (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)
 }
 
-data[, c(
+data[
+  , c(
     "Zprey_speed",
     "Zgame_duration",
     "Zcumul_xp",
     "Zprey_avg_rank"
-    ) := lapply(.SD, standardize),
-       .SDcols = c(3:6)]
+  ) := lapply(.SD, standardize),
+  .SDcols = c(3:6)
+]
 
 # ==========================================================================
 # ==========================================================================
@@ -110,11 +114,10 @@ stanvars <- stanvar(scode = stan_funs, block = "functions")
 # Model formula ------------------------------------------------------------
 
 model_formula <- brmsformula(
-  hunting_success | vint(4) ~
-    s(Zcumul_xp) +
-    s(predator_id, bs = "re") +
-    Zgame_duration +
-    Zprey_avg_rank
+    hunting_success | vint(4) ~
+        s(Zcumul_xp, predator_id, bs = "fs", m = 2) +
+        Zgame_duration +
+        Zprey_avg_rank
 )
 
 
@@ -129,9 +132,6 @@ priors <- c(
   set_prior("normal(0, 1)",
             class = "b",
             coef = "Zprey_avg_rank"),
-  set_prior("normal(0, 2)",
-            class = "b",
-            coef = "sZcumul_xp_1"),
   # prior on the intercept
   set_prior("normal(0, 0.5)",
             class = "Intercept"),
@@ -141,9 +141,7 @@ priors <- c(
   # priors on phi
   set_prior("normal(2, 0.5)",
             class = "phi")
-            )
-
-
+)
 
 # ==========================================================================
 # ==========================================================================
@@ -156,23 +154,25 @@ priors <- c(
 # 3. Run the model
 # ==========================================================================
 
[…]

### 2024-09-07 code/code_models/GAMM-IV.R

@@ -2,6 +2,7 @@
 
 #         GAMM model S - individual curves without common smoother
 #                              hunting success
+#                            Model IV in article
 
 # ==========================================================================
 
@@ -58,13 +59,15 @@ standardize <- function(x) {
   (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)
 }
 
-data[, c(
+data[
+  , c(
     "Zprey_speed",
     "Zgame_duration",
     "Zcumul_xp",
     "Zprey_avg_rank"
-    ) := lapply(.SD, standardize),
-       .SDcols = c(3:6)]
+  ) := lapply(.SD, standardize),
+  .SDcols = c(3:6)
+]
 
 # ==========================================================================
 # ==========================================================================
@@ -111,10 +114,11 @@ stanvars <- stanvar(scode = stan_funs, block = "functions")
 # Model formula ------------------------------------------------------------
 
 model_formula <- brmsformula(
-    hunting_success | vint(4) ~
-        s(Zcumul_xp, predator_id, bs = "fs", m = 2) +
-        Zgame_duration +
-        Zprey_avg_rank
+  hunting_success | vint(4) ~
+    s(Zcumul_xp, predator_id, bs = "fs", m = 2) +
+    Zgame_duration +
+    Zprey_speed +
+    Zprey_avg_rank
 )
 
 
@@ -129,6 +133,9 @@ priors <- c(
   set_prior("normal(0, 1)",
             class = "b",
             coef = "Zprey_avg_rank"),
+  set_prior("normal(-1, 0.5)",
+            class = "b",
+            coef = "Zprey_speed"),
   # prior on the intercept
   set_prior("normal(0, 0.5)",
             class = "Intercept"),
@@ -151,23 +158,25 @@ priors <- c(
 # 3. Run the model
 # ==========================================================================
 
-model_s <- brm(formula = model_formula,
-                family = beta_binomial2,
-                warmup = 500,
-                iter = 1500,
-                thin = 4,
-                chains = 4,
-                threads = threading(12),
-                backend = "cmdstanr",
-                inits = "0",
-                seed = 123,
-                prior = priors,
-                sample_prior = TRUE,
-                control = list(adapt_delta = 0.99),
-                data = data,
-                stanvars = stanvars)
+fit <- brm(
+  formula = model_formula,
+  family = beta_binomial2,
+  warmup = 500,
+  iter = 1500,
+  thin = 4,
[…]

### 2024-09-07 code/code_models/GAMM-V.R

@@ -1,6 +1,7 @@
 # ==========================================================================
 
 #           GAMM model GS - group-level smoother - hunting success
+#                           Model V in article
 
 # ==========================================================================
 
@@ -128,6 +129,9 @@ priors <- c(
   set_prior("normal(1, 0.5)",
             class = "b",
             coef = "Zgame_duration"),
+  set_prior("normal(-1, 0.5)",
+            class = "b",
+            coef = "Zprey_speed"),
   set_prior("normal(0, 1)",
             class = "b",
             coef = "Zprey_avg_rank"),
@@ -156,23 +160,25 @@ priors <- c(
 # 3. Run the model
 # ==========================================================================
 
-model_gs <- brm(formula = model_formula,
-                family = beta_binomial2,
-                warmup = 500,
-                iter = 1500,
-                thin = 4,
-                chains = 4,
-                threads = threading(12),
-                backend = "cmdstanr",
-                inits = "0",
-                seed = 123,
-                prior = priors,
-                sample_prior = TRUE,
-                control = list(adapt_delta = 0.99),
-                data = data,
-                stanvars = stanvars)
+fit <- brm(
+  formula = model_formula,
+  family = beta_binomial2,
+  warmup = 500,
+  iter = 1500,
+  thin = 4,
+  chains = 4,
+  threads = threading(12),
+  backend = "cmdstanr",
+  inits = "0",
+  seed = 123,
+  prior = priors,
+  sample_prior = TRUE,
+  control = list(adapt_delta = 0.99),
+  data = data,
+  stanvars = stanvars
+)
 
-saveRDS(model_gs, file = "A2_GAMM-rank.rds")
+saveRDS(fit, file = "GAMM-V.rds")
 
 # ==========================================================================
 # ==========================================================================

### 2024-09-07 manuscript/manuscript.Rmd
transfered folder to new OS
update files based on GAMM naming changes

@@
 # Import top 2 model fits
-fit1 <- readRDS(here::here("outputs", "outputs_models", "A2_GAMM-rank.rds"))
-fit2 <- readRDS(here::here("outputs", "outputs_models", "A2_GAMM-speed-rank.rds"))
+fit1 <- readRDS(here::here("outputs", "outputs_models", "GAMM-II.rds"))
+fit2 <- readRDS(here::here("outputs", "outputs_models", "GAMM-V.rds"))
 

### 2024-09-07 code/code_figures/figure1.R

@@ -25,8 +25,8 @@ library(viridis)
 
 path <- file.path(getwd(), "outputs", "outputs_models")
 
-modA2 <- readRDS(file.path(path, "A2_GAMM-rank.rds"))
-modA2_prey <- readRDS(file.path(path, "A2_GAMM-speed-rank.rds"))
+mod1 <- readRDS(file.path(path, "GAMM-II.rds"))
+mod2 <- readRDS(file.path(path, "GAMM-V.rds"))
 
 
 
@@ -50,7 +50,7 @@ data[, predator_id := as.factor(predator_id)]
 
 # Post processing preparations for custom family ---------------------------
 
-expose_functions(modA2, vectorize = TRUE)
+expose_functions(mod1, vectorize = TRUE)
 
 # Define the log likelihood function
 log_lik_beta_binomial2 <- function(i, prep) {
@@ -92,39 +92,39 @@ posterior_epred_beta_binomial2 <- function(prep) {
 
 # Prepare the plots --------------------------------------------------------
 
-# Group-level smooths model A2
-tabA2_a <- conditional_effects(
- modA2, method = "fitted",
- effects = "Zcumul_xp:predator_id",
- robust = TRUE, re_formula = NULL
+# Group-level smooths model 1
+tab1_a <- conditional_effects(
+  mod1, method = "fitted",
+  effects = "Zcumul_xp:predator_id",
+  robust = TRUE, re_formula = NULL
 )
-tabA2_a <- data.table(tabA2_a[[1]])
+tab1_a <- data.table(tab1_a[[1]])
 
-# Cumulative XP global trend model A2
-tabA2_b <- conditional_effects(
- modA2, method = "fitted",
- robust = TRUE, re_formula = NULL,
- effects = "Zcumul_xp",
- conditions = data.frame(predator_id = NA)
+# Cumulative XP global trend model 1
+tab1_b <- conditional_effects(
+  mod1, method = "fitted",
+  robust = TRUE, re_formula = NULL,
+  effects = "Zcumul_xp",
+  conditions = data.frame(predator_id = NA)
 )
-tabA2_b <- data.table(tabA2_b[[1]])
+tab1_b <- data.table(tab1_b[[1]])
 
-# Group-level smooths model A2prey
-tabA2_prey_a <- conditional_effects(
- modA2_prey, method = "fitted",
[…]

### 2024-09-07 code/code_figures/figure2.R

@@ -25,8 +25,8 @@ library(ggplot2)
 
 path <- file.path(getwd(), "outputs", "outputs_models")
 
-fit1 <- readRDS(file.path(path, "A2_GAMM-rank.rds"))
-fit2 <- readRDS(file.path(path, "A2_GAMM-speed-rank.rds"))
+fit1 <- readRDS(file.path(path, "GAMM-II.rds"))
+fit2 <- readRDS(file.path(path, "GAMM-IV.rds"))
 
 # ==========================================================================
 # ==========================================================================

### 2024-09-07 code/code_figures/figureS2.R

@@ -25,7 +25,7 @@ library(viridis)
 # Import model -------------------------------------------------------------
 
 path <- file.path(getwd(), "outputs", "outputs_models")
-fit <- readRDS(file.path(path, "A2_GAMM-speed-rank.rds"))
+fit <- readRDS(file.path(path, "GAMM-V.rds"))
 
 
 

### 2024-10-25 manuscript/manuscript.Rmd
manuscript up to date ready for preprint

@@
 
-Antipredation responses by prey are hypothesized to drive differences among predators in their hunting success by disrupting the acquisition of hunting expertise [@wooster.etal2023].
+Predation is a fundamental biological process that acts as an agent of evolutionary change and shapes community structure [@abrams2000;@wirsing.etal2021].
+By consuming prey, predators regulate prey populations, limit the spread of diseases, and mediate energy flow across trophic levels [@Hairston.1993;@ripple.etal2014;@schmitz2017].
+These processes emerge from interactions between individual predators and their prey, lying at the core of key concepts such as predation risk and prey selection[@pettorelli.etal2015;@labarge.etal2024].
+Predators also respond to interactions with their prey, increasing their effectiveness by adapting their strategy and choosing suitable habitats to hunt [@quevedo.etal2009;@schmitz2017].
+Central theory suggests that these responses by predators may result from acquiring expertise through practicing and learning the proper skills to locate, select, and capture their prey [@woo.etal2008;@wooster.etal2023;@labarge.etal2024].
+Hence, the role of expertise remains a missing but essential link to assess of how individual characteristics shape prey consumption during predator-prey interactions.
+
 Expertise can be defined as the characteristics, skills, and knowledge that provide individuals with the ability to outperform novices on complex tasks [@dukas2017].
-In the context of hunting, predators need to acquire expertise by practising and learning the proper skills to successfully locate, select, and capture their prey, a process deemed essential for many predators to reach adulthood and survive [@woo.etal2008;@wooster.etal2023].
 Empirical studies on human and non-human hunters show that individuals optimize foraging efficiency (e.g.
@@
 
-Prey are known to use evasive tactics such as rapid escapes to avoid being captured [@walker.etal2005;@kelley.magurran2011;@herbert-read.etal2017], driving predators to develop the motor skills required to capture such prey succesfully.
-Yet, predators may be limited in their capacity to develop the necessary attributes (e.g.
-physical, physiological, neurological) for fast-paced hunting, which may impair their acquisition of hunting expertise.
+Prey use antipredator tactics such as camouflage to avoid detection and rapid escapes to evade capture [@walker.etal2005;@kelley.magurran2011;@herbert-read.etal2017].
+These strategies are hypothesized to drive differences in prey consumption among predators by disrupting the acquisition of hunting expertise [@wooster.etal2023].
+Experimental studies have shown that certain camouflage tactics of prey can impair expertise acquisition in humans and birds [@martin.etal2012;@troscianko.etal2013].
+For example, @troscianko.etal2018 found that disruptive colouration interfered with human subjects' ability to form search images, hindering improvements in detection times over repeated attempts.
+Thus, antipredator behaviour may also hold the potential to hinder the acquisition of predator expertise.
+For instance, some predators may be limited in their capacity to develop the necessary attributes (e.g.
+physical, physiological, neurological) for fast-paced hunting, which could impair their acquisition of hunting expertise.
 To our knowledge, there is no empirical evidence showing links between prey antipredator behaviour and the acquisition of expertise in human and nonhuman predators.
-However, experimental studies show that certain camouflage strategies of prey can impair expertise acquisition in humans and birds [@martin.etal2012;@troscianko.etal2013].
-For instance, @troscianko.etal2018 found that disruptive colouration interfered with the formation of search images in human subjects, hindering improvement in detection times after repeated attempts.
-Given these examples, antipredator behaviour may also hold the potential to hinder the acquisition of predator expertise.
-This represents a significant gap in our understanding of predator-prey interactions, as individual differences in predator expertise may structure food webs via changes in foraging behaviour and frequency- or trait-dependent prey selection [@allen.etal1988;@ishii.shimada2012;@wooster.etal2023].
+This represents a significant gap in our understanding of predator-prey interactions, as individual variation in predator expertise could lead to shifts in foraging behaviour and define whether predators select prey based on their frequency or their traits [@allen.etal1988;@ishii.shimada2012;@wooster.etal2023].
 
@@
 The prey can increase their chances of survival by cooperating and moving fast to escape the predator [@cere.etal2021;@fraserfranco.etal2022;@santostefano.etal2024].
-The game also elicits natural reactions in players such as freezing when predation is imminent (personal observations), corroborating another virtual ecological study showing that predation drives individual variation in risk perception [@beauchamp2020].
+The game also elicits natural reactions in players such as freezing when predation is imminent (M.F.F., personal observations), corroborating another virtual ecological study showing that predation drives individual variation in risk perception [@beauchamp2020].
 These observations outline how ecological phenomena can emerge from human interactions in virtual systems with fixed rules [@brosnan.etal2017;@kasumovic.etal2017].
@@
 
-Using a virtual predator-prey system where we monitored hunting success across experience, we provide rare empirical support for the hypothesis that prey antipredator behaviour can impair the acquisition of hunting expertise.
-Our observations show that the predator population increased their hunting success with experience.
+Using a virtual predator-prey system where we monitored predator hunting success across experience, we provide rare empirical support for the hypothesis that prey antipredator behaviour can impair the acquisition of hunting expertise.
+We show that the predator population increased its hunting success with experience.
 We found, however, important differences among individuals in expertise acquisition that were related to the speed of the prey encountered.
@@
 However, prey speed played a crucial role in shaping this pattern at the population level as encounters with faster prey resulted in lower hunting success (Figure S2).
-We previously showed in *Dead by Daylight* that faster movement is effective for prey to evade predation [@fraserfranco.etal2022], a finding consistent with studies using other systems [@walker.etal2005;@kelley.magurran2011;@martin.etal2022].
+We previously showed in *Dead by Daylight* that faster movement is effective for prey to evade predation [@fraserfranco.etal2022], in agreement with studies in other animals [@walker.etal2005;@kelley.magurran2011;@martin.etal2022].
 We suspect that experienced prey may increasingly rely on this strategy, which could explain why the relationship between hunting success and experience was concave when we did not control for prey speed (model II, Figure 1A).
@@
 For example, a predator that played 300 matches in six months might forget critical information related to prey detection or escape patterns compared to one that played 300 matches in six days.
-Investigating the impact of such time lags in future analyses may reveal important insights into the outcome of predator-prey interactions.
+While there is unequivoqual evidence that many predator species can learn quickly how to be efficient hunters, the role of the frequency of interactions remains unclear [@wooster.etal2023].
+Therefore, investigating the impact of such time lags in future analyses may reveal important insights into the outcome of predator-prey interactions.
 Another potential reason for the persistent differences among individuals is that neither the predator nor prey players' lives are at stake in the game.

### 2024-11-08 manuscript/manuscript.Rmd
update analysis title
update MS - MCMC to HMC

@@
 
-## Statistical analyses
+## Data analyses
 
@@
 We assumed that the random intercepts for the predator ID ($id$) followed a Gaussian distribution with estimated standard deviation ($id \sim N(0,\ I\ \sigma_{id}^{2})$).
-We fitted all models in R (version 4.1.2) using Markov chain Monte Carlo (MCMC) sampling with the package "brms" version 2.16.3 [@burkner2017], an R front-end for the STAN software [@StanDevelopmentTeam2023], and "cmdstanr" version 0.4.0 [@gabry.cesnovar2021] as the back-end for parameter estimation (cmdstan installation version 2.28.2).
+We fitted all models in R (version 4.1.2) using Hamiltonian Monte Carlo (HMC) sampling with the package "brms" version 2.16.3 [@burkner2017], an R front-end for the STAN software [@StanDevelopmentTeam2023], and "cmdstanr" version 0.4.0 [@gabry.cesnovar2021] as the back-end for parameter estimation (cmdstan installation version 2.28.2).
 For further details, please consult the GitHub repository of this project (https://github.com/quantitative-ecologist/predator-expertise).
@@
 
-### Markov Chain Monte Carlo settings
+### Hamiltonian Monte Carlo settings
 
-We parametrized the GAMMs to run four MCMC chains.
+We parametrized the GAMMs to run four chains.
 We ran 2500 iterations with a thinning set to eight for model I (see Table 1), and 1500 iterations with a thinning set to four for the remaining additive models (models II to V).
@@
 For each model, we obtained 1000 posterior samples per parameter.
-We assessed the convergence of the MCMC chains using trace plots, R-hat diagnostics with a threshold of <1.01, and effective sample sizes (ESS) with a threshold of >100 [@vehtari.etal2021].
+We assessed the convergence of the chains using trace plots, R-hat diagnostics with a threshold of <1.01, and effective sample sizes (ESS) with a threshold of >100 [@vehtari.etal2021].
 We also performed posterior predictive checks which showed an adequate fit of the models.

### 2024-11-08 code/code_figures/figure2.R
add test to fig2 in script

@@ -240,3 +240,55 @@ ggsave(
 
 # ==========================================================================
 # ==========================================================================
+
+
+
+
+
+
+
+
+
+
+# Test ----------------------------------------------------------------------
+
+# Testing plot 2 with less information
+
+library(ggridges)
+
+p <- ggplot(data = figdat_long, aes(x = value, y = parameter, fill = model))
+
+#p + geom_violinhalf() #+ geom_jitter(width = 0.2, shape=20)
+
+fig2_test <- p + geom_density_ridges(
+  scale=0.75, alpha = 0.5,
+  quantile_lines = TRUE,
+  quantiles = 2,
+  rel_min_height = 0.0009
+  ) +
+  #geom_vline(
+  #  data = results,
+  #  aes(xintercept = median, color = model, linetype = parameter),
+  #  #linetype = model,
+  #  show.legend = FALSE
+  #) +
+  scale_color_manual(
+    values = c("#482173FF", "#51C56AFF"),
+    labels = c("Prey rank", "Prey rank + prey speed")
+  ) +
+  scale_fill_manual(
+    values = c("#482173FF", "#51C56AFF"),
+    labels = c("Prey rank", "Prey rank + prey speed")
+  ) +
+  labs(fill = "Model:", color = "Model:") +
+  ylab("") +
+  xlab("\nStandard deviation") +
+  theme_bw() +
+  custom_theme
+
+ggsave(
+  fig2_test,
+  filename = "fig2test.png",
+  width = 8,
+  height = 6
+)

### 2024-11-08 code/code_figures/figure2.R
update fig2 test

@@ -187,8 +187,8 @@ custom_theme <- theme(
   strip.text.x = element_text(size = 15),
   panel.grid = element_blank(),
   panel.background = element_blank(),
-  legend.title = element_text(size = 15),
-  legend.text = element_text(size = 15),
+  legend.title = element_text(size = 13),
+  legend.text = element_text(size = 13),
   legend.position = "top"
 )
 

### 2024-11-15 code/code_figures/figure2.R
update fig2

@@ -276,7 +276,7 @@ fig2_test <- p + geom_density_ridges(
   #) +
   scale_fill_manual(
     values = c("#482173FF", "#51C56AFF"),
-    labels = c("Prey rank", "Prey rank + prey speed")
+    labels = c("II (Prey rank)", "V (Prey rank + prey speed)")
   ) +
   labs(fill = "Model:") +
   ylab("") +

### 2024-11-15 manuscript/manuscript.Rmd
update MS with fig.2

@@
 
-### Hamiltonian Monte Carlo settings
+### Parameter sampling settings
 
@@
 ```{r fig2, out.width = "100%", fig.cap = "**Figure 2**.
-Posterior distributions of the individual-level standard deviation parameters.
-The first two panels represent the intercept and slope standard deviations of the linear components relating hunting success to cumulative experience, and the third panel represents the standard deviation of the wiggliness component of the curves.
-The vertical dashed lines are the medians of the posterior distributions.
+Posterior distributions of the standard deviation of individual-level parameters estimated by the GAMM.
+The parameters are displayed on the y axis, and their standard deviation are displayed on the x axis.
+The intercept and slope standard deviations refer to the linear components of the estimated relationship between hunting success and cumulative experience.
+The standard deviation of the wiggliness parameter refers to the shape of the curves (i.e.
+nonlinear component).
+The vertical lines are the medians of the posterior distributions.
 For each panel, the light-coloured distributions (i.e.

### 2024-11-15 code/code_figures/figure2.R

@@ -276,9 +276,9 @@ fig2_test <- p + geom_density_ridges(
   #) +
   scale_fill_manual(
     values = c("#482173FF", "#51C56AFF"),
-    labels = c("II (Prey rank)", "V (Prey rank + prey speed)")
+    labels = c("Model II: prey rank", "Model V: prey rank + prey speed")
   ) +
-  labs(fill = "Model:") +
+  labs(fill = " ") +
   ylab("") +
   xlab("\nStandard deviation") +
   theme_classic() +
@@ -291,6 +291,6 @@ path <- file.path(getwd(), "outputs", "outputs_figures")
 ggsave(
   fig2_test,
   filename = file.path(path, "figure2.png"),
-  width = 8,
+  width = 8.5,
   height = 6
 )

### 2025-04-11 manuscript/manuscript.Rmd

@@
 By consuming prey, predators regulate prey populations, limit the spread of diseases, and mediate energy flow across trophic levels [@Hairston.1993;@ripple.etal2014;@schmitz2017].
-These processes emerge from interactions between individual predators and their prey, lying at the core of key concepts such as predation risk and prey selection[@pettorelli.etal2015;@labarge.etal2024].
+These processes emerge from interactions between individual predators and their prey, lying at the core of key concepts such as predation risk and prey selection [@pettorelli.etal2015;@labarge.etal2024].
 Predators also respond to interactions with their prey, increasing their effectiveness by adapting their strategy and choosing suitable habitats to hunt [@quevedo.etal2009;@schmitz2017].
 Central theory suggests that these responses by predators may result from acquiring expertise through practicing and learning the proper skills to locate, select, and capture their prey [@woo.etal2008;@wooster.etal2023;@labarge.etal2024].
-Hence, the role of expertise remains a missing but essential link to assess of how individual characteristics shape prey consumption during predator-prey interactions.
+Hence, the role of expertise remains a missing but essential link to assess how individual characteristics shape prey consumption during predator-prey interactions.
 
@@
 search and handling times, return rates) by associative learning, by developing search images, or by exploiting cues from their prey and their environment [@edwards.jackson1994;@morse2000;@macdonald2007;@reid.etal2010;@wilson-rankin2015].
-Through these processes, expert predators should have greater knowledge, better energy management, and acute motor skills to increase their chances of locating and capturing prey [@dukas2019].
+Through these processes, expert predators should have greater knowledge, better energy management, and acute motor skills that increase their chances of locating and capturing prey [@dukas2019].
 Therefore, differences among predators in their capacity to acquire expertise throughout their lifetime may underlie differences in hunting success.
@@
 To our knowledge, there is no empirical evidence showing links between prey antipredator behaviour and the acquisition of expertise in human and nonhuman predators.
-This represents a significant gap in our understanding of predator-prey interactions, as individual variation in predator expertise could lead to shifts in foraging behaviour and define whether predators select prey based on their frequency or their traits [@allen.etal1988;@ishii.shimada2012;@wooster.etal2023].
+This represents a significant gap in our understanding of predator-prey interactions, as individual variation in predator expertise could lead to shifts in foraging behaviour based on whether predators select prey for their frequency or their traits [@allen.etal1988;@ishii.shimada2012;@wooster.etal2023].
 
@@
 We assumed that the random intercepts for the predator ID ($id$) followed a Gaussian distribution with estimated standard deviation ($id \sim N(0,\ I\ \sigma_{id}^{2})$).
-We fitted all models in R (version 4.1.2) using Hamiltonian Monte Carlo (HMC) sampling with the package "brms" version 2.16.3 [@burkner2017], an R front-end for the STAN software [@StanDevelopmentTeam2023], and "cmdstanr" version 0.4.0 [@gabry.cesnovar2021] as the back-end for parameter estimation (cmdstan installation version 2.28.2).
+We fitted all models in R (version 4.1.2) using Hamiltonian Monte Carlo (HMC) sampling with the package "brms" version 2.17.0 [@burkner2017], an R front-end for the STAN software [@StanDevelopmentTeam2023], and "cmdstanr" version 0.4.0 [@gabry.cesnovar2021] as the back-end for parameter estimation (cmdstan installation version 2.28.2).
 For further details, please consult the GitHub repository of this project (https://github.com/quantitative-ecologist/predator-expertise).
@@
 We parametrized the GAMMs to run four chains.
-We ran 2500 iterations with a thinning set to eight for model I (see Table 1), and 1500 iterations with a thinning set to four for the remaining additive models (models II to V).
+We ran 2500 iterations with a thinning set to eight for model I, and 1500 iterations with a thinning set to four for the remaining additive models (models II to V).
 We set the first 500 iterations of each model as warm-ups.

### 2025-05-25 manuscript/manuscript.Rmd
Linux --> .rmd manuscript update up to results. need to finish results

@@
 ---
-title: "Prey movement shapes the acquisition of predator expertise in a virtual bi-trophic system"
+title: "Prey skill shapes the acquisition of predator expertise in a virtual bi-trophic system"
 output: 
@@
     "game_duration", "total_xp_pred", "cumul_xp_pred",
-    "hunting_success", "pred_speed", "prey_avg_speed"
+    "hunting_success", "pred_speed", "prey_avg_speed", 
+    "prey_avg_rank"
   )
@@
 To achieve this, predators must hone their skills and gain knowledge through repeated and extensive practice.
-On the other hand, prey may hinder the acquisition of predator expertise by employing antipredator tactics to evade detection and pursuit.
+Prey can interfere with this process and hinder the acquisition of predator expertise by employing antipredator tactics to evade detection and pursuit.
 However, empirical evidence on how predators acquire expertise through repeated encounters with their prey remains limited, largely due to the challenges of monitoring direct interactions in the wild.
@@
 We show that predators optimized prey consumption as they gained experience, indicating that they acquired expertise through extensive practice.
-At the population-level, we found that faster prey impaired the acquisition of expertise by reducing hunting success.
-Prey speed was also an important mediator of this relationship at the individual level, driving differences among predators in the acquisition of expertise.
-Our study outlines how prey antipredator behaviour can mediate the acquisition of expertise in predator populations.
+At the population-level, we found that more skillful prey impaired the acquisition of expertise by reducing hunting success.
+Prey skill was also an important mediator of this relationship at the individual level, driving differences among predators in the acquisition of expertise.
+More skillful prey moved faster, and predators were less successful against faster prey.
+Our results suggest that skilled prey use behaviours to outmanoeuver their predators, thereby mediating how predators acquire expertise at the population and individual level.
 
@@
 
-Predation is a fundamental biological process that acts as an agent of evolutionary change and shapes community structure [@abrams2000;@wirsing.etal2021].
-By consuming prey, predators regulate prey populations, limit the spread of diseases, and mediate energy flow across trophic levels [@Hairston.1993;@ripple.etal2014;@schmitz2017].
-These processes emerge from interactions between individual predators and their prey, lying at the core of key concepts such as predation risk and prey selection [@pettorelli.etal2015;@labarge.etal2024].
-Predators also respond to interactions with their prey, increasing their effectiveness by adapting their strategy and choosing suitable habitats to hunt [@quevedo.etal2009;@schmitz2017].
-Central theory suggests that these responses by predators may result from acquiring expertise through practicing and learning the proper skills to locate, select, and capture their prey [@woo.etal2008;@wooster.etal2023;@labarge.etal2024].
-Hence, the role of expertise remains a missing but essential link to assess how individual characteristics shape prey consumption during predator-prey interactions.
+Predation is a fundamental biological process acting as an agent of evolutionary change that regulates prey populations, limits the spread of diseases, and mediates energy flow across trophic levels [@hairston1993;@abrams2000;@ripple.etal2014;@wirsing.etal2021].
+These processes are driven by changes in predation risk and prey selection across landscapes as predators adjust their strategy and select suitable habitats to hunt [@quevedo.etal2009;@pettorelli.etal2015;@labarge.etal2024;@schmitz2017].
+Central theory suggests that predators become more efficient hunters by acquiring expertise through practicing and learning the proper skills to locate, select, and capture their prey [@woo.etal2008;@wooster.etal2023;@labarge.etal2024].
+Therefore, differences in hunting success between predators may be attributed to differences in the acquisition of hunting expertise.
+Unraveling how this expertise emerges remains a missing, but essential, link to assess how foraging behaviour shapes prey consumption during predator-prey interactions.
 
@@
 Through these processes, expert predators should have greater knowledge, better energy management, and acute motor skills that increase their chances of locating and capturing prey [@dukas2019].
-Therefore, differences among predators in their capacity to acquire expertise throughout their lifetime may underlie differences in hunting success.
-This is an important topic of study since predator cognition can maintain behavioural variation, shape prey phenotypic composition, and destabilize predator-prey systems [@kondoh2010;@skelhornandcandy2016;@kikuchi.etal2023].
+Theory predicts that predator cognition can maintain behavioural variation, shape prey phenotypic composition, and destabilize predator-prey systems [@kondoh2010;@skelhornandcandy2016;@kikuchi.etal2023].
+Therefore, individual differences among predators in their capacity to acquire expertise over time may be important for the ecological consequences of predation.
 
@@
 For example, @troscianko.etal2018 found that disruptive colouration interfered with human subjects' ability to form search images, hindering improvements in detection times over repeated attempts.
-Thus, antipredator behaviour may also hold the potential to hinder the acquisition of predator expertise.
+Antipredator behaviour may also hold the potential to hinder the acquisition of predator expertise.
 For instance, some predators may be limited in their capacity to develop the necessary attributes (e.g.
-physical, physiological, neurological) for fast-paced hunting, which could impair their acquisition of hunting expertise.
-To our knowledge, there is no empirical evidence showing links between prey antipredator behaviour and the acquisition of expertise in human and nonhuman predators.
-This represents a significant gap in our understanding of predator-prey interactions, as individual variation in predator expertise could lead to shifts in foraging behaviour based on whether predators select prey for their frequency or their traits [@allen.etal1988;@ishii.shimada2012;@wooster.etal2023].
+physical, physiological, neurological) for fast-paced hunting, which could impair their acquisition of hunting expertise if they chase faster prey.
+Moreover, prey antipredator behaviour can improve with experience through continued exposure to predators [@turner2006;@kelley.magurran2011;@lonnstedt.etal2012].
+Indeed, animal reintroduction programs show that individuals trained to recognize predators display more antipredator behaviours and are 1.5 times more likely to survive than predator-naive individuals [@tetzalf.etal2019].
+To our knowledge, there is little empirical evidence showing links between prey skill, antipredator behaviour, and the acquisition of predator expertise in human and nonhuman predators, representing a significant gap in our understanding of predator-prey interactions.
 
@@
 Here we mitigate these challenges by using *Dead by Daylight* as our study system, a videogame where four prey players must forage for resources while avoiding predation by a fifth player.
-Similar to agent-based simulations, *Dead by Daylight* provides controlled virtual environments to test ecological hypotheses [see @lymbery.etal2023 for an example with *Age of Empires II*], but with the advantage of having real players that interact in the virtual space.
-In this game, the predator population comprises individuals that either ambush or hunt at high speeds (i.e.
-mean movement speed along a slow-fast continuum), and their success is driven by the movement of the prey [@fraserfranco.etal2022].
+Similar to agent-based simulations, *Dead by Daylight* provides controlled virtual environments to test ecological hypotheses [see @montiglio.etal2025 for a review], with the advantage of having real players that interact in the virtual space.
+In this game, the predator population comprises individuals that either ambush or hunt at high speeds (i.e., mean movement speed along an ambush-cursorial continuum of tactics), and their success is driven by the movement of the prey [@fraserfranco.etal2022].
 The prey can increase their chances of survival by cooperating and moving fast to escape the predator [@cere.etal2021;@fraserfranco.etal2022;@santostefano.etal2024].
-The game also elicits natural reactions in players such as freezing when predation is imminent (M.F.F., personal observations), corroborating another virtual ecological study showing that predation drives individual variation in risk perception [@beauchamp2020].
+The game also elicits natural reactions in players such as freezing when predation is imminent (M.F.F., personal observations), corroborating virtual ecological studies showing that predation drives individual variation in risk perception [@beauchamp2020].
 These observations outline how ecological phenomena can emerge from human interactions in virtual systems with fixed rules [@brosnan.etal2017;@kasumovic.etal2017].
@@
 the probability of capturing all prey) and repeated experience (i.e.
-cumulated atempts).
+cumulated matches).
 First, we test the hypothesis that predator success will increase with experience up to some point at which it will stabilize [@dukas2019].
-However, we expect this pattern to change depending on the movement of the prey encountered.
-We hypothesize that prey will influence expertise acquisition, and predict that faster prey will reduce the gain in expertise.
-Therefore, we also investigate how prey movement influences expertise acquisition at the individual level.
-If prey speed does not influence hunting success, we predict that the gain in expertise will be similar among individuals.
-Alternatively, if prey speed influences hunting success, then the acquisition of expertise will vary among individuals.
+However, we hypothesize that prey skill and speed will influence predator expertise acquisition, and predict that faster and more skillful prey will reduce the gain in expertise at the population level.
+Prey with greater skills should move at faster speeds.
+Moreover, if prey skill and speed influence hunting success at the population level, then the acquisition of predator expertise should vary among individuals given these two prey features.
+Alternatively, if prey characteristics do not influence predator hunting success at the population level, we predict that the gain in predator expertise will be similar among individuals.
 
@@
 Behaviour Interactive Inc.
-provided data that spanned six months of gameplay recorded for every player from `r toString(format(min(dat$date)))` to `r toString(format(max(dat$date)))`.
-We analyzed only matches where players did not know each other.
+provided data that spanned six months of gameplay recorded for every player from `r toString(format(min(dat$date)))` to `r toString(format(max(dat$date)))` (game build 4.4.0 to 4.7.2).
+We analyzed only matches where players did not know each other and were unable to communicate using voice-recording devices.
 We filtered any matches where players were inactive, such as when mean distances travelled per second (i.e.
@@
 
-We sampled players that played 300 matches or more and monitored all their matches from the first to a maximum of 500 matches.
+Our study population comprised `r toString(format(length(unique(dat$predator_id))))` players that played at least 300 matches.
+We monitored all their matches from the first to a maximum of 500 matches, with a total record of `r toString(formatC(length(unique(dat$match_id)), big.mark = " "))` matches overall.
+The predator-players' experience in our population sample varied between `r toString(format(min(dat$total_xp_pred)))` and `r toString(format(max(dat$total_xp_pred)))` matches played.
+These matches lasted between `r toString(format(round(min(dat$game_duration / 60))), digits = 2)` and `r toString(format(round(max(dat$game_duration / 60))), digits = 2)` min (mean = `r toString(format(round(mean(dat$game_duration / 60))), digits = 2)` min).
 We recognize that we could have introduced a bias by retaining only those individuals, as they might already be seasoned video game enthusiasts and exhibit expert-level performance in their early matches in *Dead by Daylight*.
-Thus, we verified that our sample was not biased by comparing a random sample of players that played either 20 to 50 matches, 51 to 100 matches, or 101 to 300 matches during the same timeframe as our sampled population.
-We then took the first 20 matches played by these players, including those from our sampled population, and compared their median hunting success using a Bayesian hierarchical linear model.
-We found that all four groups had similar success as predators (Appendix 1: Table S1 and Figure S1), suggesting an absence of bias due to data sampling.
+Thus, we tested sample bias by comparing a random sample of players that played either 20 to 50 matches, 51 to 100 matches, or 101 to 300 matches during the same timeframe as our population sample.
+We then took the first 20 matches played by these players, including those from our population sample, and compared their mean hunting success using a Bayesian hierarchical linear model.
+We found that all four groups had similar success as predators in their first 20 matches (Appendix 1: Table S1 and Figure S1), suggesting an absence of bias due to data sampling.
 
-Our population consists of `r toString(format(length(unique(dat$predator_id))))` players who played as the predator, with a total record of `r toString(formatC(length(unique(dat$match_id)), big.mark = " "))` matches.
-The predator-players' experience varied between `r toString(format(min(dat$total_xp_pred)))` and `r toString(format(max(dat$total_xp_pred)))` matches played.
-These matches lasted between `r toString(format(round(min(dat$game_duration / 60))), digits = 2)` and `r toString(format(round(max(dat$game_duration / 60))), digits = 2)` min (mean = `r toString(format(round(mean(dat$game_duration / 60))), digits = 2)` min).
-The following information is collected and reported for every match : the player's anonymous ID, its avatar (i.e.
-the predator character chosen with its specific skill-gameplay mechanics), the game environment, the predator-player's experience, the mean speed of the groups of prey that the predator player encountered, and the mean rank of the prey encountered (a proxy for prey skill).
-The ranking system in *Dead by Daylight* was implemented by the company to pair players in a match based on their skill (https://deadbydaylight.fandom.com/wiki/Rank), and failing to account for it would prevent us from detecting a change in the predator's foraging success with experience.
+We collected the following information for every match in our population sample : the player's anonymous ID, the predator player's hunting success, the predator player's cumulated experience, and the mean rank and speed of the the group of prey that the predator player encountered.
+We defined hunting success as the number of prey consumed during the match (min = `r toString(format(min(dat$hunting_success)))`, max = `r toString(format(max(dat$hunting_success)))`).
+We defined the predator's cumulative experience as the number of matches played as the predator prior to the match being monitored.
+For example, the first match of a player would have a cumulative experience value of 0, while the tenth match would have a value of 9.
+Laslty, we measured the prey's rank as the mean rank of the four individual prey in a match (mean = `r toString(format(round(mean(dat$prey_avg_rank, na.rm = T), digits = 2), nsmall = 2))` ± `r toString(format(round(sd(dat$prey_avg_rank, na.rm = T), digits = 2), nsmall = 2))`), and the prey's speed as the mean travel speed of the four individual prey in a match (mean = `r toString(format(round(mean(dat$prey_avg_speed, na.rm = T), digits = 2), nsmall = 2))` ± `r toString(format(round(sd(dat$prey_avg_speed, na.rm = T), digits = 2), nsmall = 2))` m/s).
 
-We analyzed the mean speed of the prey group encountered by the predator.
-We measured the prey's speed as the mean travel speed of the four individual prey in a match (mean = `r toString(format(round(mean(dat$prey_avg_speed, na.rm = T), digits = 2), nsmall = 2))` ± `r toString(format(round(sd(dat$prey_avg_speed, na.rm = T), digits = 2), nsmall = 2))` m/s).
-We defined hunting success as the number of prey consumed during the match (min = `r toString(format(min(dat$hunting_success)))`, max = `r toString(format(max(dat$hunting_success)))`).
-Lastly, we defined the predator's cumulative experience as the number of matches played as the predator prior to the match being monitored.
-For example, the first match of a player would have a cumulative experience value of 0, while the tenth match would have a value of 9.
-We did not account for matches where predators played as the prey.
+We used the mean rank of the prey as a proxy of their skill.
+The ranking system in *Dead by Daylight* varies between `r toString(format(min(dat$prey_avg_rank)))` and `r toString(format(max(dat$prey_avg_rank)))` (1 indicating the highest skill), and was implemented by the company to pair players in a match based on their skill (https://deadbydaylight.fandom.com/wiki/Rank).
+The skill of a player increases based on their performance from match to match.
+While the ranking system represents an approximation of prey skill, it was the most readily available metric we could access to determine the skill of the prey group.
+In addition, the pairing system is subject to variation depending on factors such as player availability, which can result in predator-prey groups with unbalanced skill.
+This allowed us to evaluate the effect of the prey's skill on predator success.
 
@@
 
-We tested how predators developed their expertise by computing five Bayesian generalized additive mixed models (GAMM) with thin-plate regression splines, all of which estimated the relationship between hunting success (i.e.
-number of prey consumed) and the predators' cumulative experience (i.e.
-number of matches played before the current match).
+We tested how predators developed their expertise by computing five Bayesian generalized additive mixed models (GAMM) with thin-plate regression splines, all of which estimated the relationship between hunting success and the predators' cumulative experience.
 We parametrized the models following the method of @pedersen.etal2019.
-The first model (I) was the simplest, with a common global smoothing function and random intercepts for the predator ID.
-In this model, we assume that predators have the same acquisition of expertise, with the model estimating a trend for the average individual (i.e.
-global smoother).
-The second model (II) included varying individual smoothers for the predator ID.
-Here, we assume that individual predators share a similar relationship between success and experience, but that this relationship can vary among them (e.g.
+For the first and second models (I and II), we assume that individual predators share a similar relationship between success and experience, but that this relationship can vary among them (e.g.
 predator 1 has a steeper curve than predator 2).
 This enabled us to test whether predators differed in the development of their expertise.
+Both models included a common global smoothing function (i.e., population effects) and random smooths for the predator ID (i.e., individual effects).
+Model I included the standardized match duration as a covariate, while model II included the standardized match duration and the average rank of the prey group, allowing us to test if the variation among predators in expertise acquisition was driven by prey skill.
 In the third model (III), we kept the individual smoothers for the predators, but removed the global smoother.
 This model assumes that predators do not share a common relationship between success and experience.
-The fourth (IV) and fifth (V) models were expansions of the third and second models respectively, where we included the prey speed to assess its effect on the relationship between success and experience.
-We included the standardized match duration and prey rank as covariates in all five models.
+In this model, we control for the standardized match duration and the average rank of the prey group.
+The fourth (IV) and fifth (V) models were expansions of the model III and model II respectively, where we included the standardized prey speed to assess its effect on the relationship between success and experience.
 
-We computed the five models using a modified version of the beta-binomial distribution.
-Hunting success was estimated as the probability of consuming the four prey ($\mu_i$), drawn from a Beta distribution ($Beta(\mu_i, \phi)$) with mean ($\mu \in [0,1]$) and precision ($\phi > 0$) parameters.
-We used a logit link function to estimate $\mu_i$ where $\mu_i = \frac{exp(\eta_i)}{exp(\eta_i) + 1}$ and $\eta_i$ is the linear predictor, while the precision parameter ($\phi$) was estimated with an identity link.
+We computed the five models using a modified beta-binomial distribution.
+Hunting success was estimated as the number of prey consumed out of four, with probability of success drawn from a Beta distribution ($Beta(\mu_i, \phi)$) with mean ($\mu \in [0,1]$) and precision ($\phi > 0$) parameters.
+We used a logit link function to estimate $\mu_i$ where $\mu_i = \frac{exp(\eta_i)}{exp(\eta_i) + 1}$ and $\eta_i$ is the linear predictor.
+The precision parameter ($\phi$) was estimated using a log-link function.
 We used ten basis functions (K = 10) for the models to estimate the relationship between hunting success and experience.
 We assumed that the random intercepts for the predator ID ($id$) followed a Gaussian distribution with estimated standard deviation ($id \sim N(0,\ I\ \sigma_{id}^{2})$).
-We fitted all models in R (version 4.1.2) using Hamiltonian Monte Carlo (HMC) sampling with the package "brms" version 2.17.0 [@burkner2017], an R front-end for the STAN software [@StanDevelopmentTeam2023], and "cmdstanr" version 0.4.0 [@gabry.cesnovar2021] as the back-end for parameter estimation (cmdstan installation version 2.28.2).
-For further details, please consult the GitHub repository of this project (https://github.com/quantitative-ecologist/predator-expertise).
+We compared the predictive accuracy of all five models using approximate leave-one-out cross-validation with Pareto-smoothed importance sampling [@vehtari.etal2017;@piironen.vehtari2017;@vehtari.etal2022].
 
-We used weakly informative Gaussian priors for the intercept ($N(0,\ 0.5^{2})$) and the global trend of cumulative experience ($N(0,\ 2^{2}))$).
-Following @fraserfranco.etal2022, we defined a positive Gaussian prior on the precision parameter ($N(2,\ 0.5^{2})$), a positive Gaussian prior ($N(1,\ 0.5^{2}))$) on the game duration because longer trials lead to greater success, and a negative Gaussian prior on prey speed ($N(-1,\ 0.5^{2})$) because encountering faster prey is associated with lower success in this system.
+We defined weakly informative Gaussian priors for the intercept ($N(0,\ 0.5^{2})$) and the global trend of cumulative experience ($N(0,\ 2^{2}))$).
+Following @fraserfranco.etal2022, we defined a positive Gaussian prior ($N(1,\ 0.5^{2}))$) for the slope of the game duration because longer trials lead to greater success, a Gaussian prior ($N(0,\ 1^{2}))$) for the slope of the prey rank, and a negative Gaussian prior on prey speed ($N(-1,\ 0.5^{2})$) because encountering faster prey is associated with lower success in this system [@fraserfranco.etal2022].
+We defined a positive Gaussian prior on the precision parameter ($N(2,\ 0.5^{2})$).
 We employed weakly informative half-Gaussian priors on all the standard deviation parameters ($N(0,\ 0.5^{2})$).
-We compared the predictive accuracy of all five models using approximate leave-one-out cross-validation with Pareto-smoothed importance sampling [@vehtari.etal2017;@piironen.vehtari2017;@vehtari.etal2022].
+
+To test the hypothesis that prey speed increases with prey skill, we fitted a Bayesian linear regression estimating the relationship between prey speed and the standardized prey rank.
+We controlled for the game duration, using it as a linear covariate in the model, and included the predator identity as a random effect.
+We used a weakly informative Gaussian prior on the slope of the game duration ($N(0,\ 0.5^{2})$) and a weakly informative negative Gaussian prior the slope of the prey rank ($N(-0.5,\ 1^{2})$) because prey of greater skill (i.e., rank 1 is the highest skill) should move faster.
+Lastly, we used a weakly informative Gaussian prior on the intercept ($N(2,\ 0.5^{2})$) because prey should move at around two meters per second, and half-Gaussian priors on the standard deviation parameters ($N(0,\ 0.5^{2})$).
 
@@
 
-We parametrized the GAMMs to run four chains.
-We ran 2500 iterations with a thinning set to eight for model I, and 1500 iterations with a thinning set to four for the remaining additive models (models II to V).
+We parametrized the GAMMs to run four chains of 1500 iterations each, sampling one parameter value every four iterations.
 We set the first 500 iterations of each model as warm-ups.
-For each model, we obtained 1000 posterior samples per parameter.
+Similarly, the linear model ran four chains of 1500 iterations each, sampling a parameter value at every iteration, and discarding the first 500 iterations.
+For GAMM model, we obtained 1000 posterior samples per parameter, and 4000 values for the linear model.
 We assessed the convergence of the chains using trace plots, R-hat diagnostics with a threshold of <1.01, and effective sample sizes (ESS) with a threshold of >100 [@vehtari.etal2021].
@@
 
+We fitted all models in R (version 4.1.2) using Hamiltonian Monte Carlo (HMC) sampling with the package "brms" version 2.22.0 [@burkner2017], an R front-end for the STAN software [@StanDevelopmentTeam2023], and "cmdstanr" version 0.8.0 [@gabry.cesnovar2021] as the back-end for parameter estimation (cmdstan installation version 2.35.0).
+For further details on how to reproduce the results, please consult the GitHub repository of this project (https://github.com/quantitative-ecologist/predator-expertise).
+
 ### Hypothesis testing
 
-We tested the hypothesis that antipredator behaviour impairs the acquisition of expertise at the population level by visually comparing the global trends of model II (not controlling for prey speed) and model V (controlling for prey speed) relating hunting success to cumulated experience (Figure 1).
-At the individual level, we tested our hypothesis that antipredator behaviour generates differences among predators in expertise acquisition by comparing individual-level parameters of model II and model V.
+We tested the hypothesis that prey skill and antipredator behaviour impairs the acquisition of expertise at the population level by visually comparing the global trends of model I (only controlling for game duration), model II (controlling for game duration and prey rank) and model V (controlling game duration, prey rank, and prey speed) (Figure 1).
+At the individual level, we tested our hypothesis that prey skill and antipredator behaviour generate differences among predators in expertise acquisition by comparing the individual-level variance parameters.
 Specifically, we compared the standard deviations of 1) the random intercepts (i.e.
@@
 
-Out of all five GAMM models, the two that accounted for the prey group's rank and speed were the best at predicting the data with similar predictive accuracies (Table 1).
-Models in which prey rank was not accounted for resulted in no change in hunting success with experience for the average individual (i.e.
-no gain in expertise, results not shown).
-Accounting for the prey rank resulted in a concave-shaped relationship (model II), with the highest success ranging between ~200 and ~300 matches (Figure 1A).
-
-We found strong evidence of a negative relationship between hunting success and prey speed (Figure S2).
-As predicted, the effect of experience on hunting success for the average individual followed a diminishing returns curve when controlling for prey speed (model V), with predators optimizing their success after playing ~300 matches (Figure 1C).
-The curve shows a 38% increase in the probability of consuming all prey for the average individual between the first and the ~330$^{th}$ match, where success reached a plateau (Figure 1C).
+Out of all five GAMM models, those that included a global smoother were better at predicting the data compared to those without a global smoother, indicating that predators shared a similar relationship between prey capture and experience (Table 1).
+Model V, which accounted for the prey group's rank and speed was the best at predicting the data (Table 1).
 
@@
 
-## Acquisition of expertise at the individual level
-
-Prey speed did not influence among-individual differences in average hunting success as the posterior distributions of the standard deviations of individual intercepts were almost completely overlapping (Figure 2, median = `r toString(round(sum1[2,]$Estimate, digits = 2))` vs median = `r toString(round(sum2[2,]$Estimate, digits = 2))`).
-However, individuals differed in the acquisition of their hunting expertise (Figure 1B-D).
-We found strong evidence that the speed of the prey mediated among-individual differences in the linear relationship between success and experience, as there were substantial differences in the standard deviations of the individual slopes between the two models (Figure 2, median = `r toString(round(sum1[3,]$Estimate, digits = 2))` vs median = `r toString(round(sum2[3,]$Estimate, digits = 2))`).
-Differences among individuals in the direction of the linear relationship between success and experience were `r toString(round(sum1[3,]$Estimate / sum2[3,]$Estimate, digits = 2))` times lower when we removed the effect of prey speed (i.e.
-accounting for it in the model).
+Models in which prey rank was not accounted for resulted in no change in hunting success with experience for the average individual (i.e.
+no gain in expertise, Figure 1A).
+As we predicted, accounting for the prey rank resulted in a diminishing returns curve (model II), with predators optimizing their success after playing ~400 matches (Figure 1B).
+The predictions indicate a 40% increase in the probability of consuming all prey for the average individual between the first and the ~400$^{th}$ match, where success reached a plateau (Figure 1B).
+Controling for prey speed did not influence the results at the population level (Figure 1C).
+Yet, we found strong evidence of a negative relationship between hunting success and prey speed, indicating that hunting faster prey reduced the predator's chances of capture (Figure S2).
+Moreover, prey with more expertise moved at higher speeds compared to less skilled prey (Figure 2).
 
@@
 
+## Acquisition of expertise at the individual level
+
+Prey speed did not influence among-individual differences in average hunting success as the posterior distributions of the standard deviations of individual intercepts were almost completely overlapping (Figure 3, median = `r toString(round(sum1[2,]$Estimate, digits = 2))` vs median = `r toString(round(sum2[2,]$Estimate, digits = 2))`).
+However, individuals differed in the acquisition of their hunting expertise (Figure 1B-D).
+We found strong evidence that the speed of the prey mediated among-individual differences in the linear relationship between success and experience, as there were substantial differences in the standard deviations of the individual slopes between the two models (Figure 3, median = `r toString(round(sum1[3,]$Estimate, digits = 2))` vs median = `r toString(round(sum2[3,]$Estimate, digits = 2))`).
+Differences among individuals in the direction of the linear relationship between success and experience were `r toString(round(sum1[3,]$Estimate / sum2[3,]$Estimate, digits = 2))` times lower when we removed the effect of prey speed (i.e.
+accounting for it in the model).
+
+```{r fig2, out.width = "100%", fig.cap = "**Figure 2**.
+Median posterior predictions of the relationship between prey speed and prey skill.
+The average speed of the prey group is on the y-axis, and the average rank of the prey group is on the x-axis.
+The colour gradient showcases the density of observations, with lighter colours indicating lower densities, and darker colours indicating higher densities of observations.
+Lower rank values indicate that the prey group is more skilled, while higher values indicate that the prey are less skilled"}
+
+knitr::include_graphics(
+  here::here("outputs", "outputs_figures", "figurex.png")
+)
+```
+
 We also found strong evidence that the speed of the prey mediated the form of the relationship between hunting success and experience at the individual level.
@@
 
-```{r fig2, out.width = "100%", fig.cap = "**Figure 2**.
+```{r fig3, out.width = "100%", fig.cap = "**Figure 3**.
 Posterior distributions of the standard deviation of individual-level parameters estimated by the GAMM.

### 2025-05-25 code/code_figures/figure2.R
Windows --> rename from fig2 to fig3

@@ -1,222 +1,0 @@
-# ==========================================================================
-
-#                         Plot GAMM sds parameters
-
-# ==========================================================================
-
-# Figure of the posterior parameters for:
-#- Intercept standard deviation
-#- Slope standard deviation
-#- Wiggliness standard deviation
-
-
-
-# ==========================================================================
-# 1. Prepare script
-# ==========================================================================
-
-
-# Load libraries and model -------------------------------------------------
-
-library(brms)
-library(data.table)
-library(ggplot2)
-library(ggridges)
-
-path <- file.path(getwd(), "outputs", "outputs_models")
-
-fit1 <- readRDS(file.path(path, "GAMM-I.rds"))
-fit2 <- readRDS(file.path(path, "GAMM-II.rds"))
-fit5 <- readRDS(file.path(path, "GAMM-V.rds"))
-
-# ==========================================================================
-# ==========================================================================
-
-
-
-
-
-# ==========================================================================
-# 2. Extract posterior draws into a tidy table
-# ==========================================================================
-
-# Extract posterior draws fit
-post1 <- data.table(
-  as_draws_df(
-    fit1,
-    variable = c(
-      "sds_sZcumul_xppredator_id_1",
-      "sds_sZcumul_xppredator_id_2",
-      "sds_sZcumul_xppredator_id_3"
-    )
-  )
-)
-
-
-# Extract posterior draws fit1
-post2 <- data.table(
-  as_draws_df(
-    fit2,
-    variable = c(
-      "sds_sZcumul_xppredator_id_1",
-      "sds_sZcumul_xppredator_id_2",
-      "sds_sZcumul_xppredator_id_3"
-    )
-  )
-)
-
-# Extract posterior draws fit2
-post5 <- data.table(
-  as_draws_df(
-    fit5,
-    variable = c(
-      "sds_sZcumul_xppredator_id_1",
-      "sds_sZcumul_xppredator_id_2",
-      "sds_sZcumul_xppredator_id_3"
-    )
-  )
-)
-
-# Clear memory
-rm(fit1)
-rm(fit2)
-rm(fit5)
-
-# Combine posterior draws
-figdat <- rbind(
-  post1,
-  post2,
-  post5
-)
-
-# Add model variable
-figdat[, model := c(rep("fit1", 4000), rep("fit2", 4000), rep("fit5", 4000))]
-
-# ==========================================================================
-# ==========================================================================
-
-
-
-
-
[…]

### 2025-05-25 code/code_figures/figure3.R

@@ -1,0 +1,222 @@
+# ==========================================================================
+
+#                         Plot GAMM sds parameters
+
+# ==========================================================================
+
+# Figure of the posterior parameters for:
+#- Intercept standard deviation
+#- Slope standard deviation
+#- Wiggliness standard deviation
+
+
+
+# ==========================================================================
+# 1. Prepare script
+# ==========================================================================
+
+
+# Load libraries and model -------------------------------------------------
+
+library(brms)
+library(data.table)
+library(ggplot2)
+library(ggridges)
+
+path <- file.path(getwd(), "outputs", "outputs_models")
+
+fit1 <- readRDS(file.path(path, "GAMM-I.rds"))
+fit2 <- readRDS(file.path(path, "GAMM-II.rds"))
+fit5 <- readRDS(file.path(path, "GAMM-V.rds"))
+
+# ==========================================================================
+# ==========================================================================
+
+
+
+
+
+# ==========================================================================
+# 2. Extract posterior draws into a tidy table
+# ==========================================================================
+
+# Extract posterior draws fit
+post1 <- data.table(
+  as_draws_df(
+    fit1,
+    variable = c(
+      "sds_sZcumul_xppredator_id_1",
+      "sds_sZcumul_xppredator_id_2",
+      "sds_sZcumul_xppredator_id_3"
+    )
+  )
+)
+
+
+# Extract posterior draws fit1
+post2 <- data.table(
+  as_draws_df(
+    fit2,
+    variable = c(
+      "sds_sZcumul_xppredator_id_1",
+      "sds_sZcumul_xppredator_id_2",
+      "sds_sZcumul_xppredator_id_3"
+    )
+  )
+)
+
+# Extract posterior draws fit2
+post5 <- data.table(
+  as_draws_df(
+    fit5,
+    variable = c(
+      "sds_sZcumul_xppredator_id_1",
+      "sds_sZcumul_xppredator_id_2",
+      "sds_sZcumul_xppredator_id_3"
+    )
+  )
+)
+
+# Clear memory
+rm(fit1)
+rm(fit2)
+rm(fit5)
+
+# Combine posterior draws
+figdat <- rbind(
+  post1,
+  post2,
+  post5
+)
+
+# Add model variable
+figdat[, model := c(rep("fit1", 4000), rep("fit2", 4000), rep("fit5", 4000))]
+
+# ==========================================================================
+# ==========================================================================
+
+
+
+
+
[…]

### 2025-05-25 code/code_figures/figure_2.R
Windows --> figx now fig2

@@ -215,7 +215,7 @@ path <- file.path(getwd(), "outputs", "outputs_figures")
 
 ggsave(
   p,
-  filename = file.path(path, "figurex.png"),
+  filename = file.path(path, "figure2.png"),
   units = "in",
   dpi = 300,
   width = 5,

### 2025-05-26 manuscript/manuscript.Rmd
Windows --> figx now fig2
Windows --> correct fig2
load proper loo file

@@
 path <- file.path(here::here("outputs", "outputs_model-validation"))
-loo_tab <- data.table(readRDS(file.path(path, "GAMM_loo-cv-table.rds")))
+loo_tab <- data.table(readRDS(file.path(path, "loo-cv-table_GAMM.rds")))
 
@@
 knitr::include_graphics(
-  here::here("outputs", "outputs_figures", "figurex.png")
+  here::here("outputs", "outputs_figures", "figure2.png")
 )
@@
 knitr::include_graphics(
-  here::here("outputs", "outputs_figures", "figure2.png")
+  here::here("outputs", "outputs_figures", "figure3.png")
 )

### 2025-06-01 code/code_figures/figure1.R
Windows --> Fix fig1

@@ -501,7 +501,7 @@ figure <- ggarrange(
   ncol = 6, nrow = 2,
   labels = c(
     "(A)", "", "(B)", "", "(C)", "",
-    "(C)", "", "(D)", "", "(E)", ""
+    "(D)", "", "(E)", "", "(F)", ""
   ),
   widths = c(
     0.15, 1.5, 0.15, 1.5, 0.15, 1.5,

### 2025-06-01 code/code_figures/figure1.R
fix: fig1 artefact

@@ -334,9 +334,6 @@ predicted_values <- tab2_b$estimate__
 # Define the x-axis range
 x <- tab2_b$Zcumul_xp
 
-# Fit a spline to the predicted values
-spline_fit <- splinefun(x, predicted_values)
-
 # Calculate the first derivative using finite differences
 dx <- mean(diff(x))
 derivatives <- diff(predicted_values) / dx
@@ -412,9 +409,6 @@ predicted_values <- tab5_b$estimate__
 
 # Define the x-axis range
 x <- tab5_b$Zcumul_xp
-
-# Fit a spline to the predicted values
-spline_fit <- splinefun(x, predicted_values)
 
 # Calculate the first derivative using finite differences
 dx <- mean(diff(x))

### 2025-06-01 manuscript/manuscript.Rmd
Windows --·> finished results section

@@
 First, we test the hypothesis that predator success will increase with experience up to some point at which it will stabilize [@dukas2019].
-However, we hypothesize that prey skill and speed will influence predator expertise acquisition, and predict that faster and more skillful prey will reduce the gain in expertise at the population level.
-Prey with greater skills should move at faster speeds.
-Moreover, if prey skill and speed influence hunting success at the population level, then the acquisition of predator expertise should vary among individuals given these two prey features.
-Alternatively, if prey characteristics do not influence predator hunting success at the population level, we predict that the gain in predator expertise will be similar among individuals.
+However, we hypothesize that prey skill and speed will influence predator expertise acquisition by modulating the rate of gain in expertise at the population level.
+Second, we test the hypothesis that prey with more skill should have improved antipredator behaviour, and predict that prey with greater skills will move at faster speeds.
+Third, if prey skill and speed influence hunting success at the population level, then we expect that the acquisition of predator expertise will vary among individuals given these two prey features.
 
@@
 Model I included the standardized match duration as a covariate, while model II included the standardized match duration and the average rank of the prey group, allowing us to test if the variation among predators in expertise acquisition was driven by prey skill.
-In the third model (III), we kept the individual smoothers for the predators, but removed the global smoother.
+The third model had the same structure as model II but without a global smoother.
 This model assumes that predators do not share a common relationship between success and experience.
 In this model, we control for the standardized match duration and the average rank of the prey group.
-The fourth (IV) and fifth (V) models were expansions of the model III and model II respectively, where we included the standardized prey speed to assess its effect on the relationship between success and experience.
+The fourth (IV) and fifth (V) models were expansions of model II, but included the standardized prey speed to assess its effect on the relationship between success and experience.
+Model IV does not have a standard global smoother, while mother V includes one.
 
@@
 We used a logit link function to estimate $\mu_i$ where $\mu_i = \frac{exp(\eta_i)}{exp(\eta_i) + 1}$ and $\eta_i$ is the linear predictor.
-The precision parameter ($\phi$) was estimated using a log-link function.
+The precision parameter ($\phi$) was estimated using a log link function.
 We used ten basis functions (K = 10) for the models to estimate the relationship between hunting success and experience.
@@
 For GAMM model, we obtained 1000 posterior samples per parameter, and 4000 values for the linear model.
-We assessed the convergence of the chains using trace plots, R-hat diagnostics with a threshold of <1.01, and effective sample sizes (ESS) with a threshold of >100 [@vehtari.etal2021].
+We assessed the convergence of the chains using trace plots, R-hat diagnostics with a threshold set between 1.00 and 1.02, and effective sample sizes (ESS) with a threshold of >100 [@vehtari.etal2021].
 We also performed posterior predictive checks which showed an adequate fit of the models.
 
-We fitted all models in R (version 4.1.2) using Hamiltonian Monte Carlo (HMC) sampling with the package "brms" version 2.22.0 [@burkner2017], an R front-end for the STAN software [@StanDevelopmentTeam2023], and "cmdstanr" version 0.8.0 [@gabry.cesnovar2021] as the back-end for parameter estimation (cmdstan installation version 2.35.0).
-For further details on how to reproduce the results, please consult the GitHub repository of this project (https://github.com/quantitative-ecologist/predator-expertise).
+We fitted all models in R (version 4.4.0) using Hamiltonian Monte Carlo (HMC) sampling with the package "brms" version 2.22.12 [@burkner2017], an R front-end for the STAN software [@StanDevelopmentTeam2023], and "cmdstanr" version 0.9.0 [@gabry.cesnovar2021] as the back-end for parameter estimation (cmdstan installation version 2.36.0).
+For further details on how to reproduce our results, please consult the GitHub repository (https://github.com/quantitative-ecologist/predator-expertise).
 
@@
 ```{r summary, cache = TRUE}
-# Import top 2 model fits
-fit1 <- readRDS(here::here("outputs", "outputs_models", "GAMM-II.rds"))
-fit2 <- readRDS(here::here("outputs", "outputs_models", "GAMM-V.rds"))
+# Import top 3 model fits
+fit1 <- readRDS(here::here("outputs", "outputs_models", "GAMM-I.rds"))
+fit2 <- readRDS(here::here("outputs", "outputs_models", "GAMM-II.rds"))
+fit5 <- readRDS(here::here("outputs", "outputs_models", "GAMM-V.rds"))
 
@@
 sum2 <- summary(fit2, robust = TRUE)$splines
+sum5 <- summary(fit5, robust = TRUE)$splines
 
@@
 rm(fit2)
+rm(fit5)
 ```
@@
 
-Out of all five GAMM models, those that included a global smoother were better at predicting the data compared to those without a global smoother, indicating that predators shared a similar relationship between prey capture and experience (Table 1).
-Model V, which accounted for the prey group's rank and speed was the best at predicting the data (Table 1).
+Among models that shared the same covariate structure (i.e., model II and III; model IV and V), those that included a global smoother generally performed better at predicting hunting success compared than those without one (Table 1).
+While we detected differences among predators in the form and shape of the relationship between success and experience (Figure 1), this suggests that there is a common underlying structure in this relationship across the population.
+Overall, model V, which included all prey features, was the best at predicting predator hunting success (Table 1).
 
@@
 loo_tab[
-  , model := c(
+  , model1 := c(
     "(V) predator xp + ID smoothers + prey rank + prey speed",
     "(IV) ID smoothers + prey rank + prey speed",
+    "(II) predator xp + ID smoothers + prey rank",
     "(III) ID smoothers + prey rank",
-    "(II) predator xp + ID smoothers + prey rank",
-    "(I) predator xp + ID intercepts + prey rank"
+    "(I) predator xp + ID smoothers"
   )
@@
 # Reorder columns
-loo_tab <- loo_tab[, c(9, 1:8)]
+loo_tab <- loo_tab[, c(10, 2:9)]
 
@@
   col_keys = c(
-    "model",
+    "model1",
     "elpd_diff",
@@
 tab1 <- loo_tab %>%
-  select(model, elpd_diff, se_diff, elpd_loo, se_elpd_loo) %>%
+  select(model1, elpd_diff, se_diff, elpd_loo, se_elpd_loo) %>%
   flextable(col_keys = my_header$col_keys) %>%
@@
   set_caption(
-    caption = "Table 1. Leave-one-out cross-validation table of the five GAMMs relating hunting success to predator experience.",
+    caption = "Table 1.
+Leave-one-out cross-validation results comparing the predictive performance of the five GAMMs estimating the relationship between hunting success and predator experience.
+All models reported include the game duration as a control.
+Only differences in smoothing functions and prey covariates are reported.",
     style = "Table Caption"
@@
 
-Models in which prey rank was not accounted for resulted in no change in hunting success with experience for the average individual (i.e.
+Models in which we did not account for the prey rank resulted in almost no change in hunting success with experience for the average individual (i.e.
 no gain in expertise, Figure 1A).
-As we predicted, accounting for the prey rank resulted in a diminishing returns curve (model II), with predators optimizing their success after playing ~400 matches (Figure 1B).
-The predictions indicate a 40% increase in the probability of consuming all prey for the average individual between the first and the ~400$^{th}$ match, where success reached a plateau (Figure 1B).
-Controling for prey speed did not influence the results at the population level (Figure 1C).
-Yet, we found strong evidence of a negative relationship between hunting success and prey speed, indicating that hunting faster prey reduced the predator's chances of capture (Figure S2).
-Moreover, prey with more expertise moved at higher speeds compared to less skilled prey (Figure 2).
+As we expected, accounting for the prey rank (model II) influenced the form of the relationship, resulting in a diminishing returns curve with predators optimizing their success after playing ~450 matches (Figure 1B).
+The predictions indicate that the population would improve their hunting success by ~40% from the first to the ~450$^{th}$ match by hunting prey with average skill (Figure 1B).
+
+Controlling for prey speed did not alter the form of the relationship between success and experience at the population level.
+However, as predicted, the rate of expertise acquisition shifted slightly: the predator population optimized its hunting success earlier, shortly after 300 matches (Figure 1C).
+We also found strong evidence of a negative relationship between hunting success and prey speed, indicating that hunting faster prey reduced the predator's chances of capture (Figure S2).
+Moreover, prey groups composed of individuals with greater skill moved faster than those composed of individuals with lower skill (Figure 2).
 
@@
 the number of matches played before each observation) is on the x-axis.
-Panels A and C show the acquisition of expertise for the average individual with the vertical dashed lines on the left representing the lowest predicted values.
+Panels A to C show the acquisition of expertise for the average individual with the vertical dashed lines on the left representing the lowest predicted values.
 For panel A, the right-side vertical dashed line shows the highest predicted success.
-For panel C, the right-side dashed line represents the point on the curve where success was optimized, which we calculated using the finite differences method to obtain the first derivative of the predicted values.
-Panels B and D show among individual differences in the acquisition of expertise, with each curve representing an individual predator.
-(A-B) model II where we control for the prey rank (C-D) model V where we control for the prey rank and the speed of the prey group."}
+For panels B and C, the right-side dashed line represents the point on the curve where success was optimized, which we calculated using the finite differences method to obtain the first derivative of the predicted values.
+Panels D to F show among individual differences in the acquisition of expertise, with each curve representing an individual predator.
+(A-D) Model I where we control for the game duration exclusively.
+(B-E) Model II where we control for the game duration and the prey rank exclusively.
+(C-F) Model V where we control for game duration as well as the rank and speed of the prey group."}
 
@@
 
-Prey speed did not influence among-individual differences in average hunting success as the posterior distributions of the standard deviations of individual intercepts were almost completely overlapping (Figure 3, median = `r toString(round(sum1[2,]$Estimate, digits = 2))` vs median = `r toString(round(sum2[2,]$Estimate, digits = 2))`).
-However, individuals differed in the acquisition of their hunting expertise (Figure 1B-D).
-We found strong evidence that the speed of the prey mediated among-individual differences in the linear relationship between success and experience, as there were substantial differences in the standard deviations of the individual slopes between the two models (Figure 3, median = `r toString(round(sum1[3,]$Estimate, digits = 2))` vs median = `r toString(round(sum2[3,]$Estimate, digits = 2))`).
-Differences among individuals in the direction of the linear relationship between success and experience were `r toString(round(sum1[3,]$Estimate / sum2[3,]$Estimate, digits = 2))` times lower when we removed the effect of prey speed (i.e.
-accounting for it in the model).
+Our results show that individual predators varied in their acquisition of expertise, regardless of whether we controlled for prey features (Figure 1 D-F and Figure 3).
+However, the extent of the variation in the relationship between success and experience was shaped by the prey's skill, as shown by the the posterior distributions of individual-level parameters in the model accounting for the rank of the prey (model II, Figure 2).
+Specifically, the standard deviations of individual intercepts (median model I = `r toString(round(sum1[2,]$Estimate, digits = 2))` vs median model II = `r toString(round(sum2[2,]$Estimate, digits = 2))`), slopes (median model I = `r toString(round(sum1[3,]$Estimate, digits = 2))` vs median model II = `r toString(round(sum2[3,]$Estimate, digits = 2))`), and wiggliness (median model I = `r toString(round(sum1[4,]$Estimate, digits = 2))` vs median model II = `r toString(round(sum2[4,]$Estimate, digits = 2))`) were all larger in model II than in model I.
+These results indicate that if predators hunted prey with average skill levels, differences among individual predators in expertise acquisition would be greater.
 
@@
 
-We also found strong evidence that the speed of the prey mediated the form of the relationship between hunting success and experience at the individual level.
-We detected large differences in the standard deviations of the wiggliness (Figure 2, median = `r toString(round(sum1[4,]$Estimate, digits = 2))` vs median = `r toString(round(sum2[4,]$Estimate, digits = 2))`).
-The lower standard deviation for the model where we accounted for prey speed (model V) suggests that the form of the relationship between success and experience was more similar among individuals.
+Prey speed did not influence among-individual differences in average hunting success (median model II = `r toString(round(sum2[2,]$Estimate, digits = 2))` vs median model V = `r toString(round(sum5[2,]$Estimate, digits = 2))`), nor in expertise acquisition (median model II = `r toString(round(sum2[3,]$Estimate, digits = 2))` vs median model V = `r toString(round(sum5[3,]$Estimate, digits = 2))`), as the posterior distributions of the standard deviations were almost completely overlapping (Figure 3).
 
@@
 The vertical lines are the medians of the posterior distributions.
-For each panel, the light-coloured distributions (i.e.
-green) are for the model with a shared trend where we did not account for prey speed (model II), and the darker-coloured distributions (i.e.
-purple) are for the model with a shared trend where we accounted for prey speed (model IV)."}
+The yellow distributions are for the model accounting only for the game duration (model I), the purple distributions are for the model accounting for game duration and the prey rank (model II), and the green distributions are for the model accounting for the game duration, prey rank, and prey speed (model V)."}
 

### 2025-06-15 code/code_figures/figure1.R
Windows-->updated fig1

@@ -119,8 +119,9 @@ tab5_b <- data.table(tab5_b[[1]])
 # Transform values --------------------------------------------------------
 
 # Back transform x-axis values
-sequence <- (seq(0, 500, 100) - mean(data$cumul_xp_pred))
+mean_xp <- mean(data$cumul_xp_pred)
 standev <- sd(data$cumul_xp_pred)
+sequence <- (seq(0, 500, 100) - mean_xp)
 scaled_breaks <- sequence / standev
 
 # List the tables
@@ -329,20 +330,21 @@ plot1_b <- ggplot(
 # Model 2 rank only --------------------------------------------------------
 
 # Predicted values from the GAMM
-predicted_values <- tab2_b$estimate__
+predicted_values2 <- tab2_b$estimate__
 
 # Define the x-axis range
-x <- tab2_b$Zcumul_xp
+x2 <- tab2_b$Zcumul_xp
 
 # Calculate the first derivative using finite differences
-dx <- mean(diff(x))
-derivatives <- diff(predicted_values) / dx
+dx2 <- mean(diff(x))
+derivatives2 <- diff(predicted_values2) / dx2
 
 # Establish the treshold at a value very close to 0 since its optimization
 threshold <- 0.0088
 # Find the point where the slope is close to zero
-stabilized_point <- x[which(abs(derivatives) <= threshold)][1]
+stabilized_point2 <- x2[which(abs(derivatives2) <= threshold)][1]
 
+(stabilized_point2 * standev) + mean_xp
 
 plot2_b <- ggplot(
   tab2_b,
@@ -362,7 +364,7 @@ plot2_b <- ggplot(
     size = 5
   ) +
   geom_vline(
-    xintercept = tab2_b[Zcumul_xp == stabilized_point]$Zcumul_xp,
+    xintercept = tab2_b[Zcumul_xp == stabilized_point2]$Zcumul_xp,
     lty = "dashed",
     color = "#440154"
   ) +
@@ -370,12 +372,12 @@ plot2_b <- ggplot(
     aes(label = paste(
       "y =",
       format(
-        round(tab2_b[Zcumul_xp == stabilized_point]$estimate__ / 4, digits = 2),
+        round(tab2_b[Zcumul_xp == stabilized_point2]$estimate__ / 4, digits = 2),
         nsmall = 2
       )
     ),
[…]

### 2025-06-15 manuscript/manuscript.Rmd
Windows--> updated full MS up to discussion

@@
 To achieve this, predators must hone their skills and gain knowledge through repeated and extensive practice.
-Prey can interfere with this process and hinder the acquisition of predator expertise by employing antipredator tactics to evade detection and pursuit.
+It is hypothesized that prey interfere with this process and hinder the acquisition of predator expertise by employing antipredator tactics to evade detection and pursuit.
 However, empirical evidence on how predators acquire expertise through repeated encounters with their prey remains limited, largely due to the challenges of monitoring direct interactions in the wild.
-Here, we use a virtual predator-prey system (the game *Dead by Daylight*) to investigate how experience shapes individual and population hunting success in human predators across repeated interactions with their prey.
-We show that predators optimized prey consumption as they gained experience, indicating that they acquired expertise through extensive practice.
-At the population-level, we found that more skillful prey impaired the acquisition of expertise by reducing hunting success.
-Prey skill was also an important mediator of this relationship at the individual level, driving differences among predators in the acquisition of expertise.
+Here, we use the game *Dead by Daylight*, a virtual predator-prey system where a human predator chases four human prey, to investigate how experience shapes individual and population hunting success in predators across repeated interactions with their prey.
+Our analyses reveal that the predator population should optimize prey consumption over repeated encounters with prey of average skill, indicating that they acquire expertise through extensive practice.
+We observed that more skillful prey impaired the acquisition of expertise by reducing hunting success.
+Prey skill was also an important mediator of expertise acquisition at the individual level, driving differences among predators in the relationship between hunting success and experience.
 More skillful prey moved faster, and predators were less successful against faster prey.
-Our results suggest that skilled prey use behaviours to outmanoeuver their predators, thereby mediating how predators acquire expertise at the population and individual level.
+Our results suggest that skilled prey use behaviours such as fast movement to outmanoeuver their predators, thereby mediating how predators acquire expertise at the population and individual level.
 
@@
 These processes are driven by changes in predation risk and prey selection across landscapes as predators adjust their strategy and select suitable habitats to hunt [@quevedo.etal2009;@pettorelli.etal2015;@labarge.etal2024;@schmitz2017].
-Central theory suggests that predators become more efficient hunters by acquiring expertise through practicing and learning the proper skills to locate, select, and capture their prey [@woo.etal2008;@wooster.etal2023;@labarge.etal2024].
+Central theory suggests that predators become more efficient hunters by acquiring expertise through practicing and learning the proper skills and tactics to locate, select, and capture their prey [@woo.etal2008;@wooster.etal2023;@labarge.etal2024].
 Therefore, differences in hunting success between predators may be attributed to differences in the acquisition of hunting expertise.
@@
 Through these processes, expert predators should have greater knowledge, better energy management, and acute motor skills that increase their chances of locating and capturing prey [@dukas2019].
-Theory predicts that predator cognition can maintain behavioural variation, shape prey phenotypic composition, and destabilize predator-prey systems [@kondoh2010;@skelhornandcandy2016;@kikuchi.etal2023].
-Therefore, individual differences among predators in their capacity to acquire expertise over time may be important for the ecological consequences of predation.
+Differences among predators in these attributes modulate behavioural variation and prey phenotypic composition, contributing in the dynamics of predator-prey systems [@kondoh2010;@skelhornandcandy2016;@kikuchi.etal2023].
 
-Prey use antipredator tactics such as camouflage to avoid detection and rapid escapes to evade capture [@walker.etal2005;@kelley.magurran2011;@herbert-read.etal2017].
-These strategies are hypothesized to drive differences in prey consumption among predators by disrupting the acquisition of hunting expertise [@wooster.etal2023].
-Experimental studies have shown that certain camouflage tactics of prey can impair expertise acquisition in humans and birds [@martin.etal2012;@troscianko.etal2013].
-For example, @troscianko.etal2018 found that disruptive colouration interfered with human subjects' ability to form search images, hindering improvements in detection times over repeated attempts.
-Antipredator behaviour may also hold the potential to hinder the acquisition of predator expertise.
+Prey respond to predators by using antipredator tactics such as camouflage to avoid detection and rapid escapes to evade capture [@walker.etal2005;@kelley.magurran2011;@herbert-read.etal2017].
+These strategies are hypothesized to drive differences in prey consumption among predators by disrupting their capacity to acquire hunting expertise [@wooster.etal2023].
+This has been supported by experimental studies showing that certain camouflage tactics of prey can impair expertise acquisition in humans and birds [@martin.etal2012;@troscianko.etal2013].
+For example, using a virtual experiment, @troscianko.etal2018 found in human subjects that disruptive colouration interfered with their ability to form search images, hindering improvements in detection times over repeated attempts.
+In that sense, antipredator behaviour may also hold the potential to hinder the acquisition of predator expertise.
 For instance, some predators may be limited in their capacity to develop the necessary attributes (e.g.
@@
 Moreover, prey antipredator behaviour can improve with experience through continued exposure to predators [@turner2006;@kelley.magurran2011;@lonnstedt.etal2012].
-Indeed, animal reintroduction programs show that individuals trained to recognize predators display more antipredator behaviours and are 1.5 times more likely to survive than predator-naive individuals [@tetzalf.etal2019].
+Indeed, animal reintroduction programs show that individuals trained to recognize predators display more antipredator behaviours and are 1.5 times more likely to survive than individuals that were never exposed to predators [@tetzalf.etal2019].
 To our knowledge, there is little empirical evidence showing links between prey skill, antipredator behaviour, and the acquisition of predator expertise in human and nonhuman predators, representing a significant gap in our understanding of predator-prey interactions.
@@
 Behaviour Interactive Inc.
-provided data that spanned six months of gameplay recorded for every player from `r toString(format(min(dat$date)))` to `r toString(format(max(dat$date)))` (game build 4.4.0 to 4.7.2).
+provided data that spanned six months of gameplay recorded for every player from `r toString(format(min(dat$date)))` to `r toString(format(max(dat$date)))` (game builds 4.4.0 to 4.7.2).
 We analyzed only matches where players did not know each other and were unable to communicate using voice-recording devices.
@@
 We then took the first 20 matches played by these players, including those from our population sample, and compared their mean hunting success using a Bayesian hierarchical linear model.
-We found that all four groups had similar success as predators in their first 20 matches (Appendix 1: Table S1 and Figure S1), suggesting an absence of bias due to data sampling.
+We found that all four groups had similar success as predators in their first 20 matches (Appendix 1: Table S1 and Figure S1), giving us confidence that we did not introduce a sampling bias.
 
-We collected the following information for every match in our population sample : the player's anonymous ID, the predator player's hunting success, the predator player's cumulated experience, and the mean rank and speed of the the group of prey that the predator player encountered.
+We collected the following information for every match in our population sample: the player's anonymous ID, the predator player's hunting success, the predator player's cumulated experience, and the mean rank and speed of the group of prey that the predator player encountered.
 We defined hunting success as the number of prey consumed during the match (min = `r toString(format(min(dat$hunting_success)))`, max = `r toString(format(max(dat$hunting_success)))`).
@@
 We used the mean rank of the prey as a proxy of their skill.
-The ranking system in *Dead by Daylight* varies between `r toString(format(min(dat$prey_avg_rank)))` and `r toString(format(max(dat$prey_avg_rank)))` (1 indicating the highest skill), and was implemented by the company to pair players in a match based on their skill (https://deadbydaylight.fandom.com/wiki/Rank).
+Player ranks vary between `r toString(format(min(dat$prey_avg_rank)))` and `r toString(format(max(dat$prey_avg_rank)))`, with a rank of 1 indicating the highest skill.
+The ranking system in *Dead by Daylight* was implemented by the company to pair players in a match based on their skill (https://deadbydaylight.fandom.com/wiki/Rank).
 The skill of a player increases based on their performance from match to match.
-While the ranking system represents an approximation of prey skill, it was the most readily available metric we could access to determine the skill of the prey group.
+While the ranking system represents an approximation of prey skill, it was the most readily available metric we could access to determine the skill of the prey group, as we did not have access to the prior matches of the prey.
 In addition, the pairing system is subject to variation depending on factors such as player availability, which can result in predator-prey groups with unbalanced skill.
@@
 We tested how predators developed their expertise by computing five Bayesian generalized additive mixed models (GAMM) with thin-plate regression splines, all of which estimated the relationship between hunting success and the predators' cumulative experience.
-We parametrized the models following the method of @pedersen.etal2019.
+We parametrised the models following the method of @pedersen.etal2019.
 For the first and second models (I and II), we assume that individual predators share a similar relationship between success and experience, but that this relationship can vary among them (e.g.
@@
 
-We computed the five models using a modified beta-binomial distribution.
-Hunting success was estimated as the number of prey consumed out of four, with probability of success drawn from a Beta distribution ($Beta(\mu_i, \phi)$) with mean ($\mu \in [0,1]$) and precision ($\phi > 0$) parameters.
+We configured the five models to use a modified beta-binomial distribution.
+Hunting success was modeled as the number of prey consumed out of four, with the probability of success estimated from a Beta distribution ($Beta(\mu_i, \phi)$) characterized by a mean ($\mu \in [0,1]$) and a precision ($\phi > 0$) parameter.
 We used a logit link function to estimate $\mu_i$ where $\mu_i = \frac{exp(\eta_i)}{exp(\eta_i) + 1}$ and $\eta_i$ is the linear predictor.
@@
 We controlled for the game duration, using it as a linear covariate in the model, and included the predator identity as a random effect.
-We used a weakly informative Gaussian prior on the slope of the game duration ($N(0,\ 0.5^{2})$) and a weakly informative negative Gaussian prior the slope of the prey rank ($N(-0.5,\ 1^{2})$) because prey of greater skill (i.e., rank 1 is the highest skill) should move faster.
+We used a weakly informative Gaussian prior on the slope of the game duration ($N(0,\ 0.5^{2})$) and a weakly informative negative Gaussian prior on the slope of the prey rank ($N(-0.5,\ 1^{2})$) because prey of greater skill (i.e., rank 1 is the highest skill) should move faster.
 Lastly, we used a weakly informative Gaussian prior on the intercept ($N(2,\ 0.5^{2})$) because prey should move at around two meters per second, and half-Gaussian priors on the standard deviation parameters ($N(0,\ 0.5^{2})$).
@@
 
-We parametrized the GAMMs to run four chains of 1500 iterations each, sampling one parameter value every four iterations.
-We set the first 500 iterations of each model as warm-ups.
+We parametrised the GAMMs to run four chains of 1500 iterations each, discarding the first 500 iterations of each chain as warm-up, and sampling one parameter value per iteration thereafter.
 Similarly, the linear model ran four chains of 1500 iterations each, sampling a parameter value at every iteration, and discarding the first 500 iterations.
-For GAMM model, we obtained 1000 posterior samples per parameter, and 4000 values for the linear model.
+We thus obtained 4000 posterior samples per parameter in each model.
 We assessed the convergence of the chains using trace plots, R-hat diagnostics with a threshold set between 1.00 and 1.02, and effective sample sizes (ESS) with a threshold of >100 [@vehtari.etal2021].
@@
 
-We fitted all models in R (version 4.4.0) using Hamiltonian Monte Carlo (HMC) sampling with the package "brms" version 2.22.12 [@burkner2017], an R front-end for the STAN software [@StanDevelopmentTeam2023], and "cmdstanr" version 0.9.0 [@gabry.cesnovar2021] as the back-end for parameter estimation (cmdstan installation version 2.36.0).
-For further details on how to reproduce our results, please consult the GitHub repository (https://github.com/quantitative-ecologist/predator-expertise).
+We fitted all models on a remote computer cluster (https://docs.alliancecan.ca/wiki/Cedar) using R (version 4.4.0).
+Model fitting was performed via Hamiltonian Monte Carlo (HMC) sampling with the "brms" package version 2.22.12 [@burkner2017], an R front-end for the STAN software [@StanDevelopmentTeam2023].
+We used "cmdstanr" version 0.9.0 [@gabry.cesnovar2021] as the computational back-end for parameter estimation, with CmdStan installation version 2.36.0.
+We provide a detailed workflow for reproducing our results in the following GitHub repository (https://github.com/quantitative-ecologist/predator-expertise).
 
@@
 
-We tested the hypothesis that prey skill and antipredator behaviour impairs the acquisition of expertise at the population level by visually comparing the global trends of model I (only controlling for game duration), model II (controlling for game duration and prey rank) and model V (controlling game duration, prey rank, and prey speed) (Figure 1).
-At the individual level, we tested our hypothesis that prey skill and antipredator behaviour generate differences among predators in expertise acquisition by comparing the individual-level variance parameters.
+We tested the hypothesis that prey skill and antipredator behaviour impairs the acquisition of expertise at the population level by comparing the global trends of model I (only controlling for game duration), model II (controlling for game duration and prey rank) and model V (controlling game duration, prey rank, and prey speed) using the finite differences method (Figure 1).
+This allowed us to locate the point at which hunting success was optimized.
+At the individual level, we tested the hypothesis that prey skill and antipredator behaviour generate differences among predators in expertise acquisition by comparing the individual-level variance parameters.
 Specifically, we compared the standard deviations of 1) the random intercepts (i.e.
@@
 
-Among models that shared the same covariate structure (i.e., model II and III; model IV and V), those that included a global smoother generally performed better at predicting hunting success compared than those without one (Table 1).
-While we detected differences among predators in the form and shape of the relationship between success and experience (Figure 1), this suggests that there is a common underlying structure in this relationship across the population.
+When comparing models with the same covariate structure (i.e., model II and III; model IV and V), those that included a global smoother generally performed better at predicting hunting success compared than those without one (Table 1).
+This improvement in performance suggests that there is a common underlying structure in the relationship between success and experience across the population despite differences among predators (Figure 1).
 Overall, model V, which included all prey features, was the best at predicting predator hunting success (Table 1).
-
-\newpage
 
@@
 no gain in expertise, Figure 1A).
-As we expected, accounting for the prey rank (model II) influenced the form of the relationship, resulting in a diminishing returns curve with predators optimizing their success after playing ~450 matches (Figure 1B).
-The predictions indicate that the population would improve their hunting success by ~40% from the first to the ~450$^{th}$ match by hunting prey with average skill (Figure 1B).
+As we expected, accounting for the prey rank (model II) influenced the form of the relationship, resulting in a diminishing returns curve with predators optimizing their success after playing 454 matches (Figure 1B).
+The predictions indicate that the population would improve their hunting success by ~40% from the first to the 454$^{th}$ match by hunting prey with average skill (Figure 1B).
 
-Controlling for prey speed did not alter the form of the relationship between success and experience at the population level.
-However, as predicted, the rate of expertise acquisition shifted slightly: the predator population optimized its hunting success earlier, shortly after 300 matches (Figure 1C).
+Controlling for prey speed did not qualitatively alter the form of the relationship between success and experience at the population level.
+However, the rate of expertise acquisition shifted slightly: the predator population optimized its hunting success earlier, after 323 matches (Figure 1C).
 We also found strong evidence of a negative relationship between hunting success and prey speed, indicating that hunting faster prey reduced the predator's chances of capture (Figure S2).
@@
 However, the extent of the variation in the relationship between success and experience was shaped by the prey's skill, as shown by the the posterior distributions of individual-level parameters in the model accounting for the rank of the prey (model II, Figure 2).
-Specifically, the standard deviations of individual intercepts (median model I = `r toString(round(sum1[2,]$Estimate, digits = 2))` vs median model II = `r toString(round(sum2[2,]$Estimate, digits = 2))`), slopes (median model I = `r toString(round(sum1[3,]$Estimate, digits = 2))` vs median model II = `r toString(round(sum2[3,]$Estimate, digits = 2))`), and wiggliness (median model I = `r toString(round(sum1[4,]$Estimate, digits = 2))` vs median model II = `r toString(round(sum2[4,]$Estimate, digits = 2))`) were all larger in model II than in model I.
+Specifically, the standard deviations of individual intercepts (median model I = `r toString(round(sum1[2,]$Estimate, digits = 2))` vs median model II = `r toString(round(sum2[2,]$Estimate, digits = 2))`), slopes (median model I = `r toString(round(sum1[3,]$Estimate, digits = 2))` vs median model II = `r toString(round(sum2[3,]$Estimate, digits = 2))`), and wiggliness (median model I = `r toString(round(sum1[4,]$Estimate, digits = 2))` vs median model II = `r toString(round(sum2[4,]$Estimate, digits = 2))`) were all larger in model II compared to model I.
 These results indicate that if predators hunted prey with average skill levels, differences among individual predators in expertise acquisition would be greater.
@@
 
-Using a virtual predator-prey system where we monitored predator hunting success across experience, we provide rare empirical support for the hypothesis that prey antipredator behaviour can impair the acquisition of hunting expertise.
+Using a virtual predator-prey system where we monitored predator hunting success across experience, we provide rare empirical support for the hypothesis that prey skill and behaviour can modulate the acquisition of hunting expertise.
 We show that the predator population increased its hunting success with experience.
-We found, however, important differences among individuals in expertise acquisition that were related to the speed of the prey encountered.
+While we could not show that prey skill directly impaired expertise acquisition, we found that the predator population optimized its success sooner if the prey moved at average speeds over repeated encounters.
+This is most likely due to the fact that faster prey are more difficult to hunt since predators had lower success against such prey, and that prey with greater skill also move at faster speeds.
+We also observed important differences among individuals in expertise acquisition that were apparent only after accounting for prey skill.
+While this most likely represents an artifact of the game pairing system, it highlights that many other factors can be at play which we further discuss below.
 
@@
 These observations are consistent with empirical studies of expertise in both humans and nonhuman animals [reviewed in @dukas2019].
-However, prey speed played a crucial role in shaping this pattern at the population level as encounters with faster prey resulted in lower hunting success (Figure S2).
-We previously showed in *Dead by Daylight* that faster movement is effective for prey to evade predation [@fraserfranco.etal2022], in agreement with studies in other animals [@walker.etal2005;@kelley.magurran2011;@martin.etal2022].
-We suspect that experienced prey may increasingly rely on this strategy, which could explain why the relationship between hunting success and experience was concave when we did not control for prey speed (model II, Figure 1A).
+Prey speed played a role in shaping expertise acquisition at the population level.
+First, encounters with faster prey resulted in lower hunting success (Figure S2).
+We previously showed in *Dead by Daylight* that faster movement is effective for prey to evade predation [@fraserfranco.etal2022], a pattern observed in studies involving other animals [@walker.etal2005;@kelley.magurran2011;@martin.etal2022].
+Second, we observed that if we controlled for the speed of the prey, predators optimized their hunting success sooner, after 323 matches played compared to 454 matches when we didn't account for it.
+
+prey of greater skill increasingly relied on this strategy, which could explain why the relationship between hunting success and experience was concave when we did not control for prey speed (model II, Figure 1A).
+
 Thus, our results suggest that predators can gain expertise and maintain success when they encounter prey that move at speeds lower than or closer to the population-average.

### 2025-06-23 manuscript/manuscript.Rmd
Linux --> changed the discussion, conclusions to finish. update refs

@@
 Using a virtual predator-prey system where we monitored predator hunting success across experience, we provide rare empirical support for the hypothesis that prey skill and behaviour can modulate the acquisition of hunting expertise.
-We show that the predator population increased its hunting success with experience.
-While we could not show that prey skill directly impaired expertise acquisition, we found that the predator population optimized its success sooner if the prey moved at average speeds over repeated encounters.
-This is most likely due to the fact that faster prey are more difficult to hunt since predators had lower success against such prey, and that prey with greater skill also move at faster speeds.
-We also observed important differences among individuals in expertise acquisition that were apparent only after accounting for prey skill.
-While this most likely represents an artifact of the game pairing system, it highlights that many other factors can be at play which we further discuss below.
+We show that the predator population is expected to increase its hunting success with experience if it hunts prey were of average skill.
+While we could not show that prey skill or behaviour directly impaired expertise acquisition, we found that the predator population would optimize its success sooner if the prey move at average speeds over repeated encounters.
+This is most likely due to faster prey being more difficult to hunt, which predators had lower success against.
+We also observed that prey with greater skill moved faster.
+At the individual level, predators displayed important differences in expertise acquisition that were apparent only after accounting for prey skill.
+While this most likely due the game pairing system, it highlights how prey encounters can shape the relationship between hunting success and experience in predators.
 
-Our results suggest that predators hone their hunting expertise through extensive practice.
-The predator population displayed an asymptotic relationship between experience and success, wherein initial gains in success were significant but gradually stabilized as experience accumulated.
-These observations are consistent with empirical studies of expertise in both humans and nonhuman animals [reviewed in @dukas2019].
-Prey speed played a role in shaping expertise acquisition at the population level.
+The matchmaking algorithm in *Dead by Daylight* is intended to pair players with similar skill levels within matches, and our results show that the system is effective, as the population maintained consistent hunting success across experience (Figure 1A).
+However, when we controlled for the prey's skill, the gain in expertise changed at both the population and individual levels: the relationship between success and experience became asymptotic, wherein initial gains in success were significant and gradually stabilized as experience accumulated.
+This suggests that an increase in hunting success would emerge if predators hunted prey of average skill.
+This pattern of expertise acquisition is consistent with empirical studies in both humans and nonhuman animals [reviewed in @dukas2019], but provides additional information on how it may change given the difficulty of the task.
+For example, the population may have had a reduced gain in expertise if the skill of the prey encountered was consistently higher.
+
+Prey behaviour also played a role in shaping expertise acquisition at the population level.
 First, encounters with faster prey resulted in lower hunting success (Figure S2).
 We previously showed in *Dead by Daylight* that faster movement is effective for prey to evade predation [@fraserfranco.etal2022], a pattern observed in studies involving other animals [@walker.etal2005;@kelley.magurran2011;@martin.etal2022].
-Second, we observed that if we controlled for the speed of the prey, predators optimized their hunting success sooner, after 323 matches played compared to 454 matches when we didn't account for it.
-
-prey of greater skill increasingly relied on this strategy, which could explain why the relationship between hunting success and experience was concave when we did not control for prey speed (model II, Figure 1A).
-
-Thus, our results suggest that predators can gain expertise and maintain success when they encounter prey that move at speeds lower than or closer to the population-average.
-
-Prey speed also mediated differences among predator players in the acquisition of expertise, suggesting that individual predators varied in their capacity to adjust to challenging prey.
-Animals are expected to have limited attention spans, which restricts diet choice and the formation of search images [@dukas.etal2001].
-Hunting faster prey demands specialized cognitive abilities and coordination that are energetically costly [@kelley.magurran2011].
-Thus, predators that failed to develop counter-strategies for detecting or chasing faster prey were likely at a disadvantage.
+Prey of greater skill increasingly relied on this strategy, and predators were less successful against skillful prey, suggesting that the effect of prey skill on the gain in expertise by predators was in part mediated by changes in prey behaviour following previous predator encounters[@tetzalf.etal2019].
+Second, we observed that predators optimized their hunting success sooner when we controlled for the prey's speed — after 323 matches played, compared to 454 matches when prey speed was not accounted for.
 Parallel observations have been outlined in studies of prey camouflage strategies.
 For example, @troscianko.etal2018 showed in a computer experiment involving humans that disruptive camouflage was efficient at preventing the acquisition of expertise during search image formation.
-Human subjects exposed to a restricted set of strategies were also less efficient compared to those exposed to a variety of strategies.
-Therefore, our observations show that prey antipredator behaviour can also impair predator expertise acquisition.
+Hence, our analyses provide evidence that prey antipredator behaviour could mediate and potentially impair the acquisition of hunting expertise in predators at the population level.
 
-Despite adjusting for the prey's speed (i.e.
-model V), noticeable differences in expertise acquisition among predators persited.
-One possible explanation is that the predators' hunting tactics may indirectly shape their own expertise through changes in prey behaviour.
-Predators tend to specialize as cursorial or ambush hunters in *Dead by Daylight* [@fraserfranco.etal2022].
-Consequently, those employing a cursorial tactic may push prey to move faster, which, in turn, may hinder their own expertise acquisition due to the increased difficulty of hunting such prey.
-An alternative explanation is that longer time intervals between hunting events are hypothesized to hinder or delay the acquisition of expertise because individuals may forget information when delays are longer [@endler1991;@wright.etal2022].
-For example, a predator that played 300 matches in six months might forget critical information related to prey detection or escape patterns compared to one that played 300 matches in six days.
-While there is unequivoqual evidence that many predator species can learn quickly how to be efficient hunters, the role of the frequency of interactions remains unclear [@wooster.etal2023].
-Therefore, investigating the impact of such time lags in future analyses may reveal important insights into the outcome of predator-prey interactions.
+Our analyses also revealed important differences among individual predators in the acquisition of expertise, indicating that predators varied in their capacity to improve over time.
+However, none of the prey features under study appeared to influence among-individual differences in expertise.
+The large indivual variation in expertise acquisition appeared when we controlled for the prey skill and persisted when we controlled for the prey's speed, suggesting that it would be observed even if predators hunted prey of average skill and speed.
+While this may represent an artifact of the game's pairing system, it reveals the potential role of other ecological factors.
+For instance, animals have limited attention spans when they hunt, which restricts diet choice and the formation of search images when they need to focus on multiple prey simultaneously [@dukas.etal2001;troscianko.etal2018].
+Chasing prey demands specialized cognitive abilities and coordination that are energetically costly [@kelley.magurran2011].
+Hence, individuals with neurological or physical attributes allowing them to better concentrate on the hunting task may have shown greater improvement in success over time [@dukas2019].
+Alternatively, predators that failed to develop counter-strategies to other prey antipredator behaviours were likely at a disadvantage.
+A study has shown that prey in *Dead by Daylight* display altruistic behaviours that increase their chances of escaping predation [@cere.etal2024].
+Predators may also shape their own expertise via their hunting tactic.
+For example, faster hunters may push their prey to move faster, which, in turn, may hinder their own expertise acquisition due to the increased difficulty of hunting such prey.
+Lastly, longer time intervals between hunting events are hypothesized to hinder or delay the acquisition of expertise because of information loss [@endler1991;@wright.etal2022;wooster.etal2023].
+For example, a predator that played 300 matches within six months might forget critical information related to prey detection or escape patterns compared to one that played 300 matches within six weeks.
+Therefore, investigating the impact of the aforementioned elements may reveal important insights into the outcome of predator-prey interactions, but integrating them into a unified framework remains challenging due to computational constraints and limitations in current modeling approaches.
+
+## Conclusions
+
+We found support for the hypothesis that prey antipredator behaviour can drive the acquisition of hunting expertise using a human predator population in the game *Dead by Daylight*.
+However, the effect of antipredator behaviour was observed only at the population level.
+While there were important differences among individuals in expertise acquisition, we did not find evidence that they were mediated by either the prey's skill or prey movement behaviour.
+
+Future analyses should investigate how antipredator tactics developp with experience, as it may reveal important insights on the eco-evolutionary dynamics of predator-prey interactions.
+
+Our study demonstrating that prey antipredator behaviour can impair the acquisition of hunting expertise adds to a growing body of research showing how virtual systems can be used to test hypotheses on ecological interactions [@beauchamp2020;@cere.etal2021;@fraserfranco.etal2022;@lymbery.etal2023;@santostefano.etal2024].
+
+We therefore hope that our study will inspire more collaborations between scientists and the videogame industry to tackle fundamental questions in ecology.
+
 Another potential reason for the persistent differences among individuals is that neither the predator nor prey players' lives are at stake in the game.
@@
 
-## Conclusions
-
-We found support of our hypothesis that prey antipredator behaviour drives individual differences in expertise acquisition in a human predator population in the game *Dead by Daylight*.
-Future analyses should investigate how antipredator tactics developp with experience, as it may reveal important insights on the eco-evolutionary dynamics of predator-prey interactions.
-Our study demonstrating that prey antipredator behaviour can impair the acquisition of hunting expertise adds to a growing body of research showing how virtual systems can be used to test hypotheses on ecological interactions [@beauchamp2020;@cere.etal2021;@fraserfranco.etal2022;@lymbery.etal2023;@santostefano.etal2024].
-We therefore hope that our study will inspire more collaborations between scientists and the videogame industry to tackle fundamental questions in ecology.
-
 # REFERENCES {.unnumbered}

### 2025-06-24 manuscript/manuscript.Rmd
linux--> update title and abstract

@@
 ---
-title: "Prey skill shapes the acquisition of predator expertise in a virtual bi-trophic system"
+title: "Implications of prey skill and antipredator behaviour in the acquisition of predator expertise: a virtual ecological study"
 output: 
@@
 To achieve this, predators must hone their skills and gain knowledge through repeated and extensive practice.
-It is hypothesized that prey interfere with this process and hinder the acquisition of predator expertise by employing antipredator tactics to evade detection and pursuit.
+It is hypothesized that prey can interfere in this process and hinder the acquisition of predator expertise by employing antipredator tactics to evade detection and pursuit.
 However, empirical evidence on how predators acquire expertise through repeated encounters with their prey remains limited, largely due to the challenges of monitoring direct interactions in the wild.
-Here, we use the game *Dead by Daylight*, a virtual predator-prey system where a human predator chases four human prey, to investigate how experience shapes individual and population hunting success in predators across repeated interactions with their prey.
-Our analyses reveal that the predator population should optimize prey consumption over repeated encounters with prey of average skill, indicating that they acquire expertise through extensive practice.
-We observed that more skillful prey impaired the acquisition of expertise by reducing hunting success.
-Prey skill was also an important mediator of expertise acquisition at the individual level, driving differences among predators in the relationship between hunting success and experience.
-More skillful prey moved faster, and predators were less successful against faster prey.
-Our results suggest that skilled prey use behaviours such as fast movement to outmanoeuver their predators, thereby mediating how predators acquire expertise at the population and individual level.
+Here, we use the game *Dead by Daylight*, a virtual predator-prey system where a human predator chases four human prey, to investigate how experience shapes population and individual hunting success in predators across repeated interactions with their prey.
+We observed that predators were less successful when hunting faster and more skillful prey.
+As the prey gained skill, they also moved faster.
+Our analyses reveal that the predator population should optimize prey consumption over repeated encounters with prey of average skill, and optimize their success sooner if they consistently encounter prey moving at average speeds.
+However, both prey features did not mediate differences in expertise acquisition among individual predators.
+
+Our results suggest that skilled prey use behaviours such as fast movement to outmanoeuver their predators, thereby impacting how predators acquire expertise at the population level.
+
+impaired the acquisition of expertise by reducing hunting success
 

### 2025-07-01 manuscript/manuscript.Rmd
Windows --> manuscript edits complete + update README

@@
 Predation is a fundamental biological process acting as an agent of evolutionary change that regulates prey populations, limits the spread of diseases, and mediates energy flow across trophic levels [@hairston1993;@abrams2000;@ripple.etal2014;@wirsing.etal2021].
-These processes are driven by changes in predation risk and prey selection across landscapes as predators adjust their strategy and select suitable habitats to hunt [@quevedo.etal2009;@pettorelli.etal2015;@labarge.etal2024;@schmitz2017].
+These processes are driven by changes in predation risk and prey selection across landscapes as predators select suitable habitats to hunt and adjust their strategies accordingly  [@quevedo.etal2009;@pettorelli.etal2015;@labarge.etal2024;@schmitz2017].
 Central theory suggests that predators become more efficient hunters by acquiring expertise through practicing and learning the proper skills and tactics to locate, select, and capture their prey [@woo.etal2008;@wooster.etal2023;@labarge.etal2024].
-Therefore, differences in hunting success between predators may be attributed to differences in the acquisition of hunting expertise.
-Unraveling how this expertise emerges remains a missing, but essential, link to assess how foraging behaviour shapes prey consumption during predator-prey interactions.
+This raises the hypothesis that differences in hunting success between predators may be attributed to differences in the capacity to acquire hunting expertise.
+Hence, unraveling how this expertise emerges remains a missing, but essential, link to assess how foraging behaviour shapes prey consumption during predator-prey interactions.
 
@@
 Through these processes, expert predators should have greater knowledge, better energy management, and acute motor skills that increase their chances of locating and capturing prey [@dukas2019].
-Differences among predators in these attributes modulate behavioural variation and prey phenotypic composition, contributing in the dynamics of predator-prey systems [@kondoh2010;@skelhornandcandy2016;@kikuchi.etal2023].
+Differences among predators in these attributes modulate behavioural variation and prey phenotypic composition, contributing to the dynamics of predator-prey systems [@kondoh2010;@skelhornandcandy2016;@kikuchi.etal2023].
 
@@
 For example, using a virtual experiment, @troscianko.etal2018 found in human subjects that disruptive colouration interfered with their ability to form search images, hindering improvements in detection times over repeated attempts.
-In that sense, antipredator behaviour may also hold the potential to hinder the acquisition of predator expertise.
+Given these observations, antipredator behaviour may also hold the potential to hinder the acquisition of predator expertise.
 For instance, some predators may be limited in their capacity to develop the necessary attributes (e.g.
-physical, physiological, neurological) for fast-paced hunting, which could impair their acquisition of hunting expertise if they chase faster prey.
-Moreover, prey antipredator behaviour can improve with experience through continued exposure to predators [@turner2006;@kelley.magurran2011;@lonnstedt.etal2012].
+physical, physiological, neurological) for fast-paced hunting, which could impair their acquisition of hunting expertise if they hunt faster prey.
+Prey antipredator behaviour can also improve with experience through continued exposure to predators [@turner2006;@kelley.magurran2011;@lonnstedt.etal2012].
 Indeed, animal reintroduction programs show that individuals trained to recognize predators display more antipredator behaviours and are 1.5 times more likely to survive than individuals that were never exposed to predators [@tetzalf.etal2019].
@@
 
-In this study, we assess how repeated encounters with prey shapes predator hunting success using data from players in *Dead by Daylight*.
-We quantify expertise acquisition as the relationship between hunting success (i.e.
-the probability of capturing all prey) and repeated experience (i.e.
-cumulated matches).
+In this study, we assess how predator expertise emerges over repeated encounters with prey using data from players in *Dead by Daylight*.
+We quantify expertise acquisition as the relationship between hunting success (i.e., prey capture) and repeated experience (i.e., cumulated matches).
 First, we test the hypothesis that predator success will increase with experience up to some point at which it will stabilize [@dukas2019].
-However, we hypothesize that prey skill and speed will influence predator expertise acquisition by modulating the rate of gain in expertise at the population level.
+However, we hypothesize that prey speed will influence predator expertise acquisition by modulating the rate of gain in expertise at the population level.
 Second, we test the hypothesis that prey with more skill should have improved antipredator behaviour, and predict that prey with greater skills will move at faster speeds.
-Third, if prey skill and speed influence hunting success at the population level, then we expect that the acquisition of predator expertise will vary among individuals given these two prey features.
+Third, if prey speed influences hunting success at the population level, then we expect that the acquisition of predator expertise will vary among individuals.
 
@@
 
-*Dead by Daylight* is a survival asymmetric (i.e.
-gameplay mechanics differ between two groups) multiplayer online game  developed by Behaviour Interactive Inc., in which players can play either as predators or prey.
+*Dead by Daylight* is a survival asymmetric (i.e.,, gameplay mechanics differ between two groups) multiplayer online game  developed by Behaviour Interactive Inc., in which players can play either as predators or prey.
 Every match includes only one predator and four prey.
@@
 A skill-based matchmaking algorithm determines the composition of the predator and prey group in a match.
-A match ends when the predator kills all the prey available (i.e.
-that have not escaped), or when the last remaining prey escapes the virtual environment.
+A match ends when the predator kills all the prey available (i.e., that have not escaped), or when the last remaining prey escapes the virtual environment.
 
@@
 We analyzed only matches where players did not know each other and were unable to communicate using voice-recording devices.
-We filtered any matches where players were inactive, such as when mean distances travelled per second (i.e.
-speed) were equal to, or very close to, zero.
+We filtered any matches where players were inactive, such as when mean distances travelled per second (i.e., speed) were equal to, or very close to, zero.
 Moreover, we used our knowledge of the game to remove any matches where players were potentially hacking, or not playing the game as intended.
@@
 
-We tested the hypothesis that prey skill and antipredator behaviour impairs the acquisition of expertise at the population level by comparing the global trends of model I (only controlling for game duration), model II (controlling for game duration and prey rank) and model V (controlling game duration, prey rank, and prey speed) using the finite differences method (Figure 1).
+We tested the hypothesis that antipredator behaviour impairs the acquisition of predator expertise at the population level by comparing the global trends of model I (only controlling for game duration), model II (controlling for game duration and prey rank) and model V (controlling game duration, prey rank, and prey speed) using the finite differences method (Figure 1).
 This allowed us to locate the point at which hunting success was optimized.
-At the individual level, we tested the hypothesis that prey skill and antipredator behaviour generate differences among predators in expertise acquisition by comparing the individual-level variance parameters.
-Specifically, we compared the standard deviations of 1) the random intercepts (i.e.
-mean differences in hunting success), 2) the random slopes (i.e.
-linear component relating hunting success with experience), and 3) the curve wiggliness (i.e.
-nonlinear component relating hunting success with experience).
+At the individual level, we tested the hypothesis that antipredator behaviour generate differences among predators in expertise acquisition by comparing the individual-level variance parameters.
+Specifically, we compared the standard deviations of 1) the random intercepts (i.e., mean differences in hunting success), 2) the random slopes (i.e., linear component relating hunting success with experience), and 3) the curve wiggliness (i.e., nonlinear component relating hunting success with experience).
 
@@
 
-Models in which we did not account for the prey rank resulted in almost no change in hunting success with experience for the average individual (i.e.
-no gain in expertise, Figure 1A).
-As we expected, accounting for the prey rank (model II) influenced the form of the relationship, resulting in a diminishing returns curve with predators optimizing their success after playing 454 matches (Figure 1B).
-The predictions indicate that the population would improve their hunting success by ~40% from the first to the 454$^{th}$ match by hunting prey with average skill (Figure 1B).
-
-Controlling for prey speed did not qualitatively alter the form of the relationship between success and experience at the population level.
-However, the rate of expertise acquisition shifted slightly: the predator population optimized its hunting success earlier, after 323 matches (Figure 1C).
-We also found strong evidence of a negative relationship between hunting success and prey speed, indicating that hunting faster prey reduced the predator's chances of capture (Figure S2).
-Moreover, prey groups composed of individuals with greater skill moved faster than those composed of individuals with lower skill (Figure 2).
+As we expected, models in which we did not account for the prey rank resulted in almost no change in hunting success with experience for the average individual (i.e., no gain in expertise, Figure 1A) due to the matchmaking algorithm pairing predators and prey with similar skill levels.
+Accounting for the prey rank (model II) influenced the form of the relationship, revealing a diminishing returns curve with predators optimizing their success after playing 454 matches (Figure 1B).
+The predictions indicate that the population would improve their hunting success by ~40% from the first to the 454$^{th}$ match by hunting prey of average skill (Figure 1B).
 
@@
 Median posterior predictions of the acquisition of predator hunting expertise.
-The predators' hunting success (i.e.
-the probability of consuming the four prey) is on the y-axis, and the predators' cumulative experience (i.e.
-the number of matches played before each observation) is on the x-axis.
+The predators' hunting success (i.e., the probability of consuming the four prey) is on the y-axis, and the predators' cumulative experience (i.e., the number of matches played before each observation) is on the x-axis.
 Panels A to C show the acquisition of expertise for the average individual with the vertical dashed lines on the left representing the lowest predicted values.
@@
 
-## Acquisition of expertise at the individual level
-
-Our results show that individual predators varied in their acquisition of expertise, regardless of whether we controlled for prey features (Figure 1 D-F and Figure 3).
-However, the extent of the variation in the relationship between success and experience was shaped by the prey's skill, as shown by the the posterior distributions of individual-level parameters in the model accounting for the rank of the prey (model II, Figure 2).
-Specifically, the standard deviations of individual intercepts (median model I = `r toString(round(sum1[2,]$Estimate, digits = 2))` vs median model II = `r toString(round(sum2[2,]$Estimate, digits = 2))`), slopes (median model I = `r toString(round(sum1[3,]$Estimate, digits = 2))` vs median model II = `r toString(round(sum2[3,]$Estimate, digits = 2))`), and wiggliness (median model I = `r toString(round(sum1[4,]$Estimate, digits = 2))` vs median model II = `r toString(round(sum2[4,]$Estimate, digits = 2))`) were all larger in model II compared to model I.
-These results indicate that if predators hunted prey with average skill levels, differences among individual predators in expertise acquisition would be greater.
+Controlling for prey speed did not qualitatively alter the form of the relationship between success and experience at the population level.
+However, the rate of expertise acquisition shifted slightly: the predator population optimized its hunting success earlier, after 323 matches (Figure 1C).
+We also found strong evidence of a negative relationship between hunting success and prey speed, indicating that hunting faster prey reduced the predator's chances of capture (Figure S2).
+Moreover, prey groups composed of individuals with greater skill moved faster than those composed of individuals with lower skill (Figure 2).
 
@@
 
+## Acquisition of expertise at the individual level
+
+Our results show that individual predators varied in their acquisition of expertise, regardless of whether we controlled for prey features (Figure 1 D-F and Figure 3).
+However, the extent of the variation in the relationship between success and experience was shaped by the prey's skill, as shown by the the posterior distributions of individual-level parameters in the model accounting for the rank of the prey (model II, Figure 2).
+Specifically, the standard deviations of individual intercepts (median model I = `r toString(round(sum1[2,]$Estimate, digits = 2))` vs median model II = `r toString(round(sum2[2,]$Estimate, digits = 2))`), slopes (median model I = `r toString(round(sum1[3,]$Estimate, digits = 2))` vs median model II = `r toString(round(sum2[3,]$Estimate, digits = 2))`), and wiggliness (median model I = `r toString(round(sum1[4,]$Estimate, digits = 2))` vs median model II = `r toString(round(sum2[4,]$Estimate, digits = 2))`) were all larger in model II compared to model I.
+Controlling for prey rank thus reveals how predators may differ expertise acquisition if they hunt prey that are not of the same skill level as them.
+
 Prey speed did not influence among-individual differences in average hunting success (median model II = `r toString(round(sum2[2,]$Estimate, digits = 2))` vs median model V = `r toString(round(sum5[2,]$Estimate, digits = 2))`), nor in expertise acquisition (median model II = `r toString(round(sum2[3,]$Estimate, digits = 2))` vs median model V = `r toString(round(sum5[3,]$Estimate, digits = 2))`), as the posterior distributions of the standard deviations were almost completely overlapping (Figure 3).
@@
 The intercept and slope standard deviations refer to the linear components of the estimated relationship between hunting success and cumulative experience.
-The standard deviation of the wiggliness parameter refers to the shape of the curves (i.e.
-nonlinear component).
+The standard deviation of the wiggliness parameter refers to the shape of the curves (i.e., nonlinear component).
 The vertical lines are the medians of the posterior distributions.
@@
 
-Using a virtual predator-prey system where we monitored predator hunting success across experience, we provide rare empirical support for the hypothesis that prey skill and behaviour can modulate the acquisition of hunting expertise.
+Using a virtual predator-prey system where we monitored predator hunting success across experience, we provide rare empirical support for the hypothesis that prey behaviour can modulate the acquisition of hunting expertise at the population level.
 We show that the predator population is expected to increase its hunting success with experience if it hunts prey were of average skill.
-While we could not show that prey skill or behaviour directly impaired expertise acquisition, we found that the predator population would optimize its success sooner if the prey move at average speeds over repeated encounters.
+While we could not show that prey behaviour directly impaired expertise acquisition, we found that the predator population would optimize its success sooner if the prey moved at average speeds over repeated encounters.
 This is most likely due to faster prey being more difficult to hunt, which predators had lower success against.
@@
 At the individual level, predators displayed important differences in expertise acquisition that were apparent only after accounting for prey skill.
-While this most likely due the game pairing system, it highlights how prey encounters can shape the relationship between hunting success and experience in predators.
+While this most likely due the game pairing system, it highlights how prey can shape the predator hunting success over time.
 
@@
 This pattern of expertise acquisition is consistent with empirical studies in both humans and nonhuman animals [reviewed in @dukas2019], but provides additional information on how it may change given the difficulty of the task.
-For example, the population may have had a reduced gain in expertise if the skill of the prey encountered was consistently higher.
+For example, the population may have had a slower gain in expertise if the skill of the prey encountered was consistently higher.
 
@@
 For example, @troscianko.etal2018 showed in a computer experiment involving humans that disruptive camouflage was efficient at preventing the acquisition of expertise during search image formation.
-Hence, our analyses provide evidence that prey antipredator behaviour could mediate and potentially impair the acquisition of hunting expertise in predators at the population level.
+Hence, our analyses provide evidence that prey antipredator behaviour could mediate the acquisition of hunting expertise in predators by increasing or slowing its rate at the population level.
 
 Our analyses also revealed important differences among individual predators in the acquisition of expertise, indicating that predators varied in their capacity to improve over time.
-However, none of the prey features under study appeared to influence among-individual differences in expertise.
-The large indivual variation in expertise acquisition appeared when we controlled for the prey skill and persisted when we controlled for the prey's speed, suggesting that it would be observed even if predators hunted prey of average skill and speed.
-While this may represent an artifact of the game's pairing system, it reveals the potential role of other ecological factors.
-For instance, animals have limited attention spans when they hunt, which restricts diet choice and the formation of search images when they need to focus on multiple prey simultaneously [@dukas.etal2001;troscianko.etal2018].
+However, prey speed did not influence among-individual differences in expertise.
+The large indivual variation in expertise acquisition emerged when we controlled for the prey skill and persisted when we controlled for the prey's speed, suggesting that it would be observed even if predators hunted prey of average speed.
+This suggests that other proximal or ecological factors may be at play.
+For example, predators have limited attention spans that restrict diet choice and the formation of search images when they need to focus on hunting multiple prey simultaneously [@dukas.etal2001;@troscianko.etal2018].
 Chasing prey demands specialized cognitive abilities and coordination that are energetically costly [@kelley.magurran2011].
 Hence, individuals with neurological or physical attributes allowing them to better concentrate on the hunting task may have shown greater improvement in success over time [@dukas2019].
-Alternatively, predators that failed to develop counter-strategies to other prey antipredator behaviours were likely at a disadvantage.
-A study has shown that prey in *Dead by Daylight* display altruistic behaviours that increase their chances of escaping predation [@cere.etal2024].
-Predators may also shape their own expertise via their hunting tactic.
-For example, faster hunters may push their prey to move faster, which, in turn, may hinder their own expertise acquisition due to the increased difficulty of hunting such prey.
-Lastly, longer time intervals between hunting events are hypothesized to hinder or delay the acquisition of expertise because of information loss [@endler1991;@wright.etal2022;wooster.etal2023].
+Alternatively, predators may also shape their own expertise via their hunting tactic.
+For example, faster hunters may push their prey to move faster and display altruistic behaviours, which, in turn, may hinder their own expertise acquisition due to the increased difficulty of hunting such prey [@fraserfranco.etal2022;@cere.etal2024].
+Lastly, longer time intervals between hunting events are hypothesized to hinder or delay the acquisition of expertise due to information loss [@endler1991;@wright.etal2022;@wooster.etal2023].
 For example, a predator that played 300 matches within six months might forget critical information related to prey detection or escape patterns compared to one that played 300 matches within six weeks.
@@
 While there were important differences among individuals in expertise acquisition, we did not find evidence that they were mediated by either the prey's skill or prey movement behaviour.
-
-Future analyses should investigate how antipredator tactics developp with experience, as it may reveal important insights on the eco-evolutionary dynamics of predator-prey interactions.
-
-Our study demonstrating that prey antipredator behaviour can impair the acquisition of hunting expertise adds to a growing body of research showing how virtual systems can be used to test hypotheses on ecological interactions [@beauchamp2020;@cere.etal2021;@fraserfranco.etal2022;@lymbery.etal2023;@santostefano.etal2024].
-
-We therefore hope that our study will inspire more collaborations between scientists and the videogame industry to tackle fundamental questions in ecology.
-
-Another potential reason for the persistent differences among individuals is that neither the predator nor prey players' lives are at stake in the game.
+Future analyses should integrate physical and behavioural antipredator tactics into a unified framework to assess their influence on the acquisition of predator expertise, which may reveal important insights on the eco-evolutionary dynamics of predator-prey interactions.
+One limitation and criticism of using videogames to study behaviour which may partly explain the persistence of differences among individuals in expertise is that neither the predator nor the prey players' lives are at stake in the game.
 As a result, emerging patterns may be driven more by the players' motivation to win rather than "true" survival.
-For example, some players could experiment with the game out of boredom, which could also contribute in shaping how expertise is honed in this particular system.
+Indeed, some players may experiment with the game out of boredom, which could contribute in shaping how expertise is honed in this particular system.
+Yet, motivation and perseverance have been outlined as core components that increase the gain in expertise in human and nonhuman hunters [@dukas2019].
+We believe that failing to account for motivation represents a limitation of our study more than in the use of videogames to study ecological interactions.
+Therefore, our demonstration of how prey antipredator behaviour can mediate the acquisition of hunting expertise adds to a growing body of research showing how virtual systems can be used to test hypotheses on ecological interactions [@beauchamp2020;@cere.etal2021;@fraserfranco.etal2022;@lymbery.etal2023;@santostefano.etal2024].
+We hope that this will inspire more collaborations between scientists and the videogame industry to tackle fundamental questions in ecology.
 

### 2025-08-19 manuscript/manuscript.Rmd
updated intro and results adding space covered

@@
   file.path(
-    path, "FraserFrancoetal2023-data.csv"
+    path, "FraserFrancoetal2025-data.csv"
   ),
@@
     "hunting_success", "pred_speed", "prey_avg_speed", 
-    "prey_avg_rank"
+    "prey_avg_rank", "prey_avg_amount_tiles_visited"
   )
@@
 # ABSTRACT {.unnumbered}
+
+**à compléter**
 
@@
 
-Predation is a fundamental biological process acting as an agent of evolutionary change that regulates prey populations, limits the spread of diseases, and mediates energy flow across trophic levels [@hairston1993;@abrams2000;@ripple.etal2014;@wirsing.etal2021].
-These processes are driven by changes in predation risk and prey selection across landscapes as predators select suitable habitats to hunt and adjust their strategies accordingly  [@quevedo.etal2009;@pettorelli.etal2015;@labarge.etal2024;@schmitz2017].
-Central theory suggests that predators become more efficient hunters by acquiring expertise through practicing and learning the proper skills and tactics to locate, select, and capture their prey [@woo.etal2008;@wooster.etal2023;@labarge.etal2024].
+Predation is an agent of evolutionary change that regulates prey populations, limits the spread of diseases, and mediates energy flow across trophic levels [@hairston1993;@abrams2000;@ripple.etal2014;@wirsing.etal2021].
+These processes are driven by changes in predation risk and prey capture across landscapes as predators select suitable habitats to hunt and adjust their strategies accordingly  [@quevedo.etal2009;@pettorelli.etal2015;@labarge.etal2024;@schmitz2017].
+Theory suggests that predators become more efficient hunters by acquiring expertise through practicing and learning the proper skills and tactics to locate, select, and capture their prey [@woo.etal2008;@wooster.etal2023;@labarge.etal2024].
 This raises the hypothesis that differences in hunting success between predators may be attributed to differences in the capacity to acquire hunting expertise.
-Hence, unraveling how this expertise emerges remains a missing, but essential, link to assess how foraging behaviour shapes prey consumption during predator-prey interactions.
+Hence, unraveling how expertise emerges is essential to better assess how foraging behaviour shapes prey consumption during predator-prey interactions.
 
@@
 These strategies are hypothesized to drive differences in prey consumption among predators by disrupting their capacity to acquire hunting expertise [@wooster.etal2023].
-This has been supported by experimental studies showing that certain camouflage tactics of prey can impair expertise acquisition in humans and birds [@martin.etal2012;@troscianko.etal2013].
+This has been supported by experimental studies showing that prey camouflage can impair expertise acquisition in humans and birds [@martin.etal2012;@troscianko.etal2013].
 For example, using a virtual experiment, @troscianko.etal2018 found in human subjects that disruptive colouration interfered with their ability to form search images, hindering improvements in detection times over repeated attempts.
-Given these observations, antipredator behaviour may also hold the potential to hinder the acquisition of predator expertise.
-For instance, some predators may be limited in their capacity to develop the necessary attributes (e.g.
-physical, physiological, neurological) for fast-paced hunting, which could impair their acquisition of hunting expertise if they hunt faster prey.
-Prey antipredator behaviour can also improve with experience through continued exposure to predators [@turner2006;@kelley.magurran2011;@lonnstedt.etal2012].
-Indeed, animal reintroduction programs show that individuals trained to recognize predators display more antipredator behaviours and are 1.5 times more likely to survive than individuals that were never exposed to predators [@tetzalf.etal2019].
-To our knowledge, there is little empirical evidence showing links between prey skill, antipredator behaviour, and the acquisition of predator expertise in human and nonhuman predators, representing a significant gap in our understanding of predator-prey interactions.
+Beyond prey camouflage, there is limited evidence to suggest that antipredator behaviour hinders the acquisition of predator expertise.
+Yet, some predators may be limited in their capacity to develop the necessary attributes (e.g.
+physical, physiological, neurological) for fast-paced hunting, which could impair their acquisition of hunting expertise when facing faster prey.
+Prey antipredator behaviour can also improve through continued exposure to predators [@turner2006;@kelley.magurran2011;@lonnstedt.etal2012].
+For instance, prey alter their activity patterns under increasing predation risk [@wirsing2021].
+Animal reintroduction programs show that individuals trained to recognize predators display more antipredator behaviours and are 1.5 times more likely to survive than individuals that were never exposed to predators [@tetzalf.etal2019].
+Hence, encountering prey with greater skill or experience may also impair expertise acquisition in predators.
+To our knowledge, the relationships among prey skill, antipredator behaviour, and the acquisition of expertise by human and nonhuman hunters remain unclear, representing a significant gap in our understanding of predator-prey interactions.
 
-A recurring challenge impeding research on predator-prey behavioural interactions is the need to collect data simultaneously on both the predator and the prey.
+A recurring challenge impeding research on predator-prey behavioural interactions is the need to collect data simultaneously on both the predator and the prey over successive interactions.
 Here we mitigate these challenges by using *Dead by Daylight* as our study system, a videogame where four prey players must forage for resources while avoiding predation by a fifth player.
@@
 First, we test the hypothesis that predator success will increase with experience up to some point at which it will stabilize [@dukas2019].
-However, we hypothesize that prey speed will influence predator expertise acquisition by modulating the rate of gain in expertise at the population level.
-Second, we test the hypothesis that prey with more skill should have improved antipredator behaviour, and predict that prey with greater skills will move at faster speeds.
-Third, if prey speed influences hunting success at the population level, then we expect that the acquisition of predator expertise will vary among individuals.
+However, we hypothesize that prey speed and space coverage will influence predator expertise acquisition by modulating the rate of gain in expertise at the population level.
+Second, we test the hypothesis that prey with more skill should have improved antipredator behaviour, and we predict that prey with greater predator-avoidance skills will move at faster speeds and cover more space in the virtual environment.
+Third, if prey antipredator behavior influences the hunting success of predators at the population level, then we expect that the acquisition of predator expertise will vary among individuals.
 
@@
 
-*Dead by Daylight* is a survival asymmetric (i.e.,, gameplay mechanics differ between two groups) multiplayer online game  developed by Behaviour Interactive Inc., in which players can play either as predators or prey.
+*Dead by Daylight* is a survival asymmetric (i.e., gameplay mechanics differ between two groups) multiplayer online game developed by Behaviour Interactive Inc.
+(Montreal, Canada), in which players can play either as a predator or prey.
 Every match includes only one predator and four prey.
@@
 
-Our study population comprised `r toString(format(length(unique(dat$predator_id))))` players that played at least 300 matches.
-We monitored all their matches from the first to a maximum of 500 matches, with a total record of `r toString(formatC(length(unique(dat$match_id)), big.mark = " "))` matches overall.
+Our study population comprised `r toString(format(length(unique(dat$predator_id))))` players who played at least 300 matches.
+We monitored all their matches from the first to a maximum of 500 matches, with a total record of `r toString(formatC(length(unique(dat$match_id)), big.mark = ","))` matches overall.
 The predator-players' experience in our population sample varied between `r toString(format(min(dat$total_xp_pred)))` and `r toString(format(max(dat$total_xp_pred)))` matches played.
@@
 We recognize that we could have introduced a bias by retaining only those individuals, as they might already be seasoned video game enthusiasts and exhibit expert-level performance in their early matches in *Dead by Daylight*.
-Thus, we tested sample bias by comparing a random sample of players that played either 20 to 50 matches, 51 to 100 matches, or 101 to 300 matches during the same timeframe as our population sample.
+Thus, we tested sample bias by comparing the results from our dataset to those from a random sample of players who played either 20 to 50 matches, 51 to 100 matches, or 101 to 300 matches during the same timeframe.
 We then took the first 20 matches played by these players, including those from our population sample, and compared their mean hunting success using a Bayesian hierarchical linear model.
-We found that all four groups had similar success as predators in their first 20 matches (Appendix 1: Table S1 and Figure S1), giving us confidence that we did not introduce a sampling bias.
+We found that all three groups of players had very similar success as the predator in their first 20 matches among them and compared to those used as our focal sample (Appendix 1: Table S1 and Figure S1), giving us confidence that we did not introduce a significant sampling bias.
 
-We collected the following information for every match in our population sample: the player's anonymous ID, the predator player's hunting success, the predator player's cumulated experience, and the mean rank and speed of the group of prey that the predator player encountered.
+We collected the following information for every match in our population sample: the player's anonymous ID, the predator player's hunting success, the predator player's cumulated experience, and the mean rank, speed, and space coverage of the group of prey that the predator player encountered.
 We defined hunting success as the number of prey consumed during the match (min = `r toString(format(min(dat$hunting_success)))`, max = `r toString(format(max(dat$hunting_success)))`).
@@
 For example, the first match of a player would have a cumulative experience value of 0, while the tenth match would have a value of 9.
-Laslty, we measured the prey's rank as the mean rank of the four individual prey in a match (mean = `r toString(format(round(mean(dat$prey_avg_rank, na.rm = T), digits = 2), nsmall = 2))` ± `r toString(format(round(sd(dat$prey_avg_rank, na.rm = T), digits = 2), nsmall = 2))`), and the prey's speed as the mean travel speed of the four individual prey in a match (mean = `r toString(format(round(mean(dat$prey_avg_speed, na.rm = T), digits = 2), nsmall = 2))` ± `r toString(format(round(sd(dat$prey_avg_speed, na.rm = T), digits = 2), nsmall = 2))` m/s).
 
-We used the mean rank of the prey as a proxy of their skill.
+We used the mean rank of the prey as a proxy of their predator-avoidance skill.
 Player ranks vary between `r toString(format(min(dat$prey_avg_rank)))` and `r toString(format(max(dat$prey_avg_rank)))`, with a rank of 1 indicating the highest skill.
@@
 While the ranking system represents an approximation of prey skill, it was the most readily available metric we could access to determine the skill of the prey group, as we did not have access to the prior matches of the prey.
-In addition, the pairing system is subject to variation depending on factors such as player availability, which can result in predator-prey groups with unbalanced skill.
-This allowed us to evaluate the effect of the prey's skill on predator success.
+In addition, the pairing system is subject to variation depending on factors such as player availability, which can result in predator-prey groups with unbalanced skill, allowing us to evaluate the effect of the prey's skill on predator success.
+For a given match, we measured the prey's rank as the mean rank of the four individual prey (mean = `r toString(format(round(mean(dat$prey_avg_rank, na.rm = T), digits = 2), nsmall = 2))` ± `r toString(format(round(sd(dat$prey_avg_rank, na.rm = T), digits = 2), nsmall = 2))`), and the prey's speed as the mean travel speed of the four individual prey in a match (mean = `r toString(format(round(mean(dat$prey_avg_speed, na.rm = T), digits = 2), nsmall = 2))` ± `r toString(format(round(sd(dat$prey_avg_speed, na.rm = T), digits = 2), nsmall = 2))` m/s).
+Lastly, the prey space coverage represents the number of unique 16x16 tiles entered by the predator in a grid that covers the whole virtual environment (mean = `r toString(format(round(mean(dat$prey_avg_amount_tiles_visited, na.rm = T), digits = 2), nsmall = 2))` ± `r toString(format(round(sd(dat$prey_avg_amount_tiles_visited, na.rm = T), digits = 2), nsmall = 2))` tile visits).
 
@@
 
-We tested how predators developed their expertise by computing five Bayesian generalized additive mixed models (GAMM) with thin-plate regression splines, all of which estimated the relationship between hunting success and the predators' cumulative experience.
+We tested how predators developed their expertise by computing six Bayesian generalized additive mixed models (GAMM) with thin-plate regression splines, all of which estimated the relationship between hunting success and the predators' cumulative experience.
 We parametrised the models following the method of @pedersen.etal2019.
@@
 The fourth (IV) and fifth (V) models were expansions of model II, but included the standardized prey speed to assess its effect on the relationship between success and experience.
-Model IV does not have a standard global smoother, while mother V includes one.
+Model IV does not have a standard global smoother, while model V includes one.
+Lastly, model VI was an extension of model V and included all the aforementioned variables with the addition of the standardized space covered by the prey.
 
-We configured the five models to use a modified beta-binomial distribution.
+We configured the six models to use a modified beta-binomial distribution.
 Hunting success was modeled as the number of prey consumed out of four, with the probability of success estimated from a Beta distribution ($Beta(\mu_i, \phi)$) characterized by a mean ($\mu \in [0,1]$) and a precision ($\phi > 0$) parameter.
@@
 We assumed that the random intercepts for the predator ID ($id$) followed a Gaussian distribution with estimated standard deviation ($id \sim N(0,\ I\ \sigma_{id}^{2})$).
-We compared the predictive accuracy of all five models using approximate leave-one-out cross-validation with Pareto-smoothed importance sampling [@vehtari.etal2017;@piironen.vehtari2017;@vehtari.etal2022].
+We compared the predictive accuracy of all six models using approximate leave-one-out cross-validation with Pareto-smoothed importance sampling [@vehtari.etal2017;@piironen.vehtari2017;@vehtari.etal2022].
 
 We defined weakly informative Gaussian priors for the intercept ($N(0,\ 0.5^{2})$) and the global trend of cumulative experience ($N(0,\ 2^{2}))$).
-Following @fraserfranco.etal2022, we defined a positive Gaussian prior ($N(1,\ 0.5^{2}))$) for the slope of the game duration because longer trials lead to greater success, a Gaussian prior ($N(0,\ 1^{2}))$) for the slope of the prey rank, and a negative Gaussian prior on prey speed ($N(-1,\ 0.5^{2})$) because encountering faster prey is associated with lower success in this system [@fraserfranco.etal2022].
+Following @fraserfranco.etal2022, we defined a positive Gaussian prior ($N(1,\ 0.5^{2}))$) for the slope of the game duration, a Gaussian prior ($N(0,\ 1^{2}))$) for the slope of the prey rank, and negative Gaussian priors on prey speed and space coverage ($N(-1,\ 0.5^{2})$).
 We defined a positive Gaussian prior on the precision parameter ($N(2,\ 0.5^{2})$).
@@
 
-To test the hypothesis that prey speed increases with prey skill, we fitted a Bayesian linear regression estimating the relationship between prey speed and the standardized prey rank.
-We controlled for the game duration, using it as a linear covariate in the model, and included the predator identity as a random effect.
-We used a weakly informative Gaussian prior on the slope of the game duration ($N(0,\ 0.5^{2})$) and a weakly informative negative Gaussian prior on the slope of the prey rank ($N(-0.5,\ 1^{2})$) because prey of greater skill (i.e., rank 1 is the highest skill) should move faster.
-Lastly, we used a weakly informative Gaussian prior on the intercept ($N(2,\ 0.5^{2})$) because prey should move at around two meters per second, and half-Gaussian priors on the standard deviation parameters ($N(0,\ 0.5^{2})$).
+To test the hypothesis that prey speed and space coverage increases with prey skill, we fitted Bayesian linear regression models estimating the relationship between 1) prey speed and the standardized prey rank and 2) prey space coverage and the standardized prey rank.
+In both models, we controlled for the game duration, using it as a linear covariate, and included the predator identity as a random effect.
+We used weakly informative Gaussian priors on the slopes of the game duration ($N(0,\ 0.5^{2})$) and weakly informative negative Gaussian priors on the slopes of the prey rank ($N(-0.5,\ 1^{2})$) because prey of greater skill (i.e., rank 1 is the highest skill) should move faster and cover more space in the environment.
+Lastly, we used weakly informative Gaussian priors on the intercepts of prey speed ($N(2,\ 0.5^{2})$) and space coverage ($N(15,\ 0.5^{2})$) as we assumed that prey should move at around two meters per second and cover at least 15 tiles in the virtual environment.
+Lastly, we employed half-Gaussian priors on the standard deviation parameters ($N(0,\ 0.5^{2})$) in both models.
 
@@
 We parametrised the GAMMs to run four chains of 1500 iterations each, discarding the first 500 iterations of each chain as warm-up, and sampling one parameter value per iteration thereafter.
-Similarly, the linear model ran four chains of 1500 iterations each, sampling a parameter value at every iteration, and discarding the first 500 iterations.
+Similarly, the linear models ran four chains of 1500 iterations each, sampling a parameter value at every iteration, and discarding the first 500 iterations.
 We thus obtained 4000 posterior samples per parameter in each model.
 We assessed the convergence of the chains using trace plots, R-hat diagnostics with a threshold set between 1.00 and 1.02, and effective sample sizes (ESS) with a threshold of >100 [@vehtari.etal2021].
-We also performed posterior predictive checks which showed an adequate fit of the models.
+We also performed posterior predictive checks, which showed an adequate fit of the models.
 
@@
 
-We tested the hypothesis that antipredator behaviour impairs the acquisition of predator expertise at the population level by comparing the global trends of model I (only controlling for game duration), model II (controlling for game duration and prey rank) and model V (controlling game duration, prey rank, and prey speed) using the finite differences method (Figure 1).
+We tested the hypothesis that antipredator behaviour impairs the acquisition of predator expertise at the population level by comparing the global trends of model I (only controlling for game duration), model II (controlling for game duration and prey rank), model V (controlling for game duration, prey rank, and prey speed), and model VI (controlling for game duration, prey rank, prey speed, and prey space coverage) using the finite differences method (Figure 1).
 This allowed us to locate the point at which hunting success was optimized.
-At the individual level, we tested the hypothesis that antipredator behaviour generate differences among predators in expertise acquisition by comparing the individual-level variance parameters.
-Specifically, we compared the standard deviations of 1) the random intercepts (i.e., mean differences in hunting success), 2) the random slopes (i.e., linear component relating hunting success with experience), and 3) the curve wiggliness (i.e., nonlinear component relating hunting success with experience).
+At the individual level, we tested the hypothesis that antipredator behaviour generates differences among predators in expertise acquisition by comparing the individual-level variance parameters.
+Specifically, we compared the posterior distributions of standard deviations for 1) the random intercepts (i.e., mean differences in hunting success), 2) the random slopes (i.e., linear component relating hunting success with experience), and 3) the curve wiggliness (i.e., nonlinear component relating hunting success with experience).
 
@@
   , model1 := c(
+    "(VI) predator xp + ID smoothers + prey rank + prey speed",
     "(V) predator xp + ID smoothers + prey rank + prey speed",
@@
     caption = "Table 1.
-Leave-one-out cross-validation results comparing the predictive performance of the five GAMMs estimating the relationship between hunting success and predator experience.
+Leave-one-out (loo) cross-validation results comparing the predictive performance of the five GAMMs estimating the relationship between hunting success and predator experience.
 All models reported include the game duration as a control.
@@
 Lastly, longer time intervals between hunting events are hypothesized to hinder or delay the acquisition of expertise due to information loss [@endler1991;@wright.etal2022;@wooster.etal2023].
-For example, a predator that played 300 matches within six months might forget critical information related to prey detection or escape patterns compared to one that played 300 matches within six weeks.
+For example, a predator who played 300 matches within six months might forget critical information related to prey detection or escape patterns compared to one who played 300 matches within six weeks.
 Therefore, investigating the impact of the aforementioned elements may reveal important insights into the outcome of predator-prey interactions, but integrating them into a unified framework remains challenging due to computational constraints and limitations in current modeling approaches.

### 2025-08-19 code/code_figures/figureS2.R
Switch figure S2 and S3 names

@@ -1,10 +1,8 @@
 # ==========================================================================
 
-#                               Figure S2
+#                 Plot GAMM model without rank - Figure S2
 
 # ==========================================================================
-
-# Figure of the relationship between success and prey speed and space
 
 
 
@@ -13,20 +11,21 @@
 # ==========================================================================
 
 
-# Load libraries -----------------------------------------------------------
 
-library(data.table)
+# Load libraries and model -------------------------------------------------
+
+options(mc.cores = parallel::detectCores())
+
+library(parallel)
 library(brms)
+library(data.table)
 library(ggplot2)
 library(ggpubr)
 library(viridis)
 
-
-
-# Import model -------------------------------------------------------------
-
 path <- file.path(getwd(), "outputs", "outputs_models")
-fit <- readRDS(file.path(path, "GAMM-VI.rds"))
+
+mod <- readRDS(file.path(path, "GAMM-I.rds"))
 
 
 
@@ -46,88 +45,84 @@ data <- fread(
 )
 
 data[, predator_id := as.factor(predator_id)]
-# ==========================================================================
-# ==========================================================================
-
-
-
-
 
 # ==========================================================================
-# 2. Compute the predictions
 # ==========================================================================
 
 
-# Predictions --------------------------------------------------------------
 
-preds <- conditional_effects(
-  x = fit,
-  method = "fitted",
-  #effects = "Zprey_speed",
-  effects = c("Zprey_speed", "Zprey_space"),
-  prob = 0.89,
-  robust = TRUE,
-  re_formula = NULL,
+
+
+# ==========================================================================
+# 2. Prepare the data for the plots
+# ==========================================================================
+
+
+
+# Prepare the plots --------------------------------------------------------
+
+
+# Group-level smooths model 2
+tab_a <- conditional_effects(
+  mod, method = "fitted",
+  effects = "Zcumul_xp:predator_id",
+  robust = TRUE, re_formula = NULL
+)
+tab_a <- data.table(tab_a[[1]])
+
+# Cumulative XP global trend model 2
+tab_b <- conditional_effects(
[…]

### 2025-08-19 code/code_figures/figureS3.R

@@ -1,8 +1,10 @@
 # ==========================================================================
 
-#                 Plot GAMM model without rank - Figure S3
+#                               Figure S3
 
 # ==========================================================================
+
+# Figure of the relationship between success and prey speed and space
 
 
 
@@ -11,21 +13,20 @@
 # ==========================================================================
 
 
+# Load libraries -----------------------------------------------------------
 
-# Load libraries and model -------------------------------------------------
-
-options(mc.cores = parallel::detectCores())
-
-library(parallel)
-library(brms)
 library(data.table)
+library(brms)
 library(ggplot2)
 library(ggpubr)
 library(viridis)
 
-path <- file.path(getwd(), "outputs", "outputs_models")
 
-mod <- readRDS(file.path(path, "GAMM-I.rds"))
+
+# Import model -------------------------------------------------------------
+
+path <- file.path(getwd(), "outputs", "outputs_models")
+fit <- readRDS(file.path(path, "GAMM-VI.rds"))
 
 
 
@@ -45,7 +46,6 @@ data <- fread(
 )
 
 data[, predator_id := as.factor(predator_id)]
-
 # ==========================================================================
 # ==========================================================================
 
@@ -54,75 +54,79 @@ data[, predator_id := as.factor(predator_id)]
 
 
 # ==========================================================================
-# 2. Prepare the data for the plots
+# 2. Compute the predictions
 # ==========================================================================
 
 
+# Predictions --------------------------------------------------------------
 
-# Prepare the plots --------------------------------------------------------
-
-
-# Group-level smooths model 2
-tab_a <- conditional_effects(
-  mod, method = "fitted",
-  effects = "Zcumul_xp:predator_id",
-  robust = TRUE, re_formula = NULL
-)
-tab_a <- data.table(tab_a[[1]])
-
-# Cumulative XP global trend model 2
-tab_b <- conditional_effects(
-  mod, method = "fitted",
-  robust = TRUE, re_formula = NULL,
-  effects = "Zcumul_xp",
+preds <- conditional_effects(
+  x = fit,
+  method = "fitted",
+  effects = c("Zprey_speed", "Zprey_space"),
+  prob = 0.89,
[…]

### 2025-08-19 manuscript/manuscript.Rmd
updated MS up to discussion

@@
 Prey antipredator behaviour can also improve through continued exposure to predators [@turner2006;@kelley.magurran2011;@lonnstedt.etal2012].
-For instance, prey alter their activity patterns under increasing predation risk [@wirsing2021].
+For instance, prey alter their activity patterns under increasing predation risk [@wirsing.etal2021].
 Animal reintroduction programs show that individuals trained to recognize predators display more antipredator behaviours and are 1.5 times more likely to survive than individuals that were never exposed to predators [@tetzalf.etal2019].
@@
 fit5 <- readRDS(here::here("outputs", "outputs_models", "GAMM-V.rds"))
+fit6 <- readRDS(here::here("outputs", "outputs_models", "GAMM-VI.rds"))
 
@@
 sum5 <- summary(fit5, robust = TRUE)$splines
+sum6 <- summary(fit6, robust = TRUE)$splines
 
@@
 rm(fit5)
+rm(fit6)
 ```
@@
 
-When comparing models with the same covariate structure (i.e., model II and III; model IV and V), those that included a global smoother generally performed better at predicting hunting success compared than those without one (Table 1).
-This improvement in performance suggests that there is a common underlying structure in the relationship between success and experience across the population despite differences among predators (Figure 1).
-Overall, model V, which included all prey features, was the best at predicting predator hunting success (Table 1).
+When comparing models with the same covariate structure (i.e., model II and III; model IV and V), those that included a global smoother generally performed better at predicting hunting success than those without one (Table 1).
+This suggests that there is a common underlying structure in the relationship between success and experience across the population, despite differences among individual predators (Figure 1).
+Overall, model VI, which included all prey features, was the best at predicting predator hunting success (Table 1).
 
@@
   , model1 := c(
-    "(VI) predator xp + ID smoothers + prey rank + prey speed",
-    "(V) predator xp + ID smoothers + prey rank + prey speed",
+    "(VI) experience + ID smoothers + prey rank + prey speed + prey space",
+    "(V) experience + ID smoothers + prey rank + prey speed",
     "(IV) ID smoothers + prey rank + prey speed",
-    "(II) predator xp + ID smoothers + prey rank",
+    "(II) experience + ID smoothers + prey rank",
     "(III) ID smoothers + prey rank",
-    "(I) predator xp + ID smoothers"
+    "(I) experience + ID smoothers"
   )
@@
   select(model1, elpd_diff, se_diff, elpd_loo, se_elpd_loo) %>%
-  flextable(col_keys = my_header$col_keys) %>%
-  set_header_df(mapping = my_header, key = "col_keys") %>%
+
+  flextable() %>%
+
+  set_header_labels(
+    model1      = "model",
+    elpd_diff   = "elpd \ndifference",
+    se_diff     = "sd \ndifference",
+    elpd_loo    = "elpd loo \nvalue",
+    se_elpd_loo = "elpd loo \nstandard error"
+  ) %>%
+
+  compose(
+    part = "header", j = "elpd_diff",
+    value = as_paragraph(
+      as_chunk("elpd \ndifference"),  # line break kept
+      as_sup("a")
+    )
+  ) %>%
+
   my_theme() %>%
   align(align = "left",   part = "all", j = 1) %>%
-  align(align = "center", part = "all", j = 2) %>%
-  align(align = "center", part = "all", j = 3) %>%
-  align(align = "center", part = "all", j = 4) %>%
-  align(align = "center", part = "all", j = 5) %>%
+  align(align = "center", part = "all", j = 2:5) %>%
+
   footnote(
-    i = 1, j = 1,
+    i = 1, j = "elpd_diff",
     part = "header",
@@
       as_sup("a "),
-      "'elpd' refers to the expected log pointwise density and is the value chosen to select the best model.",
-      as_sup("\nb "),
-      "'xp' is an acronym for experience"),
+      "'elpd' refers to the expected log pointwise density and is the value chosen to select the best model."
+    ),
     ref_symbols = " "
@@
 
-  # Table caption
   set_caption(
@@
   ) %>%
-
-  # Font options
   font(fontname = "Times New Roman", part = "all") %>%
@@
 tab1
-# ===========================================================================
-# ===========================================================================
+
 ```
 
-As we expected, models in which we did not account for the prey rank resulted in almost no change in hunting success with experience for the average individual (i.e., no gain in expertise, Figure 1A) due to the matchmaking algorithm pairing predators and prey with similar skill levels.
-Accounting for the prey rank (model II) influenced the form of the relationship, revealing a diminishing returns curve with predators optimizing their success after playing 454 matches (Figure 1B).
-The predictions indicate that the population would improve their hunting success by ~40% from the first to the 454$^{th}$ match by hunting prey of average skill (Figure 1B).
+As we expected, models in which we did not account for the prey rank resulted in almost no change in hunting success with experience for the average individual (i.e., no gain in expertise, Figure S2) due to the matchmaking algorithm pairing predators and prey with similar skill levels.
+Accounting for the prey rank (model II) influenced the form of the relationship, revealing a diminishing returns curve with predators optimizing their success after playing 454 matches (Figure 1A).
+Our results indicate that the average predator improves its hunting success by ~40% from the first to the 454$^{th}$ match by hunting prey of average skill (Figure 1A).
 
@@
 The predators' hunting success (i.e., the probability of consuming the four prey) is on the y-axis, and the predators' cumulative experience (i.e., the number of matches played before each observation) is on the x-axis.
-Panels A to C show the acquisition of expertise for the average individual with the vertical dashed lines on the left representing the lowest predicted values.
+Panels A to C show the acquisition of expertise for the average individual, with the vertical dashed lines on the left representing the lowest predicted values.
 For panel A, the right-side vertical dashed line shows the highest predicted success.
 For panels B and C, the right-side dashed line represents the point on the curve where success was optimized, which we calculated using the finite differences method to obtain the first derivative of the predicted values.
-Panels D to F show among individual differences in the acquisition of expertise, with each curve representing an individual predator.
-(A-D) Model I where we control for the game duration exclusively.
-(B-E) Model II where we control for the game duration and the prey rank exclusively.
-(C-F) Model V where we control for game duration as well as the rank and speed of the prey group."}
+Panels D to F show among-individual differences in the acquisition of expertise, with each curve representing an individual predator.
+All models displayed include the game duration as a linear covariate.
+(A-D) Model II, where we control for the game duration and the prey rank exclusively.
+(B-E) Model V, where we control for game duration as well as the rank and speed of the prey group.
+(C-F) Model VI, where we control for game duration as well as the rank, speed, and space coverage of the prey group."}
 
@@
 
-Controlling for prey speed did not qualitatively alter the form of the relationship between success and experience at the population level.
-However, the rate of expertise acquisition shifted slightly: the predator population optimized its hunting success earlier, after 323 matches (Figure 1C).
-We also found strong evidence of a negative relationship between hunting success and prey speed, indicating that hunting faster prey reduced the predator's chances of capture (Figure S2).
-Moreover, prey groups composed of individuals with greater skill moved faster than those composed of individuals with lower skill (Figure 2).
+Controlling for prey speed and space coverage did not qualitatively alter the form of the relationship between success and experience at the population level.
+However, the rate of expertise acquisition shifted slightly: the predator population reached its maximal hunting success earlier, after 323 matches (Figure 1B-C).
+We also found strong evidence that hunting success declined when predators pursued faster prey or prey occupying more space in the virtual environment (Figure S3A-B).
+Moreover, prey groups composed of individuals with greater skill moved faster than those composed of individuals with lower skill (Figure 2A).
+In contrast, there was no change in prey space coverage with increasing skill (Figure 2B).
 
 ```{r fig2, out.width = "100%", fig.cap = "**Figure 2**.
-Median posterior predictions of the relationship between prey speed and prey skill.
-The average speed of the prey group is on the y-axis, and the average rank of the prey group is on the x-axis.
+Median posterior predictions of the relationship between prey behaviour and prey skill.
+The average behaviour of the prey group is on the y-axis, and the average rank of the prey group is on the x-axis.
 The colour gradient showcases the density of observations, with lighter colours indicating lower densities, and darker colours indicating higher densities of observations.
-Lower rank values indicate that the prey group is more skilled, while higher values indicate that the prey are less skilled"}
+Lower rank values indicate that the prey group is more skilled, while higher values indicate that the prey are less skilled.
+Panel A is for the linear model of prey speed, and panel B is for the linear model of prey space coverage."}
 
@@
 Our results show that individual predators varied in their acquisition of expertise, regardless of whether we controlled for prey features (Figure 1 D-F and Figure 3).
-However, the extent of the variation in the relationship between success and experience was shaped by the prey's skill, as shown by the the posterior distributions of individual-level parameters in the model accounting for the rank of the prey (model II, Figure 2).
+However, the extent of the variation in the relationship between success and experience was shaped by the prey's skill, as shown by the posterior distributions of individual-level parameters in the model accounting for the rank of the prey (model II, Figure 2).
 Specifically, the standard deviations of individual intercepts (median model I = `r toString(round(sum1[2,]$Estimate, digits = 2))` vs median model II = `r toString(round(sum2[2,]$Estimate, digits = 2))`), slopes (median model I = `r toString(round(sum1[3,]$Estimate, digits = 2))` vs median model II = `r toString(round(sum2[3,]$Estimate, digits = 2))`), and wiggliness (median model I = `r toString(round(sum1[4,]$Estimate, digits = 2))` vs median model II = `r toString(round(sum2[4,]$Estimate, digits = 2))`) were all larger in model II compared to model I.
-Controlling for prey rank thus reveals how predators may differ expertise acquisition if they hunt prey that are not of the same skill level as them.
+Controlling for prey rank thus reveals how predators may differ in expertise acquisition if they hunt prey that are not of the same skill level as themselves.
 
-Prey speed did not influence among-individual differences in average hunting success (median model II = `r toString(round(sum2[2,]$Estimate, digits = 2))` vs median model V = `r toString(round(sum5[2,]$Estimate, digits = 2))`), nor in expertise acquisition (median model II = `r toString(round(sum2[3,]$Estimate, digits = 2))` vs median model V = `r toString(round(sum5[3,]$Estimate, digits = 2))`), as the posterior distributions of the standard deviations were almost completely overlapping (Figure 3).
+Among-individual differences in average hunting success remained unchanged regardless of whether we controlled for prey speed or prey space coverage (median model II = `r toString(round(sum2[2,]$Estimate, digits = 2))` vs median model V = `r toString(round(sum5[2,]$Estimate, digits = 2))` vs median model VI = `r toString(round(sum6[2,]$Estimate, digits = 2))`).
+Similarly, controlling for both behaviours did not alter expertise acquisition (median model II = `r toString(round(sum2[3,]$Estimate, digits = 2))` vs median model V = `r toString(round(sum5[3,]$Estimate, digits = 2))` vs median model VI = `r toString(round(sum6[3,]$Estimate, digits = 2))`), as the posterior distributions of the standard deviations showed nearly complete overlap (Figure 3).
 
@@
 Posterior distributions of the standard deviation of individual-level parameters estimated by the GAMM.
-The parameters are displayed on the y axis, and their standard deviation are displayed on the x axis.
+The parameters are displayed on the y-axis, and their standard deviation are displayed on the x-axis.
 The intercept and slope standard deviations refer to the linear components of the estimated relationship between hunting success and cumulative experience.
@@
 The vertical lines are the medians of the posterior distributions.
-The yellow distributions are for the model accounting only for the game duration (model I), the purple distributions are for the model accounting for game duration and the prey rank (model II), and the green distributions are for the model accounting for the game duration, prey rank, and prey speed (model V)."}
+The yellow (lightest) distributions are for the model accounting only for the game duration (model I), the pink (lighter) distributions are for the model accounting for game duration and the prey rank (model II), the purple (darker) distributions are for the model accounting for the game duration, prey rank, and prey speed (model V), and the dark purple (darkest) distributions are for the model accounting for the game duration and all prey features (model VI).
+The standard deviations of individual intercepts and wiggliness parameters for models II, V, and VI overlap."}
 

### 2025-08-20 download_models.R
update download-models script and makefile

@@ -11,7 +11,9 @@ fitted_models <- c(
   "GAMM-III.rds", 
   "GAMM-IV.rds", 
   "GAMM-V.rds",
+  "GAMM-VI.rds",
   "LM-PreySpeed.rds",
+  "LM-PreySpace.rds",
   "GLM-RandomSample.rds"
 )
 
@@ -39,7 +41,7 @@ if (length(missing_files) > 0) {
   model_files <- files[files$name %in% fitted_models, ]
   
   # Download each model file
-  downloaded_files <- osf_download(model_files, path = local_dir, verbose = TRUE)
+  downloaded_files <- osf_download(model_files, path = local_dir, verbose = TRUE, conflicts="skip")
   
   # Verify that each file was successfully downloaded
   for (i in seq_along(fitted_models)) {

### 2025-08-20 manuscript/appendix.Rmd
transfered folder to new OS
fixed and updated the appendix

@@
 title: |
-     Prey movement shapes the development of predator expertise in a virtual bi-trophic system: \
+     Implications of prey skill and antipredator behaviour in the acquisition of predator expertise: a virtual ecological study \
      Appendix
@@
 fit <- readRDS(
-  here::here("outputs", "outputs_models", "DHMLM-RandomSample.rds")
+  here::here("outputs", "outputs_models", "GLM-RandomSample.rds")
 )
@@
 
-# Trait variable (to remove after)
-results[parameter %like% "predspeed", trait := "predator speed"][
-  parameter %like% "preyavgspeed", trait := "prey speed"][
-    parameter %like% "success", trait := "hunting success"
-  ]
-
-# mu parameters table
-tab <- results[
-    !(parameter %like% "sigma"),
-    c(6, 1, 2, 4, 5)
-]
-tab <- tab[trait == "hunting success"]
-
 # Rename parameter values ---------------------------------------------------
@@
 # Rename
-tab[,
+results[,
   parameter := 
@@
         "cumulative experience",
+        "prey rank",
         "group 1",
@@
 ]
+
 # Reorder rows
-tab <- tab[
-  c(1,2,3,5,6,4)
+results <- results[
+  c(1,2,3,4,6,7,5)
 ]
-tab[, trait := NULL]
 
@@
 
-tabf <- tab %>%
+tabf <- results %>%
   flextable(col_keys = my_header$col_keys) %>%
@@
   set_caption(
-    tab,
+    results,
     caption = "Table S1. Fixed effects table comparing the success of random groups of players with different amounts of matches played to the group presented in the main text.",
@@
 
-```{r figS1, out.width = "100%", fig.cap = "Figure S1.
+```{r figS1, out.width = "100%", fig.cap = "**Figure S1**.
 Posterior distributions of the median hunting success per group of players.
@@
 
-```{r figS2, out.width = "100%", fig.cap = "Figure S2.
-Median posterior predictions of the relationship between hunting success (i.e.
-the probability of capturing the four prey) and the average speed of the prey group.
-The black line represents the median posterior predictions, and the gray band the the 89% compatibility interval.
-The predictions are from the model with the highest predictive accuracy (see Table 1 in main text)."}
+```{r figS2, out.width = "100%", fig.cap = "**Figure S2**.
+Median posterior predictions of the acquisition of predator hunting expertise for model I where we control for the game duration exclusively (no prey features).
+The predators' hunting success (i.e., the probability of consuming the four prey) is on the y-axis, and the predators' cumulative experience (i.e., the number of matches played before each observation) is on the x-axis.
+Panel A shows the acquisition of expertise for the average individual.
+Panel B shows among individual differences in the acquisition of expertise, with each curve representing an individual predator."}
 
@@
 ```
+
+```{r figS3, out.width = "100%", fig.cap = "**Figure S3**.
+Median posterior predictions of the relationship between hunting success (i.e.
+the probability of capturing the four prey) and the average speed (A) and average space coverage (B) of the prey group.
+The black line represents the median posterior predictions, and the gray band the the 89% compatibility interval.
+The predictions are from the model with the highest predictive accuracy (see Table 1 in main text)."}
+
+knitr::include_graphics(
+  here::here(
+    "outputs", "outputs_figures",
+    "figureS3.png"
+  )
+)
+```

### 2025-08-20 manuscript/manuscript.Rmd
manuscript updated and ready for submission

@@
 Here, we use the game *Dead by Daylight*, a virtual predator-prey system where a human predator chases four human prey, to investigate how experience shapes population and individual hunting success in predators across repeated interactions with their prey.
-We observed that predators were less successful when hunting faster and more skillful prey.
+We observed that predators were less successful when hunting faster and more skilful prey.
 As the prey gained skill, they also moved faster.
@@
 Second, we test the hypothesis that prey with more skill should have improved antipredator behaviour, and we predict that prey with greater predator-avoidance skills will move at faster speeds and cover more space in the virtual environment.
-Third, if prey antipredator behavior influences the hunting success of predators at the population level, then we expect that the acquisition of predator expertise will vary among individuals.
+Third, if prey antipredator behaviour influences the hunting success of predators at the population level, then we expect that the acquisition of predator expertise will vary among individuals.
 
@@
 Using a virtual predator-prey system where we monitored predator hunting success across experience, we provide rare empirical support for the hypothesis that prey behaviour can modulate the acquisition of hunting expertise at the population level.
-We show that the predator population is expected to increase its hunting success with experience if it hunts prey were of average skill.
-While we could not show that prey behaviour directly impaired expertise acquisition, we found that the predator population would optimize its success sooner if the prey moved at average speeds over repeated encounters.
-This is most likely due to faster prey being more difficult to hunt, which predators had lower success against.
-We also observed that prey with greater skill moved faster.
+We show that, on average, predators are expected to increase their hunting success with experience if they hunt prey of average skill.
+While we could not show that prey behaviour directly impaired expertise acquisition in predators, we found that predators reached their maximum hunting success sooner when pursuing prey with average movement behaviour (i.e., speed and space coverage) over repeated encounters.
+This is most likely because faster prey are more difficult to chase, while those that cover a lot of space are more difficult to locate since they are always moving, resulting in predators having lower success.
+We also observed that prey with greater skill moved faster, but did not increase their space coverage.
 At the individual level, predators displayed important differences in expertise acquisition that were apparent only after accounting for prey skill.
-While this most likely due the game pairing system, it highlights how prey can shape the predator hunting success over time.
+While this is most likely due to the game pairing system, it highlights how prey can shape the predator hunting success over time.
 
-The matchmaking algorithm in *Dead by Daylight* is intended to pair players with similar skill levels within matches, and our results show that the system is effective, as the population maintained consistent hunting success across experience (Figure 1A).
-However, when we controlled for the prey's skill, the gain in expertise changed at both the population and individual levels: the relationship between success and experience became asymptotic, wherein initial gains in success were significant and gradually stabilized as experience accumulated.
-This suggests that an increase in hunting success would emerge if predators hunted prey of average skill.
-This pattern of expertise acquisition is consistent with empirical studies in both humans and nonhuman animals [reviewed in @dukas2019], but provides additional information on how it may change given the difficulty of the task.
+The matchmaking algorithm in *Dead by Daylight* is intended to pair players with similar skill levels within matches, and our results show that the system is effective, as the predators maintained a near constant hunting success over repeated encounters (Figure S2).
+However, when we controlled for the prey's skill, the gain in expertise by the predator changed at both the population and individual levels: the relationship between success and experience became asymptotic, wherein initial gains in success were significant and gradually stabilized as experience accumulated.
+This suggests that hunting success would be larger for experienced predators if they were not matched to highly skilled prey, but instead, to prey of average skill levels.
+This pattern of expertise acquisition is consistent with empirical studies in both humans and nonhuman animals [reviewed in @dukas2019] but provides additional information on how it may change given the difficulty of the task.
 For example, the population may have had a slower gain in expertise if the skill of the prey encountered was consistently higher.
 
-Prey behaviour also played a role in shaping expertise acquisition at the population level.
-First, encounters with faster prey resulted in lower hunting success (Figure S2).
+Prey behaviour also played a role in shaping expertise acquisition by predators at the population level.
+First, encounters with faster prey and prey covering more space resulted in lower hunting success (Figure S3).
 We previously showed in *Dead by Daylight* that faster movement is effective for prey to evade predation [@fraserfranco.etal2022], a pattern observed in studies involving other animals [@walker.etal2005;@kelley.magurran2011;@martin.etal2022].
-Prey of greater skill increasingly relied on this strategy, and predators were less successful against skillful prey, suggesting that the effect of prey skill on the gain in expertise by predators was in part mediated by changes in prey behaviour following previous predator encounters[@tetzalf.etal2019].
-Second, we observed that predators optimized their hunting success sooner when we controlled for the prey's speed — after 323 matches played, compared to 454 matches when prey speed was not accounted for.
+Prey of greater skill increasingly relied on faster movement, and predators were less successful against skilful prey, suggesting that the effect of prey skill on the gain in expertise by predators was in part mediated by changes in prey behaviour following previous predator encounters[@tetzalf.etal2019].
+Second, we observed that predators reached their maximum hunting success sooner when we controlled for the prey's speed, after 323 matches played, compared to 454 matches when prey speed was not accounted for.
 Parallel observations have been outlined in studies of prey camouflage strategies.
@@
 Our analyses also revealed important differences among individual predators in the acquisition of expertise, indicating that predators varied in their capacity to improve over time.
-However, prey speed did not influence among-individual differences in expertise.
-The large indivual variation in expertise acquisition emerged when we controlled for the prey skill and persisted when we controlled for the prey's speed, suggesting that it would be observed even if predators hunted prey of average speed.
+However, neither prey speed nor prey space coverage influenced among-individual differences in expertise.
+The large individual variation in expertise acquisition emerged when we controlled for prey skill and persisted when we controlled for the prey's behaviour, suggesting that it would be observed even if predators hunted prey of average speed or space coverage.
 This suggests that other proximal or ecological factors may be at play.
@@
 Chasing prey demands specialized cognitive abilities and coordination that are energetically costly [@kelley.magurran2011].
-Hence, individuals with neurological or physical attributes allowing them to better concentrate on the hunting task may have shown greater improvement in success over time [@dukas2019].
+Hence, some individuals may have shown greater improvement in success over time if they possessed neurological or physical attributes that allow them to better concentrate on the hunting task  [@dukas2019].
+Longer time intervals between hunting events are also hypothesized to hinder or delay the acquisition of expertise due to information loss [@endler1991;@wright.etal2022;@wooster.etal2023].
+For example, a predator who played 300 matches within six months might forget critical information related to prey detection or escape patterns compared to one who played 300 matches within six weeks.
 Alternatively, predators may also shape their own expertise via their hunting tactic.
-For example, faster hunters may push their prey to move faster and display altruistic behaviours, which, in turn, may hinder their own expertise acquisition due to the increased difficulty of hunting such prey [@fraserfranco.etal2022;@cere.etal2024].
-Lastly, longer time intervals between hunting events are hypothesized to hinder or delay the acquisition of expertise due to information loss [@endler1991;@wright.etal2022;@wooster.etal2023].
-For example, a predator who played 300 matches within six months might forget critical information related to prey detection or escape patterns compared to one who played 300 matches within six weeks.
-Therefore, investigating the impact of the aforementioned elements may reveal important insights into the outcome of predator-prey interactions, but integrating them into a unified framework remains challenging due to computational constraints and limitations in current modeling approaches.
+For example, faster hunters may push their prey to move faster or to display altruistic behaviours, which, in turn, may hinder their own expertise acquisition due to the increased difficulty of hunting such prey [@fraserfranco.etal2022;@cere.etal2024].
+Investigating the combined effects of proximal mechanisms with the predator's tactics may thus reveal important insights into the outcome of predator-prey interactions, but integrating them into a unified framework remains challenging due to computational constraints and limitations in current modeling approaches.
+
+One limitation and criticism of using videogames to study behaviour, is that neither the predator nor the prey players' lives are at stake in the game.
+This may partly explain the observed persistence of differences among individuals in expertise acquisition.
+Hence, the gain in expertise may have been driven by the players' motivation to win rather than "true" survival.
+For instance, some players may experiment with the game out of boredom, which could contribute to shaping how expertise is honed in this particular system.
+Yet, motivation and perseverance have been outlined as core components that increase the gain in expertise in human and nonhuman hunters [@dukas2019].
+Thus, we believe that failing to account for motivation represents a limitation of our study more than the use of videogames to study ecological interactions.
 
@@
 However, the effect of antipredator behaviour was observed only at the population level.
-While there were important differences among individuals in expertise acquisition, we did not find evidence that they were mediated by either the prey's skill or prey movement behaviour.
-Future analyses should integrate physical and behavioural antipredator tactics into a unified framework to assess their influence on the acquisition of predator expertise, which may reveal important insights on the eco-evolutionary dynamics of predator-prey interactions.
-One limitation and criticism of using videogames to study behaviour which may partly explain the persistence of differences among individuals in expertise is that neither the predator nor the prey players' lives are at stake in the game.
-As a result, emerging patterns may be driven more by the players' motivation to win rather than "true" survival.
-Indeed, some players may experiment with the game out of boredom, which could contribute in shaping how expertise is honed in this particular system.
-Yet, motivation and perseverance have been outlined as core components that increase the gain in expertise in human and nonhuman hunters [@dukas2019].
-We believe that failing to account for motivation represents a limitation of our study more than in the use of videogames to study ecological interactions.
-Therefore, our demonstration of how prey antipredator behaviour can mediate the acquisition of hunting expertise adds to a growing body of research showing how virtual systems can be used to test hypotheses on ecological interactions [@beauchamp2020;@cere.etal2021;@fraserfranco.etal2022;@lymbery.etal2023;@santostefano.etal2024].
+While there were important differences among individuals in expertise acquisition, we did not find evidence that they were mediated by either the prey's skill or movement behaviour.
+Future analyses should integrate physical and behavioural antipredator tactics into a unified framework to assess their influence on the acquisition of predator expertise, which may reveal important insights into the eco-evolutionary dynamics of predator-prey interactions.
+Our demonstration of how prey antipredator behaviour can mediate the acquisition of hunting expertise adds to a growing body of research showing how virtual systems can be used to test hypotheses on ecological interactions [@beauchamp2020;@cere.etal2021;@fraserfranco.etal2022;@lymbery.etal2023;@santostefano.etal2024].
 We hope that this will inspire more collaborations between scientists and the videogame industry to tackle fundamental questions in ecology.

### 2025-08-22 code/code_figures/figure2.R
moved figure s3 to main text

@@ -1,7 +1,8 @@
 # ==========================================================================
 
-#                          Plot regression models
-#                         of prey speed and space
+#               Figure of the relationship between success 
+#                     and prey speed, space, and rank
+#                               Figure 2
 
 # ==========================================================================
 
@@ -12,23 +13,20 @@
 # ==========================================================================
 
 
+# Load libraries -----------------------------------------------------------
 
-# Load libraries and model -------------------------------------------------
-
-options(mc.cores = parallel::detectCores())
-
-library(parallel)
-library(brms)
-library(rstan)
 library(data.table)
+library(brms)
 library(ggplot2)
-library(viridis)
 library(ggpubr)
+library(viridis)
+
+
+
+# Import model -------------------------------------------------------------
 
 path <- file.path(getwd(), "outputs", "outputs_models")
-
-mod_speed <- readRDS(file.path(path, "LM-PreySpeed.rds"))
-mod_space <- readRDS(file.path(path, "LM-PreySpace.rds"))
+fit <- readRDS(file.path(path, "GAMM-VI.rds"))
 
 
 
@@ -36,32 +34,97 @@ mod_space <- readRDS(file.path(path, "LM-PreySpace.rds"))
 
 data <- fread(
   "./data/FraserFrancoetal2025-data.csv",
-  select = c(
-    "predator_id",
-    "game_duration",
-    "prey_avg_speed",
-    "prey_avg_amount_tiles_visited",
-    "prey_avg_rank"
-  )
+  select = c("predator_id",
+             "game_duration",
+             "pred_speed",
+             "prey_avg_speed",
+             "prey_avg_amount_tiles_visited",
+             "prey_avg_rank",
+             "cumul_xp_pred",
+             "total_xp_pred",
+             "hunting_success")
 )
 
-# Predator ID as factor
 data[, predator_id := as.factor(predator_id)]
+# ==========================================================================
+# ==========================================================================
 
-# Remove any NAs
-data <- data[complete.cases(data)]
[…]

### 2025-08-22 code/code_figures/figure3.R

@@ -1,13 +1,10 @@
 # ==========================================================================
 
-#                         Plot GAMM sds parameters
+#                          Plot regression models
+#                         of prey speed and space
+#                                 Figure 3
 
 # ==========================================================================
-
-# Figure of the posterior parameters for:
-#- Intercept standard deviation
-#- Slope standard deviation
-#- Wiggliness standard deviation
 
 
 
@@ -16,129 +13,56 @@
 # ==========================================================================
 
 
+
 # Load libraries and model -------------------------------------------------
 
+options(mc.cores = parallel::detectCores())
+
+library(parallel)
 library(brms)
+library(rstan)
 library(data.table)
 library(ggplot2)
-library(ggridges)
+library(viridis)
+library(ggpubr)
 
 path <- file.path(getwd(), "outputs", "outputs_models")
 
-fit1 <- readRDS(file.path(path, "GAMM-I.rds"))
-fit2 <- readRDS(file.path(path, "GAMM-II.rds"))
-fit5 <- readRDS(file.path(path, "GAMM-V.rds"))
-fit6 <- readRDS(file.path(path, "GAMM-VI.rds"))
-
-# ==========================================================================
-# ==========================================================================
+mod_speed <- readRDS(file.path(path, "LM-PreySpeed.rds"))
+mod_space <- readRDS(file.path(path, "LM-PreySpace.rds"))
 
 
 
+# Load the data ------------------------------------------------------------
 
-
-# ==========================================================================
-# 2. Extract posterior draws into a tidy table
-# ==========================================================================
-
-# Extract posterior draws fit
-post1 <- data.table(
-  as_draws_df(
-    fit1,
-    variable = c(
-      "sds_sZcumul_xppredator_id_1",
-      "sds_sZcumul_xppredator_id_2",
-      "sds_sZcumul_xppredator_id_3"
-    )
+data <- fread(
+  "./data/FraserFrancoetal2025-data.csv",
+  select = c(
+    "predator_id",
+    "game_duration",
+    "prey_avg_speed",
+    "prey_avg_amount_tiles_visited",
+    "prey_avg_rank"
   )
 )
 
+# Predator ID as factor
+data[, predator_id := as.factor(predator_id)]
[…]

### 2025-08-22 code/code_figures/figureS3.R

@@ -1,235 +1,0 @@
-# ==========================================================================
-
-#                               Figure S3
-
-# ==========================================================================
-
-# Figure of the relationship between success and prey speed and space
-
-
-
-# ==========================================================================
-# 1. Prepare script
-# ==========================================================================
-
-
-# Load libraries -----------------------------------------------------------
-
-library(data.table)
-library(brms)
-library(ggplot2)
-library(ggpubr)
-library(viridis)
-
-
-
-# Import model -------------------------------------------------------------
-
-path <- file.path(getwd(), "outputs", "outputs_models")
-fit <- readRDS(file.path(path, "GAMM-VI.rds"))
-
-
-
-# Load the data ------------------------------------------------------------
-
-data <- fread(
-  "./data/FraserFrancoetal2025-data.csv",
-  select = c("predator_id",
-             "game_duration",
-             "pred_speed",
-             "prey_avg_speed",
-             "prey_avg_amount_tiles_visited",
-             "prey_avg_rank",
-             "cumul_xp_pred",
-             "total_xp_pred",
-             "hunting_success")
-)
-
-data[, predator_id := as.factor(predator_id)]
-# ==========================================================================
-# ==========================================================================
-
-
-
-
-
-# ==========================================================================
-# 2. Compute the predictions
-# ==========================================================================
-
-
-# Predictions --------------------------------------------------------------
-
-preds <- conditional_effects(
-  x = fit,
-  method = "fitted",
-  effects = c("Zprey_speed", "Zprey_space"),
-  prob = 0.89,
-  robust = TRUE,
-  re_formula = NULL,
-  conditions = data.frame(predator_id = NA)
-)
-preds_speed <- data.table(preds[[1]])
-preds_space <- data.table(preds[[2]])
-
-
-
-# Transform values --------------------------------------------------------
-
-# Back transform x-axis values
-range_speed <- seq(
-  min(data$prey_avg_speed, na.rm = TRUE),
-  max(data$prey_avg_speed, na.rm = TRUE),
-  length.out = 5
-)
-range_space <- seq(
[…]

### 2025-08-22 manuscript/appendix.Rmd
updated MS and appendix with new figure

@@
 ```
-
-```{r figS3, out.width = "100%", fig.cap = "**Figure S3**.
-Median posterior predictions of the relationship between hunting success (i.e.
-the probability of capturing the four prey) and the average speed (A) and average space coverage (B) of the prey group.
-The black line represents the median posterior predictions, and the gray band the the 89% compatibility interval.
-The predictions are from the model with the highest predictive accuracy (see Table 1 in main text)."}
-
-knitr::include_graphics(
-  here::here(
-    "outputs", "outputs_figures",
-    "figureS3.png"
-  )
-)
-```

### 2025-08-22 manuscript/manuscript.Rmd
updated MS and appendix with new figure

@@
 
-**à compléter**
-
 The acquisition of expertise is crucial for predators to be successful hunters.
@@
 Here, we use the game *Dead by Daylight*, a virtual predator-prey system where a human predator chases four human prey, to investigate how experience shapes population and individual hunting success in predators across repeated interactions with their prey.
-We observed that predators were less successful when hunting faster and more skilful prey.
+We observed that predators were less successful when hunting faster prey or prey that covered more space in the virtual environment.
 As the prey gained skill, they also moved faster.
@@
 
+\newpage
+
 ```{r table1, ft.align = "left"}
@@
 However, the rate of expertise acquisition shifted slightly: the predator population reached its maximal hunting success earlier, after 323 matches (Figure 1B-C).
-We also found strong evidence that hunting success declined when predators pursued faster prey or prey occupying more space in the virtual environment (Figure S3A-B).
-Moreover, prey groups composed of individuals with greater skill moved faster than those composed of individuals with lower skill (Figure 2A).
-In contrast, there was no change in prey space coverage with increasing skill (Figure 2B).
+There was also strong evidence that hunting success declined when predators pursued faster prey or prey occupying more space in the virtual environment (Figure 2A-B).
+Predators had lower hunting success when hunting prey of greater skill (i.e.
+lower rank values, Figure 2C).
 
 ```{r fig2, out.width = "100%", fig.cap = "**Figure 2**.
+Median posterior predictions of the relationship between hunting success (i.e., the probability of consuming the four prey) and A) the average speed of the prey group, (B) the average space coverage of the prey group, and (C) the average rank of the prey group.
+The black line represents the median posterior predictions, and the gray band the the 89% compatibility interval.
+The predictions are from Model VI, which had the highest predictive accuracy."}
+
+knitr::include_graphics(
+  here::here(
+    "outputs", "outputs_figures",
+    "figure2.png"
+  )
+)
+```
+
+Prey groups composed of individuals with greater skill moved faster than those composed of individuals with lower skill (Figure 3A).
+In contrast, there was no change in prey space coverage with increasing skill (Figure 3B).
+
+```{r fig3, out.width = "100%", fig.cap = "**Figure 3**.
 Median posterior predictions of the relationship between prey behaviour and prey skill.
@@
 
-Our results show that individual predators varied in their acquisition of expertise, regardless of whether we controlled for prey features (Figure 1 D-F and Figure 3).
-However, the extent of the variation in the relationship between success and experience was shaped by the prey's skill, as shown by the posterior distributions of individual-level parameters in the model accounting for the rank of the prey (model II, Figure 2).
+Our results show that individual predators displayed large differences in their acquisition of expertise, regardless of whether we controlled for prey behaviour (Figure 1D-F and Figure 4).
+However, the extent of the variation in the relationship between success and experience was shaped by the prey's skill, as observed by comparing the posterior distributions of individual-level parameters among models with and without the prey rank (Figure 4).
 Specifically, the standard deviations of individual intercepts (median model I = `r toString(round(sum1[2,]$Estimate, digits = 2))` vs median model II = `r toString(round(sum2[2,]$Estimate, digits = 2))`), slopes (median model I = `r toString(round(sum1[3,]$Estimate, digits = 2))` vs median model II = `r toString(round(sum2[3,]$Estimate, digits = 2))`), and wiggliness (median model I = `r toString(round(sum1[4,]$Estimate, digits = 2))` vs median model II = `r toString(round(sum2[4,]$Estimate, digits = 2))`) were all larger in model II compared to model I.
@@
 
-Among-individual differences in average hunting success remained unchanged regardless of whether we controlled for prey speed or prey space coverage (median model II = `r toString(round(sum2[2,]$Estimate, digits = 2))` vs median model V = `r toString(round(sum5[2,]$Estimate, digits = 2))` vs median model VI = `r toString(round(sum6[2,]$Estimate, digits = 2))`).
-Similarly, controlling for both behaviours did not alter expertise acquisition (median model II = `r toString(round(sum2[3,]$Estimate, digits = 2))` vs median model V = `r toString(round(sum5[3,]$Estimate, digits = 2))` vs median model VI = `r toString(round(sum6[3,]$Estimate, digits = 2))`), as the posterior distributions of the standard deviations showed nearly complete overlap (Figure 3).
+Among-individual differences in average hunting success remained unchanged regardless of whether we controlled for prey speed and prey space coverage (median model II = `r toString(round(sum2[2,]$Estimate, digits = 2))` vs median model V = `r toString(round(sum5[2,]$Estimate, digits = 2))` vs median model VI = `r toString(round(sum6[2,]$Estimate, digits = 2))`).
+Similarly, controlling for both behaviours did not alter expertise acquisition (median model II = `r toString(round(sum2[3,]$Estimate, digits = 2))` vs median model V = `r toString(round(sum5[3,]$Estimate, digits = 2))` vs median model VI = `r toString(round(sum6[3,]$Estimate, digits = 2))`), as the posterior distributions of the standard deviations showed nearly complete overlap (Figure 4).
 
-```{r fig3, out.width = "100%", fig.cap = "**Figure 3**.
+```{r fig4, out.width = "100%", fig.cap = "**Figure 4**.
 Posterior distributions of the standard deviation of individual-level parameters estimated by the GAMM.
@@
 knitr::include_graphics(
-  here::here("outputs", "outputs_figures", "figure3.png")
+  here::here("outputs", "outputs_figures", "figure4.png")
 )
@@
 For example, the population may have had a slower gain in expertise if the skill of the prey encountered was consistently higher.
+This is further supported by our results showing reduced success against prey of greater skill.
 
 Prey behaviour also played a role in shaping expertise acquisition by predators at the population level.
-First, encounters with faster prey and prey covering more space resulted in lower hunting success (Figure S3).
-We previously showed in *Dead by Daylight* that faster movement is effective for prey to evade predation [@fraserfranco.etal2022], a pattern observed in studies involving other animals [@walker.etal2005;@kelley.magurran2011;@martin.etal2022].
+First, encounters with faster prey and prey covering more space resulted in lower hunting success (Figure 2).
+We previously showed in *Dead by Daylight* that faster movement and space coverage is effective for prey to evade predation [@fraserfranco.etal2022], a pattern observed in studies involving other animals [@walker.etal2005;@kelley.magurran2011;@martin.etal2022].
 Prey of greater skill increasingly relied on faster movement, and predators were less successful against skilful prey, suggesting that the effect of prey skill on the gain in expertise by predators was in part mediated by changes in prey behaviour following previous predator encounters[@tetzalf.etal2019].
@@
 The large individual variation in expertise acquisition emerged when we controlled for prey skill and persisted when we controlled for the prey's behaviour, suggesting that it would be observed even if predators hunted prey of average speed or space coverage.
-This suggests that other proximal or ecological factors may be at play.
+This implies that other proximal or ecological factors may be at play.
 For example, predators have limited attention spans that restrict diet choice and the formation of search images when they need to focus on hunting multiple prey simultaneously [@dukas.etal2001;@troscianko.etal2018].
@@
 For example, faster hunters may push their prey to move faster or to display altruistic behaviours, which, in turn, may hinder their own expertise acquisition due to the increased difficulty of hunting such prey [@fraserfranco.etal2022;@cere.etal2024].
-Investigating the combined effects of proximal mechanisms with the predator's tactics may thus reveal important insights into the outcome of predator-prey interactions, but integrating them into a unified framework remains challenging due to computational constraints and limitations in current modeling approaches.
+However, integrating the effects of proximal mechanisms, the predator's tactics, and the prey's behaviour into a unified framework to investigate expertise acquisition remains challenging, due to computational constraints and limitations in current modeling approaches.
 
-One limitation and criticism of using videogames to study behaviour, is that neither the predator nor the prey players' lives are at stake in the game.
+One criticism of using videogames to study behaviour is that neither the predator's nor the prey player's life is at stake in the game.
 This may partly explain the observed persistence of differences among individuals in expertise acquisition.
@@
 However, the effect of antipredator behaviour was observed only at the population level.
-While there were important differences among individuals in expertise acquisition, we did not find evidence that they were mediated by either the prey's skill or movement behaviour.
-Future analyses should integrate physical and behavioural antipredator tactics into a unified framework to assess their influence on the acquisition of predator expertise, which may reveal important insights into the eco-evolutionary dynamics of predator-prey interactions.
+While there were important differences among individuals in expertise acquisition, we did not find evidence that they were mediated by the prey's movement behaviour.
+Future analyses should incorporate cognitive, physical and behavioural attributes of the prey to assess their influence on the acquisition of predator expertise, which may reveal important insights into the eco-evolutionary dynamics of predator-prey interactions.
 Our demonstration of how prey antipredator behaviour can mediate the acquisition of hunting expertise adds to a growing body of research showing how virtual systems can be used to test hypotheses on ecological interactions [@beauchamp2020;@cere.etal2021;@fraserfranco.etal2022;@lymbery.etal2023;@santostefano.etal2024].

### 2025-08-22 manuscript/manuscript.Rmd
manuscript ready for submission

@@
 Here, we use the game *Dead by Daylight*, a virtual predator-prey system where a human predator chases four human prey, to investigate how experience shapes population and individual hunting success in predators across repeated interactions with their prey.
-We observed that predators were less successful when hunting faster prey or prey that covered more space in the virtual environment.
-As the prey gained skill, they also moved faster.
-Our analyses reveal that the predator population should optimize prey consumption over repeated encounters with prey of average skill, and optimize their success sooner if they consistently encounter prey moving at average speeds.
-However, both prey features did not mediate differences in expertise acquisition among individual predators.
+We observed that predators were less successful when hunting faster prey, prey that covered more space in the virtual environment, and prey that had greater skill.
+As the prey gained skill, they also moved faster, suggesting an improvement in antipredator behaviour.
+However, prey mpvement behaviour did not mediate differences in expertise acquisition among individual predators.
+Our results suggest that skilled prey use behaviours such as fast movement to outmanoeuver their predators, thereby impacting how predators acquire expertise at the population level.
+The predator population should optimize prey consumption over repeated encounters with prey of average skill, and optimize their success sooner if they consistently encounter prey moving at average speeds.
 
-Our results suggest that skilled prey use behaviours such as fast movement to outmanoeuver their predators, thereby impacting how predators acquire expertise at the population level.
-
-impaired the acquisition of expertise by reducing hunting success
-
-Keywords: predator-prey, experience, learning, antipredator behaviour, virtual ecology, Dead by Daylight
+Keywords: predator-prey interactions, experience, learning, movement, virtual ecology, videogames
 

### 2025-08-22 manuscript/manuscript.Rmd
updated title for BeHeco

@@
 ---
-title: "Implications of prey skill and antipredator behaviour in the acquisition of predator expertise: a virtual ecological study"
+title: "Prey skill and behaviour in predator expertise acquisition: insights from a virtual ecosystem"
 output: 
@@
 Here, we use the game *Dead by Daylight*, a virtual predator-prey system where a human predator chases four human prey, to investigate how experience shapes population and individual hunting success in predators across repeated interactions with their prey.
-We observed that predators were less successful when hunting faster prey, prey that covered more space in the virtual environment, and prey that had greater skill.
+We observed that predators were less successful when hunting prey that were faster, occupied more space in the virtual environment, and possessed greater skill.
 As the prey gained skill, they also moved faster, suggesting an improvement in antipredator behaviour.
-However, prey mpvement behaviour did not mediate differences in expertise acquisition among individual predators.
+However, prey movement behaviour did not mediate differences in expertise acquisition among individual predators.
 Our results suggest that skilled prey use behaviours such as fast movement to outmanoeuver their predators, thereby impacting how predators acquire expertise at the population level.

### 2025-08-25 manuscript/manuscript.Rmd
added lay summary

@@
 
+\newpage
+# LAY SUMMARY {.unnumbered}
+
+Predators acquire hunting expertise over successive encouters with their prey, allowing them to be more successful.
+Using a multiplayer videogame that simulates a predator-prey interaction, we show that the acquisition of predator expertise changes as a function of prey behaviour and skill.
+Our findings suggest that prey antipredator behaviour could mediate how predators learn and gain skills over time.
+
+\newpage
 # ABSTRACT {.unnumbered}
@@
 
+\newpage
 # INTRODUCTION {.unnumbered}
@@
 
+# Data availability
+
+The data and code used to produce our analyses are fully available on this GitHub repository: https://github.com/quantitative-ecologist/predator-expertise.
+
 # REFERENCES {.unnumbered}

### 2025-08-27 manuscript/manuscript.Rmd
fix mini typos

@@
 We previously showed in *Dead by Daylight* that faster movement and space coverage is effective for prey to evade predation [@fraserfranco.etal2022], a pattern observed in studies involving other animals [@walker.etal2005;@kelley.magurran2011;@martin.etal2022].
-Prey of greater skill increasingly relied on faster movement, and predators were less successful against skilful prey, suggesting that the effect of prey skill on the gain in expertise by predators was in part mediated by changes in prey behaviour following previous predator encounters[@tetzalf.etal2019].
+Prey of greater skill increasingly relied on faster movement, and predators were less successful against skilful prey, suggesting that the effect of prey skill on the gain in expertise by predators was in part mediated by changes in prey behaviour following previous predator encounters [@tetzalf.etal2019].
 Second, we observed that predators reached their maximum hunting success sooner when we controlled for the prey's speed, after 323 matches played, compared to 454 matches when prey speed was not accounted for.

### 2025-11-13 manuscript/appendix.Rmd
harmonize appendix title with ms

--- 2024-07-10 manuscript/appendix.Rmd
@@
 title: |
-     Prey movement shapes the development of predator expertise in a virtual bi-trophic system: \
+     "Implications of prey skill and antipredator behaviour in the acquisition of predator expertise: a virtual ecological study" \
      Appendix

### 2025-12-11 code/code_models/asym_models/check.R
asym models encoded and moved out old gamm files

@@ -1,321 +1,0 @@
-getwd()
-library(brms)
-library(data.table)
-library(ggplot2)
-library(viridis)
-library(ggpubr)
-
-custom_theme <- theme(
-  # Plot title
-  plot.title = element_text(size = 12),
-  # axis values size
-  axis.text = element_text(face = "plain",
-                           size = 14,
-                           color = "black"),
-  # axis ticks lenght
-  axis.ticks.length = unit(.15, "cm"),
-  # axis ticks width
-  axis.ticks = element_line(linewidth = 0.90,
-                            color = "black"),
-  # axis titles size
-  axis.title = element_text(size = 16,
-                            face = "plain",
-                            color = "black"),
-  axis.line = element_line(linewidth = 0.95,
-                           color = "black"),
-  legend.position = "none",
-  panel.grid = element_blank(),
-  panel.background = element_blank()
-)
-
-data <- fread(
-  file.path(getwd(), "data","FraserFrancoetal2025-data.csv"),
-  select = c("predator_id",
-             "hunting_success",
-             "prey_avg_speed",
-             "game_duration",
-             "cumul_xp_pred",
-             "prey_avg_rank",
-             "prey_avg_amount_tiles_visited")
-)
-
-# Predator id as factor
-data[, predator_id := as.factor(predator_id)]
-
-# Remove any NAs
-data <- data[complete.cases(data)]
-
-standardize <- function(x) {
-  (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)
-}
-
-data[
-  ,
-  c("Zprey_speed", "Zgame_duration", "Zcumul_xp", "Zprey_avg_rank", "Zprey_space") := lapply(
-    .SD, standardize
-  ),
-  .SDcols = c("prey_avg_speed", "game_duration", "cumul_xp_pred", "prey_avg_rank", "prey_avg_amount_tiles_visited")
-]
-
-
-path <- file.path(getwd(), "outputs", "outputs_models")
-fit1 <- readRDS(file.path(path, "asym-I.rds"))
-fit2 <- readRDS(file.path(path, "asym-II.rds"))
[…]

### 2025-12-11 code/code_models/asym_models/fig4_asym.R

@@ -1,220 +1,0 @@
-# ==========================================================================
-
-#                         Plot GAMM sds parameters
-#                                 Figure 4
-
-# ==========================================================================
-
-# Figure of the posterior parameters for:
-#- Baseline
-#- Rate
-#- Asymptote
-
-
-
-# ==========================================================================
-# 1. Prepare script
-# ==========================================================================
-
-
-# Load libraries and model -------------------------------------------------
-
-library(brms)
-library(data.table)
-library(ggplot2)
-library(ggridges)
-
-path <- file.path(getwd(), "outputs", "outputs_models")
-
-fit3 <- readRDS(file.path(path, "asym-III.rds"))
-fit4 <- readRDS(file.path(path, "asym-IV.rds"))
-fit5 <- readRDS(file.path(path, "asym-V.rds"))
-
-# ==========================================================================
-# ==========================================================================
-
-
-
-
-
-# ==========================================================================
-# 2. Extract posterior draws into a tidy table
-# ==========================================================================
-
-# Function to extract posterior draws
-extract_posterior_draws <- function(model, variables) {
-  post <- data.table(
-    as_draws_df(
-      model,
-      variable = variables
-    )
-  )
-  post
-}
-
-# Variables
-variables <- c(
-  "sd_predator_id__a_Intercept",
-  "sd_predator_id__b_Intercept",
-  "sd_predator_id__c_Intercept"
-)
-
-models <- list(fit3, fit4, fit5)
-
-posterior_results <- list()
-
-for (i in seq_along(models)) {
-  posterior_results[[paste0("post", i)]] <- extract_posterior_draws(
-    models[[i]],
-    variables
-  )
-}
-
-# Clear memory
-#rm(fit1)
-#rm(fit2)
-#rm(fit5)
-#rm(fit6)
-
-# Combine posterior draws
-figdat <- rbind(
-  #post2,
-  posterior_results$post1,
-  posterior_results$post2,
-  posterior_results$post3
-)
-
-# Add model variable
-#figdat[, model := c(rep("fit1", 4000), rep("fit2", 4000), rep("fit5", 4000), rep("fit6", 4000))]
-
-figdat[, model := c(rep("fit3", 12000), rep("fit4", 12000), rep("fit5", 12000))]
-
-# ==========================================================================
-# ==========================================================================
-
-
-
-
-
[…]

### 2025-12-11 code/code_models/asym_models/fig_fixef.R

@@ -1,185 +1,0 @@
-# ==========================================================================
-
-#                         Plot fixed effects parameters
-#                                 Figure X
-
-# ==========================================================================
-
-
-
-
-# ==========================================================================
-# 1. Prepare script
-# ==========================================================================
-
-
-# Load libraries and model -------------------------------------------------
-
-library(brms)
-library(data.table)
-library(ggplot2)
-library(ggridges)
-
-path <- file.path(getwd(), "outputs", "outputs_models")
-
-fit3 <- readRDS(file.path(path, "asym-III.rds"))
-fit4 <- readRDS(file.path(path, "asym-IV.rds"))
-fit5 <- readRDS(file.path(path, "asym-V.rds"))
-
-# ==========================================================================
-# ==========================================================================
-
-
-
-
-
-# ==========================================================================
-# 2. Extract posterior draws into a tidy table
-# ==========================================================================
-
-# Function
-make_ce_df <- function(fit, nlpar, effects, labels,
-                       prob = 0.89,
-                       method = "fitted",
-                       robust = TRUE,
-                       re_formula = NULL) {
-  if (length(effects) != length(labels)) {
-    stop("effects and labels must have the same length.")
-  }
-
-  out_list <- vector("list", length(effects))
-
-  for (i in seq_along(effects)) {
-    ce <- conditional_effects(
-      x = fit,
-      effects = effects[i],
-      nlpar = nlpar,
-      prob = prob,
-      robust = robust,
-      re_formula = re_formula,
-      method = method
-    )[[1]]
-
-    # variable label
-    ce$variable <- labels[i]
-
-    # Columns for plotting
-    ce <- ce[, c("variable", "effect1__", "estimate__", "lower__", "upper__")]
-
-    out_list[[i]] <- ce
-  }
-
-  # bind into one data frame
-  do.call(rbind, out_list)
-}
-
-# Select traits ------------------------------------------------------------
-
-effects_prey  <- c("Zprey_avg_rank", "Zprey_speed", "Zprey_space")
-labels_prey <- c("prey_rank", "prey_speed", "prey_space")
-
-
-
[…]

### 2025-12-11 code/code_models/asym_models/fig_ranefs.R

@@ -1,162 +1,0 @@
-# ==========================================================================
-
-#                         Plot random effects means distribution
-#                                 Figure X
-
-# ==========================================================================
-
-
-
-
-# ==========================================================================
-# 1. Prepare script
-# ==========================================================================
-
-
-# Load libraries and model -------------------------------------------------
-
-library(brms)
-library(data.table)
-library(ggplot2)
-library(ggridges)
-
-path <- file.path(getwd(), "outputs", "outputs_models")
-
-fit3 <- readRDS(file.path(path, "asym-III.rds"))
-fit4 <- readRDS(file.path(path, "asym-IV.rds"))
-fit5 <- readRDS(file.path(path, "asym-V.rds"))
-
-# ==========================================================================
-# ==========================================================================
-
-
-
-
-
-# ==========================================================================
-# 2. Extract posterior draws into a tidy table
-# ==========================================================================
-
-draws <- as.data.table(as_draws_df(fit5))
-
-summarise_param_dt <- function(draws_dt, par_short) {
-  # par_short: "a", "b", or "c"
-
-  # Global (population-level) intercept column, e.g. "b_a_Intercept"
-  global_col <- paste0("b_", par_short, "_Intercept")
-
-  # Random-effect columns for this parameter, e.g. "r_predator_id__a[ ... ]"
-  re_cols <- grep(
-    paste0("^r_predator_id__", par_short, "\\["),
-    names(draws_dt),
-    value = TRUE
-  )
-
-  # Keep only needed columns + a draw index
-  sub_dt <- draws_dt[, c(global_col, re_cols), with = FALSE]
-  sub_dt[, .draw := .I]
-
-  # Melt to long format: one row per (draw, predator)
-  long_dt <- melt(
-    sub_dt,
-    id.vars = c(".draw", global_col),
-    measure.vars = re_cols,
-    variable.name = "re_name",
-    value.name = "re"
-  )
-
-  # Extract predator_id from column names:
-  # e.g. "r_predator_id__a[pred143150,Intercept]" -> "pred143150"
-  long_dt[
-    , predator_id := sub(
-        paste0("r_predator_id__", par_short, "\\[([^,]+),Intercept\\]"),
-        "\\1",
-        re_name
-    )
-  ]
-
-  # Predator-specific parameter value = global intercept + random effect
[…]

### 2025-12-11 code/code_models/asym_models/fit_asym-III.sh

@@ -1,16 +1,0 @@
-#!/bin/bash
-#SBATCH --account=def-monti
-#SBATCH --nodes=1
-#SBATCH --cpus-per-task=32
-#SBATCH --mem=64000M
-#SBATCH --time=00-18:00
-#SBATCH --mail-user=fraser_franco.maxime@courrier.uqam.ca
-#SBATCH --mail-type=ALL
-
-module load StdEnv/2023
-module load gcc/12.3
-module load r-bundle-bioconductor/3.20
-module load r/4.4.0
-
-export R_LIBS="/home/maxime11/R/x86_64-pc-linux-gnu-library/4.4"
-Rscript fit_asym-III.R

### 2025-12-11 code/code_models/asym_models/fit_asym-V.R

@@ -1,194 +1,0 @@
-# ==========================================================================
-
-#                   Asymptotic model - Hunting success
-#                               Model IV
-#               - rank as fixed effect of hunting success
-#               - rank as fixed effect of a and c
-#               - prey speed as fixed effect of hunting success
-#               - prey speed as fixed effect of a and c
-
-# ==========================================================================
-
-
-
-
-
-# ==========================================================================
-# 1. Prepare the data
-# ==========================================================================
-
-
-
-# Load packages ------------------------------------------------------------
-
-# Detect cores
-options(mc.cores = parallel::detectCores())
-
-# Load libraries
-library(data.table)
-library(brms)
-library(parallel)
-
-
-
-# Load data ----------------------------------------------------------------
-
-# Folder path Compute Canada
-folder <- file.path("/home", "maxime11", "projects", "def-monti", 
-                    "maxime11", "phd_project", "data")
-
-# Load data on compute canada
-data <- fread(
-  file.path(folder, "FraserFrancoetal2025-data.csv"),
-  select = c("predator_id",
-             "hunting_success",
-             "prey_avg_speed",
-             "prey_avg_amount_tiles_visited",
-             "game_duration",
-             "cumul_xp_pred",
-             "prey_avg_rank")
-)
-
-# Predator id as factor
-data[, predator_id := as.factor(predator_id)]
-
-# Remove any NAs
-data <- data[complete.cases(data)]
-
-
-
-# Standardise the variables (Z-scores) -------------------------------------
-
-standardize <- function(x) {
-  (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)
-}
-
-data[
-  ,
-  c("Zprey_speed", "Zprey_space", "Zgame_duration", 
-    "Zcumul_xp", "Zprey_avg_rank") := lapply(
-    .SD, standardize
-  ),
-  .SDcols = c(
-    "prey_avg_speed", "prey_avg_amount_tiles_visited",
-    "game_duration", "cumul_xp_pred", "prey_avg_rank"
-  )
-]
-
-# ==========================================================================
-# ==========================================================================
-
-
-
-
-
[…]

### 2025-12-11 code/code_models/asym_models/fit_asym-V.sh

@@ -1,16 +1,0 @@
-#!/bin/bash
-#SBATCH --account=def-monti
-#SBATCH --nodes=1
-#SBATCH --cpus-per-task=32
-#SBATCH --mem=64000M
-#SBATCH --time=00-18:00
-#SBATCH --mail-user=fraser_franco.maxime@courrier.uqam.ca
-#SBATCH --mail-type=ALL
-
-module load StdEnv/2023
-module load gcc/12.3
-module load r-bundle-bioconductor/3.20
-module load r/4.4.0
-
-export R_LIBS="/home/maxime11/R/x86_64-pc-linux-gnu-library/4.4"
-Rscript fit_asym-V.R

### 2025-12-11 code/code_models/fit_GAMM-I.R

@@ -1,165 +1,0 @@
-# ==========================================================================
-
-#           GAMM model GS - group level smoother - Hunting success
-#                     Model I without rank in article
-
-# ==========================================================================
-
-
-
-
-
-# ==========================================================================
-# 1. Prepare the data
-# ==========================================================================
-
-
-
-# Load packages ------------------------------------------------------------
-
-# Detect cores
-options(mc.cores = parallel::detectCores())
-
-# Load libraries
-library(data.table)
-library(brms)
-library(parallel)
-
-
-
-# Load data ----------------------------------------------------------------
-
-# Folder path Compute Canada
-folder <- file.path("/home", "maxime11", "projects", "def-monti", 
-                    "maxime11", "phd_project", "data")
-
-# Load data on compute canada
-data <- fread(
-  file.path(folder, "FraserFrancoetalXXXX-data.csv"),
-  select = c("predator_id",
-             "hunting_success",
-             "prey_avg_speed",
-             "game_duration",
-             "cumul_xp_pred",
-             "prey_avg_rank")
-)
-
-# Predator id as factor
-data[, predator_id := as.factor(predator_id)]
-
-# Remove any NAs
-data <- data[complete.cases(data)]
-
-
-
-# Standardise the variables (Z-scores) -------------------------------------
-
-standardize <- function(x) {
-  (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)
-}
-
-data[
-  ,
-  c("Zprey_speed", "Zgame_duration", "Zcumul_xp", "Zprey_avg_rank") := lapply(
-    .SD, standardize
-  ),
-  .SDcols = c("prey_avg_speed", "game_duration", "cumul_xp_pred", "prey_avg_rank")
-]
-
-# ==========================================================================
-# ==========================================================================
-
-
-
-
-
-# ==========================================================================
-# 2. Parametrize the model
-# ==========================================================================
-
-
-
-# Model formula ------------------------------------------------------------
-
-model_formula <- brmsformula(
-  hunting_success | trials(4) ~
-    s(Zcumul_xp) +
-    s(Zcumul_xp, predator_id, bs = "fs") +
-    Zgame_duration
-)
-
-
-
-# Inspect priors -----------------------------------------------------------
-
-get_prior(
-  formula = model_formula,
-  family = beta_binomial(),
-  data = data
-)
-
-
-
[…]

### 2025-12-11 code/code_models/fit_GAMM-I.sh

@@ -1,16 +1,0 @@
-#!/bin/bash
-#SBATCH --account=def-monti
-#SBATCH --nodes=1
-#SBATCH --cpus-per-task=32
-#SBATCH --mem=64000M
-#SBATCH --time=05-00:00
-#SBATCH --mail-user=fraser_franco.maxime@courrier.uqam.ca
-#SBATCH --mail-type=ALL
-
-module load StdEnv/2023
-module load gcc/12.3
-module load r-bundle-bioconductor/3.20
-module load r/4.4.0
-
-export R_LIBS="/home/maxime11/R/x86_64-pc-linux-gnu-library/4.4"
-Rscript fit_GAMM-I.R

### 2025-12-11 code/code_models/fit_GAMM-II.R

@@ -1,171 +1,0 @@
-# ==========================================================================
-
-#          GAMM model GS - group-level smoother - hunting success
-#                           Model II in article
-
-# ==========================================================================
-
-
-
-
-
-# ==========================================================================
-# 1. Prepare the data
-# ==========================================================================
-
-
-
-# Load packages ------------------------------------------------------------
-
-# Detect cores
-options(mc.cores = parallel::detectCores())
-
-# Load libraries
-library(data.table)
-library(brms)
-library(parallel)
-
-
-
-# Load data ----------------------------------------------------------------
-
-# Folder path Compute Canada
-folder <- file.path("/home", "maxime11", "projects", "def-monti", 
-                    "maxime11", "phd_project", "data")
-
-# Load data
-data <- fread(
-  file.path(folder, "FraserFrancoetalXXXX-data.csv"),
-  select = c("predator_id",
-             "hunting_success",
-             "prey_avg_speed",
-             "game_duration",
-             "cumul_xp_pred",
-             "prey_avg_rank")
-)
-
-# Predator id as factor
-data[, predator_id := as.factor(predator_id)]
-
-# Remove any NAs
-data <- data[complete.cases(data)]
-
-
-
-# Standardise the variables (Z-scores) -------------------------------------
-
-standardize <- function(x) {
-  (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)
-}
-
-data[
-  ,
-  c("Zprey_speed", "Zgame_duration", "Zcumul_xp", "Zprey_avg_rank") := lapply(
-    .SD, standardize
-  ),
-  .SDcols = c("prey_avg_speed", "game_duration", "cumul_xp_pred", "prey_avg_rank")
-]
-
-
-# ==========================================================================
-# ==========================================================================
-
-
-
-
-
-# ==========================================================================
-# 2. Parametrize the model
-# ==========================================================================
-
-
-# Model formula ------------------------------------------------------------
-
-model_formula <- brmsformula(
-  hunting_success | trials(4) ~
-    s(Zcumul_xp) +
-    s(Zcumul_xp, predator_id, bs = "fs") +
-    Zgame_duration +
-    Zprey_avg_rank
-)
-
-
-
-# Inspect priors -----------------------------------------------------------
-
-get_prior(
-  formula = model_formula,
-  family = beta_binomial(),
[…]

### 2025-12-11 code/code_models/fit_GAMM-III.R

@@ -1,165 +1,0 @@
-# ==========================================================================
-
-#         GAMM model S - individual curves without common smoother
-#                              hunting success
-#                           Model III in article
-
-# ==========================================================================
-
-
-
-
-
-# ==========================================================================
-# 1. Prepare the data
-# ==========================================================================
-
-
-
-# Load packages ------------------------------------------------------------
-
-# Detect cores
-options(mc.cores = parallel::detectCores())
-
-# Load libraries
-library(data.table)
-library(brms)
-library(parallel)
-
-
-
-# Load data ----------------------------------------------------------------
-
-# Folder path Compute Canada
-folder <- file.path("/home", "maxime11", "projects", "def-monti", 
-                    "maxime11", "phd_project", "data")
-
-# Load data
-data <- fread(
-  file.path(folder, "FraserFrancoetalXXXX-data.csv"),
-  select = c("predator_id",
-             "hunting_success",
-             "prey_avg_speed",
-             "game_duration",
-             "cumul_xp_pred",
-             "prey_avg_rank")
-)
-
-# Predator id as factor
-data[, predator_id := as.factor(predator_id)]
-
-# Remove any NAs
-data <- data[complete.cases(data)]
-
-
-
-# Standardise the variables (Z-scores) -------------------------------------
-
-standardize <- function(x) {
-  (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)
-}
-
-data[
-  ,
-  c("Zprey_speed", "Zgame_duration", "Zcumul_xp", "Zprey_avg_rank") := lapply(
-    .SD, standardize
-  ),
-  .SDcols = c("prey_avg_speed", "game_duration", "cumul_xp_pred", "prey_avg_rank")
-]
-
-# ==========================================================================
-# ==========================================================================
-
-
-
-
-
-# ==========================================================================
-# 2. Parametrize the model
-# ==========================================================================
-
-
-# Model formula ------------------------------------------------------------
-
-model_formula <- brmsformula(
-    hunting_success | trials(4) ~
-    s(Zcumul_xp, predator_id, bs = "fs", m = 2) +
-    Zgame_duration +
-    Zprey_avg_rank
-)
-
-
-
-# Inspect priors -----------------------------------------------------------
-
-get_prior(
-  formula = model_formula,
-  family = beta_binomial(),
-  data = data
-)
-
-
-
[…]

### 2025-12-11 code/code_models/fit_GAMM-III.sh

@@ -1,16 +1,0 @@
-#!/bin/bash
-#SBATCH --account=def-monti
-#SBATCH --nodes=1
-#SBATCH --cpus-per-task=32
-#SBATCH --mem=64000M
-#SBATCH --time=05-00:00
-#SBATCH --mail-user=fraser_franco.maxime@courrier.uqam.ca
-#SBATCH --mail-type=ALL
-
-module load StdEnv/2023
-module load gcc/12.3
-module load r-bundle-bioconductor/3.20
-module load r/4.4.0
-
-export R_LIBS="/home/maxime11/R/x86_64-pc-linux-gnu-library/4.4"
-Rscript GAMM-III.R

### 2025-12-11 code/code_models/fit_GAMM-IV.R

@@ -1,171 +1,0 @@
-# ==========================================================================
-
-#         GAMM model S - individual curves without common smoother
-#                              hunting success
-#                            Model IV in article
-
-# ==========================================================================
-
-
-
-
-
-# ==========================================================================
-# 1. Prepare the data
-# ==========================================================================
-
-
-
-# Load packages ------------------------------------------------------------
-
-# Detect cores
-options(mc.cores = parallel::detectCores())
-
-# Load libraries
-library(data.table)
-library(brms)
-library(parallel)
-
-
-
-# Load data ----------------------------------------------------------------
-
-# Folder path Compute Canada
-folder <- file.path("/home", "maxime11", "projects", "def-monti", 
-                    "maxime11", "phd_project", "data")
-
-# Load data
-data <- fread(
-  file.path(folder, "FraserFrancoetalXXXX-data.csv"),
-  select = c("predator_id",
-             "hunting_success",
-             "prey_avg_speed",
-             "game_duration",
-             "cumul_xp_pred",
-             "prey_avg_rank")
-)
-
-# Predator id as factor
-data[, predator_id := as.factor(predator_id)]
-
-# Remove any NAs
-data <- data[complete.cases(data)]
-
-
-
-# Standardise the variables (Z-scores) -------------------------------------
-
-standardize <- function(x) {
-  (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)
-}
-
-data[
-  ,
-  c("Zprey_speed", "Zgame_duration", "Zcumul_xp", "Zprey_avg_rank") := lapply(
-    .SD, standardize
-  ),
-  .SDcols = c("prey_avg_speed", "game_duration", "cumul_xp_pred", "prey_avg_rank")
-]
-
-# ==========================================================================
-# ==========================================================================
-
-
-
-
-
-# ==========================================================================
-# 2. Parametrize the model
-# ==========================================================================
-
-
-# Model formula ------------------------------------------------------------
-
-model_formula <- brmsformula(
-  hunting_success | trials(4) ~
-    s(Zcumul_xp, predator_id, bs = "fs", m = 2) +
-    Zgame_duration +
-    Zprey_speed +
-    Zprey_avg_rank
-)
-
-
-
-# Inspect priors -----------------------------------------------------------
-
-get_prior(
-  formula = model_formula,
[…]

### 2025-12-11 code/code_models/fit_GAMM-IV.sh

@@ -1,16 +1,0 @@
-#!/bin/bash
-#SBATCH --account=def-monti
-#SBATCH --nodes=1
-#SBATCH --cpus-per-task=32
-#SBATCH --mem=64000M
-#SBATCH --time=05-00:00
-#SBATCH --mail-user=fraser_franco.maxime@courrier.uqam.ca
-#SBATCH --mail-type=ALL
-
-module load StdEnv/2023
-module load gcc/12.3
-module load r-bundle-bioconductor/3.20
-module load r/4.4.0
-
-export R_LIBS="/home/maxime11/R/x86_64-pc-linux-gnu-library/4.4"
-Rscript GAMM-IV.R

### 2025-12-11 code/code_models/fit_GAMM-V.R

@@ -1,176 +1,0 @@
-# ==========================================================================
-
-#           GAMM model GS - group-level smoother - hunting success
-#                           Model V in article
-
-# ==========================================================================
-
-
-
-
-
-# ==========================================================================
-# 1. Prepare the data
-# ==========================================================================
-
-
-
-# Load packages ------------------------------------------------------------
-
-# Detect cores
-options(mc.cores = parallel::detectCores())
-
-# Load libraries
-library(data.table)
-library(brms)
-library(parallel)
-
-
-
-# Load data ----------------------------------------------------------------
-
-# Folder path Compute Canada
-folder <- file.path("/home", "maxime11", "projects", "def-monti", 
-                    "maxime11", "phd_project", "data")
-
-# Load data
-data <- fread(
-  file.path(folder, "FraserFrancoetalXXXX-data.csv"),
-  select = c("predator_id",
-             "hunting_success",
-             "prey_avg_speed",
-             "game_duration",
-             "cumul_xp_pred",
-             "prey_avg_rank")
-)
-
-# Predator id as factor
-data[, predator_id := as.factor(predator_id)]
-
-# Remove any NAs
-data <- data[complete.cases(data)]
-
-
-
-# Standardise the variables (Z-scores) -------------------------------------
-
-standardize <- function(x) {
-  (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)
-}
-
-data[
-  ,
-  c("Zprey_speed", "Zgame_duration", "Zcumul_xp", "Zprey_avg_rank") := lapply(
-    .SD, standardize
-  ),
-  .SDcols = c("prey_avg_speed", "game_duration", "cumul_xp_pred", "prey_avg_rank")
-]
-
-
-# ==========================================================================
-# ==========================================================================
-
-
-
-
-
-# ==========================================================================
-# 2. Parametrize the model
-# ==========================================================================
-
-
-# Model formula ------------------------------------------------------------
-
-model_formula <- brmsformula(
-  hunting_success | trials(4) ~
-    s(Zcumul_xp) +
-    s(Zcumul_xp, predator_id, bs = "fs") +
-    Zgame_duration +
-    Zprey_speed +
-    Zprey_avg_rank
-)
-
-
-
-# Inspect priors -----------------------------------------------------------
-
-get_prior(
[…]

### 2025-12-11 code/code_models/fit_GAMM-V.sh

@@ -1,16 +1,0 @@
-#!/bin/bash
-#SBATCH --account=def-monti
-#SBATCH --nodes=1
-#SBATCH --cpus-per-task=32
-#SBATCH --mem=64000M
-#SBATCH --time=5-20:00
-#SBATCH --mail-user=fraser_franco.maxime@courrier.uqam.ca
-#SBATCH --mail-type=ALL
-
-module load StdEnv/2023
-module load gcc/12.3
-module load r-bundle-bioconductor/3.20
-module load r/4.4.0
-
-export R_LIBS="/home/maxime11/R/x86_64-pc-linux-gnu-library/4.4"
-Rscript GAMM-V.R

### 2025-12-11 code/code_models/fit_GAMM-VI.R

@@ -1,202 +1,0 @@
-# ==========================================================================
-
-#           GAMM model GS - group-level smoother - hunting success
-#                           Model VI in article
-
-# ==========================================================================
-
-
-
-
-
-# ==========================================================================
-# 1. Prepare the data
-# ==========================================================================
-
-
-
-# Load packages ------------------------------------------------------------
-
-# Detect cores
-options(mc.cores = parallel::detectCores())
-
-# Load libraries
-library(data.table)
-library(brms)
-library(parallel)
-
-
-
-# Load data ----------------------------------------------------------------
-
-# Folder path Compute Canada
-folder <- file.path("/home", "maxime11", "projects", "def-monti", 
-                    "maxime11", "phd_project", "data")
-
-#folder <- file.path(getwd(), "data")
-
-
-# Load data
-#data <- fread(
-#  file.path(folder, "FraserFrancoetalXXXX-data.csv"),
-#  select = c("predator_id",
-#             "hunting_success",
-#             "prey_avg_speed",
-#             "prey_avg_amount_tiles_visited",
-#             "game_duration",
-#             "cumul_xp_pred",
-#             "prey_avg_rank")
-#)
-
-data <- fread(
-  file.path(folder, "FraserFrancoetal2025-data.csv"),
-  select = c("predator_id",
-             "hunting_success",
-             "prey_avg_speed",
-             "prey_avg_amount_tiles_visited",
-             "game_duration",
-             "cumul_xp_pred",
-             "prey_avg_rank")
-)
-
-
-# Predator id as factor
-data[, predator_id := as.factor(predator_id)]
-
-# Remove any NAs
-data <- data[complete.cases(data)]
-
-
-
-# Standardise the variables (Z-scores) -------------------------------------
-
-standardize <- function(x) {
-  (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)
-}
-
-data[
-  ,
-  c("Zprey_speed", "Zprey_space", "Zgame_duration", 
-    "Zcumul_xp", "Zprey_avg_rank") := lapply(
-    .SD, standardize
-  ),
-  .SDcols = c(
-    "prey_avg_speed", "prey_avg_amount_tiles_visited",
-    "game_duration", "cumul_xp_pred", "prey_avg_rank"
-  )
-]
-
-
-# ==========================================================================
-# ==========================================================================
-
-
-
-
-
[…]

### 2025-12-11 code/code_models/fit_GAMM-VI.sh

@@ -1,16 +1,0 @@
-#!/bin/bash
-#SBATCH --account=def-monti
-#SBATCH --nodes=1
-#SBATCH --cpus-per-task=32
-#SBATCH --mem=64000M
-#SBATCH --time=4-20:00
-#SBATCH --mail-user=fraser_franco.maxime@courrier.uqam.ca
-#SBATCH --mail-type=ALL
-
-module load StdEnv/2023
-module load gcc/12.3
-module load r-bundle-bioconductor/3.20
-module load r/4.4.0
-
-export R_LIBS="/home/maxime11/R/x86_64-pc-linux-gnu-library/4.4"
-Rscript GAMM-VI.R

### 2025-12-11 code/code_models/fit_asym-I.R

@@ -59,10 +59,10 @@ standardize <- function(x) {
 
 data[
   ,
-  c("Zprey_speed", "Zgame_duration", "Zcumul_xp") := lapply(
+  c("Zprey_speed", "Zgame_duration") := lapply(
     .SD, standardize
   ),
-  .SDcols = c("prey_avg_speed", "game_duration", "cumul_xp_pred")
+  .SDcols = c("prey_avg_speed", "game_duration")
 ]
 
 # ==========================================================================
@@ -82,7 +82,7 @@ data[
 
 model_formula <- brmsformula(
   hunting_success | trials(4) ~
-    a - (a - b) * exp(-exp(c) * Zcumul_xp) +
+    a - (a - b) * exp(-exp(c) * cumul_xp_pred) +
       betaduration * Zgame_duration,
 
   a ~ 1 + (1 | p | predator_id), # (on logit scale)
@@ -116,7 +116,7 @@ priors <- c(
   # Intercept on a, b, c
   prior(normal(1.386294, 0.7), nlpar = "a", class = "b", coef = "Intercept"),
   prior(normal(-1.098612, 0.7), nlpar = "b", class = "b", coef = "Intercept"),
-  prior(normal(0, 0.5), nlpar = "c", class = "b", coef = "Intercept"),
+  prior(normal(-5.5, 1), nlpar = "c", class = "b", coef = "Intercept"),
 
   # Random effects on a, b, c
   prior(normal(0, 1), nlpar = "a", class = "sd"),

### 2025-12-11 code/code_models/fit_asym-II.R

@@ -1,11 +1,8 @@
 # ==========================================================================
 
 #                   Asymptotic model - Hunting success
-#                               Model IV
+#                               Model II
 #               - rank as fixed effect of hunting success
-#               - rank as fixed effect of a and c
-#               - prey speed as fixed effect of hunting success
-#               - prey speed as fixed effect of a and c
 
 # ==========================================================================
 
@@ -64,10 +61,10 @@ standardize <- function(x) {
 
 data[
   ,
-  c("Zprey_speed", "Zgame_duration", "Zcumul_xp", "Zprey_avg_rank") := lapply(
+  c("Zprey_speed", "Zgame_duration", "Zprey_avg_rank") := lapply(
     .SD, standardize
   ),
-  .SDcols = c("prey_avg_speed", "game_duration", "cumul_xp_pred", "prey_avg_rank")
+  .SDcols = c("prey_avg_speed", "game_duration", "prey_avg_rank")
 ]
 
 # ==========================================================================
@@ -87,18 +84,16 @@ data[
 
 model_formula <- brmsformula(
   hunting_success | trials(4) ~
-    a - (a - b) * exp(-exp(c) * Zcumul_xp) +
+    a - (a - b) * exp(-exp(c) * cumul_xp_pred) +
       betaduration * Zgame_duration +
-      betarank * Zprey_avg_rank +
-      betaspeed * Zprey_speed,
+      betarank * Zprey_avg_rank,
 
-  a ~ 1 + Zprey_avg_rank + Zprey_speed + (1 | p | predator_id),
+  a ~ 1 + (1 | p | predator_id),
   b ~ 1 + (1 | p | predator_id),
-  c ~ 1 + Zprey_avg_rank + Zprey_speed + (1 | p | predator_id),
+  c ~ 1 + (1 | p | predator_id),
 
   betaduration ~ 1,
   betarank ~ 1,
-  betaspeed ~ 1,
 
   nl = TRUE
 )
@@ -125,13 +120,7 @@ priors <- c(
   # Intercept on a, b, c
   prior(normal(1.386294, 0.7), nlpar = "a", class = "b", coef = "Intercept"),
   prior(normal(-1.098612, 0.7), nlpar = "b", class = "b", coef = "Intercept"),
-  prior(normal(0, 0.5), nlpar = "c", class = "b", coef = "Intercept"),
-
-  # Fixed effects on a, c
[…]

### 2025-12-11 code/code_models/fit_asym-III.R

@@ -1,9 +1,9 @@
 # ==========================================================================
 
 #                   Asymptotic model - Hunting success
-#                             Model III 
+#                               Model III
 #               - rank as fixed effect of hunting success
-#               - rank as fixed effect of a and c
+#               - prey speed as fixed effect of a and c
 
 # ==========================================================================
 
@@ -62,10 +62,10 @@ standardize <- function(x) {
 
 data[
   ,
-  c("Zprey_speed", "Zgame_duration", "Zcumul_xp", "Zprey_avg_rank") := lapply(
+  c("Zprey_speed", "Zgame_duration", "Zprey_avg_rank") := lapply(
     .SD, standardize
   ),
-  .SDcols = c("prey_avg_speed", "game_duration", "cumul_xp_pred", "prey_avg_rank")
+  .SDcols = c("prey_avg_speed", "game_duration", "prey_avg_rank")
 ]
 
 # ==========================================================================
@@ -85,13 +85,13 @@ data[
 
 model_formula <- brmsformula(
   hunting_success | trials(4) ~
-    a - (a - b) * exp(-exp(c) * Zcumul_xp) +
+    a - (a - b) * exp(-exp(c) * cumul_xp_pred) +
       betaduration * Zgame_duration +
       betarank * Zprey_avg_rank,
 
-  a ~ 1 + Zprey_avg_rank + (1 | p | predator_id),
+  a ~ 1 + Zprey_speed + (1 | p | predator_id),
   b ~ 1 + (1 | p | predator_id),
-  c ~ 1 + Zprey_avg_rank + (1 | p | predator_id),
+  c ~ 1 + Zprey_speed + (1 | p | predator_id),
 
   betaduration ~ 1,
   betarank ~ 1,
@@ -121,11 +121,11 @@ priors <- c(
   # Intercept on a, b, c
   prior(normal(1.386294, 0.7), nlpar = "a", class = "b", coef = "Intercept"),
   prior(normal(-1.098612, 0.7), nlpar = "b", class = "b", coef = "Intercept"),
-  prior(normal(0, 0.5), nlpar = "c", class = "b", coef = "Intercept"),
+  prior(normal(-5.5, 1), nlpar = "c", class = "b", coef = "Intercept"),
 
   # Fixed effects on a, c
[…]

### 2025-12-11 code/code_models/fit_asym-III.sh

@@ -3,7 +3,7 @@
 #SBATCH --nodes=1
 #SBATCH --cpus-per-task=32
 #SBATCH --mem=64000M
-#SBATCH --time=05-00:00
+#SBATCH --time=00-24:00
 #SBATCH --mail-user=fraser_franco.maxime@courrier.uqam.ca
 #SBATCH --mail-type=ALL
 
@@ -13,4 +13,4 @@ module load r-bundle-bioconductor/3.20
 module load r/4.4.0
 
 export R_LIBS="/home/maxime11/R/x86_64-pc-linux-gnu-library/4.4"
-Rscript GAMM-II.R
+Rscript fit_asym-III.R

### 2025-12-11 code/code_models/fit_asym-IV.R

@@ -1,8 +1,9 @@
 # ==========================================================================
 
 #                   Asymptotic model - Hunting success
-#                               Model II
+#                               Model IV
 #               - rank as fixed effect of hunting success
+#               - prey speed and space as fixed effects of a and c
 
 # ==========================================================================
 
@@ -31,7 +32,7 @@ library(parallel)
 # Load data ----------------------------------------------------------------
 
 # Folder path Compute Canada
-folder <- file.path("/home", "maxime11", "projects", "def-monti", 
+folder <- file.path("/home", "maxime11", "projects", "def-monti",
                     "maxime11", "phd_project", "data")
 
 # Load data on compute canada
@@ -40,6 +41,7 @@ data <- fread(
   select = c("predator_id",
              "hunting_success",
              "prey_avg_speed",
+             "prey_avg_amount_tiles_visited",
              "game_duration",
              "cumul_xp_pred",
              "prey_avg_rank")
@@ -61,10 +63,14 @@ standardize <- function(x) {
 
 data[
   ,
-  c("Zprey_speed", "Zgame_duration", "Zcumul_xp", "Zprey_avg_rank") := lapply(
+  c("Zprey_speed", "Zprey_space", "Zgame_duration",
+    "Zprey_avg_rank") := lapply(
     .SD, standardize
   ),
-  .SDcols = c("prey_avg_speed", "game_duration", "cumul_xp_pred", "prey_avg_rank")
+  .SDcols = c(
+    "prey_avg_speed", "prey_avg_amount_tiles_visited",
+    "game_duration", "prey_avg_rank"
+  )
 ]
 
 # ==========================================================================
@@ -84,13 +90,13 @@ data[
 
 model_formula <- brmsformula(
   hunting_success | trials(4) ~
-    a - (a - b) * exp(-exp(c) * Zcumul_xp) +
+    a - (a - b) * exp(-exp(c) * cumul_xp_pred) +
       betaduration * Zgame_duration +
       betarank * Zprey_avg_rank,
 
-  a ~ 1 + (1 | p | predator_id),
+  a ~ 1 + Zprey_speed + Zprey_space + (1 | p | predator_id),
   b ~ 1 + (1 | p | predator_id),
[…]

### 2025-12-11 code/code_models/fit_asym-IV.sh

@@ -3,7 +3,7 @@
 #SBATCH --nodes=1
 #SBATCH --cpus-per-task=32
 #SBATCH --mem=64000M
-#SBATCH --time=00-18:00
+#SBATCH --time=00-24:00
 #SBATCH --mail-user=fraser_franco.maxime@courrier.uqam.ca
 #SBATCH --mail-type=ALL
 

### 2025-12-11 manuscript/manuscript.Rmd
some notes on  intro, improved intro

@@
 
+**to update**
+
 \newpage
@@
 
+**to update**
+
 Keywords: predator-prey interactions, experience, learning, movement, virtual ecology, videogames
@@
 These processes are driven by changes in predation risk and prey capture across landscapes as predators select suitable habitats to hunt and adjust their strategies accordingly  [@quevedo.etal2009;@pettorelli.etal2015;@labarge.etal2024;@schmitz2017].
-Theory suggests that predators become more efficient hunters by acquiring expertise through practicing and learning the proper skills and tactics to locate, select, and capture their prey [@woo.etal2008;@wooster.etal2023;@labarge.etal2024].
+Theory suggests that predators become proficient hunters by acquiring expertise through learning key skills and practicing the tactics required to locate, select, and capture their prey [@woo.etal2008;@wooster.etal2023;@labarge.etal2024].
 This raises the hypothesis that differences in hunting success between predators may be attributed to differences in the capacity to acquire hunting expertise.
-Hence, unraveling how expertise emerges is essential to better assess how foraging behaviour shapes prey consumption during predator-prey interactions.
+Unraveling how this expertise emerges during predator-prey interactions would thus allow for a better assessment of the direct effects of predators on prey.
 
 Expertise can be defined as the characteristics, skills, and knowledge that provide individuals with the ability to outperform novices on complex tasks [@dukas2017].
-Empirical studies on human and non-human hunters show that individuals optimize foraging efficiency (e.g.
-search and handling times, return rates) by associative learning, by developing search images, or by exploiting cues from their prey and their environment [@edwards.jackson1994;@morse2000;@macdonald2007;@reid.etal2010;@wilson-rankin2015].
-Through these processes, expert predators should have greater knowledge, better energy management, and acute motor skills that increase their chances of locating and capturing prey [@dukas2019].
+The acquisition of expertise is measured by quantifying how the success of a skill, such as hunting success, varies as a function of time or repeated attempts [@dukas2019].
+It is usually characterised by a gradual improvement in performance over time [@dukas2008;@rutz2012;@resende.etal2008].
+Different mechanisms can be involded in this process.
+For example, in human and non-human hunters, individuals optimize foraging efficiency over time by associative learning, by developing search images, or by exploiting cues from their prey and their environment [@edwards.jackson1994;@morse2000;@macdonald2007;@reid.etal2010;@wilson-rankin2015].
+When expertise is optimized, expert predators should have greater knowledge, better energy management, and acute motor skills that increase their chances of locating and capturing prey [@dukas2019].
 Differences among predators in these attributes modulate behavioural variation and prey phenotypic composition, contributing to the dynamics of predator-prey systems [@kondoh2010;@skelhornandcandy2016;@kikuchi.etal2023].
@@
 Animal reintroduction programs show that individuals trained to recognize predators display more antipredator behaviours and are 1.5 times more likely to survive than individuals that were never exposed to predators [@tetzalf.etal2019].
-Hence, encountering prey with greater skill or experience may also impair expertise acquisition in predators.
-To our knowledge, the relationships among prey skill, antipredator behaviour, and the acquisition of expertise by human and nonhuman hunters remain unclear, representing a significant gap in our understanding of predator-prey interactions.
+Hence, encountering prey with greater skill at using antipredator tactics may also impair expertise acquisition in predators.
+To our knowledge, the relationships among prey antipredator behaviour and the acquisition of expertise by human and nonhuman predators are poorly studied, representing a significant gap in our understanding of predator-prey interactions.
 
@@
 Here we mitigate these challenges by using *Dead by Daylight* as our study system, a videogame where four prey players must forage for resources while avoiding predation by a fifth player.
-Similar to agent-based simulations, *Dead by Daylight* provides controlled virtual environments to test ecological hypotheses [see @montiglio.etal2025 for a review], with the advantage of having real players that interact in the virtual space.
+*Dead by Daylight* allows us to test ecological hypotheses by providing controlled virtual environments where interactions among thousands of individuals are captured in real time [see @montiglio.etal2025 for a review].
 In this game, the predator population comprises individuals that either ambush or hunt at high speeds (i.e., mean movement speed along an ambush-cursorial continuum of tactics), and their success is driven by the movement of the prey [@fraserfranco.etal2022].
 The prey can increase their chances of survival by cooperating and moving fast to escape the predator [@cere.etal2021;@fraserfranco.etal2022;@santostefano.etal2024].
-The game also elicits natural reactions in players such as freezing when predation is imminent (M.F.F., personal observations), corroborating virtual ecological studies showing that predation drives individual variation in risk perception [@beauchamp2020].
+Moreover, individuals in *Dead by Daylight* and other virtual systems display realistic behaviours in response to the fear of predation, such as freezing (M.F.F., personal observations), hiding [@santostefano.etal2024], or increased vigilance [@beauchamp2020].
 These observations outline how ecological phenomena can emerge from human interactions in virtual systems with fixed rules [@brosnan.etal2017;@kasumovic.etal2017].

### 2025-12-12 code/code_figures/figure1.R
updated code for figures

@@ -1,6 +1,7 @@
 # ==========================================================================
 
-#                           Plot GAMM models
+#                              Figure 1                                    #
+#        Asymptotic models: gain curves (global + random effects)          #
 
 # ==========================================================================
 
@@ -9,10 +10,6 @@
 # ==========================================================================
 # 1. Prepare script
 # ==========================================================================
-
-
-
-# Load libraries and model -------------------------------------------------
 
 options(mc.cores = parallel::detectCores())
 
@@ -23,484 +20,224 @@ library(ggplot2)
 library(ggpubr)
 library(viridis)
 
-path <- file.path(getwd(), "outputs", "outputs_models")
-
-mod2 <- readRDS(file.path(path, "GAMM-II.rds"))
-mod5 <- readRDS(file.path(path, "GAMM-V.rds"))
-mod6 <- readRDS(file.path(path, "GAMM-VI.rds"))
 
 
+# Load models --------------------------------------------------------------
 
-# Load the data ------------------------------------------------------------
+path_models <- file.path(getwd(), "outputs", "outputs_models")
+
+mod2 <- readRDS(file.path(path_models, "asym-II.rds")) # control for rank
+mod3 <- readRDS(file.path(path_models, "asym-III.rds")) # mod2 + Prey speed
+mod4 <- readRDS(file.path(path_models, "asym-IV.rds")) # mod2 + Prey speed + space
+
+
+
+# Load data ---------------------------------------------------------------
 
 data <- fread(
   "./data/FraserFrancoetal2025-data.csv",
-  select = c("predator_id",
-             "game_duration",
-             "pred_speed",
-             "prey_avg_speed",
-             "prey_avg_amount_tiles_visited",
-             "prey_avg_rank",
-             "cumul_xp_pred",
-             "total_xp_pred",
-             "hunting_success")
+  select = c(
+    "predator_id",
+    "game_duration",
+    "pred_speed",
+    "prey_avg_speed",
+    "prey_avg_amount_tiles_visited",
+    "prey_avg_space_rate",
+    "prey_avg_rank",
+    "cumul_xp_pred",
+    "total_xp_pred",
+    "hunting_success"
+  )
 )
 
 data[, predator_id := as.factor(predator_id)]
 
[…]

### 2025-12-12 code/code_figures/figure2.R

@@ -1,10 +1,10 @@
 # ==========================================================================
 
-#               Figure of the relationship between success 
-#                     and prey speed, space, and rank
-#                               Figure 2
+#                                 Figure 2
+#             Plot fixed effects on asymptote and rate parameters
 
 # ==========================================================================
+
 
 
 
@@ -13,39 +13,16 @@
 # ==========================================================================
 
 
-# Load libraries -----------------------------------------------------------
+# Load libraries and model -------------------------------------------------
 
-library(data.table)
 library(brms)
+library(data.table)
 library(ggplot2)
-library(ggpubr)
-library(viridis)
-
-
-
-# Import model -------------------------------------------------------------
+library(ggridges)
 
 path <- file.path(getwd(), "outputs", "outputs_models")
-fit <- readRDS(file.path(path, "GAMM-VI.rds"))
+fit5 <- readRDS(file.path(path, "asym-VII.rds"))
 
-
-
-# Load the data ------------------------------------------------------------
-
-data <- fread(
-  "./data/FraserFrancoetal2025-data.csv",
-  select = c("predator_id",
-             "game_duration",
-             "pred_speed",
-             "prey_avg_speed",
-             "prey_avg_amount_tiles_visited",
-             "prey_avg_rank",
-             "cumul_xp_pred",
-             "total_xp_pred",
-             "hunting_success")
-)
-
-data[, predator_id := as.factor(predator_id)]
 # ==========================================================================
 # ==========================================================================
 
@@ -54,221 +31,275 @@ data[, predator_id := as.factor(predator_id)]
 
 
 # ==========================================================================
-# 2. Compute the predictions
+# 2. Extract posterior draws of predictions into a tidy table
 # ==========================================================================
 
-
-# Predictions --------------------------------------------------------------
-
-preds <- conditional_effects(
-  x = fit,
[…]

### 2025-12-12 code/code_figures/figure3.R

@@ -1,10 +1,10 @@
 # ==========================================================================
 
-#                          Plot regression models
-#                         of prey speed and space
 #                                 Figure 3
+#                   Plot random effect means distributions
 
 # ==========================================================================
+
 
 
 
@@ -13,56 +13,105 @@
 # ==========================================================================
 
 
-
 # Load libraries and model -------------------------------------------------
 
-options(mc.cores = parallel::detectCores())
-
-library(parallel)
 library(brms)
-library(rstan)
 library(data.table)
 library(ggplot2)
-library(viridis)
-library(ggpubr)
+library(ggridges)
 
 path <- file.path(getwd(), "outputs", "outputs_models")
 
-mod_speed <- readRDS(file.path(path, "LM-PreySpeed.rds"))
-mod_space <- readRDS(file.path(path, "LM-PreySpace.rds"))
+fit3 <- readRDS(file.path(path, "asym-II.rds"))
+fit4 <- readRDS(file.path(path, "asym-III.rds"))
+fit5 <- readRDS(file.path(path, "asym-IV.rds"))
+
+# ==========================================================================
+# ==========================================================================
 
 
 
-# Load the data ------------------------------------------------------------
 
-data <- fread(
-  "./data/FraserFrancoetal2025-data.csv",
-  select = c(
-    "predator_id",
-    "game_duration",
-    "prey_avg_speed",
-    "prey_avg_amount_tiles_visited",
-    "prey_avg_rank"
+
+# ==========================================================================
+# 2. Extract posterior draws into a tidy table
+# ==========================================================================
+
+
+# Function to process draws ------------------------------------------------
+
+summarise_param_dt <- function(draws_dt, param) {
+
+  # Global intercept column
+  global_col <- paste0("b_", param, "_Intercept")
+
+  # Random-effect columns for this parameter
+  re_cols <- grep(
+    paste0("^r_predator_id__", param, "\\["),
+    names(draws_dt),
+    value = TRUE
   )
-)
 
-# Predator ID as factor
-data[, predator_id := as.factor(predator_id)]
[…]

### 2025-12-12 code/code_figures/figure4.R

@@ -1,14 +1,14 @@
 # ==========================================================================
 
-#                         Plot GAMM sds parameters
 #                                 Figure 4
+#                 Plot differences in parameters among models
 
 # ==========================================================================
 
 # Figure of the posterior parameters for:
-#- Intercept standard deviation
-#- Slope standard deviation
-#- Wiggliness standard deviation
+#- Baseline
+#- Rate
+#- Asymptote
 
 
 
@@ -26,10 +26,9 @@ library(ggridges)
 
 path <- file.path(getwd(), "outputs", "outputs_models")
 
-fit1 <- readRDS(file.path(path, "GAMM-I.rds"))
-fit2 <- readRDS(file.path(path, "GAMM-II.rds"))
-fit5 <- readRDS(file.path(path, "GAMM-V.rds"))
-fit6 <- readRDS(file.path(path, "GAMM-VI.rds"))
+fit2 <- readRDS(file.path(path, "asym-II.rds"))
+fit3 <- readRDS(file.path(path, "asym-III.rds"))
+fit4 <- readRDS(file.path(path, "asym-IV.rds"))
 
 # ==========================================================================
 # ==========================================================================
@@ -42,72 +41,49 @@ fit6 <- readRDS(file.path(path, "GAMM-VI.rds"))
 # 2. Extract posterior draws into a tidy table
 # ==========================================================================
 
-# Extract posterior draws fit
-post1 <- data.table(
-  as_draws_df(
-    fit1,
-    variable = c(
-      "sds_sZcumul_xppredator_id_1",
-      "sds_sZcumul_xppredator_id_2",
-      "sds_sZcumul_xppredator_id_3"
+# Function to extract posterior draws
+extract_posterior_draws <- function(model, variables) {
+  post <- data.table(
+    as_draws_df(
+      model,
+      variable = variables
     )
   )
+  post
+}
+
+# Variables
+variables <- c(
+  "sd_predator_id__a_Intercept",
+  #"sd_predator_id__b_Intercept",
+  "sd_predator_id__c_Intercept"
 )
 
+models <- list(fit2, fit3, fit4)
 
-# Extract posterior draws fit
-post2 <- data.table(
-  as_draws_df(
-    fit2,
-    variable = c(
-      "sds_sZcumul_xppredator_id_1",
-      "sds_sZcumul_xppredator_id_2",
[…]

### 2025-12-12 code/code_figures/figureS2.R
fig S2 is now regression model of prey

@@ -1,6 +1,8 @@
 # ==========================================================================
 
-#                 Plot GAMM model without rank - Figure S2
+#                          Plot regression models
+#                         of prey speed and space
+#                                Figure S2
 
 # ==========================================================================
 
@@ -18,14 +20,16 @@ options(mc.cores = parallel::detectCores())
 
 library(parallel)
 library(brms)
+library(rstan)
 library(data.table)
 library(ggplot2)
-library(ggpubr)
 library(viridis)
+library(ggpubr)
 
 path <- file.path(getwd(), "outputs", "outputs_models")
 
-mod <- readRDS(file.path(path, "GAMM-I.rds"))
+mod_speed <- readRDS(file.path(path, "LM-PreySpeed.rds"))
+mod_space <- readRDS(file.path(path, "LM-PreySpace.rds"))
 
 
 
@@ -33,18 +37,32 @@ mod <- readRDS(file.path(path, "GAMM-I.rds"))
 
 data <- fread(
   "./data/FraserFrancoetal2025-data.csv",
-  select = c("predator_id",
-             "game_duration",
-             "pred_speed",
-             "prey_avg_speed",
-             "prey_avg_amount_tiles_visited",
-             "prey_avg_rank",
-             "cumul_xp_pred",
-             "total_xp_pred",
-             "hunting_success")
+  select = c(
+    "predator_id",
+    "game_duration",
+    "prey_avg_speed",
+    "prey_avg_amount_tiles_visited",
+    "prey_avg_rank"
+  )
 )
 
+# Predator ID as factor
 data[, predator_id := as.factor(predator_id)]
+
+# Remove any NAs
+data <- data[complete.cases(data)]
+
+
+# Standardise the variables (Z-scores) -------------------------------------
+
+standardize <- function(x) {
+  (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)
+}
+
+data[,
+  c("Zgame_duration", "Zprey_avg_rank") := lapply(.SD, standardize),
+  .SDcols = c(2, 5)
+]
 
 # ==========================================================================
 # ==========================================================================
@@ -61,69 +79,50 @@ data[, predator_id := as.factor(predator_id)]
 
 # Prepare the plots --------------------------------------------------------
 
+# To unscale prey rank
[…]

### 2025-12-12 code/code_models/fit_asym-I.R
remove impossible rates

@@ -39,8 +39,10 @@ data <- fread(
   select = c("predator_id",
              "hunting_success",
              "prey_avg_speed",
+             "prey_avg_space_rate",
              "game_duration",
-             "cumul_xp_pred")
+             "cumul_xp_pred",
+             "prey_avg_rank")
 )
 
 # Predator id as factor
@@ -49,6 +51,8 @@ data[, predator_id := as.factor(predator_id)]
 # Remove any NAs
 data <- data[complete.cases(data)]
 
+# Space rate within 0.2
+data <- data[prey_avg_space_rate <= 0.2]
 
 
 # Standardise the variables (Z-scores) -------------------------------------

### 2025-12-12 code/code_models/fit_asym-II.R

@@ -40,6 +40,7 @@ data <- fread(
   select = c("predator_id",
              "hunting_success",
              "prey_avg_speed",
+             "prey_avg_space_rate",
              "game_duration",
              "cumul_xp_pred",
              "prey_avg_rank")
@@ -50,6 +51,9 @@ data[, predator_id := as.factor(predator_id)]
 
 # Remove any NAs
 data <- data[complete.cases(data)]
+
+# Space rate within 0.2
+data <- data[prey_avg_space_rate <= 0.2]
 
 
 

### 2025-12-12 code/code_models/fit_asym-III.R

@@ -41,6 +41,7 @@ data <- fread(
   select = c("predator_id",
              "hunting_success",
              "prey_avg_speed",
+             "prey_avg_space_rate",
              "game_duration",
              "cumul_xp_pred",
              "prey_avg_rank")
@@ -51,6 +52,9 @@ data[, predator_id := as.factor(predator_id)]
 
 # Remove any NAs
 data <- data[complete.cases(data)]
+
+# Space rate within 0.2
+data <- data[prey_avg_space_rate <= 0.2]
 
 
 

### 2025-12-12 code/code_models/fit_asym-IV.R

@@ -41,7 +41,7 @@ data <- fread(
   select = c("predator_id",
              "hunting_success",
              "prey_avg_speed",
-             "prey_avg_amount_tiles_visited",
+             "prey_avg_space_rate",
              "game_duration",
              "cumul_xp_pred",
              "prey_avg_rank")
@@ -52,6 +52,9 @@ data[, predator_id := as.factor(predator_id)]
 
 # Remove any NAs
 data <- data[complete.cases(data)]
+
+# Space rate within 0.2
+data <- data[prey_avg_space_rate <= 0.2]
 
 
 
@@ -68,7 +71,7 @@ data[
     .SD, standardize
   ),
   .SDcols = c(
-    "prey_avg_speed", "prey_avg_amount_tiles_visited",
+    "prey_avg_speed", "prey_avg_space_rate",
     "game_duration", "prey_avg_rank"
   )
 ]
